### 2024-06-25 manuscript/manuscript.tex
Try to add large files to git hub

\documentclass[a4paper,twoside,10pt]{article}

%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{parskip}

%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}

%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\onehalfspacing       %% 1,5-spacing
\usepackage{placeins}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begingroup  
  \centering
\textbf{BYLOT: Long-term abundance time-series of a High Arctic terrestrial vertebrate community}\\[1.5em]
\endgroup

\textbf{Authors}:\\
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Pierre Legagneux\textsuperscript{3,4}, Gilles Gauthier\textsuperscript{5}, Éliane Duchesne\textsuperscript{1}, Marie-Christine Cadieux \textsuperscript{3}, Dominique Berteaux\textsuperscript{1}, Dominique Fauteux\textsuperscript{6} and Joël Bêty\textsuperscript{1}\\[1.5em]

\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France \\
\textsuperscript{5} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{6} Centre d’études nordiques, Centre de connaissance et d’exploration de l’Arctique, Musée canadien de la nature, Ottawa, ON, Canada
\\[1.5em]

\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]

\section*{Abstract}
Arctic ecosystems present unique opportunities for community-wide monitoring, in part due to their relatively low species richness. However, conducting research in these remote environments poses significant logistical challenges, resulting in long-term monitoring being exceedingly rare. Here, we focus on the long-term, intensive ecological monitoring efforts conducted on the south plain of Bylot Island (almost 400 km\textsuperscript{2}, Nunavut, Canada), which has generated a remarkable dataset spanning up to 30 years, a rarity in tundra ecosystems. Our goal is to synthesize this dataset and upscale vertebrate abundance data at the landscape level, a prerequisite to conduct community-level analyses. We have standardized data obtained with different field methods to provide readily usable long-term time series of abundance for 35 vertebrate species (30 birds and 5 mammals) present in the study system. Monitoring data includes intensive capture-mark-recapture density estimates of lemmings on trapping grids, systematic or opportunistic nest monitoring conducted across the entire study area or within specific plots for all bird species, transects of vertebrate counts distributed throughout the study area, daily incidental observations of vertebrates and satellite tracking of fox movements. Annual abundance of species was estimated at the landscape level, accounting for spatial variations. Furthermore, we provide body masses for each species, derived from empirical onsite measurements for 18 species and from the literature for the remaining species. Body mass is essential to convert species abundance into biomass for studies of trophic fluxes and ecosystem processes. Our dataset provides a unique opportunity for holistic empirical studies of ecological communities, allowing a deeper understanding of community structure and dynamics. Considering that the study site is a pristine and protected area that has experienced minimal anthropogenic impact, it can also provide an ideal baseline for investigating the impacts of global changes on high-latitude terrestrial ecosystems.

\section*{Introduction}
The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology. Community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}. Additionally, community composition is essential for modeling the dynamics of ecological communities. Dynamic community modelling allows addressing important issues and questions in ecology, such as: determining the relative strength of top-down versus bottom-up forces in communities \citep{krebs2003,legagneux2014}, assessing the ecological resilience of communities under climate change \citep{griffith2019} and evaluating the cascading effects of invasive species in food web \citep{david2017, goto2020}. Dynamic community modelling can also be applied to address practical challenges, including fishery management (Plagányi, 2007) and the planning of protected areas \citep{okey2004, dahood2020}.

Modeling ecological communities requires adjusting trophic flows based on species' functional responses, which necessitates time series data on the abundance of all species within a community. However, accurately determining the abundance of all species is rarely achievable. Consequently, empirical community models often reduce taxonomic resolution by grouping species into large functional or taxonomic categories. Additionally, food webs consist of species with varying body sizes depending on their trophic level, with top-level species often being highly mobile and having large home ranges \citep{mccann2005}. Therefore, community models must use landscape-wide estimates of species abundance to accurately represent trophic fluxes. Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.

The composition of ecological communities is influenced by various factors acting at different temporal and spatial scales, leading to noisy data and emphasizing the need for long-term data sets \citep{magurran2010, lindenmayer2012}. Species abundances are influenced by stochastic effects \citep{hubbell2001}, environmental changes (e.g., climate warming), and species interactions, contributing to data variability. For instance, a community composition could be driven simultaneously by the intra-annual seasonal variations in the environment, the multi-year dynamics of small mammals cycles and a relatively slow but directional climate warming. Therefore, to untangle the relative effects of diverse abiotic and biotic factors on community composition, long-term data sets are required \citep{magurran2010, lindenmayer2012}.

Arctic environments are highly valuable systems for studying community patterns and processes due to their relatively low species richness \citep{payer2013, legagneux2014}. However, logistical challenges in the Arctic limit the number of long-term biodiversity monitoring programs. Hence, the small number of Arctic communities with long-term monitoring serve as highly valuable sites for holistic and empirical community studies. Datasets on terrestrial communities are notably scarce, and this scarcity extends to Arctic communities as well \citep{ims2013}.

Within terrestrial Arctic sites, the south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive biodiversity monitoring programs \citep{gauthier2024a}. Monitoring on Bylot Island began in 1989 with a focus on the snow goose and it gradually expanded to other species over time. Currently, the program encompasses all significant vertebrate species in the community with continuous monitoring spanning more than a decade \citep{gauthier2024a}. Monitoring is also conducted at multiple spatial scales, including intensive and systematic observations conducted across a landscape spanning approximately 400 km\textsuperscript{2}. This approach allows the scaling of local density measurements to the landscape level when required and facilitates the estimation of abundance for less common and rare species.

Previous work based on the tundra community of Bylot Island  has already produced several influential papers \citep{gauthier2011, legagneux2012, legagneux2014,hutchison2020, duchesne2021, gauthier2024b}. These studies showed that tundra communities may experience stronger top-down regulation than bottom-up regulation \citep{legagneux2012, legagneux2014}. They also revealed a heterogeneous response of trophic levels to climate warming \citep{gauthier2013} and highlighted the effects of indirect trophic interactions on the occurrence of species across the landscape \citep{duchesne2021}. However, those earlier papers were built on data from relatively short time series, they were not always scaled at the landscape level, and some species or functional groups were lacking abundance estimates. With over a decade of additional community-wide monitoring compared to earlier studies, our goal is to synthesize and upscale the data collected on the Bylot Island community since the 1990s to the landscape level. This synthesis aims to provide readily accessible annual time series (or mean values in some cases) of abundance and biomass for all vertebrate species in a tundra landscape, covering approximately 400 km\textsuperscript{2}.

\begin{figure}[H]
\centering
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island (figure adapted from \citet{gauthier2011}.}
\label{figure:food_web}
\end{figure}

\subsection*{Objective}
Our objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, but excludes non-breeding individuals that stop for only a few days during their migration. Additionally, our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments.
\newpage

\section{Study area}
The study area (389 km\textsuperscript{2}) is located on the south plain of Bylot Island, Nunavut, Canada. It comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%). Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation. The mean temperature in July is 6\textdegree C, and the area typically remains free of snow from mid-June to late September \citep{gauthier2013}. An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (Vulpes vulpes) was ambiguous. While the presence of a breeding pair has been confirmed in the study area \citep{lai2022}, the extent of the population's establishment remains unclear, and sightings are rare, therefore, we decide to exclude this species.

The study area is divided into 9 zones based on the sampling method and the level of field effort applied in each zone (\textbf{Figure \ref{figure:zones}}). Long-term monitoring of the community began in the 1990s in the Qarlikturvik valley \citep{gauthier2013, gauthier2024a}, which represents the zone of the study area with the highest annual sampling effort. Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river, where the main research infrastructure is situated. Another zone with extensive sampling efforts is Camp 2, located at the core of the snow goose colony, where the primary focus is to monitor snow goose nests. However, nests of many other avian species are also monitored within and around the colony in this zone. Camp 3, Pointe Dufour, Goose Point, and Malaview are zones where intensive sampling efforts are conducted annually, albeit for a relatively brief period (approximately one week) during the breeding season of most species \citep{gauthier2024a}. The upland zones in the study area (defined as areas approximately 300 meters above sea level or more) are the Black Plateau, Southern Plateau, and Camp 3 Plateau. These zones are primarily visited to assess raptor nesting activity \citep{beardsell2016}. The zone between the Qarlikturvik valley and Camp 3 received very little sampling effort and is therefore excluded from the study area.

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
   \vspace{-35pt} % Reduce space between figure and caption
   \caption{Map of the different zones (colored polygons) of the 389 km\textsuperscript{2} study area located on the south plain of Bylot Island, Nunavut Canada. The perimeter of the snow goose colony is delineated by white dashes; we highlighted the perimeter in 2017 since it represents the average colony area.}
  \label{figure:zones}
\end{figure}
\newpage

\input{tables/species_name_strategy.tex}
%label-> table:species_name_strategy
\newpage

\section{Data collection}
\subsection{Avian nest monitoring}
Avian nest monitoring was conducted annually except in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic. Systematic nest monitoring refers here to a systematic sampling approach aimed at documenting all nests within a specified area. Monitoring is considered opportunistic when there is a chance that some nests might not have been detected within a specific area. Nest densities derived from nest sampling could be underestimated due to early nest failure (i.e., failure that happened before our sampling period).

\input{tables/species_year_monitoring.tex}
%label-> table:species_year_monitoring
\newpage

\subsubsection{Systematic nest monitoring}
\subsubsection*{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}
Since 2004, systematic searches of wetland areas has been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area. This sampling aimed to find all nests of the cackling goose and the glaucous gull. Nest locations of other large wetland-nesting species, including the tundra swan, the red-throated loon and the Pacific loon, were also noted, as these species nest in the same habitat \citep{duchesne2021,gauthier2024a}. Each year, all known or potential nesting sites were revisited. Observers detected nests by walking and scanning around ponds and lakeshores to identify any active nesting sites. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species given the open landscape.

\subsubsection*{Snow goose}
Snow geese nest in a large colony in the study area (\textbf{Figure \ref{figure:zones}}), but also in small aggregations distributed on the island, especially in years when snowy owls are nesting \citep{lepage1996,reed2002}. Since 1994, goose nests were systematically monitored on a 0.24 km\textsuperscript{2} wetland at the center of the colony. Since 1999, nests were also systematically monitored on a variable number of plots, measuring 0.01 km\textsuperscript{2} in wetland habitat and 0.04 km\textsuperscript{2} in mesic habitat, randomly distributed throughout the goose colony \citep{gauthier2020goose}. The total area covered by the randomly distributed plots averaged 0.79 \textpm{0.37} km\textsuperscript{2} per year. From 2010 onwards, except in 2020 and 2021, we traced the approximate boundary of the goose colony using a GPS receiver aboard a helicopter flying along the colony border \citep{duchesne2021}.

\subsubsection*{Rough-legged hawk, peregrine falcon and common raven}
Peregrine falcons, rough-legged hawks and common ravens nest on cliffs, near ravines, and on large rocky outcrops and tend to reuse the same nesting sites from one year to the next \citep{beardsell2016}. Systematic monitoring of every known or potential nesting site has been carried out in the Qarlikturvik valley, Black plateau and Southern plateau since 2007 and throughout the study area since 2013 \citep{beardsell2016, gauthier2020avianpred}. Observers walked along ridges and scan surrounding areas from vantage points to detect nesting birds. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. They can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species. Each year the observers use slightly different paths to sample the areas, but locate the nests in the same positions, which supports a high probability of detection for these species. Most nesting sites were located in the upland zones of the study area, which are the Black Plateau, Southern Plateau and Camp 3 Plateau.

\subsubsection*{Snowy owl}
Snowy owls predominantly nest in habitats similar to other raptors, favoring ridges in mountainous or hilly regions, although they can occasionally be found nesting on mounds in lowland areas \citep{seyer2020}. Since 1996, searches for snowy owl nests have been conducted concurrently with searches for other raptor nests in the Black and Southern plateaus, as well as during searches for jaeger nests on the southern side of the glacial river in the Qarlikturvik Valley. Additionally, since 2012, nests have been recorded across the entire study area by scanning the landscape from hills and ridges during the nesting period \citep{duchesne2021}. Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.

\subsubsection*{Long-tailed jaeger and parasitic jaeger}
Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period. The aim of those transects was to record nests of long-tailed jaegers, parasitic jaegers, and sandhill cranes. Observers listened for alarm calls to detect territorial birds, and then located nests by observing the birds returning to their nests from elevated vantage points. We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability. We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances.

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/qarlikturvik_valley.pdf}
  \vspace{-70pt} % Reduce space between figure and caption
  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}
 \label{figure:qarlikturvik_valley}
\end{figure}

\subsubsection*{Common-ringed plover}
Between 2015 and 2019, observers conducted surveys of the primary nesting areas of the common-ringed plover. The survey involved walking in stony and sandy shores and gravel bars with scarce vegetation along rivers. Nests were found by detecting individuals exhibiting reproductive behaviors, such as incubation, alarm calls, or distraction displays. The sampling effort was particularly intensive between 2015 and 2017. Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.

\subsubsection*{Lapland longspur and Baird's sandpiper}
Since 2005, nests of passerines and sandpipers have been extensively monitored across an 8 km\textsuperscript{2} (4x2 km) area in the Qarlikturvik valley. We considered the sampling to be most systematic within a core 2 km\textsuperscript{2} (2x1 km) plot in this area (\textbf{Figure \ref{figure:qarlikturvik_valley}}). An observer conducted systematic searches of this plot during the entire breeding season to locate and monitor as many passerine and shorebird nests as possible. Assuming this observer can detect all nests within 5 or 10 meters, this single observer sampled annually at minimum an area of 0.72 \textpm{0.12} or 1.09 \textpm{0.17} km\textsuperscript{2} (n= 3). Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically. We excluded relatively large water bodies (0.26 km\textsuperscript{2}) to calculate nest density in the plot due to the presence of a large lake, which leaves an area of 1.74 km\textsuperscript{2} available for nesting.

\input{tables/interannual_nest_density_variation.tex}
%label-> table:interannual_nest_density_variation
\newpage

\subsubsection{Opportunistic nest monitoring}
Since 2005, we also noted the nest location of any other bird species encountered opportunistically during travel or while carrying out the protocols for the previously described species. The sampling was particularly intensive in the defined 8 km\textsuperscript{2} (2x4 km plot) area in the Qarlikturvik valley. The accuracy of nest monitoring in this plot thus depends on the species detection probability. We are confident to obtain a realistic order of magnitude for the number of nests present for relatively large bodied species in this area (i.e., sandhill crane, rock ptarmigan, long-tailed duck and king eider). Additionally, starting in 2009, a significant effort has been made each year, though not systematically, to visit known nesting territories of parasitic jaegers throughout the study area.

\subsection{Observation of individuals}
\subsubsection{Vertebrate count transects}
From 2010 to 2023, observers walked 500-meter linear transects where all vertebrate individuals observed within 150 meters on either side were counted (146 to 320 transects per year). Transects were distributed across all lowland zones of the study area, typically in mesic habitats, and were carried out during the nesting period (between June 21 and July 14; \citet{duchesne2021}; \textbf{Figure \ref{figure:transects}}). Furthermore, specifically for American golden-plovers, we measured the distance of each observed individual to the transect path.

\begin{figure}
\centering
  \includegraphics[width=0.7\textwidth, angle=0]{figures/transects.pdf}
  \vspace{-35pt} % Reduce space between figure and caption
  \caption{Spatial distribution of vertebrate count transects (white lines) on the south plain of Bylot Island (146 to 320 transects per year).}
 \label{figure:transects}
\end{figure}

\subsubsection{Snow goose point count}
At the start, middle, and end of each vertebrate count transect, a point count with a radius of 125 meters was conducted to determine the number of snow goose breeding pairs. On average, 613 \textpm{142} point counts were sampled each year, covering an area of 30 \textpm{7} km\textsuperscript{2}.

\subsubsection{Incidental observations}
Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}. The number of hours spent in the field served as a proxy for the sampling effort. We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species. However, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:species_relative_abundance_upland}}). Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands.

\input{tables/species_relative_abundance_upland.tex}
%label-> species_relative_abundance_upland
\newpage

\subsubsection{Testimonials of ermine sightings}
There was no direct estimation of ermine abundance on Bylot Island as they are quite difficult to obtain. The density estimates for ermine were derived from an annual abundance index established by \citet{bolduc2023}, which relied on testimonials provided by observers across the whole study area from 1993 to 2019. The testimonials provided by observers were used to create an abundance index ranging from 0 to 3 \citep{bolduc2023}. In this index, a score of 0 corresponds to the absence of ermine sightings, 1 indicates a single sighting of a lone individual, 2 represents multiple sightings of lone individuals, and 3 signifies at least one sighting of a family group. Scores of individual participants were averaged annually as detailed in \citet{bolduc2023}.

\subsection{Capture of individuals}
\subsubsection{Lemming trapping}
Since 2004, brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids. Each grid is made of 144 traps separated by 30 m according to a cartesian plane, one in mesic habitat and the other in wet habitat, located in the Qarlikturvik valley \citep{fauteux2015, gauthier2020lemmings}. Density of each species was estimated at each occasion using spatially explicit capture-recapture methods (see \citet{fauteux2015} for details). From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids 
\citep{gruyer2008}. Index of abundance derived from snap-trapping were transformed in density estimates in each habitat for the period 1995-2003 using the equation provided by \citet{fauteux2018} based on the period of overlap between the two sampling methods (2004 to 2016).

\subsubsection{Arctic fox movement tracking}
Between 2008 and 2016, 109 Arctic foxes were fitted with Argos Platform Transmitter Terminals mounted on collars \citep{lai2015,christin2015, dulude2023}. Foxes were captured between May and August across the study area, within and outside the goose colony \citep{dulude2023}. Sampling of animal locations was set for an interval of 1 or 2 days and only locations between May 1 and October 30 were retained \citep{dulude2023}.

\subsubsection{Parasitic jaeger banding}
In 2009, a significant effort was made to band as many parasitic jaegers as possible within the study area. This effort resulted in the banding of 17 adult individuals (Therrien and Gauthier, unpublished data).
\newpage

\subsection{Species body mass}
All vertebrate individuals captured for marking purposes were systematically weighed (snow goose (unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers\citep{lamarre2021},  common-ringed plovers \citep{leandri2019}, other shorebirds (Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (Gauthier and Therrien, unpublished data), Lapland longspurs (Gauthier and Bêty, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}). Mean body mass of vertebrate species was determined in priority with measurements from Bylot Island. When not available, we extracted mean body mass from the literature \citep{wilman2014}.

\input{tables/body_mass.tex}
%label-> table:body_mass
\newpage

\section{Data preparation}
We estimated the abundance of breeding individuals for most species, but there were a few exceptions. For common ravens, parasitic jaegers, long-tailed ducks, and king eiders, we suspect the presence of a significant number of non-breeding individuals in the study area. Therefore, the estimates we provided for these species include both breeding and potentially non-breeding individuals. Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, ermine, and Arctic hare. The methods used for each species are summarized in (\textbf{Table \ref{table:summary_methods}}).

R software (version 4.3.2) was used to process the data on a linux-gnu machine. Most of data manipulation was carried out with the \textit{dplyr} \citep{dplyr} and \textit{sf} packages \citep{sf}. All codes are available in the Dryad repository XXXXXXXXXXXXXX.
\newpage

\input{tables/table_summary_methods.tex}
%label-> table:summary_methods
\newpage

\subsection{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}
Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022. We transformed the number of nests to abundance of individuals by multiplying it by a factor of two.

We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4). Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12406x + 0.13775$). However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series. Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area  and the indices of relative abundance (i.e., daily incidental observations and transects). We thus fitted an exponential model using the number of nests found annually over two distinct periods: in 1996 when the first nest was discovered, and then from 2017 to 2023 when sampling effort was systematic across the whole study area (\textbf{Figure \ref{figure:cackling}}). We used the fitted model to estimate abundance between 1996 and 2016 when monitoring was less systematic, which could potentially underestimate observed abundance as seen on \textbf{Figure \ref{figure:cackling}}. We multiplied nest density by two to obtain the abundance.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
\caption{Number of cackling goose nests found across the study area over time. The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016. The model presents a strong fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area (R\textsuperscript{2}= 0.74, p= 0.15, n= 5).}
\label{figure:cackling}
\end{figure}

\subsection{Snow goose}
Between 1999 and 2023, we assessed the abundance of snow geese in the study area through a multi-step process. We calculated the mean annual density of snow goose nests separately in the mesic and wetland habitats of the area occupied by the goose colony annually. We made slight adjustments to the goose colony perimeter defined from helicopter surveys to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}). To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled. The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts. Systematic nest searches were highly precise, but covered a relatively small area, whereas transects and snow goose point counts were less precise but covered larger areas. For each method, we calculated the mean density of breeding individuals in mesic habitat by dividing the number of birds (or nests) recorded by the area sampled. Despite methodological differences, the three approaches showed similar inter-annual variations, supporting the use of a mean values to estimate nest density in mesic habitat \textbf{Figure \ref{figure:sngo_interannual_variation}}). Lastly, to transform the densities in total abundance, we determined the annual proportion of wetland and mesic habitats within the goose colony and multiplied the area of each habitat by the density of breeding individuals. For the period 1999 to 2009, we used the average limits of the colony over the period 2010 to 2023 because we did not conduct aerial survey of the colony. Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\linewidth]{figures/map_geese_colony_2017.pdf}
\caption{Map showing the region occupied by the snow goose colony in 2017 (green polygon) as an example. The perimeter was first defined from a helicopter surveys and then slightly adjusted to include all snow goose point counts where at least one breeding pair had been observed in that year (green dots). Snow goose point counts where no breeding geese were observed in that year are presented as red dots.}
\label{figure:sngo_goose_colony}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figures/snow_goose_mesic_density.pdf}
\caption{Estimates of breeding goose density in mesic habitat of the Bylot Island study area using three independent methods: Nest sampling plot, Snow goose point count, Vertebrate count transect.}
\label{figure:sngo_interannual_variation}
\end{figure}

In 1993, 1998, 2003, and 2008, the Canadian Wildlife Service conducted aerial surveys over the southern plain of Bylot Island during the brood-rearing period to estimate the abundance of breeding snow geese (Reed et al., 2002). We corrected these estimates with the annual nesting success, as these surveys were conducted after the hatching period. For 2003 and 2008, we used in priority estimates of abundance derived from aerial surveys instead of nest sampling.

\subsection{King eider and long-tailed duck}
We estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley. We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}). We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two. We acknowledge that the opportunistic monitoring of these species likely underestimated their true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident to obtain a realistic order of magnitude for the number of nests present. Because duck sightings are frequent, yet only a few nests are found, we believe there may be non-breeding individuals. Therefore, we employed an additional method to estimate the overall duck populations without differentiating between breeding and non-breeding individuals.

As an alternative approach, we estimated the abundance of ducks based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) in duck and loon species. We therefore derived the absolute abundance of long-tailed ducks and king eiders from their relative abundances using the ratio between relative and absolute abundances of red-throated loons as a reference.

\subsection{Rough-legged hawk, peregrine falcon and snowy owl}
We estimated the abundance of breeding rough-legged hawks, peregrine falcons and snowy owls based on systematic nest monitoring conducted throughout the study area for these species. To convert the number of nests into breeding abundance, we multiplied it by two. For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.99; p < 0.0001, n= 10). We used the same approach for rough-legged hawks ($y=0.49851x$, R\textsuperscript{2}= 0.99, p < 0.0001, n= 8) to extend the time series from 2007 to 2012. We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.44, p= 0.27, n=8).

\subsection{Rock ptarmigan}
We estimated the abundance of rock ptarmigans based on the annual nest density measured in the 8 km\textsuperscript{2} extensive nest search area of the Qarlikturvik valley. While we acknowledge that the opportunistic monitoring of this species likely underestimates nest density, the extensive sampling effort deployed annually within this area gives us confidence in obtaining a realistic number of nests. We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:species_relative_abundance_upland}}). Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat. To convert the number of nests into breeding abundance, we multiplied it by two.

\subsection{Sandhill crane}
We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}). In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed. We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present. Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2. We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).

\begin{figure}[H]
\centering
\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values.}
\label{figure:sacr}
\end{figure}

\subsection{American golden-plover and black-bellied plover}
We applied a distance sampling approach to the observations of American golden-plovers made along vertebrate count transects to estimate the abundance of individuals in the mesic habitat of the study area between 2014 and 2023. We used perpendicular distance between detected individuals and transect (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}. We applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered). We selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term. We excluded observations of more than four individuals, as these likely indicated groups of non-breeders passing through the area. We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}. We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).

To determine the abundance of black-bellied plovers, we used the mean number of black-bellied plovers and American golden-plovers observed per transect as an index of relative abundance. We assumed that the ratios of between relative and actual abundance are the same (i.e., similar detection probability) among those species. This assumption is realistic as those species present similarities in size, color, and reproductive behavior. We therefore derived the absolute abundance of black-bellied plovers from their relative abundance using the ratio between relative and absolute abundances of American golden-plover as a reference.

As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.

\subsection{Common-ringed plover}
To estimate the abundance of common-ringed plovers in the study area, we relied on the total number of nests recorded annually from 2015 to 2017, during which the primary nesting sites underwent intensive sampling. We multiplied the total nest count by two to represent the abundance of breeding individuals.

\subsection{Lapland longspur and Baird's sandpiper}
We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}). For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected. We made this adjustment because this species was less frequently observed. In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located. This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically. Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2. We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}). We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.

\begin{figure}[H]
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.95\linewidth]{figures/lapland_longspur_nest_transect.pdf} 
\caption{}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=0.95\linewidth]{figures/bairds_sandpiper_nest_transect.pdf}
\caption{}
\end{subfigure}
\caption{a) Linear regression between the nest density of Lapland longspurs and the number of individuals observed per transect (nest density= 2.3422 x number of individuals per transect; regression was forced through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values. b) Linear regression between the nest density of Baird’s sandpiper and the proportion of transect with at least one individual observed (nest density= 34.9248 x proportion of transects with at least one individual; regression was forced through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values.}
\label{figure:lalo_basa}
\end{figure}

\subsection{Other passerines and sandpipers}
We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section 3.9). We assumed here a similar detection probability for all species. We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section 3.9). For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones. This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}). Nest density was then converted in number of individuals by multiplying by the area and a factor 2.

As an alternative approach, we estimated the abundance of other passerines and sandpipers based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among both passerine and sandpiper species. We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.

\subsection{Long-tailed jaeger}
We determined the annual nest density of long-tailed jaegers from the systematic nest sampling between 2004 and 2023 on the southern side of the glacial river in the Qarlikturvik valley. We determined nest density by dividing the annual number of nests recorded by the area of the surveyed zone (33 km\textsuperscript{2}). As long-tailed jaegers typically nest in mesic habitat \citep{andersson1971}, we multiplied the area occupied by mesic habitat across the study area by the nest density obtained in the surveyed zone and by two to obtain the total abundance of breeding individuals.

\subsection{Parasitic jaeger}
Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 3.75 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}. This suggests that there may be non-breeding individuals present at the study site, or alternatively, individuals may regularly travel long distances, potentially from outside the study area, to forage during the breeding season. Due to limited data availability for estimating the abundance of non-breeding parasitic jaegers, we relied on the maximum number of adults banded during a single year (17 individuals in 2009; Therrien, unpublished data) as the minimum abundance on the study area. This corresponds to a density of 0.04 individuals/km\textsuperscript{2}. For comparison, \citet{taylor1974} measured a density of 0.06 individual/km\textsuperscript{2} on Bathurst Island.

\subsection{Common raven}
Despite the intensive nest searches for raptors in upland zones, we never found more than one common raven nest each year, a small number considering the frequent raven observations at the study site \citep{gauthier2024a}. This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging. Therefore, we opted for alternative approaches based on individual counts to estimate the abundance of both breeding and non-breeding ravens.

As a first approach, we based our estimate of ravens on the number of glaucous gulls observed per transect. We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species. This assumption is reasonable as those species present similarities in size and foraging strategy. We therefore derived the absolute abundance of common ravens from their relative abundance using the ratio between relative and absolute abundances of glaucous gulls as a reference.

Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.

\subsection{Brown and collared lemming}
Between 1995 and 2003, we used the density estimates derived from the snap-trapping indices obtained in late July in each habitat. Between 2004 and 2007, annual abundance of each lemming species was based on the late-July density estimates on trapping grid in wet and mesic habitats. However, starting from 2008, estimates were derived from the mean density recorded in mid-July and mid-August, except for two instances: 2019 and 2021. In 2019, due to an exceptionally early snowmelt and thus an early decline in lemmings during the summer, we only retained value from mid-July. In 2021, we relied solely on data gathered in August because it was the only trapping period carried out that year. To scale the estimated densities from the wet and mesic grids to the entire study area, we used the proportions of mesic habitats (64\%) and wet habitats (10\%) measured within the study area.

\subsection{Arctic hare}
Arctic hares are primarily observed in the upland zones of the study area, where sampling effort is limited. We thus derived abundance of hares from the estimated abundance of Arctic foxes based on indices of relative abundance presented in \citep{gauthier2024a}, which were derived from incidental daily observations. We doubled the density of Arctic hares in the upland zones (i.e., plateaus), as twice as many individuals were observed per hour of fieldwork there compared to lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}). However, it is worth noting that assuming a similar detection probability between foxes and hares might lead to an overestimation of hare detection probability due to behavioral differences between the species. Therefore, we most likely underestimate the actual abundance of Arctic hares in the study area.

\subsection{Ermine}
We estimated the annual abundance of ermines by transforming the annual index of relative abundance provided in \citet{bolduc2023} into individual density. Annual values ranged from 0, indicating no ermine sighting, to 2.88, which signifies that nearly all observers observed at least one family group during their field season. We independtly obtained measures of minimum (0.02 ind./km\textsuperscript{2}) and maximum (0.4 ind./km\textsuperscript{2}) ermine density, which were determined from estimates of individual home range obtained from radio-tracking data, observations on Bylot Island, and existing literature \citep{legagneux2012, bilodeau2013}. We associated the minimum and maximum scores of relative abundance with the minimum and maximum density of individuals, respectively. Ultimately, we calculated the ermine density by linearly interpolating between these two density extremes using the annual index of relative abundance.

\subsection{Arctic fox} 
We estimated the abundance of Arctic foxes in the study area based on their estimated home range size inside and outside the goose colony. We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size. However, here, we did not account for annual variations in lemming density as presented in \citet{dulude2023} in order to obtain mean fox home range. Given that foxes are territorial and exhibit an average spatial overlap with adjacent territories of 18\% \citep{clermont2021}, we converted home range size into individual density using the following formula: $\text{\textit{density of individuals}}= \frac{2}{0.82\times \text{\textit{home range}}}$. We used two as numerator because we assumed each territory was held by a pair of fox, either breeding or non-breeding fox pair, without accounting for potential nomadic or transient individuals. We used values of 12.26 km\textsuperscript{2} to represent the mean home range of foxes within the goose colony and 20.02 km\textsuperscript{2} for foxes outside the goose colony. We estimated the mean density of foxes in each zone of the study area according to the mean proportion of the zone covered by the goose colony. We derived the mean annual proportion of each zone covered by the goose colony from the colony outline between 2010 and 2023. We estimated a mean density of 0.14 individuals/km\textsuperscript{2} for the study area. Previously, the minimum density of foxes in the study area was estimated to be between 0.03 and 0.13 individuals per km\textsuperscript{2} based on camera traps \citep{royerboutin2015}. 
\newpage

\section{Species abundance}

\subsection{Uncertainty on estimates}
We did not quantitatively assessed uncertainty on each estimate of species abundance. However, we provide a qualitative measure of uncertainty based on the level of confidence derived from field experts, and in some cases, the fit of statistical models to estimate density (\textbf{Table \ref{table:confidence_estimate_abundance}}).

\input{tables/confidence_estimate_abundance.tex}
%label -> table:confidence_estimate_abundance
\newpage

\subsection{Abundance estimates}
\input{tables/species_mean_abudance.tex}
%label -> table:species_mean_abudance
\newpage

The following figures present the time series of species abundances. If different methods were used, we present the estimates derived from each method with different colors (see \textbf{Table \ref{table:confidence_estimate_abundance}} for the corresponding method and confidence in estimates).

\begin{figure}[h]
  \centering
  \begin{subfigure}{0.35\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/legend.pdf}
  \end{subfigure} 
  \begin{subfigure}{0.40\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/brown_lemming.pdf}
  \end{subfigure}
    \hfill
    \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/collared_lemming.pdf}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\linewidth]{figures/species_temporal_series/ermine.pdf}
  \end{subfigure}
\end{figure}
\newpage

\begin{figure}[ht]
\centering
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/snowy_owl.pdf}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/rough-legged_hawk.pdf}
\end{subfigure}
\hfill
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/peregrine_falcon.pdf}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/glaucous_gull.pdf}
\end{subfigure}
\hfill
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/long-tailed_jaeger.pdf}
\end{subfigure}
\end{figure}
\newpage

\begin{figure}[H]
\centering
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/snow_goose.pdf}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/cackling_goose.pdf}
\end{subfigure}
\hfill
\begin{subfigure}{0.45\textwidth}
\includegraphics[width=\linewidth]{figures/species_temporal_series/American_golden-plover.pdf}
\end{subfigure}
\label{figure:other}
\end{figure} 
\newpage

\bibliography{references}

\end{document}

### 2024-06-25 manuscript/tables/confidence_estimate_abundance.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:29 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{llll}
\caption{Confidence level on the various estimates of annual or mean abundance of vertebrate species on Bylot Island.} \\ 
  \hline
Species & Estimate & Confidence & Reason \\ 
  \hline
king eider & mean & moderate & Intensive, but opportunistic nest monitoring (8 km2) \\ 
  long-tailed duck & mean & moderate & Intensive, but opportunistic nest monitoring (8 km2) \\ 
  Pacific loon & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  red-throated loon & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  cackling goose & 2004-2016 & moderate & \makecell[tl]{Extrapolation from exponential model of growth (R2=0.74,\\p=0.15, n=5)} \\ 
  cackling goose & 2017-2019, 2022-2023 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  snow goose & 1999-2009 & low & \makecell[tl]{Density based only on nest monitoring and uncertainty\\on goose colony area (mean)} \\ 
  snow goose & 2010-2019, 2022-2023 & high & \makecell[tl]{Intensive study area-wide monitoring based on several\\methods and annual colony outline} \\ 
  tundra swan & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  rough-legged hawk & 2007-2012 & high & \makecell[tl]{Extrapolation from intensive nest monitoring (111 km2,\\R2=0.99, p<0.0001, n=8)} \\ 
  rough-legged hawk & 2013-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  peregrine falcon & 2013-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  snowy owl & 1996-2011 & high & \makecell[tl]{Extrapolation from intensive nest monitoring (111 km2,\\R2=0.99, p<0.0001, n=10)} \\ 
  snowy owl & 2012-2019, 2022-2023 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  rock ptarmigan & mean & low & \makecell[tl]{Intensive, but opportunistic nest monitoring (8 km2)\\and prime nesting habitat not well sampled} \\ 
  sandhill crane & mean & moderate & \makecell[tl]{Extrapolation from intensive nest monitoring (8 km2)\\and uncertain relation with large scale indices} \\ 
  American golden-plover & 2014-2019, 2022-2023 & high & Distance sampling throughout lowland (313 km2) \\ 
  black-bellied plover & mean & moderate & Derived from the abundance estimate of another species \\ 
  common-ringed plover & 2015-2017 & moderate & Nest monitoring on the main breeding sites \\ 
  ruddy turnstone & mean & low & Derived from the abundance estimate of another species \\ 
  red knot & mean & low & Derived from the abundance estimate of another species \\ 
  pectoral sandpiper & mean & low & Derived from the abundance estimate of another species \\ 
  Baird's sandpiper & mean & moderate & \makecell[tl]{Extrapolation from intensive nest monitoring (2 km2)\\and uncertain relation with large scale indices} \\ 
  white-rumped sandpiper & mean & low & Derived from the abundance estimate of another species \\ 
  buff-breasted sandpiper & mean & low & Derived from the abundance estimate of another species \\ 
  red phalarope & mean & low & Derived from the abundance estimate of another species \\ 
  glaucous gull & 2004-2016 & high & \makecell[tl]{Extrapolation from intensive nest monitoring (111 km2,\\R2=0.84, p=0.16, n=4)} \\ 
  glaucous gull & 2017-2019, 2022 & high & Intensive study area-wide nest monitoring (389 km2) \\ 
  long-tailed jaeger & 2004-2019, 2022 & high & \makecell[tl]{Extrapolation based on intensive nest monitoring (33\\km2)} \\ 
  parasitic jaeger & mean & low & \makecell[tl]{Opportunistic nest monitoring and uncertainty about number\\of non-breeding individuals } \\ 
  common raven & mean & low & \makecell[tl]{Low density and uncertainty about number of non-breeding\\individuals} \\ 
  horned lark & mean & moderate & Derived from the abundance estimate of another species \\ 
  American pipit & mean & low & \makecell[tl]{Derived from the abundance estimate of another species\\and prime nesting habitat not well sampled} \\ 
  Lapland longspur & mean & moderate & \makecell[tl]{Extrapolation from intensive nest monitoring (2 km2)\\and uncertain relation with large scale indices} \\ 
  snow bunting & mean & low & \makecell[tl]{Derived from the abundance estimate of another species\\and prime nesting habitat not well sampled} \\ 
  brown lemming & 1995-2019, 2021-2022 & moderate & \makecell[tl]{Rigorous density estimates but at a small spatial scale\\(0.22 km2)} \\ 
  collared lemming & 1995-2019, 2021-2022 & moderate & \makecell[tl]{Rigorous density estimates but at a small spatial scale\\(0.22 km2)} \\ 
  Arctic hare & mean & low & Prime habitat not well sampled \\ 
  ermine & 1993-2019 & moderate & \makecell[tl]{Based on indirect indices and uncertainty on ermine home\\range size estimates} \\ 
  Arctic fox & mean & moderate & Derived from extensive fox home range size studies (n=109) \\ 
   \hline
\hline
\label{table:confidence_estimate_abundance}
\end{longtable}
\endgroup

### 2024-06-25 manuscript/tables/interannual_nest_density_variation.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:27 2024
\begin{table}[ht]
\centering
\caption{Annual nest density (nests/km2) of selected avian species estimated on different zones of Bylot Island.} 
\label{table:interannual_nest_density_variation}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.9\textwidth}{lllr}
  \hline
Species & Zone & Mean ± SD & Number of years \\ 
  \hline
Baird's sandpiper & qarlikturvik (2x1 km plot) & 5.000 ± 3.558 &  17 \\ 
  Lapland longspur & qarlikturvik (2x1 km plot) & 13.559 ± 5.849 &  17 \\ 
  king eider & qarlikturvik (4x2 km plot) & 0.115 ± 0.138 &  16 \\ 
  long-tailed duck & qarlikturvik (4x2 km plot) & 0.092 ± 0.138 &  16 \\ 
  rock ptarmigan & qarlikturvik (4x2 km plot) & 0.031 ± 0.055 &  16 \\ 
  sandhill crane & qarlikturvik (4x2 km plot) & 0.161 ± 0.087 &  16 \\ 
  long-tailed jaeger & qarlikturvik valley & 0.362 ± 0.380 &  17 \\ 
  Pacific loon & whole study area & 0.005 ± 0.004 &   4 \\ 
  red-throated loon & whole study area & 0.082 ± 0.019 &   4 \\ 
  cackling goose & whole study area & 0.177 ± 0.064 &   5 \\ 
  tundra swan & whole study area & 0.001 ± 0.001 &   4 \\ 
  rough-legged hawk & whole study area & 0.157 ± 0.151 &   8 \\ 
  peregrine falcon & whole study area & 0.053 ± 0.007 &   8 \\ 
  snowy owl & whole study area & 0.022 ± 0.058 &  10 \\ 
  common ringed plover & whole study area & 0.070 ± 0.012 &   3 \\ 
  glaucous gull & whole study area & 0.091 ± 0.011 &   4 \\ 
  parasitic jaeger & whole study area & 0.010 ± 0.004 &  12 \\ 
  common raven & whole study area & 0.003 ± 0.003 &   8 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-06-25 manuscript/tables/species_mean_abudance.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:29 2024
\begin{table}[ht]
\centering
\caption{Mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2). In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty.} 
\label{table:species_mean_abudance}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllrrr}
  \hline
Species & Method & Mean ± sd & n & Minimum & Maximum \\ 
  \hline
king eider & nest sampling (extrapolation habitat) & 25 &  &  &  \\ 
  king eider & incidental observations (relative abundance) & 106 &  &  &  \\ 
  long-tailed duck & nest sampling (extrapolation habitat) & 20 &  &  &  \\ 
  long-tailed duck & incidental observations (relative abundance) & 191 &  &  &  \\ 
  Pacific loon & nest sampling & 4 ± 3 &   4 &   0 &   6 \\ 
  red-throated loon & nest sampling & 64 ± 15 &   4 &  42 &  76 \\ 
  cackling goose & nest sampling & 50 ± 59 &  28 &   2 & 214 \\ 
  snow goose & combined methods & 25289 ± 13327 &  23 & 2505 & 49076 \\ 
  tundra swan & nest sampling & 1 ± 1 &   4 &   0 &   2 \\ 
  rough-legged hawk & nest sampling & 25 ± 24 &  16 &   0 &  66 \\ 
  peregrine falcon & nest sampling & 10 ± 1 &   8 &   8 &  12 \\ 
  snowy owl & nest sampling & 16 ± 33 &  26 &   0 & 144 \\ 
  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  &  \\ 
  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  &  \\ 
  American golden-plover & distance sampling & 774 ± 275 &   8 & 306 & 1040 \\ 
  black-bellied plover & transects (relative abundance) & 22 &  &  &  \\ 
  black-bellied plover & incidental observations (relative abundance) & 61 &  &  &  \\ 
  common-ringed plover & nest sampling & 55 ± 9 &   3 &  44 &  62 \\ 
  ruddy turnstone & transects (relative abundance) & 40 &  &  &  \\ 
  ruddy turnstone & incidental observations (relative abundance) & 54 &  &  &  \\ 
  red knot & transects (relative abundance) & 64 &  &  &  \\ 
  red knot & incidental observations (relative abundance) & 237 &  &  &  \\ 
  pectoral sandpiper & transects (relative abundance) & 79 &  &  &  \\ 
  pectoral sandpiper & incidental observations (relative abundance) & 260 &  &  &  \\ 
  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  &  \\ 
  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  &  \\ 
  white-rumped sandpiper & incidental observations (relative abundance) & 1156 &  &  &  \\ 
  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  &  \\ 
  buff-breasted sandpiper & incidental observations (relative abundance) & 8 &  &  &  \\ 
  red phalarope & transects (relative abundance) & 156 &  &  &  \\ 
  red phalarope & incidental observations (relative abundance) & 276 &  &  &  \\ 
  glaucous gull & nest sampling & 73 ± 7 &  17 &  59 &  80 \\ 
  long-tailed jaeger & nest sampling & 272 ± 285 &  17 &   0 & 900 \\ 
  parasitic jaeger & banding & 17 &  &  &  \\ 
  parasitic jaeger & nest sampling & 8 &  &  &  \\ 
  common raven & transects (relative abundance) & 14 &  &  &  \\ 
  common raven & incidental observations (relative abundance) & 31 &  &  &  \\ 
  horned lark & transects (relative abundance) & 362 &  &  &  \\ 
  horned lark & incidental observations (relative abundance) & 411 &  &  &  \\ 
  American pipit & transects (relative abundance) & 53 &  &  &  \\ 
  American pipit & incidental observations (relative abundance) & 87 &  &  &  \\ 
  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  &  \\ 
  snow bunting & transects (relative abundance) & 18 &  &  &  \\ 
  snow bunting & incidental observations (relative abundance) & 276 &  &  &  \\ 
  brown lemming & trapping & 54043 ± 93530 &  27 &   0 & 447630 \\ 
  collared lemming & trapping & 8128 ± 10334 &  27 &   0 & 39302 \\ 
  Arctic hare & incidental observations (relative abundance) & 6 &  &  &  \\ 
  ermine & testimonials & 40 ± 37 &  27 &   8 & 156 \\ 
  Arctic fox & individual home range & 53 &  &  &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-06-25 manuscript/tables/species_name_strategy.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:25 2024
\begin{table}[ht]
\centering
\caption{Species composition of the vertebrate community of Bylot Island with the corresponding annual cycle strategy (i.e., resident, partial migrant, migrant).} 
\label{table:species_name_strategy}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.95\textwidth}{llll}
  \hline
Functional group & Scientific name & Name & Annual cycle strategy \\ 
  \hline
ducks and loons  & Somateria spectabilis & king eider & migrant \\ 
  ducks and loons  & Clangula hyemalis & long-tailed duck & migrant \\ 
  ducks and loons  & Gavia pacifica & Pacific loon & migrant \\ 
  ducks and loons  & Gavia stellata & red-throated loon & migrant \\ 
  geese and swans  & Branta hutchinsii & cackling goose & migrant \\ 
  geese and swans  & Anser caerulescens & snow goose & migrant \\ 
  geese and swans  & Cygnus columbianus & tundra swan & migrant \\ 
  raptors  & Buteo lagopus & rough-legged hawk & migrant \\ 
  raptors  & Falco peregrinus & peregrine falcon & migrant \\ 
  raptors  & Bubo scandiacus & snowy owl & migrant \\ 
  ptarmigans  & Lagopus muta & rock ptarmigan & resident \\ 
  cranes  & Antigone canadensis & sandhill crane & migrant \\ 
  shorebirds  & Pluvialis dominica & American golden-plover & migrant \\ 
  shorebirds  & Pluvialis squatarola & black-bellied plover & migrant \\ 
  shorebirds  & Charadrius hiaticula & common-ringed plover & migrant \\ 
  shorebirds  & Arenaria interpres & ruddy turnstone & migrant \\ 
  shorebirds  & Calidris canutus & red knot & migrant \\ 
  shorebirds  & Calidris melanotos & pectoral sandpiper & migrant \\ 
  shorebirds  & Calidris bairdii & Baird's sandpiper & migrant \\ 
  shorebirds  & Calidris fuscicollis & white-rumped sandpiper & migrant \\ 
  shorebirds  & Calidris subruficollis & buff-breasted sandpiper & migrant \\ 
  shorebirds  & Phalaropus fulicarius & red phalarope & migrant \\ 
  gulls  & Larus hyperboreus & glaucous gull & migrant \\ 
  jaegers  & Stercorarius longicaudus & long-tailed jaeger & migrant \\ 
  jaegers  & Stercorarius parasiticus & parasitic jaeger & migrant \\ 
  ravens  & Corvus corax & common raven & partial migrant \\ 
  passerines  & Eremophila alpestris & horned lark & migrant \\ 
  passerines  & Anthus rubescens & American pipit & migrant \\ 
  passerines  & Calcarius lapponicus & Lapland longspur & migrant \\ 
  passerines  & Plectrophenax nivalis & snow bunting & migrant \\ 
  lemmings  & Lemmus trimucronatus & brown lemming & resident \\ 
  lemmings  & Dicrostonyx groenlandicus & collared lemming & resident \\ 
  hares  & Lepus arcticus & Arctic hare & resident \\ 
  ermines  & Mustela richardsonii & ermine & resident \\ 
  foxes  & Vulpes lagopus & Arctic fox & partial migrant \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-06-25 manuscript/tables/species_relative_abundance_upland.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:27 2024
\begin{table}[ht]
\centering
\caption{Index of relative abundance (i.e., number of individuals observed per hour) derived from incidental daily observations for selected vertebrate species in lowland (i.e., Qarlikturvik valley, Camp 3, Malaview, Camp 2, Goose point and Pointe Dufour) and upland (i.e., Black plateau, Southern plateau and Camp 3 plateau) zones of the Bylot Island study area. The ratio compares relative abundance indexes between these two types of zones, calculated by dividing the upland by the lowland index of relative abundance.} 
\label{table:species_relative_abundance_upland}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.55\textwidth}{lrll}
  \hline
  & \multicolumn{2}{c}{Individuals/hour} \\
 Species & Upland & Lowland & Ratio \\
 \hline
rock ptarmigan & 0.03 & 0.03 & 1 \\ 
  sandhill crane & 0.42 & 0.287 & 1.5 \\ 
  American golden-plover & 0.26 & 0.394 & 0.7 \\ 
  black-bellied plover & 0.02 & 0.032 & 0.6 \\ 
  ruddy turnstone & 0.01 & 0.007 & 1.3 \\ 
  red knot & 0.00 & 0.033 & 0 \\ 
  pectoral sandpiper & 0.02 & 0.034 & 0.6 \\ 
  Baird's sandpiper & 0.31 & 0.32 & 1 \\ 
  white-rumped sandpiper & 0.04 & 0.137 & 0.3 \\ 
  buff-breasted sandpiper & 0.00 & 0.001 & 0 \\ 
  red phalarope & 0.01 & 0.038 & 0.2 \\ 
  horned lark & 0.24 & 0.154 & 1.6 \\ 
  American pipit & 0.34 & 0.024 & 14.2 \\ 
  snow bunting & 0.59 & 0.092 & 6.4 \\ 
  Arctic hare & 0.02 & 0.009 & 2 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-06-25 manuscript/tables/species_year_monitoring.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:26 2024
\begin{table}[ht]
\centering
\caption{Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in in Gauthier et al. (2024b).} 
\label{table:species_year_monitoring}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
Species & Zone & Years & Monitoring & References \\ 
  \hline
snow goose & camp 2 & 1999-2019, 2023 & systematic & Gauthier et Cadieux, 2020 \\ 
  rough-legged hawk & qarlikt., black \& south plat. & 2007-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  peregrine falcon & qarlikt., black \& south plat. & 2007-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  snowy owl & qarlikt., black \& south plat. & 1996-2019, 2023 & systematic & Gauthier et al., 2020 \\ 
  Baird's sandpiper & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & systematic & Bêty, 2020 \\ 
  Lapland longspur & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & systematic & Gauthier and Bêty, 2020 \\ 
  king eider & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  long-tailed duck & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  rock ptarmigan & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  sandhill crane & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & opportunistic & unpublished data \\ 
  brown lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & systematic & Gauthier, 2020 \\ 
  collared lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & systematic & Gauthier, 2020 \\ 
  Pacific loon & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  red-throated loon & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  cackling goose & qarlikturvik valley & 2004-2019, 2022-2023 & systematic & unpublished data \\ 
  tundra swan & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  glaucous gull & qarlikturvik valley & 2004-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  long-tailed jaeger & qarlikturvik valley & 2004-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  parasitic jaeger & qarlikturvik valley & 2004-2019, 2022 & systematic & unpublished data \\ 
  Pacific loon & whole study area & 2017-2019, 2022 & systematic & Duchesne et al., 2021 \\ 
  red-throated loon & whole study area & 2017-2019, 2022 & systematic & Duchesne et al., 2021 \\ 
  cackling goose & whole study area & 2017-2019, 2022-2023 & systematic & Duchesne et al., 2021 \\ 
  tundra swan & whole study area & 2017-2019, 2022 & systematic & Duchesne et al., 2021 \\ 
  rough-legged hawk & whole study area & 2013-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  peregrine falcon & whole study area & 2013-2019, 2022  & systematic & Gauthier et al., 2020 \\ 
  snowy owl & whole study area & 2012-2019, 2022-2023 & systematic & Gauthier et al., 2020 \\ 
  common-ringed plover & whole study area & 2015-2017 & systematic & Bêty, 2020 \\ 
  glaucous gull & whole study area & 2017-2019, 2022 & systematic & Gauthier et al., 2020 \\ 
  parasitic jaeger & whole study area & 2009-2019, 2022 & opportunistic & unpublished data \\ 
  common raven & whole study area & 2013-2019, 2022 & systematic & unpublished data \\ 
  ermine & whole study area & 1993-2019 & opportunistic & Bolduc et al., 2023 \\ 
  Arctic fox & whole study area & 2008-2016 & systematic & Dulude-de Broin et al., 2023 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-06-25 manuscript/tables/table_summary_methods.tex
Try to add large files to git hub

% latex table generated in R 4.4.0 by xtable 1.8-4 package
% Tue Jun 25 10:25:27 2024
\begin{table}[H]
\centering
\caption{Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapoated at the landscape scale.} 
\label{table:summary_methods}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
Species & Sampling method & Reference species & Sampling scale & Spatial extrapolation \\ 
  \hline
king eider & nest sampling &  & 4x2km plot & wetlands \\ 
  long tailed duck & nest sampling &  & 4x2km plot & wetlands \\ 
  king eider & relat. abund. (incidental obs.) & red-throated loon & study area &  \\ 
  long tailed duck & relat. abund. (incidental obs.) & red-throated loon & study area &  \\ 
  Pacific loon & nest sampling &  & study area &  \\ 
  red-throated loon & nest sampling &  & study area &  \\ 
  cackling goose & nest sampling &  & study area &  \\ 
  snow goose & nest sampling &  & camp 2 & goose colony \\ 
  snow goose & vertebrate count transects &  & lowland & goose colony \\ 
  snow goose & snow goose point counts &  & lowland & goose colony \\ 
  tundra swan & nest sampling &  & study area &  \\ 
  rough-legged hawk & nest sampling &  & study area &  \\ 
  peregrine falcon & nest sampling &  & study area &  \\ 
  snowy owl & nest sampling &  & study area &  \\ 
  rock ptarmigan & nest sampling &  & 4x2km plot & all habitats \\ 
  sandhill crane & nest sampling &  & 4x2km plot & transects \\ 
  American golden-plover & distance sampling &  & lowland & transects \\ 
  black-bellied plover & relat. abund. (transects) & A. golden-plover & lowland & transects \\ 
  black-bellied plover & relat. abund. (incidental obs.) & A. golden-plover & study area &  \\ 
  common-ringed plover & nest sampling &  & study area &  \\ 
  ruddy turnstone & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  ruddy turnstone & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  red knot & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  red knot & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  pectoral sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  pectoral sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  Baird’s sandpiper & nest sampling &  & 2x1km plot & transects \\ 
  white-rumped sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  white-rumped sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  buff-breasted sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  buff-breasted sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  red phalarope & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
  red phalarope & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
  glaucous gull & nest sampling &  & study area &  \\ 
  long-tailed jaeger & nest sampling &  & qarlikturvik & mesic habitat \\ 
  parasitic jaeger & maximum banded individual &  & study area &  \\ 
  parasitic jaeger & nest sampling &  & study area &  \\ 
  common raven & nest sampling &  & study area &  \\ 
  common raven & relat. abund. (incidental obs.) & Parasitic jaeger & study area &  \\ 
  horned lark & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
  horned lark & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
  American pipit & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
  American pipit & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
  lapland longspur & nest sampling &  & 2x1km plot & transects \\ 
  snow bunting & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
  snow bunting & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
  brown lemming & capture marking recapture &  & two 11 ha grids & wet and mesic hab. \\ 
  collared lemming & capture marking recapture &  & two 11 ha grids & wet and mesic hab. \\ 
  arctic hare & incidental observations & Arctic fox & study area &  \\ 
  ermine & relat. abund. (testimonials) &  & study area &  \\ 
  arctic fox & home range (movement track.) &  & study area &  \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-06-25 script/data_cleaning/01_clean_year_monitoring_metadata.R

@@ -1,0 +1,92 @@
+#Title: Transform the table of year sampled for nest monitoring of each species into a usable data frame in R
+#Author: Louis Moisan
+#Date: 29 september 2023
+#Last update: June 11 2024
+
+
+#------------------#
+#### Librairies ####
+#------------------#
+#general data manipulation
+library(dplyr)
+library(tidyr)
+#spatial data manipulation
+library(sf)
+
+#--- Extract list of the zone of the study area
+zone_list <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  dplyr::filter(zone != "goose_colony") %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::select(zone) %>% 
+  unlist() %>% 
+  as.vector()
+
+#--- Import raw table of species year monitoring
+sp_monitoring <- read.csv("data/metadata/species_systematic_nest_monitoring_raw.csv")
+
+#extract a list with year of sampling for each species
+year_sampling <- stringr::str_split(sp_monitoring$year, ",")
+#Divide into multiple parts if species have been sampled non-continuously
+year_sampling <- lapply(year_sampling, function(x) stringr::str_split(x, "-"))
+#Assign species name to each list
+names(year_sampling) <- sp_monitoring$species
+#transform the year format from character to numeric
+for (sp in names(year_sampling)){
+  print(sp)
+  year_sampling[[sp]] <-   lapply(year_sampling[[sp]], function(x) x %>% unlist() %>% as.numeric())
+}
+
+#create an empty data frame to store the output
+year_sampling_df <- data.frame()
+#For each list with more than one year extend the interval of years by integer
+for (list in 1:length(year_sampling)){
+  for (sublist in 1:length(year_sampling[[list]])){
+    if(length(year_sampling[[list]][[sublist]]) > 1){
+      year_sampling[[list]][[sublist]] <- year_sampling[[list]][[sublist]][[1]]: year_sampling[[list]][[sublist]][[2]]}
+    #if list contain only a single value just transform to numeric
+    if(length(year_sampling[[list]][[sublist]]) == 1){
+      year_sampling[[list]][[sublist]] <- as.numeric(year_sampling[[list]][[sublist]])}
+    #transform into a data frame
+    years <- unlist(year_sampling[[list]][[sublist]])
+    year_sampling_df <- year_sampling_df %>% 
[…]

### 2024-06-25 script/extract_species_abundance/extract_species_abundance.R

@@ -1,0 +1,743 @@
+#Title: Extract the abundance of every vertebrate species from Bylot Island
+#Date: October 11 2023
+#Last update:  June 17 2024
+
+
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE) #prevent geoprocessing errors
+#Export tables as .tex
+library(xtable)
+#Personnal functions to estimate species abundance
+source("script/functions/FUNS.R")
+
+
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp")
+
+#years each species was monitored
+sampling_year <- read.csv("data/metadata/species_systematic_nest_monitoring_clean.csv")
+
+#nests
+nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
+nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
+
+#transects
+transect_data <- sf::st_read("data/raw/shapefiles/sampling/vertebrate_count_transects_2010_2023.shp")
+
+#incidental observation
+incidental_obs <- read.csv("data/raw/sampling/incidental_obs_2007_2019.csv") %>%
+  dplyr::select(-X)
+
+#Index of relative abundance in upland zones
+relative_abundance_upland <- read.csv("data/raw/other/incidental_obs_relative_abundance_upland.csv")
+
+#Index of relative abundance from Gauthier et al., 2024
+relative_abundance_gauthier <- read.csv("data/metadata/relative_abundance_Gauthier_et_al_2023.csv")
+
+#define zones to estimate abundance
+zones <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour", "study area")
+
+
+
+#--------------------------------------------------#
+#### 3.1 Pacific loon, red-throated loon, cackling goose, tundra swan, glaucous gull ####
+#-------------------------------------------------#
+#--- Extract nest density for each year and each zones
+wetlands_sp <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = c("cackling goose", "tundra swan","red throated loon", "pacific loon", "glaucous gull"),
+  zone_list= "study area",
+  year_list= c(2017:2023))
+
[…]

### 2024-06-25 script/extract_species_abundance/species/lemmings.R

@@ -1,0 +1,104 @@
+#title: Create dataframe with the time-series for collared and brown lemmings (1995-2022)
+#author: Azenor Bideault
+#date: november 8 2023
+
+#Libraries
+library(dplyr)
+library(sf)
+
+
+#-------------------#
+#### IMPORT DATA ####
+#-------------------#
+#-- For densities from 1995 to 2003
+df_lemming = read.csv('data/raw/sampling/lemming_density_1993-2019.csv', header = TRUE)
+#-- For densities from 2004 to 2022
+df_lemming2 = read.csv('data/raw/sampling/lemming_density_2004-2022.csv', header = TRUE)
+
+
+study_area_df <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::filter(zone=="study area")
+
+
+#-----------------------------------------#
+#### DENSITIES FOR THE YEARS 2004-2022 ####
+#-----------------------------------------#
+#--- Time-series is computed for the valley of camp 1 only
+# The period P2 (july) and P3 (august) are used
+# Filter the data
+filtered_data <- df_lemming2 %>% 
+  filter(site == 'Goose camp', period %in% c("P2", "P3"))
+
+#--- Compute the densities according to the good period to consider for each year
+# 2004 to 2007 -> period 3 is used
+# 2008 to 2018 and 2020 to 2022 -> mean between periods P2 and P3 is used
+# 2019 -> period P2 only is used
+# see document Document_interne_trappage_mortel_version (tables 5 and 6 for details)
+
+# Filter the data for the specified periods and years
+dens_lemming <- df_lemming2 %>% 
+  filter(
+    (year >= 2004 & year <= 2007 & period == "P3") |
+      (year >= 2008 & year <= 2018 & period %in% c("P2", "P3")) |
+      (year == 2019 & period == "P2") |
+      (year >= 2020 & year <= 2022 & period %in% c("P2", "P3"))
+  ) %>% 
+  group_by(species, habitat, grid, year) %>% 
+  summarize(mean_density = mean(density, na.rm = TRUE))
+
+#-- Compute densities for the whole study area according to the percentages of wetland and mesic habitats :
+dens_lemming_area <- dens_lemming %>% 
+  filter((grid == 'LG2' | grid == 'LG1'))  %>% 
+  group_by(species, year) %>% 
+  summarize(
[…]

### 2024-06-25 script/extract_species_abundance/species/snow_goose.R

@@ -1,0 +1,276 @@
+#Title: Estimate snow goose abundance at the scale of the study area
+#Author: Louis Moisan
+#Date: NA
+#Last update: 31 May 2024
+
+#------------------#
+#### Librairies ####
+#------------------#
+#General data manipulation
+library(dplyr) 
+library(tidyr)
+#Spatial data manipulation
+library(sf)
+sf::sf_use_s2(FALSE)
+#Visualization
+library(ggplot2)
+library(ggpubr)
+
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#Study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") 
+
+#Goose colony outline 2010-2023
+goose_colony_2010_2023 <- study_area %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::filter(zone == "goose_colony") %>% 
+  dplyr::select(-prop_up)
+#Goose colony 1999-2009 (2020-2021)
+goose_colony_1999_2009 <- goose_colony_2010_2023 %>% 
+  dplyr::summarise(area_km2= mean(area_km2), prop_wet= mean(prop_wet), prop_mesic= mean(prop_mesic),zone= "goose_colony") 
+# Replicate the row for each year and add a "year" column
+goose_colony_1999_2009 <- goose_colony_1999_2009[rep(1, length(c(1999:2009, 2020, 2021))), ] %>% 
+dplyr::mutate(year = c(1999:2009, 2020, 2021))
+#combine to extract a single data frame of the goose colony
+goose_colony <- goose_colony_2010_2023 %>% 
+  rbind(goose_colony_1999_2009)
+
+#Systematically monitored wetland plot (peksek)
+peksek <- sf::st_read("data/raw/shapefiles/sampling/snow_goose_systematic_wetland_plot.shp") %>%
+  sf::st_transform(26717) %>% 
+  dplyr::mutate(area_km2= round(as.numeric(sf::st_area(geometry)/ 1000000), digits = 4)) %>% 
+  dplyr::select(area_km2, geometry)
+#Goose nests within the peksek systematic plot
+peksek_nest <- sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp") %>%
+  dplyr::filter(species == "snow goose") %>% 
+  sf::st_intersection(peksek)
+
+#Random plots
+#Number of nest found in random plot
+random_plot <- read.csv("data/raw/sampling/snow_goose_nests_random_plot.csv") %>%
[…]

### 2024-06-25 script/functions/nested_functions/extend_temporal_series_from_single_zone_missing_coordinates.R

@@ -1,0 +1,83 @@
+# Function to extend temporal series when abundance in a zone is tighly correlated to abundance at the study area (when coordinates of nests are not accessible, but zone is)
+#Author: Louis Moisan
+#Date: June 11 2024
+
+#------------------#
+#### Librairies ####
+#------------------#
+#data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE)
+
+extend_temporal_series_from_single_zone_nest_missing_coords <- function(
+    nest_data,
+    zone_sf,
+    sp,
+    year_list,
+    year_compare,
+    zone_ref){
+  
+#--- Extract density from nest found in each zone
+sp_density <- nest_data %>% 
+  dplyr::filter(species ==sp) %>% 
+  group_by(species, zone, year) %>% 
+  dplyr::summarise(nb_nest= n()) %>% 
+  dplyr::left_join(study_area %>% dplyr::select(zone, area_km2) %>% sf::st_drop_geometry()) %>% 
+  dplyr::mutate(nest_density_km2= nb_nest/area_km2, ind_density_km2= nest_density_km2*2) %>% 
+  #Add all years-zone-species where no nest were found and assign 0
+  full_join(expand.grid(year = c(year_list,year_compare), species = sp, zone = unique(zone_sf$zone), by = c("species", "zone","year"))) %>% 
+  #Replace the NA with 0
+  mutate(nb_nest= ifelse(is.na(nb_nest), 0, nb_nest),
+    ind_density_km2= ifelse(is.na(ind_density_km2), 0, ind_density_km2)) %>% 
+  dplyr::select(species, zone, year, nb_nest,ind_density_km2) %>% 
+  unique()
+  
+#Species density at the scale of the study area
+sp_density_study_area <- sp_density %>% 
+  dplyr::filter(zone== "study area", year %in% year_compare)%>% 
+  dplyr::arrange(year) 
+
+#Species density in the reference zone
+if(length(zone_ref)>1){
+sp_density_ref <- sp_density %>% 
+  dplyr::filter(zone %in% zone_ref) %>% 
+  dplyr::left_join(study_area %>% dplyr::select(zone, area_km2) %>%  sf::st_drop_geometry(), by= "zone") %>% 
+  dplyr::group_by(species, year) %>% 
+  dplyr::summarise(sum_nb_nest= sum(nb_nest), sum_area_km2= sum(area_km2)) %>% 
[…]

### 2024-06-25 script/functions/nested_functions/extend_temporal_series_from_single_zone_nest.R

@@ -1,0 +1,79 @@
+# Function to extend temporal series when abundance in a zone is tighly correlated to abundance at the study area
+#Author: Louis Moisan
+#Date: December 5 2023
+
+#------------------#
+#### Librairies ####
+#------------------#
+#data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE)
+
+source("script/functions/standalone_functions/get_nest_density.R")
+
+extend_temporal_series_from_single_zone_nest <- function(
+    nest_data, # data frame of nest location for each species
+    zone_sf, #Sf object of the zones of the study area to estimate density in
+    sp, # species name (in lower case, no dashed, no underscore)
+    year_list, #List of year to estimate density
+    year_compare, #List of year to based the defined the regression between the density in the reference zone and the density at the study area
+    zone_ref #List of zone(s) to used as reference
+    ){
+  
+#--- Extract density from nest found in each zone
+sp_density <- get_nest_density(
+    nest_data= nest_data,
+    zone_sf= study_area,
+    species_list = sp,
+    zone_list= c(zone_ref, "study area"),
+    year_list= year_list)
+  
+#Species density at the scale of the study area
+sp_density_study_area <- sp_density %>% 
+  dplyr::filter(zone== "study area", year %in% year_compare)%>% 
+  dplyr::arrange(year) 
+
+#Species density in the reference zone
+if(length(zone_ref)>1){
+sp_density_ref <- sp_density %>% 
+  dplyr::filter(zone %in% zone_ref) %>% 
+  dplyr::left_join(study_area %>%  sf::st_drop_geometry(), by= "zone") %>% 
+  dplyr::group_by(species, year.x) %>% 
+  dplyr::summarise(sum_nb_nest= sum(nb_nest), sum_area_km2= sum(area_km2)) %>% 
+  dplyr::mutate(species= sp, year= year.x, zone= "zone reference", ind_density_km2= sum_nb_nest*2/sum_area_km2) %>% 
+  dplyr::select(species, zone, year, ind_density_km2)
+}else {
+  sp_density_ref <- sp_density %>% 
+    dplyr::filter(zone == zone_ref) %>% 
+    dplyr::arrange(year)
+}
+#Extract a data frame of the density in the reference zone with the same year as in the study area data frame
+sp_density_ref_filter <- sp_density_ref %>% 
[…]

### 2024-06-25 script/functions/nested_functions/spatial_extrapolation_species_density.R

@@ -1,0 +1,235 @@
+# Spatially extrapolate species density monitored on a local plot to other zones of the study area using a cross multiplication based on the mean number of individual per transect
+
+
+#------------------#
+#### Librairies ####
+#------------------#
+#general data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE)
+
+#Dependencies on function created in other scripts
+source("script/functions/standalone_functions/number_ind_transects.R")
+source("script/functions/standalone_functions/get_nest_density.R")
+
+
+spatial_extrapolation_species_density <- function(
+    nest_data,#Nest data (id, coordinates, species and year)
+    transect_data,#Transect observation data
+    transect_metric, #"mean number of individual per transect" OR
+    #"proportion of transect with individual"
+    zone_sf,#shapefile that contain the polygon of the zones of the study area
+    zone_reference_nest, #List of zones to use as reference for nest density
+    zone_reference_transect, #List of zones to use as reference for transects relative abundance
+    years_reference, #List of year to used to define the relation between transect and nest density
+    sampling_year, # data frame with the species, year and zone that have been sampled
+    zone_list,#list of the zones to estimate individual density
+    sp, #species of interest
+    status_list,#status of individuals that will be taken into account
+    relative_abundance_upland, # Ratio of relative abundance between plateau and non-plateau areas
+    null_density_wetlands# TRUE/FALSE, should the area of wetlands be remove from the area used to perform spatial extrapolation
+    ) {
+ 
+#--------------------------------------------------------------#
+#### Extract local individual density based on nest density ####
+#--------------------------------------------------------------#
+#If the reference zone is the 2x1 km plot, adjust the area by removing area of lakes (0.26km2) in the zone
+if (zone_reference_nest == "2x1km plot"){
+ zone_sf[zone_sf$zone== "2x1km plot",]$area_km2 <- 1.74
+}
+  
+sp_density_local <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= zone_sf,
+  species_list = sp,
+  zone_list= zone_reference_nest,
+  year_list= years_reference) %>% 
+    dplyr::arrange(year)
[…]

### 2024-06-25 script/tables_figures/table_body_mass.R

@@ -1,0 +1,70 @@
+#general data manipulation
+library(dplyr)
+library(tidyr)
+#export latex table
+library(xtable)
+
+
+#--- Import as .csv
+df_body_mass <- read.csv("data/clean/species_body_mass.csv") %>% 
+  dplyr::select(-X, -Scientific.name) %>% 
+  dplyr::rename(Species= Species.name, `Body mass g (study area)`= Body.mass.g..study.area., `Body mass g (literature)`= Body.mass.g..literature.)
+
+
+#reorder names based on defined species list
+sp_metadata <- read.csv("data/metadata/species_metadata.csv")
+# Find the indices of species in df2 in df1
+indices <- match(sp_metadata$species, df_body_mass$Species)
+# Reorder df1 based on the indices
+df_body_mass <- df_body_mass[indices, , drop = FALSE] 
+
+#update species names
+df_body_mass <- df_body_mass %>% 
+  dplyr::mutate(Species= dplyr::recode(Species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper", "lapland longspur"= "Lapland longspur"))
+
+
+#Remove decimals
+df_body_mass <- df_body_mass %>% 
+  dplyr::mutate(`Body mass g (study area)`= round(`Body mass g (study area)`, digits = 0),
+                `Body mass g (literature)`= round(`Body mass g (literature)`, digits=0),
+                n= ifelse(is.na(n), NA, paste0("(",n,")")))
+# set NA for ermine and Arctic fox
+df_body_mass[df_body_mass$Species %in% c("ermine", "Arctic fox"),]$n <- "(NA)"
+
+
+
[…]

### 2024-06-25 script/tables_figures/table_mean_abundance.R

@@ -1,0 +1,98 @@
+# Extract Bylot Island temporal series of species abundance and biomass
+#Date: November 27 2023
+
+#------------------#
+#### Librairies ####
+#------------------#
+#data manipulation
+library(dplyr)
+library(tidyr)
+library(sf)
+#export latex table
+library(xtable)
+
+
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#Read species final densities
+abundances <- read.csv("data/clean/species_abundance.csv")
+
+#Combined all data of snow goose under "combined methods
+abundances[abundances$species == "snow goose",]$data <- "combined methods"
+
+#Read study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  sf::st_drop_geometry()
+
+#------------------------------------#
+#### Table species mean abundance ####
+#------------------------------------#
+#--- Species with temporal series
+abundance_series <- 
+  abundances %>% 
+  dplyr::filter(is.na(year)== FALSE) %>% 
+  dplyr::mutate(data= gsub(" \\(.*", "", data)) %>%
+  dplyr::group_by(species, data) %>% 
+ dplyr::summarise(mean_abundance= round(mean(abundance),
+                                        digits = 0),
+                  sd_abundance= round(sd(abundance), digits=0),
+                   n= n(),
+                  min_abundance= min(abundance),
+                  max_abundance= max(abundance)) 
+
+#Species with only mean
+mean_abundance <- 
+  abundances %>% 
+  dplyr::filter(is.na(year)) %>% 
+  dplyr::select(-year, -status,-zone) %>% 
+  dplyr::rename(mean_abundance= abundance) %>% 
+  dplyr::mutate(sd_abundance= NA, n= NA, min_abundance= NA, max_abundance= NA)
+
+#--- Combined together
+table <- abundance_series %>% 
+  rbind(mean_abundance) %>% 
+  dplyr::mutate(mean_abundance= as.character(mean_abundance), sd_abundance= as.character(sd_abundance)) %>% 
+  dplyr::mutate(`Mean ± sd`= ifelse(is.na(sd_abundance),  mean_abundance, paste(mean_abundance, "±", sd_abundance))) %>% 
+  dplyr::rename(Minimum= min_abundance, Maximum= max_abundance, Data=data, Species= species) %>% 
+  dplyr::select(Species, Data, `Mean ± sd`, n, Minimum, Maximum)
+
+
+#reorder names based on defined species list
[…]

### 2024-06-25 script/tables_figures/table_species_estimate_uncertainty.R

@@ -1,0 +1,59 @@
+#Transform .csv of uncertainty on species estimate to a .tex table
+library(dplyr)
+library(xtable)
+
+
+esti_uncert <- read.csv("data/metadata/estimate_uncertainty.csv") %>% 
+  dplyr::rename(`abundance estimate`= abundance.estimate)
+
+#reorder names based on defined species list
+esti_uncert <- read.csv("data/metadata/species_metadata.csv") %>% 
+  dplyr::select(species) %>% 
+  dplyr::left_join(esti_uncert)
+
+#update species names
+esti_uncert <- esti_uncert %>% 
+  dplyr::mutate(species= dplyr::recode(species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper","lapland longspur"= "Lapland longspur"))
+
+#Rename columns 
+esti_uncert <- esti_uncert %>% 
+  dplyr::rename(Species= species, Estimate= `abundance estimate`, Confidence= confidence, Reason= reason)
+
+# test (add \newline if string if more than 55 characters)
+Reason_updated <- c()
+
+for(i in esti_uncert$Reason){
+  #space positions
+space_positions <-gregexpr(" ", i)[[1]] %>% as.vector()
+ if (max(space_positions)>= 53){
+#Select first space position that over 53
+   select_space <- min(space_positions[which(space_positions>=53)])
+#Insert newline at this location
+ string <- paste0("\\makecell[tl]{",substr(i, 1, select_space-1), "\\\\",substr(i, select_space +1, nchar(i)), "}")
+#Replace original string
+ Reason_updated <- c(Reason_updated, string)
+ }else{
+   Reason_updated <- c(Reason_updated, i)
+ }
+}
+
+esti_uncert$Reason <- Reason_updated
+
+
+#as tex
[…]

### 2024-06-25 script/tables_figures/table_species_year_monitoring.R

@@ -1,0 +1,64 @@
+#data manipulation
+library(dplyr)
+library(tidyr)
+#spatial data manipulation
+library(sf)
+#export latex table
+library(xtable)
+
+#--- Import study area 
+#list of zones for which we want to obtain a density
+zone_list <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour", "study area")
+#shapefile of the area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  sf::st_drop_geometry() %>% 
+  #keep only non-nested zones
+  dplyr::filter(zone %in% zone_list)
+
+#--- Filter nest data to keep only nest found during years of systematic survey
+sp_monitoring <- read.csv("data/metadata/species_systematic_nest_monitoring_raw.csv")
+#reorder names based on defined species list
+sp_metadata <- read.csv("data/metadata/species_metadata.csv") %>% 
+  dplyr::select(species) 
+# Reorder df1 based on the species meta data order
+sp_monitoring <- sp_metadata %>% 
+  dplyr::left_join(sp_monitoring) %>% 
+  na.omit()
+
+#update species names
+sp_monitoring <- sp_monitoring %>% 
+  dplyr::mutate(species= dplyr::recode(species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper","lapland longspur"= "Lapland longspur"))
+
+#--- Extract standardize and extract a tex table
+table_year_monitoring <- sp_monitoring %>% 
[…]

### 2024-06-25 script/tables_figures/table_summary_methods_data.R

@@ -1,0 +1,35 @@
+#Title: Create a summary table of all data and method use for each species
+#Author: Louis Moisan
+#Date: June 3 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+
+#-----------------------#
+#### Import metadata ####
+#-----------------------#
+table <- read.csv("data/metadata/summary_data_sources.csv") %>% 
+  dplyr::rename(`Sampling method`=Density.sampling.method,
+               `Reference species`=Reference.species,
+               `Sampling scale`=Sampling.scale,
+               `Spatial extrapolation`=Spatial.extrapolation)
+
+#as tex
+table_latex <- xtable(table, caption = "Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapoated at the landscape scale.",
+                      label= "table:summary_methods")
+print(table_latex,
+      include.rownames = FALSE,
+      tabular.environment = "tabularx",
+      table.placement = "H",
+      width="\\textwidth",
+      align="rXXXXXX",
+      size="\\fontsize{8pt}{10pt}\\selectfont",
+      file = "manuscript/tables/table_summary_methods.tex",
+      caption.placement = "top")
+

### 2024-06-28 manuscript/manuscript.tex
uniformize species name in the final data set

@@
 The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology.
-Community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}.
+Variations in community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}.
 Additionally, community composition is essential for modeling the dynamics of ecological communities.
@@
 Species abundances are influenced by stochastic effects \citep{hubbell2001}, environmental changes (e.g., climate warming), and species interactions, contributing to data variability.
-For instance, a community composition could be driven simultaneously by the intra-annual seasonal variations in the environment, the multi-year dynamics of small mammals cycles and a relatively slow but directional climate warming.
-Therefore, to untangle the relative effects of diverse abiotic and biotic factors on community composition, long-term data sets are required \citep{magurran2010, lindenmayer2012}.
+For instance, the composition of a community could be driven simultaneously by the intra-annual seasonal variations in the environment, the multi-year cyclic dynamics of an important prey or predator and a relatively slow but directional climate warming.
+Therefore, long-term data series are required to untangle the relative effects of diverse abiotic and biotic factors on community composition \citep{magurran2010, lindenmayer2012}.
 
@@
 \subsection*{Objective}
-Our objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August).
-This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, but excludes non-breeding individuals that stop for only a few days during their migration.
-Additionally, our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments.
+Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August).
+This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration.
+Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments.
+Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
 \newpage
@@
 Long-term monitoring of the community began in the 1990s in the Qarlikturvik valley \citep{gauthier2013, gauthier2024a}, which represents the zone of the study area with the highest annual sampling effort.
-Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river, where the main research infrastructure is situated.
+Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river, where the main research infrastructure is located.
 Another zone with extensive sampling efforts is Camp 2, located at the core of the snow goose colony, where the primary focus is to monitor snow goose nests.
@@
 \subsubsection*{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}
-Since 2004, systematic searches of wetland areas has been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area.
+Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area.
 This sampling aimed to find all nests of the cackling goose and the glaucous gull.
@@
 Each year the observers use slightly different paths to sample the areas, but locate the nests in the same positions, which supports a high probability of detection for these species.
-Most nesting sites were located in the upland zones of the study area, which are the Black Plateau, Southern Plateau and Camp 3 Plateau.
+Most nesting sites were located in the upland zones of the study area, which include the Black Plateau, Southern Plateau and Camp 3 Plateau.
 
@@
 We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability.
-We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances.
+We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).
 
@@
 We considered the sampling to be most systematic within a core 2 km\textsuperscript{2} (2x1 km) plot in this area (\textbf{Figure \ref{figure:qarlikturvik_valley}}).
+We excluded relatively large water bodies (0.26 km\textsuperscript{2}) to calculate nest density in the plot due to the presence of a large lake, which leaves an area of 1.74 km\textsuperscript{2} available for nesting.
 An observer conducted systematic searches of this plot during the entire breeding season to locate and monitor as many passerine and shorebird nests as possible.
-Assuming this observer can detect all nests within 5 or 10 meters, this single observer sampled annually at minimum an area of 0.72 \textpm{0.12} or 1.09 \textpm{0.17} km\textsuperscript{2} (n= 3).
+Assuming the observer can detect all nests within a 5 or 10 meter radius, analysis of daily GPS tracks shows that the observer covered a minimum area of 0.72 \textpm{0.12} (5 m) or 1.09 \textpm{0.17} km\textsuperscript{2} (10 m) of the core area annually (n= 3 years).
 Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.
-We excluded relatively large water bodies (0.26 km\textsuperscript{2}) to calculate nest density in the plot due to the presence of a large lake, which leaves an area of 1.74 km\textsuperscript{2} available for nesting.
 
@@
 From 2010 to 2023, observers walked 500-meter linear transects where all vertebrate individuals observed within 150 meters on either side were counted (146 to 320 transects per year).
-Transects were distributed across all lowland zones of the study area, typically in mesic habitats, and were carried out during the nesting period (between June 21 and July 14; \citet{duchesne2021}; \textbf{Figure \ref{figure:transects}}).
+Transects were distributed across all lowland zones of the study area, typically in mesic habitat, and were carried out during the nesting period (between June 21 and July 14; \citet{duchesne2021}; \textbf{Figure \ref{figure:transects}}).
 Furthermore, specifically for American golden-plovers, we measured the distance of each observed individual to the transect path.
@@
 \subsubsection{Arctic fox movement tracking}
-Between 2008 and 2016, 109 Arctic foxes were fitted with Argos Platform Transmitter Terminals mounted on collars \citep{lai2015,christin2015, dulude2023}.
+In order to assess fox abundance based on the size of their home range, 109 Arctic foxes were fitted with Argos Platform Transmitter Terminals mounted on collars between 2008 and 2016 \citep{lai2015,christin2015, dulude2023}.
 Foxes were captured between May and August across the study area, within and outside the goose colony \citep{dulude2023}.
@@
 \subsection{Species body mass}
-All vertebrate individuals captured for marking purposes were systematically weighed (snow goose (unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers\citep{lamarre2021},  common-ringed plovers \citep{leandri2019}, other shorebirds (Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (Gauthier and Therrien, unpublished data), Lapland longspurs (Gauthier and Bêty, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}).
+All vertebrate individuals captured for marking purposes were systematically weighed (snow goose (unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021},  common-ringed plovers \citep{leandri2019}, other shorebirds (Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (Gauthier and Therrien, unpublished data), Lapland longspurs (Gauthier and Bêty, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}).
 Mean body mass of vertebrate species was determined in priority with measurements from Bylot Island.
@@
 Most of data manipulation was carried out with the \textit{dplyr} \citep{dplyr} and \textit{sf} packages \citep{sf}.
-All codes are available in the Dryad repository XXXXXXXXXXXXXX.
+All codes are available in the following repository \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset.git}.
 \newpage
@@
 However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series.
-Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area  and the indices of relative abundance (i.e., daily incidental observations and transects).
+Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area.
 We thus fitted an exponential model using the number of nests found annually over two distinct periods: in 1996 when the first nest was discovered, and then from 2017 to 2023 when sampling effort was systematic across the whole study area (\textbf{Figure \ref{figure:cackling}}).

### 2024-06-28 manuscript/tables/table_summary_methods.tex
uniformize species name in the final data set

@@
 % latex table generated in R 4.4.0 by xtable 1.8-4 package
-% Tue Jun 25 10:25:27 2024
+% Fri Jun 28 12:52:36 2024
 \begin{table}[H]
 \centering
-\caption{Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapoated at the landscape scale.} 
+\caption{Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapolated at the landscape scale.
+When sampling consist in relative abundance, we assumed that the ratios between relative and actual abundance are the same between the targeted species and a given reference species (i.e., similar detection probability).
+We therefore derived the absolute abundance of the targeted species from its relative abundance using the ratio between relative and absolute abundances of the reference species.
+Additionally, here the area of each sampling scale: two 11 ha grids (0.22 km2), 2x1km plot (2 km2) 4x2km plot (8 km2), south qarlikturvik (33 km2), camp 2 (80 km2), lowland (313 km2) and whole study area (389 km2).}
 \label{table:summary_methods}

### 2024-06-28 script/tables_figures/table_mean_abundance.R

@@ -63,6 +63,7 @@ table <- abundance_series %>%
 sp_metadata <- read.csv("data/metadata/species_metadata.csv") %>% 
   dplyr::mutate(indice=1:nrow(.)) %>% 
   dplyr::select(species, indice)
+
 # Reorder df1 based on the indices
 table <- table %>% 
   dplyr::left_join(sp_metadata, by= c("Species"= "species")) %>% 
@@ -92,7 +93,8 @@ print(table_mean_abundance_latex,
 #### Export as .csv for the data set ####
 #---------------------------------------#
 abundance_dataset <- abundance_series %>% 
-rbind(mean_abundance)
+rbind(mean_abundance)%>% 
+  dplyr::mutate(species= dplyr::recode(species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper", "lapland longspur"= "Lapland longspur"))
 
 write.csv(abundance_dataset,"dataset/BYLOT-mean_species_abundance.csv",row.names = FALSE)
 

### 2024-06-28 script/tables_figures/table_species_name_strategy.R

@@ -30,7 +30,7 @@ print(sp_name_latex,
 table_name %>% 
   dplyr::rename(functional_group= `Functional group`,
                 scientific_name= `Scientific name`,
-                name= Name,
+                species= Name,
                 annual_cycle_strategy=`Annual cycle strategy`) %>% 
   write.csv("dataset/BYLOT-community_composition.csv", row.names = FALSE)
 

### 2024-07-01 manuscript/manuscript.tex
add minor changes proposed by DF

@@
   \centering
-\textbf{BYLOT: Long-term abundance time-series of a High Arctic terrestrial vertebrate community}\\[1.5em]
+\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 \endgroup
@@
 
-Modeling ecological communities requires adjusting trophic flows based on species' functional responses, which necessitates time series data on the abundance of all species within a community.
+Modeling ecological communities requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community.
 However, accurately determining the abundance of all species is rarely achievable.
@@
 The case of the red fox (Vulpes vulpes) was ambiguous.
-While the presence of a breeding pair has been confirmed in the study area \citep{lai2022}, the extent of the population's establishment remains unclear, and sightings are rare, therefore, we decide to exclude this species.
+While the presence of a breeding pair has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare.
+Therefore, we decided to exclude this species.
 

### 2024-09-18 manuscript/manuscript.tex
Combine table 7 and 8

@@
 \usepackage{makecell}
+\usepackage{pdflscape}
 %bibliography
@@
 }
+\usepackage{geometry}
 
@@
 \textbf{Authors}:\\
- Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Pierre Legagneux\textsuperscript{3,4}, Gilles Gauthier\textsuperscript{5}, Éliane Duchesne\textsuperscript{1}, Marie-Christine Cadieux \textsuperscript{3}, Dominique Berteaux\textsuperscript{1}, Dominique Fauteux\textsuperscript{6} and Joël Bêty\textsuperscript{1}\\[1.5em]
+Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1},
+Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
 
@@
 \textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
-\textsuperscript{4} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France \\
-\textsuperscript{5} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
-\textsuperscript{6} Centre d’études nordiques, Centre de connaissance et d’exploration de l’Arctique, Musée canadien de la nature, Ottawa, ON, Canada
+\textsuperscript{4} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
+\textsuperscript{5} Centre d’études nordiques, Centre de connaissance et d’exploration de l’Arctique, Musée canadien de la nature, Ottawa, ON, Canada\\
+\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
 \\[1.5em]
 
-\textbf{Corresponding Author}:\\
+\textbf{Corresponding Authors}:\\
 \textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
@@
 
-Modeling ecological communities requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community.
+Modeling food webs requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community.
 However, accurately determining the abundance of all species is rarely achievable.
@@
 Species abundances are influenced by stochastic effects \citep{hubbell2001}, environmental changes (e.g., climate warming), and species interactions, contributing to data variability.
-For instance, the composition of a community could be driven simultaneously by the intra-annual seasonal variations in the environment, the multi-year cyclic dynamics of an important prey or predator and a relatively slow but directional climate warming.
+For instance, the composition of a community could be driven simultaneously by intra-annual seasonal variations, multi-year cyclic variations (e.g., El Niño) and slow but directional long-term variations in the environment \citep{brown1990, snyder2006}.
 Therefore, long-term data series are required to untangle the relative effects of diverse abiotic and biotic factors on community composition \citep{magurran2010, lindenmayer2012}.
@@
 Systematic monitoring of every known or potential nesting site has been carried out in the Qarlikturvik valley, Black plateau and Southern plateau since 2007 and throughout the study area since 2013 \citep{beardsell2016, gauthier2020avianpred}.
-Observers walked along ridges and scan surrounding areas from vantage points to detect nesting birds.
+Observers walked along ridges and scanned surrounding areas from vantage points to detect nesting birds.
 These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest.
@@
 All codes are available in the following repository \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset.git}.
-\newpage
-
-\input{tables/table_summary_methods.tex}
-%label-> table:summary_methods
 \newpage
@@
 
+\newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
+\begin{landscape}
 \section{Species abundance}
-
-\subsection{Uncertainty on estimates}
-We did not quantitatively assessed uncertainty on each estimate of species abundance.
-However, we provide a qualitative measure of uncertainty based on the level of confidence derived from field experts, and in some cases, the fit of statistical models to estimate density (\textbf{Table \ref{table:confidence_estimate_abundance}}).
-
-\input{tables/confidence_estimate_abundance.tex}
-%label -> table:confidence_estimate_abundance
+\input{tables/table_summary_methods.tex}
+%label-> table:summary_methods
+\end{landscape}
+\restoregeometry
 \newpage
 
-\subsection{Abundance estimates}
 \input{tables/species_mean_abudance.tex}
@@
 \newpage
-
-The following figures present the time series of species abundances.
-If different methods were used, we present the estimates derived from each method with different colors (see \textbf{Table \ref{table:confidence_estimate_abundance}} for the corresponding method and confidence in estimates).
 

### 2024-09-18 manuscript/tables/species_mean_abudance.tex
uniformize species name in the final data set
Combine table 7 and 8

@@
-% latex table generated in R 4.4.0 by xtable 1.8-4 package
-% Tue Jun 25 10:25:29 2024
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Tue Sep 10 14:57:19 2024
 \begin{table}[ht]
 \centering
-\caption{Mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).
-In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty.}
+\caption{Selected methods to represent the mean (± standard deviation) of annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).} 
 \label{table:species_mean_abudance}
 \begingroup\fontsize{8pt}{10pt}\selectfont
-\begin{tabularx}{\textwidth}{lllrrr}
+\begin{tabularx}{0.85\textwidth}{llllr}
   \hline
-Species & Method & Mean ± sd & n & Minimum & Maximum \\ 
+Species & Method & Mean & sd & n \\ 
   \hline
-king eider & nest sampling (extrapolation habitat) & 25 &  &  &  \\ 
-  king eider & incidental observations (relative abundance) & 106 &  &  &  \\ 
-  long-tailed duck & nest sampling (extrapolation habitat) & 20 &  &  &  \\ 
-  long-tailed duck & incidental observations (relative abundance) & 191 &  &  &  \\ 
-  Pacific loon & nest sampling & 4 ± 3 &   4 &   0 &   6 \\ 
-  red-throated loon & nest sampling & 64 ± 15 &   4 &  42 &  76 \\ 
-  cackling goose & nest sampling & 50 ± 59 &  28 &   2 & 214 \\ 
-  snow goose & combined methods & 25289 ± 13327 &  23 & 2505 & 49076 \\ 
-  tundra swan & nest sampling & 1 ± 1 &   4 &   0 &   2 \\ 
-  rough-legged hawk & nest sampling & 25 ± 24 &  16 &   0 &  66 \\ 
-  peregrine falcon & nest sampling & 10 ± 1 &   8 &   8 &  12 \\ 
-  snowy owl & nest sampling & 16 ± 33 &  26 &   0 & 144 \\ 
-  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  &  \\ 
-  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  &  \\ 
-  American golden-plover & distance sampling & 774 ± 275 &   8 & 306 & 1040 \\ 
-  black-bellied plover & transects (relative abundance) & 22 &  &  &  \\ 
-  black-bellied plover & incidental observations (relative abundance) & 61 &  &  &  \\ 
-  common-ringed plover & nest sampling & 55 ± 9 &   3 &  44 &  62 \\ 
-  ruddy turnstone & transects (relative abundance) & 40 &  &  &  \\ 
-  ruddy turnstone & incidental observations (relative abundance) & 54 &  &  &  \\ 
-  red knot & transects (relative abundance) & 64 &  &  &  \\ 
-  red knot & incidental observations (relative abundance) & 237 &  &  &  \\ 
-  pectoral sandpiper & transects (relative abundance) & 79 &  &  &  \\ 
-  pectoral sandpiper & incidental observations (relative abundance) & 260 &  &  &  \\ 
-  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  &  \\ 
-  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  &  \\ 
-  white-rumped sandpiper & incidental observations (relative abundance) & 1156 &  &  &  \\ 
-  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  &  \\ 
-  buff-breasted sandpiper & incidental observations (relative abundance) & 8 &  &  &  \\ 
-  red phalarope & transects (relative abundance) & 156 &  &  &  \\ 
-  red phalarope & incidental observations (relative abundance) & 276 &  &  &  \\ 
-  glaucous gull & nest sampling & 73 ± 7 &  17 &  59 &  80 \\ 
-  long-tailed jaeger & nest sampling & 272 ± 285 &  17 &   0 & 900 \\ 
-  parasitic jaeger & banding & 17 &  &  &  \\ 
-  parasitic jaeger & nest sampling & 8 &  &  &  \\ 
-  common raven & transects (relative abundance) & 14 &  &  &  \\ 
-  common raven & incidental observations (relative abundance) & 31 &  &  &  \\ 
-  horned lark & transects (relative abundance) & 362 &  &  &  \\ 
-  horned lark & incidental observations (relative abundance) & 411 &  &  &  \\ 
-  American pipit & transects (relative abundance) & 53 &  &  &  \\ 
-  American pipit & incidental observations (relative abundance) & 87 &  &  &  \\ 
-  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  &  \\ 
-  snow bunting & transects (relative abundance) & 18 &  &  &  \\ 
-  snow bunting & incidental observations (relative abundance) & 276 &  &  &  \\ 
-  brown lemming & trapping & 54043 ± 93530 &  27 &   0 & 447630 \\ 
-  collared lemming & trapping & 8128 ± 10334 &  27 &   0 & 39302 \\ 
-  Arctic hare & incidental observations (relative abundance) & 6 &  &  &  \\ 
-  ermine & testimonials & 40 ± 37 &  27 &   8 & 156 \\ 
-  Arctic fox & individual home range & 53 &  &  &  \\ 
+king eider & incidental observations (relative abundance) & 106 &  &  \\ 
+  long-tailed duck & incidental observations (relative abundance) & 191 &  &  \\ 
+  Pacific loon & nest sampling & 4 & 3 &   4 \\ 
+  red-throated loon & nest sampling & 64 & 15 &   4 \\ 
+  cackling goose & nest sampling & 140 & 41 &   7 \\ 
+  snow goose & combined methods & 31852 & 12092 &  12 \\ 
+  tundra swan & nest sampling & 1 & 1 &   4 \\ 
+  rough-legged hawk & nest sampling & 25 & 24 &  16 \\ 
+  peregrine falcon & nest sampling & 10 & 1 &   8 \\ 
+  snowy owl & nest sampling & 16 & 33 &  26 \\ 
+  rock ptarmigan & nest sampling (extrapolation habitat) & 24 &  &  \\ 
+  sandhill crane & nest sampling (extrapolation transects) & 34 &  &  \\ 
+  American golden-plover & distance sampling & 774 & 275 &   8 \\ 
+  black-bellied plover & transects (relative abundance) & 22 &  &  \\ 
+  common-ringed plover & nest sampling & 55 & 9 &   3 \\ 
+  ruddy turnstone & transects (relative abundance) & 40 &  &  \\ 
+  red knot & transects (relative abundance) & 64 &  &  \\ 
+  pectoral sandpiper & transects (relative abundance) & 79 &  &  \\ 
+  Baird's sandpiper & nest sampling (extrapolation transects) & 2495 &  &  \\ 
+  white-rumped sandpiper & transects (relative abundance) & 1135 &  &  \\ 
+  buff-breasted sandpiper & transects (relative abundance) & 41 &  &  \\ 
+  red phalarope & transects (relative abundance) & 156 &  &  \\ 
+  glaucous gull & nest sampling & 73 & 7 &  17 \\ 
+  long-tailed jaeger & nest sampling (extrapolation habitat) & 272 & 285 &  17 \\ 
+  parasitic jaeger & banding & 17 &  &  \\ 
+  common raven & incidental observations (relative abundance) & 31 &  &  \\ 
+  horned lark & incidental observations (relative abundance) & 411 &  &  \\ 
+  American pipit & incidental observations (relative abundance) & 87 &  &  \\ 
+  Lapland longspur & nest sampling (extrapolation transects) & 7110 &  &  \\ 
+  snow bunting & incidental observations (relative abundance) & 276 &  &  \\ 
+  brown lemming & trapping & 54043 & 93530 &  27 \\ 
+  collared lemming & trapping & 8128 & 10334 &  27 \\ 
+  Arctic hare & incidental observations (relative abundance) & 6 &  &  \\ 
+  ermine & testimonials & 40 & 37 &  27 \\ 
+  Arctic fox & individual home range & 53 &  &  \\ 
    \hline

### 2024-09-18 manuscript/tables/table_summary_methods.tex
Combine table 7 and 8

--- 2024-06-25 manuscript/tables/table_summary_methods.tex
@@
-% latex table generated in R 4.4.0 by xtable 1.8-4 package
-% Tue Jun 25 10:25:27 2024
-\begin{table}[H]
-\centering
-\caption{Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapoated at the landscape scale.} 
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Tue Sep 10 15:05:42 2024
+\begingroup\fontsize{8pt}{10pt}\selectfont
+\begin{longtable}{|p{0.11\textwidth}|p{0.07\textwidth}|p{0.21\textwidth}|p{0.06\textwidth}|p{0.18\textwidth}|p{0.035\textwidth}|p{0.037\textwidth}|p{0.037\textwidth}|p{0.035\textwidth}|p{0.017\textwidth}|p{0.14\textwidth}|}
+\caption{Summary of the mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).
+In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty.
+We did not quantitatively assess uncertainty on each estimate of species abundance.
+However, we provide a qualitative measure of method quality based on the quality and quantity of data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.
+We also provided coarse confidence intervals derived from consultation with field experts at the study site.} \\
+  \hline
+{\textbf{Species}} & {\textbf{Estimate}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Min}} & {\textbf{Max}} & {\textbf{Mean}} & {\textbf{sd}} & {\textbf{n}} & {\textbf{Expert based confidence interval}} \\
+  \hline
+snow goose & 1999-2009 & Nest monitoring plots extrapolated to mean goose colony area & low & Relatively small sample size and uncertainty on goose colony area  & 2505 & 35404 & 18129 & 11037 &  11 & several thousands to few tens of thousands \\
+   \hline
+snow goose & 2010-2019, 2022-2023 & Intensive study area-wide monitoring based on several
+methods and annual colony outline & high & Multiple independent methods and annual colony outline & 8687 & 49076 & 31852 & 12092 &  12 & thousands \\ 
+   \hline
+snowy owl & 1996-2011 & Extrapolation from intensive nest monitoring (111 km2,
+R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &   0 &  67 & 15 & 24 &  16 & units to few tens \\ 
+   \hline
+snowy owl & 2012-2019, 2022-2023 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 & 144 & 17 & 45 &  10 & units to few tens \\ 
+   \hline
+glaucous gull & 2004-2016 & Extrapolation from intensive nest monitoring (111 km2,
+R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density &  59 &  80 & 73 & 6 &  13 & few tens \\ 
+   \hline
+glaucous gull & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  60 &  80 & 71 & 9 &   4 & few tens \\ 
+   \hline
+peregrine falcon & 2013-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   8 &  12 & 10 & 1 &   8 & units \\ 
+   \hline
+rough-legged hawk & 2007-2012 & Extrapolation from intensive nest monitoring (111 km2,
+R2=0.99, p$<$0.0001, n=8) & high & Strong correlation with study area-wide nest density &   3 &  59 & 20 & 19 &   8 & few tens \\ 
+   \hline
+rough-legged hawk & 2013-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &  66 & 30 & 29 &   8 & few tens \\ 
+   \hline
+American golden-plover & 2014-2019, 2022-2023 & Distance sampling throughout lowland (313 km2) & high & Large sample size & 306 & 1040 & 774 & 275 &   8 & hundreds \\ 
+   \hline
+cackling goose & 1996-2016 & Extrapolation from exponential model of growth (R2=0.74,
+p=0.15, n=5) & moderate & Strong correlation with opportunistic nest monitoring &   2 & 158 & 31 & 41 &  23 & units to tens \\ 
+   \hline
+cackling goose & 2017-2019, 2022-2023 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  80 & 214 & 138 & 50 &   5 & few tens \\ 
+   \hline
+Arctic fox & mean & Derived from extensive fox home range size studies (n=109) & moderate & Indirect indices, but large sample size &  &  & 53 &  &  & few tens \\ 
+   \hline
+collared lemming & 1995-2019, 2021-2022 & Rigorous density estimates at small spatial scale
+(0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 39302 & 8128 & 10334 &  27 & few thousands to tens of thousands \\ 
+   \hline
+brown lemming & 1995-2019, 2021-2022 & Rigorous density estimates at small spatial scale
+(0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 447630 & 54043 & 93530 &  27 & few thousands to few hundreds of thousands \\
+   \hline
+ermine & 1993-2019 & Indices of relative abundance derived from testimonials converted to abundance using home range size & moderate & Indirect indices and uncertainty on ermine home range size estimates &   8 & 156 & 40 & 37 &  27 & tens \\
+   \hline
+long-tailed jaeger & 2004-2019, 2022 & Intensive nest monitoring (33 km2) extrapolated by habitat & high & Relatively large spatial coverage of sampling &   0 & 900 & 272 & 285 &  17 & units to few hundreds \\
+   \hline
+red-throated loon & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  42 &  76 & 64 & 15 &   4 & few tens \\ 
+   \hline
+Pacific loon & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   6 & 4 & 3 &   4 & units \\ 
+   \hline
+tundra swan & 2017-2019, 2022 & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   2 & 1 & 1 &   4 & units \\ 
+   \hline
+common-ringed plover & 2015-2017 & Nest monitoring on the main breeding sites & moderate & Intensive monitoring, but not exhaustive to study area &  44 &  62 & 55 & 9 &   3 & few tens \\ 
+   \hline
+black-bellied plover & mean & Derived from the abundance estimate of American golden-plover using transects observations & moderate & Derived from high quality estimate of another species &  &  & 22 &  &  & few tens \\
+   \hline
+black-bellied plover & mean & Derived from the abundance estimate of American golden-plover using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 61 &  &  & few tens \\
+   \hline
+Lapland longspur & mean & Extrapolation from intensive nest monitoring (2 km2)
+and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110 &  &  & few thousands \\ 
+   \hline
+Baird's sandpiper & mean & Extrapolation from intensive nest monitoring (2 km2)
+and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2495 &  &  & several hundreds to few thousands \\ 
+   \hline
+sandhill crane & mean & Extrapolation from intensive nest monitoring (8 km2) and transect observations & moderate & Uncertain relation with large scale indices &  &  & 34 &  &  & few tens \\ 
+   \hline
+king eider & mean & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & moderate & Intensive, but opportunistic monitoring at relatively small spatial scale &  &  & 25 &  &  & tens \\
+   \hline
+king eider & mean & Derived from the abundance estimate of red-throated loon using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 106 &  &  & tens \\
+   \hline
+long-tailed duck & mean & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & moderate & Intensive, but opportunistic monitoring at relatively small spatial scale &  &  & 20 &  &  & tens \\
+   \hline
+long-tailed duck & mean & Derived from the abundance estimate of red-throated loon using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 191 &  &  & tens \\
+   \hline
+rock ptarmigan & mean & Intensive, but opportunistic nest monitoring (8 km2) extrapolated to study area & low & Intensive, but opportunistic monitoring at relatively small spatial scale and prime nesting habitat not well sampled &  &  & 24 &  &  & tens \\
+   \hline
+horned lark & mean & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362 &  &  & few hundreds \\
+   \hline
+horned lark & mean & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411 &  &  & few hundreds \\
+   \hline
+ruddy turnstone & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40 &  &  & tens \\
+   \hline
+ruddy turnstone & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 54 &  &  & tens \\
+   \hline
+red phalarope & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 156 &  &  & tens \\
+   \hline
+red phalarope & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 276 &  &  & tens \\
+   \hline
+red knot & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 64 &  &  & tens \\ 
+   \hline
+red knot & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 237 &  &  & tens \\ 
+   \hline
+white-rumped sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 1135 &  &  & hundreds \\
+   \hline
+white-rumped sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 1156 &  &  & hundreds \\
+   \hline
+buff-breasted sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 41 &  &  & tens \\
+   \hline
+buff-breasted sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 8 &  &  & tens \\
+   \hline
+pectoral sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 79 &  &  & tens \\
+   \hline
+pectoral sandpiper & mean & Derived from the abundance estimate of Bairds sandpiper using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 260 &  &  & tens \\
+   \hline
+parasitic jaeger & mean & Maximum number of nest found annually during study area-wide opportunistic nest monitoring & low & Uncertainty about number of non-breeding individuals  &  &  & 17 &  &  & few tens \\
+   \hline
+parasitic jaeger & mean & Maximum number of individuals banded in a year & low & Based on a single year and potentially not all individuals were captured &  &  & 8 &  &  & few tens \\ 
+   \hline
+common raven & mean & Derived from the abundance estimate of glaucous gull using transects observations & low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 14 &  &  & few tens \\
+   \hline
+common raven & mean & Derived from the abundance estimate of glaucous gull using incidental observations & low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 31 &  &  & few tens \\
+   \hline
+American pipit & mean & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 53 &  &  & tens to few hundreds \\
+   \hline
+American pipit & mean & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87 &  &  & tens to few hundreds \\
+   \hline
+snow bunting & mean & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 18 &  &  & few hundreds \\
+   \hline
+snow bunting & mean & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276 &  &  & few hundreds \\
+   \hline
+Arctic hare & mean & Derived from the abundance estimate of Arctic fox using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 6 &  &  & few tens \\
+   \hline
+\hline
 \label{table:summary_methods}
-\begingroup\fontsize{8pt}{10pt}\selectfont
-\begin{tabularx}{\textwidth}{lllll}
-  \hline
-Species & Sampling method & Reference species & Sampling scale & Spatial extrapolation \\ 
-  \hline
-king eider & nest sampling &  & 4x2km plot & wetlands \\ 
-  long tailed duck & nest sampling &  & 4x2km plot & wetlands \\ 
-  king eider & relat. abund. (incidental obs.) & red-throated loon & study area &  \\ 
-  long tailed duck & relat. abund. (incidental obs.) & red-throated loon & study area &  \\ 
-  Pacific loon & nest sampling &  & study area &  \\ 
-  red-throated loon & nest sampling &  & study area &  \\ 
-  cackling goose & nest sampling &  & study area &  \\ 
-  snow goose & nest sampling &  & camp 2 & goose colony \\ 
-  snow goose & vertebrate count transects &  & lowland & goose colony \\ 
-  snow goose & snow goose point counts &  & lowland & goose colony \\ 
-  tundra swan & nest sampling &  & study area &  \\ 
-  rough-legged hawk & nest sampling &  & study area &  \\ 
-  peregrine falcon & nest sampling &  & study area &  \\ 
-  snowy owl & nest sampling &  & study area &  \\ 
-  rock ptarmigan & nest sampling &  & 4x2km plot & all habitats \\ 
-  sandhill crane & nest sampling &  & 4x2km plot & transects \\ 
-  American golden-plover & distance sampling &  & lowland & transects \\ 
-  black-bellied plover & relat. abund. (transects) & A. golden-plover & lowland & transects \\ 
-  black-bellied plover & relat. abund. (incidental obs.) & A. golden-plover & study area &  \\ 
-  common-ringed plover & nest sampling &  & study area &  \\ 
-  ruddy turnstone & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
-  ruddy turnstone & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
-  red knot & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
-  red knot & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
-  pectoral sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
-  pectoral sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
-  Baird’s sandpiper & nest sampling &  & 2x1km plot & transects \\ 
-  white-rumped sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
-  white-rumped sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
-  buff-breasted sandpiper & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
-  buff-breasted sandpiper & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
-  red phalarope & relat. abund. (transects) & Baird’s sandpiper & lowland & transects \\ 
-  red phalarope & relat. abund. (incidental obs.) & Baird’s sandpiper & study area &  \\ 
-  glaucous gull & nest sampling &  & study area &  \\ 
-  long-tailed jaeger & nest sampling &  & qarlikturvik & mesic habitat \\ 
-  parasitic jaeger & maximum banded individual &  & study area &  \\ 
-  parasitic jaeger & nest sampling &  & study area &  \\ 
-  common raven & nest sampling &  & study area &  \\ 
-  common raven & relat. abund. (incidental obs.) & Parasitic jaeger & study area &  \\ 
-  horned lark & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
-  horned lark & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
-  American pipit & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
-  American pipit & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
-  lapland longspur & nest sampling &  & 2x1km plot & transects \\ 
-  snow bunting & relat. abund. (transects) & lapland longspur & lowland & transects \\ 
-  snow bunting & relat. abund. (incidental obs.) & lapland longspur & study area &  \\ 
-  brown lemming & capture marking recapture &  & two 11 ha grids & wet and mesic hab. \\ 
-  collared lemming & capture marking recapture &  & two 11 ha grids & wet and mesic hab. \\ 
-  arctic hare & incidental observations & Arctic fox & study area &  \\ 
-  ermine & relat. abund. (testimonials) &  & study area &  \\ 
-  arctic fox & home range (movement track.) &  & study area &  \\ 
-   \hline
-\end{tabularx}
+\end{longtable}
 \endgroup
-\end{table}

### 2024-09-18 script/extract_species_abundance/species/snow_goose.R

@@ -77,6 +77,7 @@ transects_sngo <- transects_2010_2013_sngo %>%
 transects_sngo_wet <- transects_sngo %>% 
   dplyr::filter(habitat== "wetland")
   
+
 #Point counts
 sngo_point <- sf::st_read("data/raw/shapefiles/sampling/snow_goose_point_counts_2010_2023.shp")
 #Identify point counts that are located in wet habitat

### 2024-09-18 script/tables_figures/table_mean_abundance.R

@@ -7,7 +7,6 @@
 #data manipulation
 library(dplyr)
 library(tidyr)
-library(sf)
 #export latex table
 library(xtable)
 
@@ -16,72 +15,52 @@ library(xtable)
 #-------------------#
 #### Import data ####
 #-------------------#
-#Read species final densities
-abundances <- read.csv("data/clean/species_abundance.csv")
-
-#Combined all data of snow goose under "combined methods
-abundances[abundances$species == "snow goose",]$data <- "combined methods"
-
-#Read study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  sf::st_drop_geometry()
-
-#------------------------------------#
-#### Table species mean abundance ####
-#------------------------------------#
-#--- Species with temporal series
-abundance_series <- 
-  abundances %>% 
-  dplyr::filter(is.na(year)== FALSE) %>% 
-  dplyr::mutate(data= gsub(" \\(.*", "", data)) %>%
-  dplyr::group_by(species, data) %>% 
- dplyr::summarise(mean_abundance= round(mean(abundance),
-                                        digits = 0),
-                  sd_abundance= round(sd(abundance), digits=0),
-                   n= n(),
-                  min_abundance= min(abundance),
-                  max_abundance= max(abundance)) 
-
-#Species with only mean
-mean_abundance <- 
-  abundances %>% 
-  dplyr::filter(is.na(year)) %>% 
-  dplyr::select(-year, -status,-zone) %>% 
-  dplyr::rename(mean_abundance= abundance) %>% 
-  dplyr::mutate(sd_abundance= NA, n= NA, min_abundance= NA, max_abundance= NA)
-
-#--- Combined together
-table <- abundance_series %>% 
-  rbind(mean_abundance) %>% 
-  dplyr::mutate(mean_abundance= as.character(mean_abundance), sd_abundance= as.character(sd_abundance)) %>% 
-  dplyr::mutate(`Mean ± sd`= ifelse(is.na(sd_abundance),  mean_abundance, paste(mean_abundance, "±", sd_abundance))) %>% 
-  dplyr::rename(Minimum= min_abundance, Maximum= max_abundance, Data=data, Species= species) %>% 
-  dplyr::select(Species, Data, `Mean ± sd`, n, Minimum, Maximum)
-
-
+#metadata
+metadata <- read.csv("data/metadata/species_method_selected.csv", sep = "\t")
 #reorder names based on defined species list
[…]

### 2024-09-18 script/tables_figures/table_species_estimate_uncertainty.R

@@ -1,59 +1,0 @@
-#Transform .csv of uncertainty on species estimate to a .tex table
-library(dplyr)
-library(xtable)
-
-
-esti_uncert <- read.csv("data/metadata/estimate_uncertainty.csv") %>% 
-  dplyr::rename(`abundance estimate`= abundance.estimate)
-
-#reorder names based on defined species list
-esti_uncert <- read.csv("data/metadata/species_metadata.csv") %>% 
-  dplyr::select(species) %>% 
-  dplyr::left_join(esti_uncert)
-
-#update species names
-esti_uncert <- esti_uncert %>% 
-  dplyr::mutate(species= dplyr::recode(species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper","lapland longspur"= "Lapland longspur"))
-
-#Rename columns 
-esti_uncert <- esti_uncert %>% 
-  dplyr::rename(Species= species, Estimate= `abundance estimate`, Confidence= confidence, Reason= reason)
-
-# test (add \newline if string if more than 55 characters)
-Reason_updated <- c()
-
-for(i in esti_uncert$Reason){
-  #space positions
-space_positions <-gregexpr(" ", i)[[1]] %>% as.vector()
- if (max(space_positions)>= 53){
-#Select first space position that over 53
-   select_space <- min(space_positions[which(space_positions>=53)])
-#Insert newline at this location
- string <- paste0("\\makecell[tl]{",substr(i, 1, select_space-1), "\\\\",substr(i, select_space +1, nchar(i)), "}")
-#Replace original string
- Reason_updated <- c(Reason_updated, string)
- }else{
-   Reason_updated <- c(Reason_updated, i)
- }
-}
-
-esti_uncert$Reason <- Reason_updated
-
-
-#as tex
[…]

### 2024-09-18 script/tables_figures/table_summary_methods_data.R

@@ -1,35 +1,0 @@
-#Title: Create a summary table of all data and method use for each species
-#Author: Louis Moisan
-#Date: June 3 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-#export .tex
-library(xtable)
-
-
-#-----------------------#
-#### Import metadata ####
-#-----------------------#
-table <- read.csv("data/metadata/summary_data_sources.csv") %>% 
-  dplyr::rename(`Sampling method`=Sampling.method,
-               `Reference species`=Reference.species,
-               `Sampling scale`=Sampling.scale,
-               `Spatial extrapolation`=Spatial.extrapolation)
-
-#as tex
-table_latex <- xtable(table, caption = "Summary of the methods and spatial scales used to estimate the density of each vertebrate species on Bylot Island, and how abundance was extrapolated at the landscape scale. When sampling consist in relative abundance, we assumed that the ratios between relative and actual abundance are the same between the targeted species and a given reference species (i.e., similar detection probability). We therefore derived the absolute abundance of the targeted species from its relative abundance using the ratio between relative and absolute abundances of the reference species. Additionally, here the area of each sampling scale: two 11 ha grids (0.22 km2), 2x1km plot (2 km2) 4x2km plot (8 km2), south qarlikturvik (33 km2), camp 2 (80 km2), lowland (313 km2) and whole study area (389 km2).",
-                      label= "table:summary_methods")
-print(table_latex,
-      include.rownames = FALSE,
-      tabular.environment = "tabularx",
-      table.placement = "H",
-      width="\\textwidth",
-      align="rXXXXXX",
-      size="\\fontsize{8pt}{10pt}\\selectfont",
-      file = "manuscript/tables/table_summary_methods.tex",
-      caption.placement = "top")
-

### 2024-09-20 MetadataS1/MetadataS1.tex
prepare manuscript for Ecology journal

--- 2024-09-18 manuscript/manuscript.tex
@@
 \usepackage{parskip}
+\usepackage{enumitem}
 
@@
 \usepackage{textcomp}
-\usepackage[table,xcdraw]{xcolor}
+\usepackage[table,xcdraw, dvipsnames]{xcolor}
 \usepackage{tabularx}
@@
 \usepackage{pdflscape}
+
 %bibliography
@@
 
+\usepackage{titlesec}
+\titlespacing{\section}{0pt}{*4}{*1.5} 
+\titlespacing{\subsection}{20pt}{*4}{*1.5}
+\titlespacing{\subsubsection}{40pt}{*4}{*1.5} 
+\titlespacing{\paragraph}{60pt}{*4}{*1.5}
+\titlespacing{\subparagraph}{80pt}{*4}{*1.5}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \centering
-\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
-\endgroup
+\textbf{Title: Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 
@@
 \textsuperscript{4} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
-\textsuperscript{5} Centre d’études nordiques, Centre de connaissance et d’exploration de l’Arctique, Musée canadien de la nature, Ottawa, ON, Canada\\
+\textsuperscript{5} Centre d’études nordiques, Centre for Arctic Knowledge and Exploration, Canadian Museum of Nature, Ottawa, ON, Canada\\
 \textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
@@
 
-\textbf{Corresponding Authors}:\\
+\textbf{Corresponding Author}:\\
 \textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
@@
 
-\section*{Abstract}
-Arctic ecosystems present unique opportunities for community-wide monitoring, in part due to their relatively low species richness.
-However, conducting research in these remote environments poses significant logistical challenges, resulting in long-term monitoring being exceedingly rare.
-Here, we focus on the long-term, intensive ecological monitoring efforts conducted on the south plain of Bylot Island (almost 400 km\textsuperscript{2}, Nunavut, Canada), which has generated a remarkable dataset spanning up to 30 years, a rarity in tundra ecosystems.
-Our goal is to synthesize this dataset and upscale vertebrate abundance data at the landscape level, a prerequisite to conduct community-level analyses.
-We have standardized data obtained with different field methods to provide readily usable long-term time series of abundance for 35 vertebrate species (30 birds and 5 mammals) present in the study system.
-Monitoring data includes intensive capture-mark-recapture density estimates of lemmings on trapping grids, systematic or opportunistic nest monitoring conducted across the entire study area or within specific plots for all bird species, transects of vertebrate counts distributed throughout the study area, daily incidental observations of vertebrates and satellite tracking of fox movements.
-Annual abundance of species was estimated at the landscape level, accounting for spatial variations.
-Furthermore, we provide body masses for each species, derived from empirical onsite measurements for 18 species and from the literature for the remaining species.
-Body mass is essential to convert species abundance into biomass for studies of trophic fluxes and ecosystem processes.
-Our dataset provides a unique opportunity for holistic empirical studies of ecological communities, allowing a deeper understanding of community structure and dynamics.
-Considering that the study site is a pristine and protected area that has experienced minimal anthropogenic impact, it can also provide an ideal baseline for investigating the impacts of global changes on high-latitude terrestrial ecosystems.
+\textbf{Open Research statement}:
 
-\section*{Introduction}
+%%%%%%% !!!! ADD DRYAD adress !!!! %%%%%%%%
+Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
+Raw data and codes used to extract the presented data set are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset}.
+\endgroup
+\newpage
+
+\subsection*{Introduction}
  The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology.
@@
 
-\section{Study area}
-The study area (389 km\textsuperscript{2}) is located on the south plain of Bylot Island, Nunavut, Canada.
-It comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
-Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
-The mean temperature in July is 6\textdegree C, and the area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
-An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}.
+\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
+    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
+    \section*{B. Data set identification codes}
+     \texttt{BYLOT-species\_taxonomy.csv}\\
+     \texttt{BYLOT-species\_abundance.csv}\\
+     \texttt{BYLOT-mean\_species\_abundance.csv}\\
+     \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
+     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\
+
+    \section*{C. Data set description}
+		\subsection*{1. Originators}
+		\textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
+  		\textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
+  		\textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+
+		\subsection*{2. Abstract}
+Arctic ecosystems present unique opportunities for community-wide monitoring, in part due to their relatively low species richness.
+However, conducting research in these remote environments poses significant logistical challenges, resulting in long-term monitoring being exceedingly rare.
+Here, we focus on the long-term, intensive ecological monitoring efforts conducted on the south plain of Bylot Island (almost 400 km\textsuperscript{2}, Nunavut, Canada), which has generated a remarkable dataset spanning up to 30 years, a rarity in tundra ecosystems.
+Our goal is to synthesize this dataset and upscale vertebrate abundance data at the landscape level, a prerequisite to conduct community-level analyses.
+We have standardized data obtained with different field methods to provide readily usable long-term time series of abundance for 35 vertebrate species (30 birds and 5 mammals) present in the study system.
+Monitoring data includes intensive capture-mark-recapture density estimates of lemmings on trapping grids, systematic or opportunistic nest monitoring conducted across the entire study area or within specific plots for all bird species, transects of vertebrate counts distributed throughout the study area, daily incidental observations of vertebrates and satellite tracking of fox movements.
+Annual abundance of species was estimated at the landscape level, accounting for spatial variations.
+Furthermore, we provide body masses for each species, derived from empirical onsite measurements for 18 species and from the literature for the remaining species.
+Body mass is essential to convert species abundance into biomass for studies of trophic fluxes and ecosystem processes.
+Our dataset provides a unique opportunity for holistic empirical studies of ecological communities, allowing a deeper understanding of community structure and dynamics.
+Considering that the study site is a pristine and protected area that has experienced minimal anthropogenic impact, it can also provide an ideal baseline for investigating the impacts of global changes on high-latitude terrestrial ecosystems.
+
+   \section*{D. Key words/phrases}
+ Bylot Island, Canadian Arctic, Arctic tundra, 1993-2023, long-term monitoring, biodiversity monitoring, community composition, species abundance, species density, species biomass, species body mass, food web
+  \newpage
+
+\section*{\textcolor{Blue}{Class II. Research origin descriptors}}
+    \section*{A. Overall project description}
+    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial vertebrate communities
+
+	\subsection*{2. Originators:}	
+  	\textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
+  		\textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
+  		\textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+
+	\subsection*{3. Period of study} 1989 - continuing
+
+	\subsection*{4. Objectives}
+       i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial vertebrate communities.\\
+       ii) Predict the effects of current global environmental changes on the structure and dynamics of Arctic terrestrial vertebrate communities.
+
+     \subsection*{5. Abstract}
+        Arctic terrestrial vertebrate communities present low species richness, making those relatively simple communities ideal for studying ecological patterns and dynamics in terrestrial environments.
+Although, they present complex networks of interacting species, extreme seasonal changes of environmental conditions and a large portion of migratory species, which make challenging the identification of the key factors that shape their structure and dynamics.
+In the face of rapid global environmental changes, it is crucial to have a comprehensive understanding of the key processes shaping Arctic terrestrial communities structure and dynamics in order to predict how global changes will impact them in the future.
+Our research emphasizes long-term biodiversity monitoring, a community-wide perspective and food web modeling to achieve this understanding.
+
+      \subsection*{6. Sources of funding}
+Natural Sciences and Engineering Research Council of Canada, Fonds Québécois de Recherche Nature et Technologies, Centre d'études nordiques, Natural Resources Canada (Polar Continental Shelf Program), Network of Centers of Excellence Canada (ArcticNet), Canada First Research Excellence Fund (Sentinel North Program), Polar Knowledge Canada, Environment and Climate Change Canada, Canada Foundation for Innovation, Parks Canada Agency, International Polar Year program of the Government of Canada, Crown–Indigenous Relations and Northern Affairs Canada (Northern Contaminant Program), Duck Unlimited Canada, Kenneth M.
+Molson Foundation (Kenneth M.
+Molson Foundation’s donation for wildlife research, conservation, and habitat), Garfield Weston Foundation, First Air——Canadian North, Nunavut Wildlife Management Board, Université Laval, Université du Québec à Rimouski
+\newpage
+
+    \section*{B. Specific subproject description}
+     	\subsection*{1. Site description}
+ 			\subsubsection*{a. Site type}
+  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}.
 The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}).
@@
 We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}.
-The case of the red fox (Vulpes vulpes) was ambiguous.
+The case of the red fox (\textit{Vulpes vulpes}) was ambiguous.
 While the presence of a breeding pair has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare.
@@
 
+         		\input{tables/species_name_strategy.tex}
+%label-> table:species_name_strategy
+\newpage
+            	\subsubsection*{b.
+Geography} Our 463 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
+            	\subsubsection*{c.
+Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
+            	\subsubsection*{d. Geology} see \citet{klassen1993} for a complete and detailed description of the geology of the study area.
+            	\subsubsection*{e.
+Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
+            	\subsubsection*{f. Site history} see \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
+            	\subsubsection*{g.
+Climate} The mean temperature in July is 6°C, and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
+The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds.
+\citep{gauthier2024a}.
+ \newpage
+
+      	\subsection*{2. Experimental or sampling design}
+        		\subsubsection*{a. Permanent plots}
 The study area is divided into 9 zones based on the sampling method and the level of field effort applied in each zone (\textbf{Figure \ref{figure:zones}}).
 Long-term monitoring of the community began in the 1990s in the Qarlikturvik valley \citep{gauthier2013, gauthier2024a}, which represents the zone of the study area with the highest annual sampling effort.
-Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river, where the main research infrastructure is located.
+Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river (\textbf{Figure \ref{figure:qarlikturvik_valley}}), where the main research infrastructure is located.
 Another zone with extensive sampling efforts is Camp 2, located at the core of the snow goose colony, where the primary focus is to monitor snow goose nests.
@@
   \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
-   \vspace{-35pt} % Reduce space between figure and caption
-   \caption{Map of the different zones (colored polygons) of the 389 km\textsuperscript{2} study area located on the south plain of Bylot Island, Nunavut Canada.
-The perimeter of the snow goose colony is delineated by white dashes; we highlighted the perimeter in 2017 since it represents the average colony area.}
+   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada.
+The outline of the snow goose colony is presented with white dash; we used the outline of the colony in 2015 since it represents an average colony area (67 km\textsuperscript{2}).}
   \label{figure:zones}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.8\textwidth, angle=0]{figures/qarlikturvik_valley.pdf}
+  \vspace{-70pt} % Reduce space between figure and caption
+  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}
+ \label{figure:qarlikturvik_valley}
 \end{figure}
@@
 
-\input{tables/species_name_strategy.tex}
-%label-> table:species_name_strategy
-\newpage
-
-\section{Data collection}
-\subsection{Avian nest monitoring}
+		\subsubsection*{b. Avian nest monitoring}
 Avian nest monitoring was conducted annually except in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic.
@@
 %label-> table:species_year_monitoring
-\newpage
 
-\subsubsection{Systematic nest monitoring}
-\subsubsection*{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}
+	\begin{enumerate}[label=\roman*]
+	\item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}\\
 	Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area.
@@
 Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays.
-We are confident that nest detection probability was high for these species given the open landscape.
+We are confident that nest detection probability was high for these species given the open landscape.\\
 
-\subsubsection*{Snow goose}
+	\item[] \textit{\textbf{Snow goose}}\\
 	Snow geese nest in a large colony in the study area (\textbf{Figure \ref{figure:zones}}), but also in small aggregations distributed on the island, especially in years when snowy owls are nesting \citep{lepage1996,reed2002}.
@@
 
-\subsubsection*{Rough-legged hawk, peregrine falcon and common raven}
+	\item[] \textit{\textbf{Rough-legged hawk, peregrine falcon and common raven}}\\
 	Peregrine falcons, rough-legged hawks and common ravens nest on cliffs, near ravines, and on large rocky outcrops and tend to reuse the same nesting sites from one year to the next \citep{beardsell2016}.
@@
 
-\subsubsection*{Snowy owl}
+	\item[] \textit{\textbf{Snowy owl}} \\
 	Snowy owls predominantly nest in habitats similar to other raptors, favoring ridges in mountainous or hilly regions, although they can occasionally be found nesting on mounds in lowland areas \citep{seyer2020}.
@@
 Additionally, since 2012, nests have been recorded across the entire study area by scanning the landscape from hills and ridges during the nesting period \citep{duchesne2021}.
-Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.
+Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.\\
 
-\subsubsection*{Long-tailed jaeger and parasitic jaeger}
+	\item[] \textit{\textbf{Long-tailed jaeger and parasitic jaeger}} \\
 	Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period.
@@
 We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability.
-We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).
+We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).\\
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.8\textwidth, angle=0]{figures/qarlikturvik_valley.pdf}
-  \vspace{-70pt} % Reduce space between figure and caption
-  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}
- \label{figure:qarlikturvik_valley}
-\end{figure}
-
-\subsubsection*{Common-ringed plover}
+	\item[] \textit{\textbf{Common-ringed plover}} \\
 	Between 2015 and 2019, observers conducted surveys of the primary nesting areas of the common-ringed plover.
@@
 The sampling effort was particularly intensive between 2015 and 2017.
-Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.
+Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.\\
 
-\subsubsection*{Lapland longspur and Baird's sandpiper}
+	\item[] \textit{\textbf{Lapland longspur and Baird's sandpiper}}\\
 	Since 2005, nests of passerines and sandpipers have been extensively monitored across an 8 km\textsuperscript{2} (4x2 km) area in the Qarlikturvik valley.
@@
 Assuming the observer can detect all nests within a 5 or 10 meter radius, analysis of daily GPS tracks shows that the observer covered a minimum area of 0.72 \textpm{0.12} (5 m) or 1.09 \textpm{0.17} km\textsuperscript{2} (10 m) of the core area annually (n= 3 years).
-Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.
+Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.\\
+
+	\item[] \textit{\textbf{Opportunistic nest monitoring}}\\
+	Since 2005, we also noted the nest location of any other bird species encountered opportunistically during travel or while carrying out the protocols for the previously described species.
+The sampling was particularly intensive in the defined 8 km\textsuperscript{2} (2x4 km plot) area in the Qarlikturvik valley.
+The accuracy of nest monitoring in this plot thus depends on the species detection probability.
+We are confident to obtain a realistic order of magnitude for the number of nests present for relatively large bodied species in this area (i.e., sandhill crane, rock ptarmigan, long-tailed duck and king eider).
+Additionally, starting in 2009, a significant effort has been made each year, though not systematically, to visit known nesting territories of parasitic jaegers throughout the study area.
+	\end{enumerate}
 
@@
 
-\subsubsection{Opportunistic nest monitoring}
-Since 2005, we also noted the nest location of any other bird species encountered opportunistically during travel or while carrying out the protocols for the previously described species.
-The sampling was particularly intensive in the defined 8 km\textsuperscript{2} (2x4 km plot) area in the Qarlikturvik valley.
-The accuracy of nest monitoring in this plot thus depends on the species detection probability.
-We are confident to obtain a realistic order of magnitude for the number of nests present for relatively large bodied species in this area (i.e., sandhill crane, rock ptarmigan, long-tailed duck and king eider).
-Additionally, starting in 2009, a significant effort has been made each year, though not systematically, to visit known nesting territories of parasitic jaegers throughout the study area.
-
-\subsection{Observation of individuals}
-\subsubsection{Vertebrate count transects}
+\subsubsection*{c. Observation of individuals}
+\begin{enumerate}[label=\roman*]
+\item[] \textit{\textbf{Vertebrate count transects}}\\
 From 2010 to 2023, observers walked 500-meter linear transects where all vertebrate individuals observed within 150 meters on either side were counted (146 to 320 transects per year).
@@
 Furthermore, specifically for American golden-plovers, we measured the distance of each observed individual to the transect path.
-
+\\
 \begin{figure}
@@
 \end{figure}
+\newpage
 
-\subsubsection{Snow goose point count}
+\item[] \textit{\textbf{Snow goose point count}}\\
 At the start, middle, and end of each vertebrate count transect, a point count with a radius of 125 meters was conducted to determine the number of snow goose breeding pairs.
 On average, 613 \textpm{142} point counts were sampled each year, covering an area of 30 \textpm{7} km\textsuperscript{2}.
+\\
 
-\subsubsection{Incidental observations}
+\item[] \textit{\textbf{Incidental observations}}\\
 Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}.
@@
 Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands.
+\\
 
@@
 
-\subsubsection{Testimonials of ermine sightings}
+\item[] \textit{\textbf{Testimonials of ermine sightings}}\\
 There was no direct estimation of ermine abundance on Bylot Island as they are quite difficult to obtain.
@@
 Scores of individual participants were averaged annually as detailed in \citet{bolduc2023}.
+\end{enumerate}
 
-\subsection{Capture of individuals}
-\subsubsection{Lemming trapping}
+\subsubsection*{d. Capture of individuals}
+\begin{enumerate}[label=\roman*]
+\item[] \textit{\textbf{Lemming trapping}}\\
 Since 2004, brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids.
@@
 Density of each species was estimated at each occasion using spatially explicit capture-recapture methods (see \citet{fauteux2015} for details).
-From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids
-\citep{gruyer2008}.
+From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids \citep{gruyer2008}.
 Index of abundance derived from snap-trapping were transformed in density estimates in each habitat for the period 1995-2003 using the equation provided by \citet{fauteux2018} based on the period of overlap between the two sampling methods (2004 to 2016).
+\\
 
-\subsubsection{Arctic fox movement tracking}
+\item[] \textit{\textbf{Arctic fox movement tracking}}\\
 In order to assess fox abundance based on the size of their home range, 109 Arctic foxes were fitted with Argos Platform Transmitter Terminals mounted on collars between 2008 and 2016 \citep{lai2015,christin2015, dulude2023}.
@@
 Sampling of animal locations was set for an interval of 1 or 2 days and only locations between May 1 and October 30 were retained \citep{dulude2023}.
+\\
 
-\subsubsection{Parasitic jaeger banding}
+\item[] \textit{\textbf{Parasitic jaeger banding}}\\
 In 2009, a significant effort was made to band as many parasitic jaegers as possible within the study area.
 This effort resulted in the banding of 17 adult individuals (Therrien and Gauthier, unpublished data).
-\newpage
+\end{enumerate}
 
-\subsection{Species body mass}
+\subsubsection*{e. Species body mass}
 All vertebrate individuals captured for marking purposes were systematically weighed (snow goose (unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021},  common-ringed plovers \citep{leandri2019}, other shorebirds (Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (Gauthier and Therrien, unpublished data), Lapland longspurs (Gauthier and Bêty, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}).
@@
 When not available, we extracted mean body mass from the literature \citep{wilman2014}.
-
-\input{tables/body_mass.tex}
-%label-> table:body_mass
 \newpage
 
-\section{Data preparation}
-We estimated the abundance of breeding individuals for most species, but there were a few exceptions.
-For common ravens, parasitic jaegers, long-tailed ducks, and king eiders, we suspect the presence of a significant number of non-breeding individuals in the study area.
-Therefore, the estimates we provided for these species include both breeding and potentially non-breeding individuals.
-Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, ermine, and Arctic hare.
-The methods used for each species are summarized in (\textbf{Table \ref{table:summary_methods}}).
+    \subsection*{3.  Research methods}
+    		\subsubsection*{1. Field/laboratory}
+    		\begin{enumerate}[label=\alph*.]
+        			\item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous
+gull}}
+        			\item[] \textit{\textbf{Snow goose}}
+        			\item[] \textit{\textbf{King eider and long-tailed duck}}
+        			\item[] \textit{\textbf{Rough-legged hawk, peregrine falcon and snowy owl}} 
+        			\item[] \textit{\textbf{Rock ptarmigan}}
+        			\item[] \textit{\textbf{Sandhill crane}}
+        			\item[] \textit{\textbf{American golden-plover and black-bellied plover}}
+        			\item[] \textit{\textbf{Common-ringed plover}}
+        			\item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}
+        			\item[] \textit{\textbf{Other passerines and sandpipers}}
+        			\item[] \textit{\textbf{Long-tailed jaeger}}
+        			\item[] \textit{\textbf{Parasitic jaeger}}
+        			\item[] \textit{\textbf{Common raven}}
+        			\item[] \textit{\textbf{Brown and collared lemming}}
+        			\item[] \textit{\textbf{Arctic hare}}
+        			\item[] \textit{\textbf{Ermine}}
+        		\end{enumerate}
 
-R software (version 4.3.2) was used to process the data on a linux-gnu machine.
-Most of data manipulation was carried out with the \textit{dplyr} \citep{dplyr} and \textit{sf} packages \citep{sf}.
-All codes are available in the following repository \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset.git}.
-\newpage
+        		\subsubsection*{2. Instrumentation}
 
-\subsection{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}
-Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022.
-We transformed the number of nests to abundance of individuals by multiplying it by a factor of two.
+        		\subsubsection*{3. Taxonomy and systematics}
 
-We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4).
-Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12406x + 0.13775$).
-However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series.
-Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area.
-We thus fitted an exponential model using the number of nests found annually over two distinct periods: in 1996 when the first nest was discovered, and then from 2017 to 2023 when sampling effort was systematic across the whole study area (\textbf{Figure \ref{figure:cackling}}).
-We used the fitted model to estimate abundance between 1996 and 2016 when monitoring was less systematic, which could potentially underestimate observed abundance as seen on \textbf{Figure \ref{figure:cackling}}.
-We multiplied nest density by two to obtain the abundance.
+        		\subsubsection*{4. Permit history}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
-\caption{Number of cackling goose nests found across the study area over time.
-The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016.
-The model presents a strong fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area (R\textsuperscript{2}= 0.74, p= 0.15, n= 5).}
-\label{figure:cackling}
-\end{figure}
+        		\subsubsection*{5. Legal/organizational requirements}
 
-\subsection{Snow goose}
-Between 1999 and 2023, we assessed the abundance of snow geese in the study area through a multi-step process.
-We calculated the mean annual density of snow goose nests separately in the mesic and wetland habitats of the area occupied by the goose colony annually.
-We made slight adjustments to the goose colony perimeter defined from helicopter surveys to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}).
-To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled.
-The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts.
-Systematic nest searches were highly precise, but covered a relatively small area, whereas transects and snow goose point counts were less precise but covered larger areas.
-For each method, we calculated the mean density of breeding individuals in mesic habitat by dividing the number of birds (or nests) recorded by the area sampled.
-Despite methodological differences, the three approaches showed similar inter-annual variations, supporting the use of a mean values to estimate nest density in mesic habitat \textbf{Figure \ref{figure:sngo_interannual_variation}}).
-Lastly, to transform the densities in total abundance, we determined the annual proportion of wetland and mesic habitats within the goose colony and multiplied the area of each habitat by the density of breeding individuals.
-For the period 1999 to 2009, we used the average limits of the colony over the period 2010 to 2023 because we did not conduct aerial survey of the colony.
-Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).
+        \subsubsection*{4. Project personnel: Principal and associated investigator(s), technicians, supervisors, students}
+ \section*{\textcolor{Blue}{Class III. Data set status and accessibility}}
+    \subsection*{A. Status}
+    \begin{enumerate}
+      \item Latest update: Date of last modification of data set
+      \item Latest archive date: Date of last data set archival
+      \item Metadata status: Date of last metadata update and current status
+      \item Data verification: Status of data quality assurance checking
+    \end{enumerate}
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\linewidth]{figures/map_geese_colony_2017.pdf}
-\caption{Map showing the region occupied by the snow goose colony in 2017 (green polygon) as an example.
-The perimeter was first defined from a helicopter surveys and then slightly adjusted to include all snow goose point counts where at least one breeding pair had been observed in that year (green dots).
-Snow goose point counts where no breeding geese were observed in that year are presented as red dots.}
-\label{figure:sngo_goose_colony}
-\end{figure}
+    \subsection*{B. Accessibility}
+    		\begin{enumerate}
+    			\item Storage location and medium: Pointers to where data reside (including archival required by the ESA Open Research policy and redundant archival sites)
+        		\item Contact persons: Name, address, phone, email
+        		\item Copyright restrictions: Whether copyright restrictions prohibit use of all or portions of the data set
+        		\item Proprietary restrictions: Any other restrictions that may prevent use of all or portions of data set
+        			\begin{enumerate}[label=\alph*.]
+            			\item Release date: Date when proprietary restrictions expire
+           			\item Citation: How data may be appropriately cited
+            			\item Disclaimer(s): Any disclaimers that should be acknowledged by secondary users
+            		\end{enumerate}
+        		\item Costs: Costs associated with acquiring data (may vary by size of data request, desired medium, etc.)
+    		\end{enumerate}
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\linewidth]{figures/snow_goose_mesic_density.pdf}
-\caption{Estimates of breeding goose density in mesic habitat of the Bylot Island study area using three independent methods: Nest sampling plot, Snow goose point count, Vertebrate count transect.}
-\label{figure:sngo_interannual_variation}
-\end{figure}
+ \section*{\textcolor{Blue}{Class IV. Data structural descriptors}}
+    \subsection*{A. Data set file}
+    		\begin{enumerate}
+    			\item Identity: Unique file names or codes
+        		\item Size: Number of records, record length, total number of bytes, etc.
+        		\item Format and storage mode: File type (e.g., ASCII, binary, etc.), compression schemes employed (if any), etc.
+        		\item Header information: Description of any header data or information attached to file [Note: may include elements related to variable information (IV.B.); if so, could be linked to appropriate section(s)]
+        		\item Alphanumeric attributes: Mixed, upper, or lower case
+        		\item Special characters/fields: Methods used to denote comments, flag modified or questionable data, etc.
+        		\item Authentication procedures: Digital signature, checksum, actual subset(s) of data, and other techniques for assuring accurate transmission of data to secondary users
+    		\end{enumerate}
 
-In 1993, 1998, 2003, and 2008, the Canadian Wildlife Service conducted aerial surveys over the southern plain of Bylot Island during the brood-rearing period to estimate the abundance of breeding snow geese (Reed et al., 2002).
-We corrected these estimates with the annual nesting success, as these surveys were conducted after the hatching period.
-For 2003 and 2008, we used in priority estimates of abundance derived from aerial surveys instead of nest sampling.
+    \subsection*{B. Variable information}
+    \begin{enumerate}
+    		\item Variable identity: Unique variable name or code
+        \item Variable definition: Precise definition of variables in data set
+        \item Units of measurement: Units of measurement associated with each variable
+        \item Data type
+        		\begin{enumerate}[label=\alph*.]
+            		\item Storage type: Integer, floating point, character, string, etc.
+            		\item List and definition of variable codes: Description of any codes associated with variables
+            		\item Range for numeric values: Minimum, maximum
+            		\item Missing value codes: Description of how missing values are represented in data set
+            		\item Precision: Number of significant digits
+            \end{enumerate}
+        \item Data format
+        		\begin{enumerate}[label=\alph*.]
+        			\item Fixed, variable length
+            		\item Columns: Start column, end column
+            		\item Optional number of decimal places
+        		\end{enumerate}
+    \end{enumerate}
 
-\subsection{King eider and long-tailed duck}
-We estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley.
-We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}).
-We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two.
-We acknowledge that the opportunistic monitoring of these species likely underestimated their true nest density.
-However, considering the extensive sampling effort deployed annually within this area, we are confident to obtain a realistic order of magnitude for the number of nests present.
-Because duck sightings are frequent, yet only a few nests are found, we believe there may be non-breeding individuals.
-Therefore, we employed an additional method to estimate the overall duck populations without differentiating between breeding and non-breeding individuals.
+    \subsection*{C. Data anomalies: Description of missing data, anomalous data, calibration errors, etc.}
+ \section*{\textcolor{Blue}{Class V. Supplemental descriptors}}
+    \subsection*{A. Data acquisition}
+    \begin{enumerate}
+       \item Data forms or acquisition methods: Description or examples of data forms, automated data loggers, digitizing procedures, etc.
+       \item Location of completed data forms
+       \item Data entry verification procedures: Procedures employed to verify that digital data set is free of errors
+    \end{enumerate}
 
-As an alternative approach, we estimated the abundance of ducks based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}.
-We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) in duck and loon species.
-We therefore derived the absolute abundance of long-tailed ducks and king eiders from their relative abundances using the ratio between relative and absolute abundances of red-throated loons as a reference.
+    \subsection*{B.
+Quality assurance/quality control procedures}Identification and treatment of outliers, description of quality assessments, calibration of reference standards, equipment performance results, etc.
+    \subsection*{C. Related materials} References and locations of maps, photographs, videos, GIS data layers, physical specimens, field notebooks, comments, etc.
+    \subsection*{D. Computer programs and data-processing algorithms}
+    		\begin{enumerate}
+    			\item \textbf{Program}: R version 4.3.2 (2023-10-31)
+    			\item \textbf{Operating system}: x86\_64-pc-linux-gnu (64-bit) with Ubuntu 22.04.3 LTS
+    			\item \textbf{Packages}: dplyr, RPostgreSQL, sf, stringr, tidyr, xtable, Distance, ggplot2, ggmap, png, grid
+    		\end{enumerate}
+    \subsection*{E. Archiving}
+    		\begin{enumerate}
+    		 \item \textbf{Archival procedures}: Data will be archived on Dryad upon acceptance for publication
+        	 \item \textbf{Redundant archival sites}: n/a
+    		\end{enumerate}
 
-\subsection{Rough-legged hawk, peregrine falcon and snowy owl}
-We estimated the abundance of breeding rough-legged hawks, peregrine falcons and snowy owls based on systematic nest monitoring conducted throughout the study area for these species.
-To convert the number of nests into breeding abundance, we multiplied it by two.
-For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.99; p < 0.0001, n= 10).
-We used the same approach for rough-legged hawks ($y=0.49851x$, R\textsuperscript{2}= 0.99, p < 0.0001, n= 8) to extend the time series from 2007 to 2012.
-We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.44, p= 0.27, n=8).
+    \subsection*{F. Publications and results: Electronic reprints, lists of publications resulting from or related to the study, graphical/statistical data representations, etc.}
+\textbf{List of publications that include data from the data set}:
+		\begin{enumerate}
+			\item 
+		\end{enumerate}
 
-\subsection{Rock ptarmigan}
-We estimated the abundance of rock ptarmigans based on the annual nest density measured in the 8 km\textsuperscript{2} extensive nest search area of the Qarlikturvik valley.
-While we acknowledge that the opportunistic monitoring of this species likely underestimates nest density, the extensive sampling effort deployed annually within this area gives us confidence in obtaining a realistic number of nests.
-We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:species_relative_abundance_upland}}).
-Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat.
-To convert the number of nests into breeding abundance, we multiplied it by two.
+   \subsection*{ G. History of data set usage}
+  		\begin{enumerate}
+  			\item \textbf{Data request history}: n/a
+        		\item \textbf{Data set update history}: n/a
+        		\item \textbf{Review history}: n/a
+       		\item \textbf{Questions and comments from secondary users}: n/a
+  		\end{enumerate}
 
-\subsection{Sandhill crane}
-We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}).
-In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed.
-We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density.
-However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present.
-Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2.
-We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
-\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin).
-The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
-Data points represent annual values.}
-\label{figure:sacr}
-\end{figure}
-
-\subsection{American golden-plover and black-bellied plover}
-We applied a distance sampling approach to the observations of American golden-plovers made along vertebrate count transects to estimate the abundance of individuals in the mesic habitat of the study area between 2014 and 2023.
-We used perpendicular distance between detected individuals and transect (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}.
-We applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered).
-We selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term.
-We excluded observations of more than four individuals, as these likely indicated groups of non-breeders passing through the area.
-We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}.
-We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
-
-To determine the abundance of black-bellied plovers, we used the mean number of black-bellied plovers and American golden-plovers observed per transect as an index of relative abundance.
-We assumed that the ratios of between relative and actual abundance are the same (i.e., similar detection probability) among those species.
-This assumption is realistic as those species present similarities in size, color, and reproductive behavior.
-We therefore derived the absolute abundance of black-bellied plovers from their relative abundance using the ratio between relative and absolute abundances of American golden-plover as a reference.
-
-As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.
-
-\subsection{Common-ringed plover}
-To estimate the abundance of common-ringed plovers in the study area, we relied on the total number of nests recorded annually from 2015 to 2017, during which the primary nesting sites underwent intensive sampling.
-We multiplied the total nest count by two to represent the abundance of breeding individuals.
-
-\subsection{Lapland longspur and Baird's sandpiper}
-We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}).
-For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected.
-We made this adjustment because this species was less frequently observed.
-In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located.
-This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically.
-Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2.
-We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
-We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.
-
-\begin{figure}[H]
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=0.95\linewidth]{figures/lapland_longspur_nest_transect.pdf} 
-\caption{}
-\end{subfigure}
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=0.95\linewidth]{figures/bairds_sandpiper_nest_transect.pdf}
-\caption{}
-\end{subfigure}
-\caption{a) Linear regression between the nest density of Lapland longspurs and the number of individuals observed per transect (nest density= 2.3422 x number of individuals per transect; regression was forced through the origin).
-The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
-Data points represent annual values.
-b) Linear regression between the nest density of Baird’s sandpiper and the proportion of transect with at least one individual observed (nest density= 34.9248 x proportion of transects with at least one individual; regression was forced through the origin).
-The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
-Data points represent annual values.}
-\label{figure:lalo_basa}
-\end{figure}
-
-\subsection{Other passerines and sandpipers}
-We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section 3.9).
-We assumed here a similar detection probability for all species.
-We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section 3.9).
-For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
-Nest density was then converted in number of individuals by multiplying by the area and a factor 2.
-
-As an alternative approach, we estimated the abundance of other passerines and sandpipers based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}.
-We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among both passerine and sandpiper species.
-We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.
-
-\subsection{Long-tailed jaeger}
-We determined the annual nest density of long-tailed jaegers from the systematic nest sampling between 2004 and 2023 on the southern side of the glacial river in the Qarlikturvik valley.
-We determined nest density by dividing the annual number of nests recorded by the area of the surveyed zone (33 km\textsuperscript{2}).
-As long-tailed jaegers typically nest in mesic habitat \citep{andersson1971}, we multiplied the area occupied by mesic habitat across the study area by the nest density obtained in the surveyed zone and by two to obtain the total abundance of breeding individuals.
-
-\subsection{Parasitic jaeger}
-Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 3.75 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}.
-This suggests that there may be non-breeding individuals present at the study site, or alternatively, individuals may regularly travel long distances, potentially from outside the study area, to forage during the breeding season.
-Due to limited data availability for estimating the abundance of non-breeding parasitic jaegers, we relied on the maximum number of adults banded during a single year (17 individuals in 2009; Therrien, unpublished data) as the minimum abundance on the study area.
-This corresponds to a density of 0.04 individuals/km\textsuperscript{2}.
-For comparison, \citet{taylor1974} measured a density of 0.06 individual/km\textsuperscript{2} on Bathurst Island.
-
-\subsection{Common raven}
-Despite the intensive nest searches for raptors in upland zones, we never found more than one common raven nest each year, a small number considering the frequent raven observations at the study site \citep{gauthier2024a}.
-This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging.
-Therefore, we opted for alternative approaches based on individual counts to estimate the abundance of both breeding and non-breeding ravens.
-
-As a first approach, we based our estimate of ravens on the number of glaucous gulls observed per transect.
-We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species.
-This assumption is reasonable as those species present similarities in size and foraging strategy.
-We therefore derived the absolute abundance of common ravens from their relative abundance using the ratio between relative and absolute abundances of glaucous gulls as a reference.
-
-Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.
-
-\subsection{Brown and collared lemming}
-Between 1995 and 2003, we used the density estimates derived from the snap-trapping indices obtained in late July in each habitat.
-Between 2004 and 2007, annual abundance of each lemming species was based on the late-July density estimates on trapping grid in wet and mesic habitats.
-However, starting from 2008, estimates were derived from the mean density recorded in mid-July and mid-August, except for two instances: 2019 and 2021.
-In 2019, due to an exceptionally early snowmelt and thus an early decline in lemmings during the summer, we only retained value from mid-July.
-In 2021, we relied solely on data gathered in August because it was the only trapping period carried out that year.
-To scale the estimated densities from the wet and mesic grids to the entire study area, we used the proportions of mesic habitats (64\%) and wet habitats (10\%) measured within the study area.
-
-\subsection{Arctic hare}
-Arctic hares are primarily observed in the upland zones of the study area, where sampling effort is limited.
-We thus derived abundance of hares from the estimated abundance of Arctic foxes based on indices of relative abundance presented in \citep{gauthier2024a}, which were derived from incidental daily observations.
-We doubled the density of Arctic hares in the upland zones (i.e., plateaus), as twice as many individuals were observed per hour of fieldwork there compared to lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
-However, it is worth noting that assuming a similar detection probability between foxes and hares might lead to an overestimation of hare detection probability due to behavioral differences between the species.
-Therefore, we most likely underestimate the actual abundance of Arctic hares in the study area.
-
-\subsection{Ermine}
-We estimated the annual abundance of ermines by transforming the annual index of relative abundance provided in \citet{bolduc2023} into individual density.
-Annual values ranged from 0, indicating no ermine sighting, to 2.88, which signifies that nearly all observers observed at least one family group during their field season.
-We independtly obtained measures of minimum (0.02 ind./km\textsuperscript{2}) and maximum (0.4 ind./km\textsuperscript{2}) ermine density, which were determined from estimates of individual home range obtained from radio-tracking data, observations on Bylot Island, and existing literature \citep{legagneux2012, bilodeau2013}.
-We associated the minimum and maximum scores of relative abundance with the minimum and maximum density of individuals, respectively.
-Ultimately, we calculated the ermine density by linearly interpolating between these two density extremes using the annual index of relative abundance.
-
-\subsection{Arctic fox} 
-We estimated the abundance of Arctic foxes in the study area based on their estimated home range size inside and outside the goose colony.
-We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size.
-However, here, we did not account for annual variations in lemming density as presented in \citet{dulude2023} in order to obtain mean fox home range.
-Given that foxes are territorial and exhibit an average spatial overlap with adjacent territories of 18\% \citep{clermont2021}, we converted home range size into individual density using the following formula: $\text{\textit{density of individuals}}= \frac{2}{0.82\times \text{\textit{home range}}}$.
-We used two as numerator because we assumed each territory was held by a pair of fox, either breeding or non-breeding fox pair, without accounting for potential nomadic or transient individuals.
-We used values of 12.26 km\textsuperscript{2} to represent the mean home range of foxes within the goose colony and 20.02 km\textsuperscript{2} for foxes outside the goose colony.
-We estimated the mean density of foxes in each zone of the study area according to the mean proportion of the zone covered by the goose colony.
-We derived the mean annual proportion of each zone covered by the goose colony from the colony outline between 2010 and 2023.
-We estimated a mean density of 0.14 individuals/km\textsuperscript{2} for the study area.
-Previously, the minimum density of foxes in the study area was estimated to be between 0.03 and 0.13 individuals per km\textsuperscript{2} based on camera traps \citep{royerboutin2015}.
-\newpage
-
-\newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
-\begin{landscape}
-\section{Species abundance}
-\input{tables/table_summary_methods.tex}
-%label-> table:summary_methods
-\end{landscape}
-\restoregeometry
-\newpage
-
-\input{tables/species_mean_abudance.tex}
-%label -> table:species_mean_abudance
-\newpage
-
-\begin{figure}[h]
-  \centering
-  \begin{subfigure}{0.35\textwidth}
-    \includegraphics[width=\linewidth]{figures/species_temporal_series/legend.pdf}
-  \end{subfigure} 
-  \begin{subfigure}{0.40\textwidth}
-    \includegraphics[width=\linewidth]{figures/species_temporal_series/brown_lemming.pdf}
-  \end{subfigure}
-    \hfill
-    \hfill
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\linewidth]{figures/species_temporal_series/collared_lemming.pdf}
-  \end{subfigure}
-  \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\linewidth]{figures/species_temporal_series/ermine.pdf}
-  \end{subfigure}
-\end{figure}
-\newpage
-
-\begin{figure}[ht]
-\centering
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/snowy_owl.pdf}
-\end{subfigure}
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/rough-legged_hawk.pdf}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/peregrine_falcon.pdf}
-\end{subfigure}
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/glaucous_gull.pdf}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/long-tailed_jaeger.pdf}
-\end{subfigure}
-\end{figure}
-\newpage
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/snow_goose.pdf}
-\end{subfigure}
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/cackling_goose.pdf}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{figures/species_temporal_series/American_golden-plover.pdf}
-\end{subfigure}
-\label{figure:other}
-\end{figure} 
-\newpage
-
+\clearpage
 \bibliography{references}

### 2024-09-20 script/tables_figures/table_interannual_variation_nest_density.R

@@ -118,10 +118,10 @@ table_dataset <- table %>%
   dplyr::rename(species= Species,
                 zone= Zone,
                 mean= `Mean ± SD`,
-                n=`Number of years`) %>% 
-  dplyr::mutate(mean= as.numeric(gsub(" ±.*","",table$`Mean ± SD`)),
-                sd= as.numeric(gsub(".* ±","",table$`Mean ± SD`))) %>% 
-  dplyr::select(species, zone, mean, sd, n)
+                sample_size_nest_density_km2=`Number of years`) %>% 
+  dplyr::mutate(mean_nest_density_km2= as.numeric(gsub(" ±.*","",table$`Mean ± SD`)),
+                sd_nest_density_km2= as.numeric(gsub(".* ±","",table$`Mean ± SD`))) %>% 
+  dplyr::select(species, zone, mean_nest_density_km2, sd_nest_density_km2, sample_size_nest_density_km2)
 
 
 write.csv(table_dataset,"dataset/BYLOT-interannual_variation_nest_density.csv",row.names = FALSE)

### 2024-09-20 script/tables_figures/table_mean_abundance.R

@@ -72,7 +72,7 @@ print(table_latex,
 #### Export as .csv for the data set ####
 #---------------------------------------#
 table_abundance <- table %>% 
-  dplyr::rename(species= Species, method= Method, mean_abundance= Mean, sd_abundance= sd, sample_size= n)
+  dplyr::rename(species= Species, method= Method, mean_abundance= Mean, sd_abundance= sd, sample_size_abundance= n)
 
-write.csv(table,"dataset/BYLOT-mean_species_abundance.csv",row.names = FALSE)
+write.csv(table_abundance,"dataset/BYLOT-mean_species_abundance.csv",row.names = FALSE)
 

### 2024-09-25 MetadataS1/MetadataS1.tex
Add table variables summary

@@
-\documentclass[a4paper,twoside,10pt]{article}
+\documentclass[a4paper,twoside,12pt]{article}
 
@@
 \usepackage{placeins}
+\usepackage{lineno}
+\linenumbers
 
@@
     \subsection*{3.  Research methods}
-    		\subsubsection*{1. Field/laboratory}
+    		\subsubsection*{a. Field/laboratory}
+We estimated the abundance of breeding individuals for most species, but there were a few exceptions.
+For common ravens, parasitic jaegers, long-tailed ducks, and king eiders, we suspect the presence of a significant number of non-breeding individuals in the study area.
+Therefore, the estimates we provided for these species include both breeding and potentially non-breeding individuals.
+Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, ermine, and Arctic hare.
+The methods used for each species are summarized in (\textbf{Table \ref{table:summary_methods}}).
+\newpage
+
     	\begin{enumerate}[label=\alph*.]
-        			\item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous
-gull}}
-        			\item[] \textit{\textbf{Snow goose}}
-        			\item[] \textit{\textbf{King eider and long-tailed duck}}
-        			\item[] \textit{\textbf{Rough-legged hawk, peregrine falcon and snowy owl}} 
-        			\item[] \textit{\textbf{Rock ptarmigan}}
-        			\item[] \textit{\textbf{Sandhill crane}}
-        			\item[] \textit{\textbf{American golden-plover and black-bellied plover}}
-        			\item[] \textit{\textbf{Common-ringed plover}}
-        			\item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}
-        			\item[] \textit{\textbf{Other passerines and sandpipers}}
-        			\item[] \textit{\textbf{Long-tailed jaeger}}
-        			\item[] \textit{\textbf{Parasitic jaeger}}
-        			\item[] \textit{\textbf{Common raven}}
-        			\item[] \textit{\textbf{Brown and collared lemming}}
-        			\item[] \textit{\textbf{Arctic hare}}
-        			\item[] \textit{\textbf{Ermine}}
+
+     \item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}	\\
+Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022.
+We transformed the number of nests to abundance of individuals by multiplying it by a factor of two.
+
+We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4).
+Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12406x + 0.13775$).
+However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series.
+Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area.
+We thus fitted an exponential model using the number of nests found annually over two distinct periods: in 1996 when the first nest was discovered, and then from 2017 to 2023 when sampling effort was systematic across the whole study area (\textbf{Figure \ref{figure:cackling}}).
+We used the fitted model to estimate abundance between 1996 and 2016 when monitoring was less systematic, which could potentially underestimate observed abundance as seen on \textbf{Figure \ref{figure:cackling}}.
+We multiplied nest density by two to obtain the abundance.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
+\caption{Number of cackling goose nests found across the study area over time.
+The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016.
+The model presents a strong fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area (R\textsuperscript{2}= 0.74, p= 0.15, n= 5).}
+\label{figure:cackling}
+\end{figure}
+
+      \item[] \textit{\textbf{Snow goose}}\\
+       Between 1999 and 2023, we assessed the abundance of snow geese in the study area through a multi-step process.
+We calculated the mean annual density of snow goose nests separately in the mesic and wetland habitats of the area occupied by the goose colony annually.
+We made slight adjustments to the goose colony perimeter defined from helicopter surveys to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}).
+To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled.
+The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts.
+Systematic nest searches were highly precise, but covered a relatively small area, whereas transects and snow goose point counts were less precise but covered larger areas.
+For each method, we calculated the mean density of breeding individuals in mesic habitat by dividing the number of birds (or nests) recorded by the area sampled.
+Despite methodological differences, the three approaches showed similar inter-annual variations, supporting the use of a mean values to estimate nest density in mesic habitat \textbf{Figure \ref{figure:sngo_interannual_variation}}).
+Lastly, to transform the densities in total abundance, we determined the annual proportion of wetland and mesic habitats within the goose colony and multiplied the area of each habitat by the density of breeding individuals.
+For the period 1999 to 2009, we used the average limits of the colony over the period 2010 to 2023 because we did not conduct aerial survey of the colony.
+Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).
+
+	\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.5\linewidth]{figures/map_geese_colony_2017.pdf}
+	\caption{Map showing the region occupied by the snow goose colony in 2017 (green polygon) as an example.
+The perimeter was first defined from a helicopter surveys and then slightly adjusted to include all snow goose point counts where at least one breeding pair had been observed in that year (green dots).
+Snow goose point counts where no breeding geese were observed in that year are presented as red dots.}
+	\label{figure:sngo_goose_colony}
+	\end{figure}
+
+	\begin{figure}[ht]
+	\centering
+	\includegraphics[width=\linewidth]{figures/snow_goose_mesic_density.pdf}
+	\caption{Estimates of breeding goose density in mesic habitat of the Bylot Island study area using three independent methods: Nest sampling plot, Snow goose point count, Vertebrate count transect.}
+	\label{figure:sngo_interannual_variation}
+	\end{figure}
+	\newpage
+
+        	\item[] \textit{\textbf{King eider and long-tailed duck}}\\
+		We estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley.
+We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}).
+We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two.
+We acknowledge that the opportunistic monitoring of these species likely underestimated their true nest density.
+However, considering the extensive sampling effort deployed annually within this area, we are confident to obtain a realistic order of magnitude for the number of nests present.
+Because duck sightings are frequent, yet only a few nests are found, we believe there may be non-breeding individuals.
+Therefore, we employed an additional method to estimate the overall duck populations without differentiating between breeding and non-breeding individuals.
+As an alternative approach, we estimated the abundance of ducks based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}.
+We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) in duck and loon species.
+We therefore derived the absolute abundance of long-tailed ducks and king eiders from their relative abundances using the ratio between relative and absolute abundances of red-throated loons as a reference.
+
+        	\item[] \textit{\textbf{Rough-legged hawk, peregrine falcon and snowy owl}} \\
+        	We estimated the abundance of breeding rough-legged hawks, peregrine falcons and snowy owls based on systematic nest monitoring conducted throughout the study area for these species.
+To convert the number of nests into breeding abundance, we multiplied it by two.
+For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.99; p < 0.0001, n= 10).
+We used the same approach for rough-legged hawks ($y=0.49851x$, R\textsuperscript{2}= 0.99, p < 0.0001, n= 8) to extend the time series from 2007 to 2012.
+We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.44, p= 0.27, n=8).
+
+        	\item[] \textit{\textbf{Rock ptarmigan}}\\
+        	We estimated the abundance of rock ptarmigans based on the annual nest density measured in the 8 km\textsuperscript{2} extensive nest search area of the Qarlikturvik valley.
+While we acknowledge that the opportunistic monitoring of this species likely underestimates nest density, the extensive sampling effort deployed annually within this area gives us confidence in obtaining a realistic number of nests.
+We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat.
+To convert the number of nests into breeding abundance, we multiplied it by two.
+
+        	\item[] \textit{\textbf{Sandhill crane}}\\
+        	We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}).
+In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed.
+We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density.
+However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present.
+Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2.
+We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones.
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
+\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin).
+The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
+Data points represent annual values.}
+\label{figure:sacr}
+\end{figure}
+
+        	\item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
+        	We applied a distance sampling approach to the observations of American golden-plovers made along vertebrate count transects to estimate the abundance of individuals in the mesic habitat of the study area between 2014 and 2023.
+We used perpendicular distance between detected individuals and transect (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}.
+We applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered).
+We selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term.
+We excluded observations of more than four individuals, as these likely indicated groups of non-breeders passing through the area.
+We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}.
+We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones.
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+
+To determine the abundance of black-bellied plovers, we used the mean number of black-bellied plovers and American golden-plovers observed per transect as an index of relative abundance.
+We assumed that the ratios of between relative and actual abundance are the same (i.e., similar detection probability) among those species.
+This assumption is realistic as those species present similarities in size, color, and reproductive behavior.
+We therefore derived the absolute abundance of black-bellied plovers from their relative abundance using the ratio between relative and absolute abundances of American golden-plover as a reference.
+
+As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.
+
+        	\item[] \textit{\textbf{Common-ringed plover}}\\
+        	To estimate the abundance of common-ringed plovers in the study area, we relied on the total number of nests recorded annually from 2015 to 2017, during which the primary nesting sites underwent intensive sampling.
+We multiplied the total nest count by two to represent the abundance of breeding individuals.
+
+        	\item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}\\
+        	We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}).
+For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected.
+We made this adjustment because this species was less frequently observed.
+In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located.
+This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically.
+Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2.
+We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones.
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.
+
+        	\begin{figure}[H]
+        	\begin{subfigure}{0.5\textwidth}
+        	\includegraphics[width=0.95\linewidth]{figures/lapland_longspur_nest_transect.pdf} 
+        	\caption{}
+        	\end{subfigure}
+        	\begin{subfigure}{0.5\textwidth}
+        	\includegraphics[width=0.95\linewidth]{figures/bairds_sandpiper_nest_transect.pdf}
+        	\caption{}
+        	\end{subfigure}
+        	\caption{a) Linear regression between the nest density of Lapland longspurs and the number of individuals observed per transect (nest density= 2.3422 x number of individuals per transect; regression was forced through the origin).
+The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
+Data points represent annual values.
+b) Linear regression between the nest density of Baird’s sandpiper and the proportion of transect with at least one individual observed (nest density= 34.9248 x proportion of transects with at least one individual; regression was forced through the origin).
+The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
+Data points represent annual values.}
+        	\label{figure:lalo_basa}
+        	\end{figure}
+
+        	\item[] \textit{\textbf{Other passerines and sandpipers}}\\
+        	We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section 3.9).
+We assumed here a similar detection probability for all species.
+We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section 3.9).
+For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones.
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+Nest density was then converted in number of individuals by multiplying by the area and a factor 2.
+As an alternative approach, we estimated the abundance of other passerines and sandpipers based on the indices of relative abundance (i.e., the number of individuals observed per 100 hours) presented by \citet{gauthier2024a}.
+We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among both passerine and sandpiper species.
+We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.
+
+        	\item[] \textit{\textbf{Long-tailed jaeger}}\\
+        	We determined the annual nest density of long-tailed jaegers from the systematic nest sampling between 2004 and 2023 on the southern side of the glacial river in the Qarlikturvik valley.
+We determined nest density by dividing the annual number of nests recorded by the area of the surveyed zone (33 km\textsuperscript{2}).
+As long-tailed jaegers typically nest in mesic habitat \citep{andersson1971}, we multiplied the area occupied by mesic habitat across the study area by the nest density obtained in the surveyed zone and by two to obtain the total abundance of breeding individuals.
+
+        	\item[] \textit{\textbf{Parasitic jaeger}}\\
+        	Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 3.75 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}.
+This suggests that there may be non-breeding individuals present at the study site, or alternatively, individuals may regularly travel long distances, potentially from outside the study area, to forage during the breeding season.
+Due to limited data availability for estimating the abundance of non-breeding parasitic jaegers, we relied on the maximum number of adults banded during a single year (17 individuals in 2009; Therrien, unpublished data) as the minimum abundance on the study area.
+This corresponds to a density of 0.04 individuals/km\textsuperscript{2}.
+For comparison, \citet{taylor1974} measured a density of 0.06 individual/km\textsuperscript{2} on Bathurst Island.
+
+        	\item[] \textit{\textbf{Common raven}}\\
+        	Despite the intensive nest searches for raptors in upland zones, we never found more than one common raven nest each year, a small number considering the frequent raven observations at the study site \citep{gauthier2024a}.
+This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging.
+Therefore, we opted for alternative approaches based on individual counts to estimate the abundance of both breeding and non-breeding ravens.
+As a first approach, we based our estimate of ravens on the number of glaucous gulls observed per transect.
+We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species.
+This assumption is reasonable as those species present similarities in size and foraging strategy.
+We therefore derived the absolute abundance of common ravens from their relative abundance using the ratio between relative and absolute abundances of glaucous gulls as a reference.
+Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.
+
+        \item[] \textit{\textbf{Brown and collared lemming}}\\
+        Between 1995 and 2003, we used the density estimates derived from the snap-trapping indices obtained in late July in each habitat.
+Between 2004 and 2007, annual abundance of each lemming species was based on the late-July density estimates on trapping grid in wet and mesic habitats.
+However, starting from 2008, estimates were derived from the mean density recorded in mid-July and mid-August, except for two instances: 2019 and 2021.
+In 2019, due to an exceptionally early snowmelt and thus an early decline in lemmings during the summer, we only retained value from mid-July.
+In 2021, we relied solely on data gathered in August because it was the only trapping period carried out that year.
+To scale the estimated densities from the wet and mesic grids to the entire study area, we used the proportions of mesic habitats (64\%) and wet habitats (10\%) measured within the study area.
+
+        	\item[] \textit{\textbf{Arctic hare}}\\
+        	Arctic hares are primarily observed in the upland zones of the study area, where sampling effort is limited.
+We thus derived abundance of hares from the estimated abundance of Arctic foxes based on indices of relative abundance presented in \citep{gauthier2024a}, which were derived from incidental daily observations.
+We doubled the density of Arctic hares in the upland zones (i.e., plateaus), as twice as many individuals were observed per hour of fieldwork there compared to lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+However, it is worth noting that assuming a similar detection probability between foxes and hares might lead to an overestimation of hare detection probability due to behavioral differences between the species.
+Therefore, we most likely underestimate the actual abundance of Arctic hares in the study area.
+
+        	\item[] \textit{\textbf{Ermine}}\\
+        	We estimated the annual abundance of ermines by transforming the annual index of relative abundance provided in \citet{bolduc2023} into individual density.
+Annual values ranged from 0, indicating no ermine sighting, to 2.88, which signifies that nearly all observers observed at least one family group during their field season.
+We independtly obtained measures of minimum (0.02 ind./km\textsuperscript{2}) and maximum (0.4 ind./km\textsuperscript{2}) ermine density, which were determined from estimates of individual home range obtained from radio-tracking data, observations on Bylot Island, and existing literature \citep{legagneux2012, bilodeau2013}.
+We associated the minimum and maximum scores of relative abundance with the minimum and maximum density of individuals, respectively.
+Ultimately, we calculated the ermine density by linearly interpolating between these two density extremes using the annual index of relative abundance.
+
+        	\item[] \textit{\textbf{Arctic fox}}\\
+        	We estimated the abundance of Arctic foxes in the study area based on their estimated home range size inside and outside the goose colony.
+We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size.
+However, here, we did not account for annual variations in lemming density as presented in \citet{dulude2023} in order to obtain mean fox home range.
+Given that foxes are territorial and exhibit an average spatial overlap with adjacent territories of 18\% \citep{clermont2021}, we converted home range size into individual density using the following formula: $\text{\textit{density of individuals}}= \frac{2}{0.82\times \text{\textit{home range}}}$.
+We used two as numerator because we assumed each territory was held by a pair of fox, either breeding or non-breeding fox pair, without accounting for potential nomadic or transient individuals.
+We used values of 12.26 km\textsuperscript{2} to represent the mean home range of foxes within the goose colony and 20.02 km\textsuperscript{2} for foxes outside the goose colony.
+We estimated the mean density of foxes in each zone of the study area according to the mean proportion of the zone covered by the goose colony.
+We derived the mean annual proportion of each zone covered by the goose colony from the colony outline between 2010 and 2023.
+We estimated a mean density of 0.14 individuals/km\textsuperscript{2} for the study area.
+Previously, the minimum density of foxes in the study area was estimated to be between 0.03 and 0.13 individuals per km\textsuperscript{2} based on camera traps \citep{royerboutin2015}.
+		\newpage
         		\end{enumerate}
 
-        		\subsubsection*{2. Instrumentation}
+        		\subsubsection*{b. Taxonomy and systematics}
+       Bird scientific names were obtained from the IOC World Bird List 14.2 \citep{gill2024}.
+Mammal scientific names were obtained from the Mammal species of the world: a taxonomic and geographic reference \citep{wilson2005}.
+However, we considered here \textit{Mustela erminea richardsonii} as \textit{Mustela richardsonii} due to recent genetic analysis \citep{colella2021}.
 
-        		\subsubsection*{3. Taxonomy and systematics}
+        		\subsubsection*{c. Permit history}
+      All research involving animals on Bylot Island has been approved by an institutional Animal Care Committee.
+In 1999, the study area became part of Sirmiliik National Park, managed by Parks Canada.
+Since then, all research activities in the park have been approved by a Joint Park Management Committee.
 
-        		\subsubsection*{4. Permit history}
+      		\subsubsection*{d. Project personnel}
+      		\begin{enumerate}[label=\alph*.]
+      			\item[] \textit{\textbf{Principal and associated investigators}}\\
+      			Gilles Gauthier, Eric Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux \citep{gauthier2024b}
+      			\item[] \textit{\textbf{Students}}\\
+      			By combining animal and plant ecology, 24 doctoral theses and 56 master theses have been completed in relation to the Bylot Island South Plain study area \citep{gauthier2024b}.
+      		\end{enumerate}
+\newpage
 
-        		\subsubsection*{5. Legal/organizational requirements}
+ \section*{\textcolor{Blue}{Class III. Data set status and accessibility}}
+    \section*{A. Status}
+ 		\subsection*{1. Latest update} 19th September 2024
+        \subsection*{2. Latest archive date} XXXXXX October 2024
+        \subsection*{3. Metadata status} XXXXXX October 2024
+        \subsection*{4. Data verification} The methods used to extract final species abundance estimates were subject to several rounds of revision by the principal investigators.
 
-        \subsubsection*{4. Project personnel: Principal and associated investigator(s), technicians, supervisors, students}
- \section*{\textcolor{Blue}{Class III. Data set status and accessibility}}
-    \subsection*{A. Status}
-    \begin{enumerate}
-      \item Latest update: Date of last modification of data set
-      \item Latest archive date: Date of last data set archival
-      \item Metadata status: Date of last metadata update and current status
-      \item Data verification: Status of data quality assurance checking
+   \section*{B. Accessibility}
+    		\subsection*{1. Storage location and medium}
+    		 Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
+Raw data and codes used to extract the presented data set are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset}.
+
+        \subsection*{2. Contact persons}
+        \begin{enumerate}[label=\alph*.]
+      			\item[] \textit{\textbf{Overall project}}\\
+      			Joël Bêty; \textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}; 418 723-1986 \#1701; 300 allée des Ursulines, Rimouski, Québec, Canada, G5L 3A1, Office B-002
+      			\item[] \textit{\textbf{Data and codes}}\\
+      			 Louis Moisan, \textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}
       		\end{enumerate}
 
-    \subsection*{B. Accessibility}
-    		\begin{enumerate}
-    			\item Storage location and medium: Pointers to where data reside (including archival required by the ESA Open Research policy and redundant archival sites)
-        		\item Contact persons: Name, address, phone, email
-        		\item Copyright restrictions: Whether copyright restrictions prohibit use of all or portions of the data set
-        		\item Proprietary restrictions: Any other restrictions that may prevent use of all or portions of data set
-        			\begin{enumerate}[label=\alph*.]
-            			\item Release date: Date when proprietary restrictions expire
-           			\item Citation: How data may be appropriately cited
-            			\item Disclaimer(s): Any disclaimers that should be acknowledged by secondary users
-            		\end{enumerate}
-        		\item Costs: Costs associated with acquiring data (may vary by size of data request, desired medium, etc.)
-    		\end{enumerate}
+        	\subsection*{3. Copyright restrictions}
+        	None
+
+        	\subsection*{4. Proprietary restrictions}
+        		\subsubsection*{a. Release date} None
+        		\subsubsection*{b. Citation}
+        		 Please cite this data paper when using the data.
+        		\subsubsection*{c. Disclaimer} None
+
+        	\subsection*{5. Costs}
+        	None	, the data can be used free of charge.		
+\newpage
 
  \section*{\textcolor{Blue}{Class IV. Data structural descriptors}}
-    \subsection*{A. Data set file}
-    		\begin{enumerate}
-    			\item Identity: Unique file names or codes
-        		\item Size: Number of records, record length, total number of bytes, etc.
-        		\item Format and storage mode: File type (e.g., ASCII, binary, etc.), compression schemes employed (if any), etc.
-        		\item Header information: Description of any header data or information attached to file [Note: may include elements related to variable information (IV.B.); if so, could be linked to appropriate section(s)]
-        		\item Alphanumeric attributes: Mixed, upper, or lower case
-        		\item Special characters/fields: Methods used to denote comments, flag modified or questionable data, etc.
-        		\item Authentication procedures: Digital signature, checksum, actual subset(s) of data, and other techniques for assuring accurate transmission of data to secondary users
-    		\end{enumerate}
+    \section*{A. Data set file}
+    			\subsection*{1. Identity} 
+    			a. \texttt{BYLOT-species\_taxonomy.csv}\\
+     		b. \texttt{BYLOT-species\_abundance.csv}\\
+     		c. \texttt{BYLOT-mean\_species\_abundance.csv}\\
+     		d. \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
+    			e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\
 
-    \subsection*{B. Variable information}
-    \begin{enumerate}
-    		\item Variable identity: Unique variable name or code
-        \item Variable definition: Precise definition of variables in data set
-        \item Units of measurement: Units of measurement associated with each variable
-        \item Data type
+        		\subsection*{2. Size} 
+        		a. 35 records, not including header row (2.3 kB)\\
+        		b. 273 records, not including header row  (17.0 kB)\\
+        		c. 35 records, not including header row (2.1 kB)\\
+        		d. 51 records, not including header row  (1.8 kB)\\
+        		e. 18 records, not including header row  (1.1 kB)
+
+        		\subsection*{3. Format and storage mode} All files are in a comma-separated value format (.csv).
+
+        		\subsection*{4. Header information} 
+        			\subsubsection*{a. \texttt{BYLOT-species\_taxonomy.csv}}
+        			\texttt{functional\_group}; \texttt{scientific\_name}; \texttt{species}; \texttt{annual\_cycle\_strategy}
+
+        			\subsubsection*{b. \texttt{BYLOT-species\_abundance.csv}}
+        			\texttt{species}; \texttt{year}; \texttt{status}; \texttt{method}; \texttt{abundance}
+
+        			\subsubsection*{c. \texttt{BYLOT-mean\_species\_abundance.csv}}
+        			\texttt{species}; \texttt{method}; \texttt{mean\_abundance}; \texttt{sd\_abundance}; \texttt{sample\_size\_abundance}
+
+        			\subsubsection*{d. \texttt{BYLOT-mean\_species\_body\_mass.csv}}
+        			\texttt{species}; \texttt{source}; \texttt{body\_mass\_g}
+
+        			\subsubsection*{e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}}
+        			\texttt{species}; \texttt{zone}; \texttt{mean\_nest\_density\_km2}; \texttt{sd\_nest\_density\_km2}; \texttt{sample\_size\_nest\_density\_km2}
+
+        		\subsection*{5. Alphanumeric attributes} Mixed
+
+        		\subsection*{6. Special characters/fields} Unavailable values are indicated by NA.
+
+        		\subsection*{7. Authentication procedures}
+        		Sums of the numeric columns: \\
+        		b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2291027\\
+        		c. \texttt{BYLOT-mean\_species\_abundance.csv}: mean\_abundance= 107875; sd\_abundance= 116687; sample\_size\_abundance= 207\\
+        		d. \texttt{BYLOT-mean\_species\_body\_mass.csv}: body\_mass\_g= 46223\\
+        		e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185
+
+    \section*{B. Variable information}
+    		\subsection*{1. Variable identity} See Table \ref{table:summary_variables}
+        \subsection*{2. Variable definition} See Table \ref{table:summary_variables}
+        \subsection*{3. Units of measurement} See Table \ref{table:summary_variables}
+
+        \input{tables/table_summary_variables.tex}
+
+        \subsection*{4. Data type}
         		\begin{enumerate}[label=\alph*.]
@@
             \end{enumerate}
-        \item Data format
+
+        \subsection*{5. Data format}
         		\begin{enumerate}[label=\alph*.]
@@
         		\end{enumerate}
-    \end{enumerate}
 
-    \subsection*{C. Data anomalies: Description of missing data, anomalous data, calibration errors, etc.}
+ \section*{C. Data anomalies: Description of missing data, anomalous data, calibration errors, etc.}
+
+ \newpage
+
  \section*{\textcolor{Blue}{Class V. Supplemental descriptors}}
-    \subsection*{A. Data acquisition}
+    \section*{A. Data acquisition}
     \begin{enumerate}
@@
 
-    \subsection*{B.
+    \section*{B.
 Quality assurance/quality control procedures}Identification and treatment of outliers, description of quality assessments, calibration of reference standards, equipment performance results, etc.
-    \subsection*{C. Related materials} References and locations of maps, photographs, videos, GIS data layers, physical specimens, field notebooks, comments, etc.
-    \subsection*{D. Computer programs and data-processing algorithms}
+    \section*{C. Related materials} References and locations of maps, photographs, videos, GIS data layers, physical specimens, field notebooks, comments, etc.
+    \section*{D. Computer programs and data-processing algorithms}
     		\begin{enumerate}
@@
     			\item \textbf{Operating system}: x86\_64-pc-linux-gnu (64-bit) with Ubuntu 22.04.3 LTS
-    			\item \textbf{Packages}: dplyr, RPostgreSQL, sf, stringr, tidyr, xtable, Distance, ggplot2, ggmap, png, grid
+    			\item \textbf{Packages}: dplyr, tidyr, sf, stringr, xtable, Distance, ggplot2, lme4, AICcmodavg, scales, ggmap
+    			\item \textbf{Codes}: All raw data and codes are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset.git}
     		\end{enumerate}
-    \subsection*{E. Archiving}
+
+    \section*{E. Archiving}
     		\begin{enumerate}
@@
 
-    \subsection*{F. Publications and results: Electronic reprints, lists of publications resulting from or related to the study, graphical/statistical data representations, etc.}
+    \section*{F. Publications and results: Electronic reprints, lists of publications resulting from or related to the study, graphical/statistical data representations, etc.}
 \textbf{List of publications that include data from the data set}:
@@
 
-   \subsection*{ G. History of data set usage}
+   \section*{G. History of data set usage}
   		\begin{enumerate}

### 2024-09-25 MetadataS1/tables/table_summary_variables.tex
Add table variables summary

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Sep 24 06:28:59 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.08\textwidth}|p{0.27\textwidth}|p{0.40\textwidth}|p{0.08\textwidth}|p{0.06\textwidth}|}
\caption{Summary of variable definitions, units, and data types for each data set file.
} \\ 
  \hline
{\textbf{Data file}} & {\textbf{Variable identity}} & {\textbf{Variable definition}} & {\textbf{Units}} & {\textbf{Type}} \\ 
  \hline
a. & functional\_group & Functional group for each species. The classification of species into functional groups is based on Moisan et al. (2023). &  & string \\ 
   \hline
a. & scientific\_name & Bird scientific names were obtained from the IOC World Bird List 14.2 (Gill et al., 2024). Mammal scientific names were obtained from the Mammal species of the world: a taxonomic and geographic reference (Wilson, 2005). However, we considered here Mustela erminea richardsonii as Mustela richardsonii due to recent genetic analysis (Colella et al., 2021). &  & string \\ 
   \hline
a. & species & Common species english name. &  & string \\ 
   \hline
a. & annual\_cycle\_strategy & Annual cycle strategy of each species (i.e., resident, partially migratory or migratory). The classification of species annual cycle strategy is based on Gauthier et al., (2011)  and Moisan et al. (2023). &  & string \\ 
   \hline
b. & species & Common species english name. &  & string \\ 
   \hline
b. & year & Year corresponding to the estimate of annual abundance. & years & integer \\ 
   \hline
b. & status & Reproductive status of the individuals (i.e., undetermined or breeding). &  & string \\ 
   \hline
b. & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
b. & abundance & Estimate of the annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
c. & species & Common species english name. &  & string \\ 
   \hline
c. & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
c. & mean\_abundance & Estimate of the mean annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
c. & sd\_abundance & Standard deviation of the annual species abundance. & individuals & numeric \\ 
   \hline
c. & sample\_size\_abundance & Number of years consider in the calculation of mean annual abundance. & years & integer \\ 
   \hline
d. & species & Common species english name. &  & string \\ 
   \hline
d. & source & Source from which estimate were derived (i.e., measurements from the study area or  extracted from Wilman et al. 2014). &  & string \\ 
   \hline
d. & body\_mass\_g & Mean individual body mass. & grams & numeric \\ 
   \hline
e. & species & Common species english name. &  & string \\ 
   \hline
e. & zone & Sampled zone of the study area (see figure 2 and 3). &  & string \\ 
   \hline
e. & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
of the study area. & nests per square kilometer & numeric \\ 
   \hline
e. & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
zone of the study area. & nests per square kilometer & numeric \\ 
   \hline
e. & sample\_size\_nest\_density\_km2 & Number of years consider in the calculation of the nest density. & years & integer \\ 
   \hline
\hline
\label{table:summary_variables}
\end{longtable}
\endgroup

### 2024-09-25 manuscript/manuscript.tex
Add table variables summary

@@
 
-In 1993, 1998, 2003, and 2008, the Canadian Wildlife Service conducted aerial surveys over the southern plain of Bylot Island during the brood-rearing period to estimate the abundance of breeding snow geese (Reed et al., 2002).
-We corrected these estimates with the annual nesting success, as these surveys were conducted after the hatching period.
-For 2003 and 2008, we used in priority estimates of abundance derived from aerial surveys instead of nest sampling.
-
 \subsection{King eider and long-tailed duck}

### 2024-09-25 manuscript/tables/table_summary_variables.tex
Add table variables summary

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Mon Sep 23 15:40:52 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.15\textwidth}|p{0.10\textwidth}|p{0.5\textwidth}|p{0.07\textwidth}|p{0.10\textwidth}|}
\caption{Definition of the variables includes in each data file.} \\ 
  \hline
{\textbf{Data.set.file}} & {\textbf{Variable.identity}} & {\textbf{Variable.definition}} & {\textbf{Units}} & {\textbf{Storage.type}} \\ 
  \hline
BYLOT-species\_taxonomy.csv & functional\_group & Functional group for each species. The classification of species into functional groups is based on Moisan et al. (2023). &  & string \\ 
   \hline
BYLOT-species\_taxonomy.csv & scientific\_name & Bird scientific names were obtained from the IOC World Bird List 14.2 (Gill et al., 2024). Mammal scientific names were obtained from the Mammal species of the world: a taxonomic and geographic reference (Wilson, 2005). However, we considered here Mustela erminea richardsonii as Mustela richardsonii due to recent genetic analysis (Colella et al., 2021). &  & string \\ 
   \hline
BYLOT-species\_taxonomy.csv & species & Common species english name. &  & string \\ 
   \hline
BYLOT-species\_taxonomy.csv & annual\_cycle\_strategy & Annual cycle strategy of each species (i.e., resident, partially migratory or migratory). The classification of species annual cycle strategy is based on Gauthier et al., (2011)  and Moisan et al. (2023). &  & string \\ 
   \hline
BYLOT-species\_abundance.csv & species & Common species english name. &  & string \\ 
   \hline
BYLOT-species\_abundance.csv & year & Year corresponding to the estimate of annual abundance. & years & integer \\ 
   \hline
BYLOT-species\_abundance.csv & status & Reproductive status of the individuals (i.e., undetermined or breeding). &  & string \\ 
   \hline
BYLOT-species\_abundance.csv & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
BYLOT-species\_abundance.csv & abundance & Estimate of the annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
BYLOT-mean\_species\_abundance.csv & species & Common species english name. &  & string \\ 
   \hline
BYLOT-mean\_species\_abundance.csv & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
   \hline
BYLOT-mean\_species\_abundance.csv & mean\_abundance & Estimate of the mean annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals & integer \\ 
   \hline
BYLOT-mean\_species\_abundance.csv & sd\_abundance & Standard deviation of the annual species abundance. & individuals & numeric \\ 
   \hline
BYLOT-mean\_species\_abundance.csv & sample\_size\_abundance & Number of years consider in the calculation of mean annual abundance. & years & integer \\ 
   \hline
BYLOT-mean\_species\_body\_mass.csv & species & Common species english name. &  & string \\ 
   \hline
BYLOT-mean\_species\_body\_mass.csv & source & Source from which estimate were derived (i.e., measurements from the study area or  extracted from Wilman et al. 2014). &  & string \\ 
   \hline
BYLOT-mean\_species\_body\_mass.csv & body\_mass\_g & Mean individual body mass. & grams & numeric \\ 
   \hline
BYLOT-interannual\_variation\_nest\_density.csv & species & Common species english name. &  & string \\ 
   \hline
BYLOT-interannual\_variation\_nest\_density.csv & zone & Sampled zone of the study area (see figure 2 and 3). &  & string \\ 
   \hline
BYLOT-interannual\_variation\_nest\_density.csv & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
of the study area. & nests per square kilometer & numeric \\ 
   \hline
BYLOT-interannual\_variation\_nest\_density.csv & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
zone of the study area. & nests per square kilometer & numeric \\ 
   \hline
BYLOT-interannual\_variation\_nest\_density.csv & sample\_size\_nest\_density\_km2 & Number of years consider in the calculation of the nest density. & years & integer \\ 
   \hline
\hline
\label{table:summary_variables}
\end{longtable}
\endgroup

### 2024-09-25 script/extract_species_abundance/species/snow_goose.R

@@ -14,7 +14,6 @@ library(sf)
 sf::sf_use_s2(FALSE)
 #Visualization
 library(ggplot2)
-library(ggpubr)
 
 
 #-------------------#

### 2024-09-25 script/tables_figures/table_variables_definition.R

@@ -1,0 +1,42 @@
+#Title: Create a summary table of all variables in the data set
+#Author: Louis Moisan
+#Date: September 23 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+#Import .csv format
+table <- read.csv("MetadataS1/tables/table_variables_definition.csv") %>% 
+  dplyr::rename(`Data file`= Data.file, `Variable identity`= Variable.identity, `Variable definition`= Variable.definition)
+  
+
+
+#as tex
+table_latex <- xtable(table, caption = "Summary of variable definitions, units, and data types for each data set file.
+",
+                      label= "table:summary_variables",
+                      align= c("p{0.00\\textwidth}", 
+                               "|p{0.08\\textwidth}|", #Data set file
+                               "p{0.27\\textwidth}|", #Variable identity
+                               "p{0.40\\textwidth}|", #Variable definition
+                               "p{0.08\\textwidth}|", #Units
+                               "p{0.06\\textwidth}|" #Storage type
+                               ) #CI
+) 
+#function to bold column header
+bold <-  function(x) {paste('{\\textbf{',x,'}}', sep ='')}
+
+
+print(table_latex,
+      include.rownames = FALSE,
+      tabular.environment = "longtable",
+      hline.after= c(-1,0:nrow(table)),
+      size="\\fontsize{8pt}{10pt}\\selectfont",
+      file = "MetadataS1/tables/table_summary_variables.tex",
+      sanitize.colnames.function=bold,
+      caption.placement = "top")

### 2024-09-26 MetadataS1/MetadataS1.tex
begin table of variable format

@@
 
-%%%%%%% !!!! ADD DRYAD adress !!!! %%%%%%%%
+%%%%%%% !!!! ADD DRYAD and Zenodo adress !!!! %%%%%%%%
 Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
-Raw data and codes used to extract the presented data set are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset}.
+Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
 \endgroup
@@
 \includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
-\caption{Synthetic vertebrate food web of the south plain of Bylot Island (figure adapted from \citet{gauthier2011}.}
+\caption{Synthetic vertebrate food web of the south plain of Bylot Island.}
 \label{figure:food_web}
@@
 The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}).
-The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012}.
+The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}.
 We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}.
 The case of the red fox (\textit{Vulpes vulpes}) was ambiguous.
-While the presence of a breeding pair has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare.
+While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare.
 Therefore, we decided to exclude this species.
@@
    \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada.
-The outline of the snow goose colony is presented with white dash; we used the outline of the colony in 2015 since it represents an average colony area (67 km\textsuperscript{2}).}
+The outline of the snow goose colony area is presented with white dash; we used the outline of the colony in 2017 since it represents an average colony area (74 km\textsuperscript{2}).}
   \label{figure:zones}
@@
 		\subsubsection*{b. Avian nest monitoring}
-Avian nest monitoring was conducted annually except in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic.
+Avian nest monitoring was not conducted in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic.
 Systematic nest monitoring refers here to a systematic sampling approach aimed at documenting all nests within a specified area.
@@
 
+        		\newgeometry{top=0.5cm, bottom=1.5cm, left=0.5cm, right=0.5cm}
+\begin{landscape}
+\input{tables/table_summary_methods.tex}
+%label-> table:summary_methods
+\end{landscape}
+\restoregeometry
+\newpage
+
         		\subsubsection*{b. Taxonomy and systematics}
@@
 Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
-Raw data and codes used to extract the presented data set are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset}.
+Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
 
@@
         		\subsubsection*{b. Citation}
-        		 Please cite this data paper when using the data.
+        		 Please cite this document when using the data.
         		\subsubsection*{c. Disclaimer} None
@@
      		b. \texttt{BYLOT-species\_abundance.csv}\\
-     		c. \texttt{BYLOT-mean\_species\_abundance.csv}\\
-     		d. \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
-    			e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\
+     		c. \texttt{BYLOT-species\_body\_mass.csv}\\
+    			d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\
 
@@
         		b. 273 records, not including header row  (17.0 kB)\\
-        		c. 35 records, not including header row (2.1 kB)\\
-        		d. 51 records, not including header row  (1.8 kB)\\
-        		e. 18 records, not including header row  (1.1 kB)
+        		c. 51 records, not including header row  (1.8 kB)\\
+        		d. 18 records, not including header row  (1.1 kB)
 
@@
 
-        			\subsubsection*{c. \texttt{BYLOT-mean\_species\_abundance.csv}}
-        			\texttt{species}; \texttt{method}; \texttt{mean\_abundance}; \texttt{sd\_abundance}; \texttt{sample\_size\_abundance}
-
-        			\subsubsection*{d. \texttt{BYLOT-mean\_species\_body\_mass.csv}}
+        			\subsubsection*{c. \texttt{BYLOT-species\_body\_mass.csv}}
         			\texttt{species}; \texttt{source}; \texttt{body\_mass\_g}
 
-        			\subsubsection*{e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}}
+        			\subsubsection*{d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}}
         			\texttt{species}; \texttt{zone}; \texttt{mean\_nest\_density\_km2}; \texttt{sd\_nest\_density\_km2}; \texttt{sample\_size\_nest\_density\_km2}
@@
         		b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2291027\\
-        		c. \texttt{BYLOT-mean\_species\_abundance.csv}: mean\_abundance= 107875; sd\_abundance= 116687; sample\_size\_abundance= 207\\
-        		d. \texttt{BYLOT-mean\_species\_body\_mass.csv}: body\_mass\_g= 46223\\
-        		e. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185
+        		c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 46223\\
+        		d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185
 

### 2024-09-26 MetadataS1/tables/table_summary_methods.tex
begin table of variable format

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Sep 25 17:05:28 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.10\textwidth}|p{0.27\textwidth}|p{0.06\textwidth}|p{0.23\textwidth}|p{0.09\textwidth}|p{0.09\textwidth}|p{0.085\textwidth}|p{0.04\textwidth}|p{0.10\textwidth}|}
\caption{Summary of the lowest, highest, mean and standard deviation of the estimated abundance of each vertebrate species in the vertebrate community of the southern plain of Bylot Island (389 km2). In some cases, two independent approaches have been used to estimate the abundance of the same species as a proxy for uncertainty. We provide a qualitative measure of the method quality based on data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.} \\ 
  \hline
{\textbf{Species}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Lowest abundance}} & {\textbf{Highest abundance}} & {\textbf{Mean abundance}} & {\textbf{sd}} & {\textbf{n}} \\ 
  \hline
snow goose & Nest monitoring plots extrapolated to mean goose colony area & low & Relatively small sample size and uncertainty on goose colony area  & 2505 & 35404 & 18129 & 11037 & 11 (1999-2009) \\ 
   \hline
snow goose & Intensive study area-wide monitoring based on a combination of methods (transects, point counts and nest monitoring plots) and annual colony outline & high & Multiple independent methods and annual colony outline & 8687 & 49076 & 31852 & 12092 & 12 (2010-2019, 2022-2023) \\ 
   \hline
snowy owl & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &   0 &  67 & 15 & 24 & 16 (1996-2011) \\ 
   \hline
snowy owl & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 & 144 & 17 & 45 & 10 (2012-2019, 2022-2023) \\ 
   \hline
glaucous gull & Extrapolation from intensive nest monitoring (111 km2, R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density &  59 &  80 & 73 & 6 & 13 (2004-2016) \\ 
   \hline
glaucous gull & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  60 &  80 & 71 & 9 & 4 (2017-2019, 2022) \\ 
   \hline
peregrine falcon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   8 &  12 & 10 & 1 & 8 (2013-2019, 2022) \\ 
   \hline
rough-legged hawk & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=8) & high & Strong correlation with study area-wide nest density &   3 &  59 & 20 & 19 & 8 (2007-2012) \\ 
   \hline
rough-legged hawk & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &  66 & 30 & 29 & 8 (2013-2019, 2022) \\ 
   \hline
American golden-plover & Distance sampling throughout lowland (313 km2) & high & Large sample size & 397 & 1725 & 1102 & 432 & 8 (2014-2019, 2022-2023) \\ 
   \hline
cackling goose & Extrapolation from exponential model of growth (R2=0.74, p=0.15, n=5) & moderate & Strong correlation with opportunistic nest monitoring &   2 & 158 & 31 & 41 & 23 (1996-2016, 2020-2021) \\ 
   \hline
cackling goose & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  80 & 214 & 138 & 50 & 5 (2017-2019, 2022-2023) \\ 
   \hline
Arctic fox & Derived from extensive fox home range size studies (n=109) & moderate & Indirect indices, but large sample size &  &  & 53 &  &  \\ 
   \hline
collared lemming & Rigorous density estimates at small spatial scale (0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 39302 & 8128 & 10334 & 27 (1995-2019, 2021-2022) \\ 
   \hline
brown lemming & Rigorous density estimates at small spatial scale (0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 447630 & 54043 & 93530 & 27 (1995-2019, 2021-2022) \\ 
   \hline
ermine & Indices of relative abundance derived from testimonials converted to abundance using home range size & moderate & Indirect indices and uncertainty on ermine home range size estimates &   8 & 156 & 40 & 37 & 27 (1993-2019) \\ 
   \hline
long-tailed jaeger & Intensive nest monitoring (33 km2) extrapolated by habitat & high & Relatively large spatial coverage of sampling &   0 & 900 & 272 & 285 & 17 (2004-2019, 2022) \\ 
   \hline
red-throated loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  42 &  76 & 64 & 15 & 4 (2017-2019, 2022) \\ 
   \hline
Pacific loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   6 & 4 & 3 & 4 (2017-2019, 2022) \\ 
   \hline
tundra swan & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   2 & 1 & 1 & 4 (2017-2019, 2022) \\ 
   \hline
common-ringed plover & Nest monitoring on the main breeding sites & moderate & Intensive monitoring, but not exhaustive to study area &  44 &  62 & 55 & 9 & 3 (2015-2017) \\ 
   \hline
black-bellied plover & Derived from the abundance estimate of American golden-plover using transects observations & moderate & Derived from high quality estimate of another species &  &  & 29 &  &  \\ 
   \hline
black-bellied plover & Derived from the abundance estimate of American golden-plover using incidental observations & low & Derived from high quality estimate of another species, but potentially includes transient migratory individuals &  &  & 87 &  &  \\ 
   \hline
Lapland longspur & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110 &  &  \\ 
   \hline
Baird's sandpiper & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2448 &  &  \\ 
   \hline
sandhill crane & Extrapolation from intensive nest monitoring (8 km2) and transect observations & moderate & Uncertain relation with large scale indices &  &  & 34 &  &  \\ 
   \hline
king eider & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 25 &  &  \\ 
   \hline
king eider & Derived from the abundance estimate of red-throated loon using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 106 &  &  \\ 
   \hline
long-tailed duck & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 20 &  &  \\ 
   \hline
long-tailed duck & Derived from the abundance estimate of red-throated loon using incidental observations & moderate & Derived from high quality estimate of another species &  &  & 191 &  &  \\ 
   \hline
rock ptarmigan & Intensive, but opportunistic nest monitoring (8 km2) extrapolated to study area & low & Intensive, but opportunistic monitoring at relatively small spatial scale and prime nesting habitat not well sampled &  &  & 24 &  &  \\ 
   \hline
horned lark & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362 &  &  \\ 
   \hline
horned lark & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411 &  &  \\ 
   \hline
ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40 &  &  \\ 
   \hline
ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 53 &  &  \\ 
   \hline
red phalarope & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 140 &  &  \\ 
   \hline
red phalarope & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 270 &  &  \\ 
   \hline
red knot & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 66 &  &  \\ 
   \hline
red knot & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 233 &  &  \\ 
   \hline
white-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 991 &  &  \\ 
   \hline
white-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 1134 &  &  \\ 
   \hline
buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 6 &  &  \\ 
   \hline
buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 8 &  &  \\ 
   \hline
pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 80 &  &  \\ 
   \hline
pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 255 &  &  \\ 
   \hline
parasitic jaeger & Maximum number of individuals banded in a year & low & Based on a single year and potentially not all individuals were captured &  &  & 17 &  &  \\ 
   \hline
parasitic jaeger & Maximum number of nest found annually during study area-wide opportunistic nest monitoring & very low & Monitoring does not include potential non-breeding individuals &  &  & 8 &  &  \\ 
   \hline
common raven & Derived from the abundance estimate of glaucous gull using transects observations & very low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 14 &  &  \\ 
   \hline
common raven & Derived from the abundance estimate of glaucous gull using incidental observations & very low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 31 &  &  \\ 
   \hline
American pipit & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 53 &  &  \\ 
   \hline
American pipit & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87 &  &  \\ 
   \hline
snow bunting & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 18 &  &  \\ 
   \hline
snow bunting & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276 &  &  \\ 
   \hline
Arctic hare & Derived from the abundance estimate of Arctic fox using incidental observations & very low & Derived from moderate quality estimate of another species, prime nesting habitat not well sampled and difference in detectability between species &  &  & 6 &  &  \\ 
   \hline
\hline
\label{table:summary_methods}
\end{longtable}
\endgroup

### 2024-09-26 MetadataS1/tables/table_summary_variables.tex
begin table of variable format

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Sep 24 06:28:59 2024
+% Wed Sep 25 16:53:01 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
    \hline
-c. & method & Brief overview of the method used to estimate the species abundance. &  & string \\ 
+c. & source & Source from which estimate were derived (i.e., measurements from the study area or  extracted from Wilman et al. 2014). &  & string \\ 
    \hline
-c.
-& mean\_abundance & Estimate of the mean annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August).
-This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration.
-The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults.
-& individuals & integer \\
-   \hline
-c. & sd\_abundance & Standard deviation of the annual species abundance. & individuals & numeric \\ 
-   \hline
-c. & sample\_size\_abundance & Number of years consider in the calculation of mean annual abundance. & years & integer \\ 
+c. & body\_mass\_g & Mean individual body mass. & grams & numeric \\ 
    \hline
@@
    \hline
-d. & source & Source from which estimate were derived (i.e., measurements from the study area or  extracted from Wilman et al. 2014). &  & string \\ 
+d. & zone & Sampled zone of the study area (see figure 2 and 3). &  & string \\ 
    \hline
-d. & body\_mass\_g & Mean individual body mass. & grams & numeric \\ 
-   \hline
-e. & species & Common species english name. &  & string \\ 
-   \hline
-e. & zone & Sampled zone of the study area (see figure 2 and 3). &  & string \\ 
-   \hline
-e. & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
+d. & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
 of the study area. & nests per square kilometer & numeric \\ 
    \hline
-e. & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
+d. & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
 zone of the study area. & nests per square kilometer & numeric \\ 
    \hline
-e. & sample\_size\_nest\_density\_km2 & Number of years consider in the calculation of the nest density. & years & integer \\ 
+d. & sample\_size\_nest\_density\_km2 & Number of years consider in the calculation of the nest density. & years & integer \\ 
    \hline

### 2024-09-26 manuscript/tables/table_summary_methods.tex
begin table of variable format

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Sep 10 15:05:42 2024
+% Wed Sep 25 09:14:30 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
   \hline
-{\textbf{Species}} & {\textbf{Estimate}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Min}} & {\textbf{Max}} & {\textbf{Mean}} & {\textbf{sd}} & {\textbf{n}} & {\textbf{Expert based confidence interval}} \\
+{\textbf{Species}} & {\textbf{Estimate}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Lowest abundance}} & {\textbf{Highest abundance}} & {\textbf{Mean abundance}} & {\textbf{sd}} & {\textbf{n}} & {\textbf{Expert based confidence interval}} \\
   \hline

### 2024-09-26 script/extract_species_abundance/species/snow_goose.R

@@ -176,7 +176,7 @@ ggplot(data= mesic_density, aes(x= year, y= ind_density_km2, col= data))+
   geom_line(linewidth=0.75)+
   scale_x_continuous(breaks = seq(1999, 2023, by = 3))+
   theme_classic()
-ggsave(units= "cm",width = 18,height = 10,dpi= 500, "manuscript/figures/snow_goose_mesic_density.pdf")
+ggsave(units= "cm",width = 18,height = 10,dpi= 500, "MetadataS1/figures/snow_goose_mesic_density.pdf")
 
 
 #-------------------------------#
@@ -273,4 +273,4 @@ map <-
             dplyr::filter(year== 2017, n_oie_125 == 0), aes(geometry= geometry), alpha=1, fill= "#d00000",col= "#d00000")+
   theme_void()
 
-ggsave(plot= map, units = "cm", height= 18, width=10,"manuscript/figures/map_geese_colony_2017.pdf")
+ggsave(plot= map, units = "cm", height= 18, width=10,"MetadataS1/figures/map_geese_colony_2017.pdf")

### 2024-09-26 script/functions/standalone_functions/extract_density_distance_sampling.R

@@ -76,6 +76,7 @@ sampled_year_zone <- data.frame(zone= rep(transect_area$zone, times= length(year
 sp_abundance <- data.frame()
 
 #for each year and zone combination (e.g., camp 2 in 2017)
+i= 1
 for (i in 1:nrow(sampled_year_zone)){
   #Extract the year of interest as an object
   year_select= sampled_year_zone$year[i]
@@ -96,7 +97,8 @@ obs <- transects %>%
 #--- If presence of observation estimate density
 if (nrow(obs) >0){
   #Create a object id containing a unique id for each observation
-obs_table <-  obs %>% 
+  #repeat the row of transects with multiple individuals observed the number of times corresponding to the number of individuals. Since each row correpond to a given individual detected.
+obs_table <- obs[rep(1:nrow(obs), times = obs$nb_ind), ] %>% 
   dplyr::mutate(Sample.Label= paste(transect, year), object= 1:length(year)) %>% 
   dplyr::rename(Region.Label= zone) %>% 
   sf::st_drop_geometry() %>% 

### 2024-09-26 script/tables_figures/maps.R

@@ -34,6 +34,7 @@ goose_colony <- sf::st_read("data/raw/shapefiles/study_area.shp") %>%
   dplyr::filter(year== 2017, zone== "goose_colony") %>% 
   sf::st_transform(3857)
 
+
 #color palette
 colors_palette <- c("#bd959d","#84a98c", "#94214b", "#463a49","#ad382b","#096377","#63141c", "#003049", "#d4af37", "white")
 
@@ -72,7 +73,7 @@ ggmap(map) +
   geom_sf(data = study_area, aes(fill = zone, color= zone), lwd= 0.8,inherit.aes = FALSE)+
   scale_fill_manual(values= alpha(colors_palette,0.5))+
   scale_color_manual(values= alpha(colors_palette, 0.9),
-                     labels=c("camp 2"=  expression("Camp 2 - 80 km"^2), "goose point"= expression("Goose point - 41 km"^2),"malaview"= expression("Malaview - 60 km"^2),"qarlikturvik valley"= expression("Qarlikturvik valley - 65 km"^2), "camp 3"= expression("Camp 3 - 33 km"^2), "black plateau"= expression("Black plateau - 34 km"^2), "south plateau"= expression("Southern plateau - 12 km"^2), "camp 3 plateau"= expression("Camp 3 plateau - 30 km"^2), "dufour"= expression("Pointe Dufour - 33 km"^2), "goose_colony"= expression("Snow goose colony - 74 km"^2)))+
+                     labels=c("camp 2"=  expression("Camp 2 - 80 km"^2), "goose point"= expression("Goose point - 41 km"^2),"malaview"= expression("Malaview - 60 km"^2),"qarlikturvik valley"= expression("Qarlikturvik valley - 65 km"^2), "camp 3"= expression("Camp 3 - 33 km"^2), "black plateau"= expression("Black plateau - 34 km"^2), "south plateau"= expression("Southern plateau - 12 km"^2), "camp 3 plateau"= expression("Camp 3 plateau - 30 km"^2), "dufour"= expression("Pointe Dufour - 33 km"^2), "goose_colony"= expression("")))+
   geom_sf(data = goose_colony,fill= NA, aes(color=zone), lwd= 0.8, linetype = "twodash", inherit.aes = FALSE)+
   ggplot2::coord_sf(xlim = c(-8822069.645366931, -8926709.966712607), ylim = c(12028963.853099173, 12238691.808127215))+ # set the map lat/lon limits
[…]

### 2024-09-26 script/tables_figures/table_interannual_variation_nest_density.R

@@ -106,7 +106,7 @@ print(table_latex,
       width="0.9\\textwidth",
       align="rXXXXXX",
       size="\\fontsize{10pt}{10pt}\\selectfont",
-      file = "manuscript/tables/interannual_nest_density_variation.tex",
+      file = "MetadataS1/tables/interannual_nest_density_variation.tex",
       caption.placement = "top")
 
 

### 2024-09-26 script/tables_figures/table_mean_abundance.R

@@ -1,78 +1,0 @@
-# Extract Bylot Island temporal series of species abundance and biomass
-#Date: November 27 2023
-
-#------------------#
-#### Librairies ####
-#------------------#
-#data manipulation
-library(dplyr)
-library(tidyr)
-#export latex table
-library(xtable)
-
-
-
-#-------------------#
-#### Import data ####
-#-------------------#
-#metadata
-metadata <- read.csv("data/metadata/species_method_selected.csv", sep = "\t")
-#reorder names based on defined species list
-sp_metadata <- read.csv("data/metadata/species_metadata.csv")
-# Find the indices of species in df2 in df1
-indices <- match(sp_metadata$species, metadata$species)
-# Reorder df1 based on the indices
-metadata <- metadata[indices, , drop = FALSE] 
-
-#update species names
-metadata <- metadata %>% 
-  dplyr::mutate(species= dplyr::recode(species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper", "lapland longspur"= "Lapland longspur"))
-
-
-#Read species final densities
-#Abundance data
-abundance <- read.csv("dataset/BYLOT-species_abundance.csv") %>%
-  dplyr::mutate(method= ifelse(species== "cackling goose" & year<= 2016, "nest sampling (model)", method)) %>% 
-  dplyr::mutate(year= ifelse(is.na(year), "mean", year),
-                method= dplyr::recode(method, "nest sampling (extrapolation nests)"= "nest sampling")) %>% 
-  dplyr::group_by(species, method) %>% 
-  dplyr::summarise(mean= as.character(round(mean(abundance), digits=0)), sd= as.character(round(sd(abundance), digits= 0)), n= n()) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::mutate(n= ifelse(n==1, NA, n))
-
-
-#--------------------------#
-#### Generate the table ####
-#--------------------------#
[…]

### 2024-09-26 script/tables_figures/table_species_name_strategy.R

@@ -20,7 +20,7 @@ print(sp_name_latex,
       width="0.95\\textwidth",
       align="rXXXXXX",
       size="\\fontsize{10pt}{10pt}\\selectfont",
-      file = "manuscript/tables/species_name_strategy.tex",
+      file = "MetadataS1/tables/species_name_strategy.tex",
       caption.placement = "top")
 
 

### 2024-09-26 script/tables_figures/table_species_relative_abundance_upland.R

@@ -73,5 +73,5 @@ print(table_latex,
       width="0.55\\textwidth",
       align="rXXXXXX",
       size="\\fontsize{10pt}{10pt}\\selectfont",
-      file = "manuscript/tables/species_relative_abundance_upland.tex",
+      file = "MetadataS1/tables/species_relative_abundance_upland.tex",
       caption.placement = "top")

### 2024-09-26 script/tables_figures/table_species_year_monitoring.R

@@ -50,7 +50,7 @@ table_year_monitoring[table_year_monitoring$Species == "rough-legged hawk" & tab
 table_year_monitoring <- table_year_monitoring[order(table_year_monitoring$Zone), ]
 
 #as tex
-table_year_monitoring_latex <- xtable(table_year_monitoring, caption = "Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in in Gauthier et al. (2024b).",
+table_year_monitoring_latex <- xtable(table_year_monitoring, caption = "Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in Gauthier et al. (2024b).",
                                       label= "table:species_year_monitoring")
 
 print(table_year_monitoring_latex,
@@ -59,6 +59,6 @@ print(table_year_monitoring_latex,
       width="\\textwidth",
       align="rXXXXXX",
       size="\\fontsize{8pt}{10pt}\\selectfont",
-      file = "manuscript/tables/species_year_monitoring.tex",
+      file = "MetadataS1/tables/species_year_monitoring.tex",
       caption.placement = "top")
 

### 2024-09-26 script/tables_figures/table_summary.R

@@ -35,25 +35,31 @@ table <- metadata %>%
 
 #format table
 table <- table %>% 
-  dplyr::select(species, Estimate, Method.description, Method.quality, Justification, min, max, mean, sd, n, Expert.based.confidence.interval) %>% 
-  dplyr::rename(Species= species,  Method= Method.description,  `Method quality`= Method.quality, Min= min, Max= max, Mean= mean,  `Expert based confidence interval`=Expert.based.confidence.interval)
+  dplyr::select(species, Estimate, Method.description, Method.quality, Justification, min, max, mean, sd, n) %>% 
+  #rename columns
+  dplyr::rename(Species= species,  Method= Method.description,  `Method quality`= Method.quality, `Lowest abundance`= min, `Highest abundance`= max, `Mean abundance`= mean) %>% 
+  #Set mean to NA in column estimate
+  dplyr::mutate(Estimate= ifelse(Estimate == "mean", NA,Estimate)) %>% 
+  #Add years to column n (sample size)
+  dplyr::mutate(n= paste0(n, " (",Estimate,")")) %>% 
+  dplyr::mutate(n= ifelse(n== "NA (NA)", NA, n)) %>% 
+  #Remove column Estimate
+  dplyr::select(-Estimate)
 
 
 #as tex
-table_latex <- xtable(table, caption = "Summary of the mean (± standard deviation) and minimum and maximum (when available) annual abundance of vertebrate species for the whole Bylot Island study area (389 km2).  In some cases, several independent approaches have been used to estimate the abundance of a species as a proxy for uncertainty. We did not quantitatively assess uncertainty on each estimate of species abundance. However, we provide a qualitative measure of method quality based on the quality and quantity of data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density. We also provided coarse confidence intervals derived from consultation with field experts at the study site.",
+table_latex <- xtable(table, caption = "Summary of the lowest, highest, mean and standard deviation of the estimated abundance of each vertebrate species in the vertebrate community of the southern plain of Bylot Island (389 km2). In some cases, two independent approaches have been used to estimate the abundance of the same species as a proxy for uncertainty. We provide a qualitative measure of the method quality based on data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.",
                       label= "table:summary_methods",
                       align= c("p{0.00\\textwidth}", 
-                               "|p{0.11\\textwidth}|", #species
[…]

### 2024-09-26 table_variables_format.R

@@ -1,0 +1,39 @@
+#Title: Create a table of all variables and their associated type,code definition and range of values
+#Author: Louis Moisan
+#Date: September 25 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+#Import table structure .csv format
+table <- read.csv("MetadataS1/tables/table_data_format.csv") 
+
+#Import data set
+species_taxonomy <- read.csv("dataset/BYLOT-species_taxonomy.csv")
+species_abundance <- read.csv("dataset/BYLOT-species_abundance.csv")
+species_body_mass <- read.csv("dataset/BYLOT-species_body_mass.csv")
+variation_nest_density <- read.csv("dataset/BYLOT-interannual_variation_nest_density.csv")
+
+#Assign range values
+table[table$Variable.identity == "year",]$Range <- paste0(min(as.numeric(species_abundance$year), na.rm = T), "-", max(as.numeric(species_abundance$year), na.rm = T))
+table[table$Variable.identity == "abundance",]$Range <- paste0(min(species_abundance$abundance), "-", max(species_abundance$abundance))
+table[table$Variable.identity == "body_mass_g",]$Range <- paste0(min(species_body_mass$body_mass_g), " - ", max(species_body_mass$body_mass_g))
+table[table$Variable.identity == "mean_nest_density_km2",]$Range <- paste0(min(variation_nest_density$mean_nest_density_km2), "-", max(variation_nest_density$mean_nest_density_km2))
+table[table$Variable.identity == "sd_nest_density_km2",]$Range <- paste0(min(variation_nest_density$sd_nest_density_km2), "-", max(variation_nest_density$sd_nest_density_km2))
+table[table$Variable.identity == "sample_size_nest_density_km2",]$Range <-paste0(min(variation_nest_density$sample_size_nest_density_km2), "-", max(variation_nest_density$sample_size_nest_density_km2))
+
+#number of digits
+table[table$Storage.type=="numeric",]$Number.digits <- 3
+table[table$Storage.type=="integer",]$Number.digits <- 0
+
+
+unique(species_taxonomy$annual_cycle_strategy)
+unique(species_abundance$status)
+unique(species_abundance$method)
+unique(species_body_mass$source)
+unique(variation_nest_density$zone)

### 2024-10-01 MetadataS1/MetadataS1.tex
update table variable definition and complete first draft of the entire document

@@
 \documentclass[a4paper,twoside,12pt]{article}
-
 %% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{enumitem}
-
 %% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{pdflscape}
-
 %bibliography
@@
 \usepackage{geometry}
-
 %% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{amsfonts}
-
 %% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \linenumbers
-
 \usepackage{titlesec}
@@
 \titlespacing{\subparagraph}{80pt}{*4}{*1.5}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \centering
-\textbf{Title: Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
-
-\textbf{Authors}:\\
+\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
  Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
  Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
-
 \textbf{Author Affiliations}:\\
-\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
-\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
-\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
-\textsuperscript{4} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
-\textsuperscript{5} Centre d’études nordiques, Centre for Arctic Knowledge and Exploration, Canadian Museum of Nature, Ottawa, ON, Canada\\
+\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
+\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
+\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
+\textsuperscript{4} Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
+\textsuperscript{5} Centre d’Études Nordiques, Centre de Connaissance et d’Exploration de l’Arctique, Musée Canadien de la Nature, Ottawa, ON, Canada\\
 \textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
 \\[1.5em]
-
 \textbf{Corresponding Author}:\\
@@
 \\[1.5em]
-
-\textbf{Open Research statement}:
-
-%%%%%%% !!!! ADD DRYAD and Zenodo adress !!!! %%%%%%%%
-Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
-Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
+\textbf{Open Research statement}:\\
+The data set is publicly available at \url{https://datadryad.org/} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/}.\\
 \endgroup
@@
 Dynamic community modelling allows addressing important issues and questions in ecology, such as: determining the relative strength of top-down versus bottom-up forces in communities \citep{krebs2003,legagneux2014}, assessing the ecological resilience of communities under climate change \citep{griffith2019} and evaluating the cascading effects of invasive species in food web \citep{david2017, goto2020}.
-Dynamic community modelling can also be applied to address practical challenges, including fishery management (Plagányi, 2007) and the planning of protected areas \citep{okey2004, dahood2020}.
-
+Dynamic community modelling can also be applied to address practical challenges, including fishery management \citep{plaganyi2007} and the planning of protected areas \citep{okey2004, dahood2020}.
 Modeling food webs requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community.
@@
 Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.
-
 The composition of ecological communities is influenced by various factors acting at different temporal and spatial scales, leading to noisy data and emphasizing the need for long-term data sets \citep{magurran2010, lindenmayer2012}.
@@
 Therefore, long-term data series are required to untangle the relative effects of diverse abiotic and biotic factors on community composition \citep{magurran2010, lindenmayer2012}.
-
 Arctic environments are highly valuable systems for studying community patterns and processes due to their relatively low species richness \citep{payer2013, legagneux2014}.
@@
 Datasets on terrestrial communities are notably scarce, and this scarcity extends to Arctic communities as well \citep{ims2013}.
-
 Within terrestrial Arctic sites, the south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive biodiversity monitoring programs \citep{gauthier2024a}.
@@
 This approach allows the scaling of local density measurements to the landscape level when required and facilitates the estimation of abundance for less common and rare species.
-
 Previous work based on the tundra community of Bylot Island  has already produced several influential papers \citep{gauthier2011, legagneux2012, legagneux2014,hutchison2020, duchesne2021, gauthier2024b}.
@@
 This synthesis aims to provide readily accessible annual time series (or mean values in some cases) of abundance and biomass for all vertebrate species in a tundra landscape, covering approximately 400 km\textsuperscript{2}.
-
 \begin{figure}[H]
@@
 \end{figure}
-
 \subsection*{Objective}
@@
 \newpage
-
 \section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
@@
                 \subsection*{1. Originators}
-		\textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
-  		\textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
-  		\textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+                \textbf{Gilles Gauthier}, Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
+                \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
+                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
 
@@
 Considering that the study site is a pristine and protected area that has experienced minimal anthropogenic impact, it can also provide an ideal baseline for investigating the impacts of global changes on high-latitude terrestrial ecosystems.
-
    \section*{D. Key words/phrases}
@@
     \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial vertebrate communities
-
         \subsection*{2. Originators:}   
@@
                         \subsubsection*{a. Site type}
-  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}.
-The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}).
+  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.
+An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}.
+The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:table_species_name_strategy}}; \citet{moisan2023, gauthier2024a}).
 The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}.
@@
 Therefore, we decided to exclude this species.
-
-         		\input{tables/species_name_strategy.tex}
-%label-> table:species_name_strategy
+                        \input{tables/table_species_name_strategy.tex}
+%label-> table:table_species_name_strategy
 \newpage
@@
 Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
-            	\subsubsection*{d. Geology} see \citet{klassen1993} for a complete and detailed description of the geology of the study area.
+                \subsubsection*{d. Geology} See \citet{klassen1993} for a complete and detailed description of the geology of the study area.
                 \subsubsection*{e.
 Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
-            	\subsubsection*{f. Site history} see \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
+                \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                 \subsubsection*{g.
@@
 The zone between the Qarlikturvik valley and Camp 3 received very little sampling effort and is therefore excluded from the study area.
-
 \begin{figure}
@@
 \end{figure}
-
 \begin{figure}
@@
 \newpage
-
                 \subsubsection*{b. Avian nest monitoring}
@@
 Nest densities derived from nest sampling could be underestimated due to early nest failure (i.e., failure that happened before our sampling period).
-
-\input{tables/species_year_monitoring.tex}
-%label-> table:species_year_monitoring
-
+\input{tables/table_species_year_monitoring.tex}
+%label-> table:table_species_year_monitoring
         \begin{enumerate}[label=\roman*]
@@
         \end{enumerate}
-
-\input{tables/interannual_nest_density_variation.tex}
-%label-> table:interannual_nest_density_variation
+\input{tables/table_interannual_nest_density_variation.tex}
+%label-> table:table_interannual_nest_density_variation
 \newpage
-
 \subsubsection*{c. Observation of individuals}
@@
 \newpage
-
 \item[] \textit{\textbf{Snow goose point count}}\\
@@
 \\
-
 \item[] \textit{\textbf{Incidental observations}}\\
@@
 We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species.
-However, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+However, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands.
 \\
-
-\input{tables/species_relative_abundance_upland.tex}
-%label-> species_relative_abundance_upland
+\input{tables/table_species_relative_abundance_upland.tex}
+%label-> table_species_relative_abundance_upland
 \newpage
-
 \item[] \textit{\textbf{Testimonials of ermine sightings}}\\
@@
 \end{enumerate}
-
 \subsubsection*{d. Capture of individuals}
@@
 \\
-
 \item[] \textit{\textbf{Arctic fox movement tracking}}\\
@@
 \\
-
 \item[] \textit{\textbf{Parasitic jaeger banding}}\\
@@
 \end{enumerate}
-
 \subsubsection*{e. Species body mass}
-All vertebrate individuals captured for marking purposes were systematically weighed (snow goose (unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021},  common-ringed plovers \citep{leandri2019}, other shorebirds (Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (Gauthier and Therrien, unpublished data), Lapland longspurs (Gauthier and Bêty, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}).
+All vertebrate individuals captured for marking purposes were systematically weighed: snow goose (G.
+Gauthier, M.-C.
+Cadieux and J.
+Lefebvre, unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021}, common-ringed plovers \citep{leandri2019}, other shorebirds (J.
+Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (J.-F.
+Therrien and G.
+Gauthier, unpublished data), Lapland longspurs (J.
+Bêty and G.
+Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}.
 Mean body mass of vertebrate species was determined in priority with measurements from Bylot Island.
@@
 Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, ermine, and Arctic hare.
-The methods used for each species are summarized in (\textbf{Table \ref{table:summary_methods}}).
+The methods used for each species are summarized in (\textbf{Table \ref{table:table_summary_methods_and_abundance}}).
 \newpage
-
         \begin{enumerate}[label=\alph*.]
@@
 We transformed the number of nests to abundance of individuals by multiplying it by a factor of two.
-
 We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4).
@@
 We multiplied nest density by two to obtain the abundance.
-
 \begin{figure}[h]
@@
 \end{figure}
-
       \item[] \textit{\textbf{Snow goose}}\\
@@
 Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).
-
+                                \newpage
         \begin{figure}[ht]
@@
         \end{figure}
-
         \begin{figure}[ht]
@@
 While we acknowledge that the opportunistic monitoring of this species likely underestimates nest density, the extensive sampling effort deployed annually within this area gives us confidence in obtaining a realistic number of nests.
-We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat.
@@
 We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+
 \begin{figure}[H]
@@
 \end{figure}
-
                 \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
@@
 We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 
@@
 We therefore derived the absolute abundance of black-bellied plovers from their relative abundance using the ratio between relative and absolute abundances of American golden-plover as a reference.
-
 As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.
-
                 \item[] \textit{\textbf{Common-ringed plover}}\\
@@
 We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.
@@
                 \end{figure} 
-
                 \item[] \textit{\textbf{Other passerines and sandpipers}}\\
-        	We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section 3.9).
+                We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section \textit{\textbf{Lapland longspur and Baird’s sandpiper}}).
 We assumed here a similar detection probability for all species.
-We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section 3.9).
+We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section \textit{\textbf{Lapland longspur and Baird’s sandpiper}}).
 For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 Nest density was then converted in number of individuals by multiplying by the area and a factor 2.
@@
 We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.
-
                 \item[] \textit{\textbf{Long-tailed jaeger}}\\
@@
 Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.
-
         \item[] \textit{\textbf{Brown and collared lemming}}\\
@@
 We thus derived abundance of hares from the estimated abundance of Arctic foxes based on indices of relative abundance presented in \citep{gauthier2024a}, which were derived from incidental daily observations.
-We doubled the density of Arctic hares in the upland zones (i.e., plateaus), as twice as many individuals were observed per hour of fieldwork there compared to lowland zones (\textbf{Table \ref{table:species_relative_abundance_upland}}).
+We doubled the density of Arctic hares in the upland zones (i.e., plateaus), as twice as many individuals were observed per hour of fieldwork there compared to lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 However, it is worth noting that assuming a similar detection probability between foxes and hares might lead to an overestimation of hare detection probability due to behavioral differences between the species.
@@
 \begin{landscape}
-\input{tables/table_summary_methods.tex}
-%label-> table:summary_methods
+\input{tables/table_summary_methods_and_abundance.tex}
+%label-> table:table_summary_methods_and_abundance
 \end{landscape}
 \restoregeometry
+\newpage
+\begin{figure}[h]
+\caption{Time series of the estimated annual adult abundance of vertebrate species on the southern plain of Bylot Island (389 km\textsuperscript{2}).
+Estimated abundance represents adult individuals, with the exception of lemmings, for which juveniles were also included in the estimate.
+Time series shorter than 5 years are not presented.}
+  \centering
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Cackling_goose.pdf}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Snow_goose.pdf}
+  \end{subfigure} 
+    \hfill
+    \hfill
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Rough-legged_hawk.pdf}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Peregrine_falcon.pdf}
+  \end{subfigure}
+\end{figure}
+\newpage
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Snowy_owl.pdf}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/American_golden-plover.pdf}
+  \end{subfigure} 
+    \hfill
+    \hfill
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Glaucous_gull.pdf}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Long-tailed_jaeger.pdf}
+  \end{subfigure}
+\end{figure}
+\newpage
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Nearctic_brown_lemming.pdf}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Nearctic_collared_lemming.pdf}
+  \end{subfigure} 
+    \hfill
+    \hfill
+  \begin{subfigure}{0.45\textwidth}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/Ermine.pdf}
+  \end{subfigure}
+\end{figure}
+\newpage
+\input{tables/table_expert_based_uncertainty.tex}
+%label-> table:table_expert_based_uncertainty
 \newpage
@@
 \newpage 
-
  \section*{\textcolor{Blue}{Class III. Data set status and accessibility}}
     \section*{A. Status}
- 		\subsection*{1. Latest update} 19th September 2024
+                \subsection*{1. Latest update} 30th September 2024
         \subsection*{2. Latest archive date} XXXXXX October 2024
@@
         \subsection*{4. Data verification} The methods used to extract final species abundance estimates were subject to several rounds of revision by the principal investigators.
-
    \section*{B. Accessibility}
                 \subsection*{1. Storage location and medium}
-Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
+The data set is publicly available at \url{https://datadryad.org/}.\\[1.5em]
 Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
@@
                 \end{enumerate}
-
                 \subsection*{3. Copyright restrictions}
@@
                 \subsection*{5. Costs}
-        	None	, the data can be used free of charge.		
+                None, the data can be used free of charge.          
 \newpage
@@
                         \subsection*{2. Size} 
-        		a. 35 records, not including header row (2.3 kB)\\
-        		b. 273 records, not including header row  (17.0 kB)\\
-        		c. 51 records, not including header row  (1.8 kB)\\
-        		d. 18 records, not including header row  (1.1 kB)
+                        a. 35 records, not including header row (4.3 kB)\\
+                        b. 271 records, not including header row  (34.5 kB)\\
+                        c. 53 records, not including header row  (3.5 kB)\\
+                        d. 18 records, not including header row  (961 B)
 
@@
                                 \subsubsection*{a. \texttt{BYLOT-species\_taxonomy.csv}}
-        			\texttt{functional\_group}; \texttt{scientific\_name}; \texttt{species}; \texttt{annual\_cycle\_strategy}
+                                \texttt{class}; \texttt{order}; \texttt{family}; \texttt{genus}; \texttt{species\_scientific}; \texttt{species\_en}; \texttt{species\_fr}; \texttt{species\_code}; \texttt{functional\_group}; \texttt{migratory\_status}
 
                                 \subsubsection*{b. \texttt{BYLOT-species\_abundance.csv}}
-        			\texttt{species}; \texttt{year}; \texttt{status}; \texttt{method}; \texttt{abundance}
+                                \texttt{species\_en}; \texttt{year}; \texttt{breeding\_status}; \texttt{abundance}; \texttt{method\_description}; \texttt{method\_quality}
 
                                 \subsubsection*{c. \texttt{BYLOT-species\_body\_mass.csv}}
-        			\texttt{species}; \texttt{source}; \texttt{body\_mass\_g}
+                                \texttt{species\_en}; \texttt{site};\texttt{mean\_body\_mass\_g}; \texttt{sample\_size}; \texttt{reference}
 
                                 \subsubsection*{d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}}
-        			\texttt{species}; \texttt{zone}; \texttt{mean\_nest\_density\_km2}; \texttt{sd\_nest\_density\_km2}; \texttt{sample\_size\_nest\_density\_km2}
+                                \texttt{species\_en}; \texttt{zone}; \texttt{mean\_nest\_density\_km2}; \texttt{sd\_nest\_density\_km2}; \texttt{number\_years}
 
@@
                         Sums of the numeric columns: \\
-        		b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2291027\\
-        		c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 46223\\
+                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2293389\\
+                        c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 49617; sample\_size= 13902\\
                         d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185
@@
     \section*{B. Variable information}
-    		\subsection*{1. Variable identity} See Table \ref{table:summary_variables}
-        \subsection*{2. Variable definition} See Table \ref{table:summary_variables}
-        \subsection*{3. Units of measurement} See Table \ref{table:summary_variables}
+                \subsection*{1. Variable identity} See Table \ref{table:table_variables_definition}
+        \subsection*{2. Variable definition} See Table \ref{table:table_variables_definition}
+        \subsection*{3. Units of measurement} See Table \ref{table:table_variables_definition}
+        \newpage
 
-        \input{tables/table_summary_variables.tex}
+        \input{tables/table_variables_definition.tex}
+       \newpage
 
         \subsection*{4. Data type}
-        		\begin{enumerate}[label=\alph*.]
-            		\item Storage type: Integer, floating point, character, string, etc.
-            		\item List and definition of variable codes: Description of any codes associated with variables
-            		\item Range for numeric values: Minimum, maximum
-            		\item Missing value codes: Description of how missing values are represented in data set
-            		\item Precision: Number of significant digits
-            \end{enumerate}
+                    \subsubsection*{a. Storage type} See Table \ref{table:table_variables_format}
+                    \subsubsection*{b. List and definition of variable codes} See Table \ref{table:table_variables_format}
+                    \subsubsection*{c. Range for numeric values} See Table \ref{table:table_variables_format}
+                    \subsubsection*{d. Missing value codes} Unavailable values are indicated by NA.
+                    \subsubsection*{e. Number of digits} See Table \ref{table:table_variables_format}
+            \newpage
 
-        \subsection*{5. Data format}
-        		\begin{enumerate}[label=\alph*.]
-        			\item Fixed, variable length
-            		\item Columns: Start column, end column
-            		\item Optional number of decimal places
-        		\end{enumerate}
+           \input{tables/table_variables_format.tex}
+%label-> table:table_variables_format
+			\newpage
 
  \section*{C. Data anomalies: Description of missing data, anomalous data, calibration errors, etc.}
-
- \newpage
+  If abundance of a given species has not been calculated for a series of years, but rather as a general average, then NA has been assigned as "year".
 
@@
     \section*{A. Data acquisition}
-    \begin{enumerate}
-       \item Data forms or acquisition methods: Description or examples of data forms, automated data loggers, digitizing procedures, etc.
-       \item Location of completed data forms
-       \item Data entry verification procedures: Procedures employed to verify that digital data set is free of errors
-    \end{enumerate}
+    		\subsection*{1. Data forms or acquisition methods}
+   		See Section \textbf{2. Experimental or sampling design} 
+   		\subsection*{2. Location of completed data forms}
+        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://zenodo.org/}.
+        \subsection*{3. Data entry verification procedures}
+       	The methods used to extract final species abundance estimates were subject to several rounds of revision by the authors.
 
-    \section*{B.
-Quality assurance/quality control procedures}Identification and treatment of outliers, description of quality assessments, calibration of reference standards, equipment performance results, etc.
-    \section*{C. Related materials} References and locations of maps, photographs, videos, GIS data layers, physical specimens, field notebooks, comments, etc.
-    \section*{D. Computer programs and data-processing algorithms}
-    		\begin{enumerate}
-    			\item \textbf{Program}: R version 4.3.2 (2023-10-31)
-    			\item \textbf{Operating system}: x86\_64-pc-linux-gnu (64-bit) with Ubuntu 22.04.3 LTS
-    			\item \textbf{Packages}: dplyr, tidyr, sf, stringr, xtable, Distance, ggplot2, lme4, AICcmodavg, scales, ggmap
-    			\item \textbf{Codes}: All raw data and codes are publicly available at \url{https://github.com/chaireBioNorth/BYLOT_species_abundance_dataset.git}
-    		\end{enumerate}
+    \section*{B. Quality assurance/quality control procedures}
+    Final abundance estimate were revised by the authors.
 
-    \section*{E. Archiving}
-    		\begin{enumerate}
-    		 \item \textbf{Archival procedures}: Data will be archived on Dryad upon acceptance for publication
-        	 \item \textbf{Redundant archival sites}: n/a
-    		\end{enumerate}
+    \section*{C. Computer programs and data-processing algorithms}
+         \subsection*{1. Program} R version 4.3.2 (2023-10-31)
+         \subsection*{2. Operating system} Data preparation was performed on x86\_64-pc-linux-gnu (64-bit) with Ubuntu 22.04.3 LTS
+         \subsection*{3.
+Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{Distance}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
+         \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
 
-    \section*{F. Publications and results: Electronic reprints, lists of publications resulting from or related to the study, graphical/statistical data representations, etc.}
-\textbf{List of publications that include data from the data set}:
-		\begin{enumerate}
-			\item 
-		\end{enumerate}
+    \section*{D. Archiving}
+         \subsection*{1. Archival procedures} Data are publicly available at \url{https://datadryad.org/}.
+         \subsection*{2. Redundant archival sites} None
 
-   \section*{G. History of data set usage}
-  		\begin{enumerate}
-  			\item \textbf{Data request history}: n/a
-        		\item \textbf{Data set update history}: n/a
-        		\item \textbf{Review history}: n/a
-       		\item \textbf{Questions and comments from secondary users}: n/a
-  		\end{enumerate}
+    \section*{E. Publications and results}
+The presented estimates of species abundance have not been integrated in publications to date.
+Previous estimates of species abundance on the souther plain of Bylot Island were presented by \citet{legagneux2012}, however, the temporal series presented here is longer, the methods are more refined and the taxonomic resolution is higher.
+
+   \section*{F. History of data set usage}
+       \subsection*{1.Data request history} None
+       \subsection*{2.Data set update history} None
+       \subsection*{3.Review history}  None
+       \subsection*{4.Questions and comments from secondary users} None
+
+   \section*{Acknowledgements}
+We are deeply appreciative of the extensive data gathered over the decades by generations students and researchers during their fieldwork at the Bylot Island research station, which made this project possible.
+Our gratitude also extends to the Center of Northern Studies for providing research facilities, as well as to the Polar Continental Shelf Program and Sirmilik National Park (Parks Canada) for their logistical support throughout the years.
+Additionally, we would like to express our special thanks to the Mittimatalik community and the Mittimatalik Hunters and Trappers Organization for their ongoing support of ecological monitoring on Bylot Island and for permitting us to conduct research on their land.
 
@@
 \bibliography{references}
-
 \end{document}

### 2024-10-01 MetadataS1/tables/table_interannual_nest_density_variation.tex
update table variable definition and complete first draft of the entire document

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 09:17:41 2024
\begin{table}[ht]
\centering
\caption{Annual nest density (nests/km2) of selected avian species estimated on different zones of Bylot Island.} 
\label{table:table_interannual_nest_density_variation}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.95\textwidth}{lllr}
  \hline
Species & Zone & Mean nests/km2 ± SD & Number of years \\ 
  \hline
Lapland longspur & qarlikturvik (2x1 km plot) & 13.559 ± 5.849 &  17 \\ 
  Baird's sandpiper & qarlikturvik (2x1 km plot) & 5.000 ± 3.558 &  17 \\ 
  Rock ptarmigan & qarlikturvik (4x2 km plot) & 0.031 ± 0.055 &  16 \\ 
  Long-tailed duck & qarlikturvik (4x2 km plot) & 0.092 ± 0.138 &  16 \\ 
  King eider & qarlikturvik (4x2 km plot) & 0.115 ± 0.138 &  16 \\ 
  Sandhill crane & qarlikturvik (4x2 km plot) & 0.161 ± 0.087 &  16 \\ 
  Long-tailed jaeger & qarlikturvik valley & 0.362 ± 0.380 &  17 \\ 
  Tundra swan & whole study area & 0.001 ± 0.001 &   4 \\ 
  Common raven & whole study area & 0.003 ± 0.003 &   8 \\ 
  Pacific loon & whole study area & 0.005 ± 0.004 &   4 \\ 
  Parasitic jaeger & whole study area & 0.010 ± 0.004 &  12 \\ 
  Snowy owl & whole study area & 0.022 ± 0.058 &  10 \\ 
  Peregrine falcon & whole study area & 0.053 ± 0.007 &   8 \\ 
  Common Ringed plover & whole study area & 0.070 ± 0.012 &   3 \\ 
  Red-throated loon & whole study area & 0.082 ± 0.019 &   4 \\ 
  Glaucous gull & whole study area & 0.091 ± 0.011 &   4 \\ 
  Rough-legged hawk & whole study area & 0.157 ± 0.151 &   8 \\ 
  Cackling goose & whole study area & 0.177 ± 0.064 &   5 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-10-01 MetadataS1/tables/table_species_name_strategy.tex
update table variable definition and complete first draft of the entire document

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 09:17:32 2024
\begin{table}[ht]
\centering
\caption{Species composition of the vertebrate community of Bylot Island with the corresponding migratory status (i.e., resident, partial migrant or migrant).} 
\label{table:table_species_name_strategy}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.95\textwidth}{llll}
  \hline
Functional group & Scientific name & English name & Migratory status \\ 
  \hline
Ducks and loons  & Gavia pacifica & Pacific loon & migrant \\ 
  Ducks and loons  & Gavia stellata & Red-throated loon & migrant \\ 
  Ducks and loons  & Somateria spectabilis & King eider & migrant \\ 
  Ducks and loons  & Clangula hyemalis & Long-tailed duck & migrant \\ 
  Geese and swans  & Branta hutchinsii & Cackling goose & migrant \\ 
  Geese and swans  & Anser caerulescens & Snow goose & migrant \\ 
  Geese and swans  & Cygnus columbianus & Tundra swan & migrant \\ 
  Raptors  & Buteo lagopus & Rough-legged hawk & migrant \\ 
  Raptors  & Falco peregrinus & Peregrine falcon & migrant \\ 
  Raptors  & Bubo scandiacus & Snowy owl & migrant \\ 
  Ptarmigans  & Lagopus muta & Rock ptarmigan & resident \\ 
  Cranes  & Antigone canadensis & Sandhill crane & migrant \\ 
  Shorebirds  & Pluvialis dominica & American golden-plover & migrant \\ 
  Shorebirds  & Pluvialis squatarola & Black-bellied plover & migrant \\ 
  Shorebirds  & Charadrius hiaticula & Common Ringed plover & migrant \\ 
  Shorebirds  & Arenaria interpres & Ruddy turnstone & migrant \\ 
  Shorebirds  & Calidris canutus & Red knot & migrant \\ 
  Shorebirds  & Calidris melanotos & Pectoral sandpiper & migrant \\ 
  Shorebirds  & Calidris bairdii & Baird's sandpiper & migrant \\ 
  Shorebirds  & Calidris fuscicollis & White-rumped sandpiper & migrant \\ 
  Shorebirds  & Calidris subruficollis & Buff-breasted sandpiper & migrant \\ 
  Shorebirds  & Phalaropus fulicarius & Red phalarope & migrant \\ 
  Gulls  & Larus hyperboreus & Glaucous gull & migrant \\ 
  Jaegers  & Stercorarius longicaudus & Long-tailed jaeger & migrant \\ 
  Jaegers  & Stercorarius parasiticus & Parasitic jaeger & migrant \\ 
  Ravens  & Corvus corax & Common raven & partial migrant \\ 
  Passerines & Eremophila alpestris & Horned lark & migrant \\ 
  Passerines & Anthus rubescens & American pipit & migrant \\ 
  Passerines & Calcarius lapponicus & Lapland longspur & migrant \\ 
  Passerines & Plectrophenax nivalis & Snow bunting & migrant \\ 
  Lemmings  & Lemmus trimucronatus & Nearctic brown lemming & resident \\ 
  Lemmings  & Dicrostonyx groenlandicus & Nearctic collared lemming & resident \\ 
  Hares  & Lepus arcticus & Arctic hare & resident \\ 
  Ermines  & Mustela richardsonii & Ermine & resident \\ 
  Foxes  & Vulpes lagopus & Arctic fox & partial migrant \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-10-01 MetadataS1/tables/table_species_relative_abundance_upland.tex
update table variable definition and complete first draft of the entire document

--- 2024-06-25 manuscript/tables/species_relative_abundance_upland.tex
@@
-% latex table generated in R 4.4.0 by xtable 1.8-4 package
-% Tue Jun 25 10:25:27 2024
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Tue Oct  1 09:17:37 2024
 \begin{table}[ht]
@@
 The ratio compares relative abundance indexes between these two types of zones, calculated by dividing the upland by the lowland index of relative abundance.}
-\label{table:species_relative_abundance_upland}
+\label{table:table_species_relative_abundance_upland}
 \begingroup\fontsize{10pt}{10pt}\selectfont
@@
  \hline
-rock ptarmigan & 0.03 & 0.03 & 1 \\ 
-  sandhill crane & 0.42 & 0.287 & 1.5 \\ 
+Rock ptarmigan & 0.03 & 0.03 & 1 \\ 
+  Sandhill crane & 0.42 & 0.287 & 1.5 \\ 
   American golden-plover & 0.26 & 0.394 & 0.7 \\ 
-  black-bellied plover & 0.02 & 0.032 & 0.6 \\ 
-  ruddy turnstone & 0.01 & 0.007 & 1.3 \\ 
-  red knot & 0.00 & 0.033 & 0 \\ 
-  pectoral sandpiper & 0.02 & 0.034 & 0.6 \\ 
+  Black-bellied plover & 0.02 & 0.032 & 0.6 \\ 
+  Ruddy turnstone & 0.01 & 0.007 & 1.3 \\ 
+  Red knot & 0.00 & 0.033 & 0 \\ 
+  Pectoral sandpiper & 0.02 & 0.034 & 0.6 \\ 
   Baird's sandpiper & 0.31 & 0.32 & 1 \\ 
-  white-rumped sandpiper & 0.04 & 0.137 & 0.3 \\ 
-  buff-breasted sandpiper & 0.00 & 0.001 & 0 \\ 
-  red phalarope & 0.01 & 0.038 & 0.2 \\ 
-  horned lark & 0.24 & 0.154 & 1.6 \\ 
+  White-rumped sandpiper & 0.04 & 0.137 & 0.3 \\ 
+  Buff-breasted sandpiper & 0.00 & 0.001 & 0 \\ 
+  Red phalarope & 0.01 & 0.038 & 0.2 \\ 
+  Horned lark & 0.24 & 0.154 & 1.6 \\ 
   American pipit & 0.34 & 0.024 & 14.2 \\ 
-  snow bunting & 0.59 & 0.092 & 6.4 \\ 
+  Lapland longspur & 1.93 & 2.641 & 0.7 \\ 
+  Snow bunting & 0.59 & 0.092 & 6.4 \\ 
   Arctic hare & 0.02 & 0.009 & 2 \\ 

### 2024-10-01 MetadataS1/tables/table_species_year_monitoring.tex
update table variable definition and complete first draft of the entire document

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 09:17:34 2024
\begin{table}[ht]
\centering
\caption{Summary of vertebrate species monitoring in the Bylot Island study area. In this paper, we excluded certain years for specific species due to reduced sampling efforts. As a result, duration of times series presented here may differ slightly from those in Gauthier et al. (2024b).} 
\label{table:table_species_year_monitoring}
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{tabularx}{\textwidth}{lllll}
  \hline
Species & Zone & Years & Number of years & Monitoring \\ 
  \hline
Snow goose & camp 2 & 1999-2019, 2023 & (22) & systematic \\ 
  Rough-legged hawk & qarlikt., black \& south plat. & 2007-2019, 2022 & (15) & systematic \\ 
  Peregrine falcon & qarlikt., black \& south plat. & 2007-2019, 2022 & (15) & systematic \\ 
  Snowy owl & qarlikt., black \& south plat. & 1996-2019, 2023 & (25) & systematic \\ 
  Baird's sandpiper & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
  Lapland longspur & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
  King eider & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Long-tailed duck & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Rock ptarmigan & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Sandhill crane & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
  Nearctic brown lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
  Nearctic collared lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
  Pacific loon & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Red-throated loon & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Cackling goose & qarlikturvik valley & 2004-2019, 2022-2023 & (18) & systematic \\ 
  Tundra swan & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Glaucous gull & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Long-tailed jaeger & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Parasitic jaeger & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
  Pacific loon & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Red-throated loon & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Cackling goose & whole study area & 2017-2019, 2022-2023 & (5) & systematic \\ 
  Tundra swan & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Rough-legged hawk & whole study area & 2013-2019, 2022 & (8) & systematic \\ 
  Peregrine falcon & whole study area & 2013-2019, 2022  & (8) & systematic \\ 
  Snowy owl & whole study area & 2012-2019, 2022-2023 & (10) & systematic \\ 
  Common Ringed plover & whole study area & 2015-2017 & (3) & systematic \\ 
  Glaucous gull & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
  Parasitic jaeger & whole study area & 2009-2019, 2022 & (12) & opportunistic \\ 
  Common raven & whole study area & 2013-2019, 2022 & (8) & systematic \\ 
  Ermine & whole study area & 1993-2019 & (27) & opportunistic \\ 
  Arctic fox & whole study area & 2008-2016 & (9) & systematic \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-10-01 MetadataS1/tables/table_summary_methods_and_abundance.tex
update table variable definition and complete first draft of the entire document

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 09:17:43 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.10\textwidth}|p{0.27\textwidth}|p{0.06\textwidth}|p{0.23\textwidth}|p{0.09\textwidth}|p{0.09\textwidth}|p{0.085\textwidth}|p{0.04\textwidth}|p{0.10\textwidth}|}
\caption{Summary of the lowest, highest, mean and standard deviation of the estimated abundance of each vertebrate species in the vertebrate community of the southern plain of Bylot Island (389 km2). In some cases, two independent approaches have been used to estimate the abundance of the same species as a proxy for uncertainty. We provide a qualitative measure of the method quality based on data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.} \\ 
  \hline
{\textbf{Species}} & {\textbf{Method}} & {\textbf{Method quality}} & {\textbf{Justification}} & {\textbf{Lowest abundance}} & {\textbf{Highest abundance}} & {\textbf{Mean abundance}} & {\textbf{sd}} & {\textbf{n}} \\ 
  \hline
Pacific loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   6 & 4 & 3 & 4 (2017-2019, 2022) \\ 
   \hline
Red-throated loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  42 &  76 & 64 & 15 & 4 (2017-2019, 2022) \\ 
   \hline
King eider & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & very low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 25 &  &  \\ 
   \hline
King eider & Derived from the abundance estimate of red-throated loon using incidental observations & low & Derived from high quality estimate of another species &  &  & 106 &  &  \\ 
   \hline
Long-tailed duck & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & very low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 20 &  &  \\ 
   \hline
Long-tailed duck & Derived from the abundance estimate of red-throated loon using incidental observations & low & Derived from high quality estimate of another species &  &  & 191 &  &  \\ 
   \hline
Cackling goose & Extrapolation from exponential model of growth (R2=0.74, p=0.15, n=5) & moderate & Strong correlation with opportunistic nest monitoring &   2 & 158 & 31 & 41 & 23 (1996-2016, 2020-2021) \\ 
   \hline
Cackling goose & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  80 & 214 & 138 & 50 & 5 (2017-2019, 2022-2023) \\ 
   \hline
Snow goose & Nest monitoring plots extrapolated to mean goose colony area & moderate & Relatively small sample size and uncertainty on goose colony area  & 2505 & 35404 & 18129 & 11037 & 11 (1999-2009) \\ 
   \hline
Snow goose & Intensive study area-wide monitoring based on a combination of methods (transects, point counts and nest monitoring plots) and annual colony outline & high & Multiple independent methods and annual colony outline & 8687 & 49076 & 31852 & 12092 & 12 (2010-2019, 2022-2023) \\ 
   \hline
Tundra swan & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   2 & 1 & 1 & 4 (2017-2019, 2022) \\ 
   \hline
Rough-legged hawk & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=8) & high & Strong correlation with study area-wide nest density &   3 &  59 & 22 & 21 & 6 (2007-2012) \\ 
   \hline
Rough-legged hawk & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &  66 & 30 & 29 & 8 (2013-2019, 2022) \\ 
   \hline
Peregrine falcon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   8 &  12 & 10 & 1 & 8 (2013-2019, 2022) \\ 
   \hline
Snowy owl & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &   0 &  67 & 15 & 24 & 16 (1996-2011) \\ 
   \hline
Snowy owl & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 & 144 & 17 & 45 & 10 (2012-2019, 2022-2023) \\ 
   \hline
Rock ptarmigan & Intensive, but opportunistic nest monitoring (8 km2) extrapolated to study area & very low & Intensive, but opportunistic monitoring at relatively small spatial scale and prime nesting habitat not well sampled &  &  & 24 &  &  \\ 
   \hline
Sandhill crane & Extrapolation from intensive nest monitoring (8 km2) and transect observations (R2= 0.44 p= 0.32, n=8) & moderate & Uncertain relation with large scale indices &  &  & 34 &  &  \\ 
   \hline
American golden-plover & Distance sampling throughout lowland (313 km2) & high & Large sample size & 397 & 1725 & 1102 & 432 & 8 (2014-2019, 2022-2023) \\ 
   \hline
Black-bellied plover & Derived from the abundance estimate of American golden-plover using transects observations & low & Derived from high quality estimate of another species &  &  & 29 &  &  \\ 
   \hline
Black-bellied plover & Derived from the abundance estimate of American golden-plover using incidental observations & very low & Derived from high quality estimate of another species, but potentially includes transient migratory individuals &  &  & 87 &  &  \\ 
   \hline
Common Ringed plover & Nest monitoring on the main breeding sites & moderate & Intensive monitoring, but not exhaustive to study area &  44 &  62 & 55 & 9 & 3 (2015-2017) \\ 
   \hline
Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40 &  &  \\ 
   \hline
Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 53 &  &  \\ 
   \hline
Red knot & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 66 &  &  \\ 
   \hline
Red knot & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 233 &  &  \\ 
   \hline
Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 80 &  &  \\ 
   \hline
Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 255 &  &  \\ 
   \hline
Baird's sandpiper & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2448 &  &  \\ 
   \hline
White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 991 &  &  \\ 
   \hline
White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 1134 &  &  \\ 
   \hline
Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 6 &  &  \\ 
   \hline
Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 8 &  &  \\ 
   \hline
Red phalarope & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 140 &  &  \\ 
   \hline
Red phalarope & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 270 &  &  \\ 
   \hline
Glaucous gull & Extrapolation from intensive nest monitoring (111 km2, R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density &  59 &  80 & 73 & 6 & 13 (2004-2016) \\ 
   \hline
Glaucous gull & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  60 &  80 & 71 & 9 & 4 (2017-2019, 2022) \\ 
   \hline
Long-tailed jaeger & Intensive nest monitoring (33 km2) extrapolated by habitat & high & Relatively large spatial coverage of sampling &   0 & 900 & 272 & 285 & 17 (2004-2019, 2022) \\ 
   \hline
Parasitic jaeger & Maximum number of individuals banded in a year & low & Based on a single year and potentially not all individuals were captured &  &  & 17 &  &  \\ 
   \hline
Parasitic jaeger & Maximum number of nest found annually during study area-wide opportunistic nest monitoring & very low & Monitoring does not include potential non-breeding individuals &  &  & 8 &  &  \\ 
   \hline
Common raven & Derived from the abundance estimate of glaucous gull using transects observations & very low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 14 &  &  \\ 
   \hline
Common raven & Derived from the abundance estimate of glaucous gull using incidental observations & very low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 31 &  &  \\ 
   \hline
Horned lark & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362 &  &  \\ 
   \hline
Horned lark & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411 &  &  \\ 
   \hline
American pipit & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 53 &  &  \\ 
   \hline
American pipit & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87 &  &  \\ 
   \hline
Lapland longspur & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110 &  &  \\ 
   \hline
Snow bunting & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 18 &  &  \\ 
   \hline
Snow bunting & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276 &  &  \\ 
   \hline
Nearctic brown lemming & Rigorous density estimates at small spatial scale (0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 447630 & 54043 & 93530 & 27 (1995-2019, 2021-2022) \\ 
   \hline
Nearctic collared lemming & Rigorous density estimates at small spatial scale (0.22 km2) extrapolated by habitat & moderate & Intensive sampling, but small spatial coverage and extrapolation by habitat &   0 & 39302 & 8128 & 10334 & 27 (1995-2019, 2021-2022) \\ 
   \hline
Arctic hare & Derived from the abundance estimate of Arctic fox using incidental observations & very low & Derived from moderate quality estimate of another species, prime nesting habitat not well sampled and difference in detectability between species &  &  & 6 &  &  \\ 
   \hline
Ermine & Indices of relative abundance derived from testimonials converted to abundance using home range size & moderate & Indirect indices and uncertainty on ermine home range size estimates &   8 & 156 & 40 & 37 & 27 (1993-2019) \\ 
   \hline
Arctic fox & Derived from extensive fox home range size studies (n=109) & moderate & Indirect indices, but large sample size &  &  & 53 &  &  \\ 
   \hline
\hline
\label{table:table_summary_methods_and_abundance}
\end{longtable}
\endgroup

### 2024-10-01 MetadataS1/tables/table_variables_definition.tex
update table variable definition and complete first draft of the entire document

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 10:40:29 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.08\textwidth}|p{0.21\textwidth}|p{0.50\textwidth}|p{0.08\textwidth}|}
\caption{Summary of variable definition and unit of measurement.} \\ 
  \hline
{\textbf{Data file}} & {\textbf{Variable identity}} & {\textbf{Variable definition}} & {\textbf{Units}} \\ 
  \hline
a. & class & Taxonomic class  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
   \hline
a. & order & Taxonomic order  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
   \hline
a. & family & Taxonomic family  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
   \hline
a. & genus & Taxonomic genus  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
   \hline
a. & species\_scientific & Taxonomic species  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). We used Mustela richardsonii instead of Mustela erminea richardsonii due to recent genetic analysis (Colella et al., 2021). & NA \\ 
   \hline
a. & species\_en & Common names of species in English. & NA \\ 
   \hline
a. & species\_fr & Common names of species in French. & NA \\ 
   \hline
a. & functional\_group & Functional group for each species. The classification of species into functional groups is based on Moisan et al. (2023). & NA \\ 
   \hline
a. & migratory\_status & Migratory status of each species (i.e., resident, partially migratory or migratory). The classification of species migratory status is based on Gauthier et al., (2011)  and Moisan et al. (2023). & NA \\ 
   \hline
b. & species\_en & Common names of species in English. & NA \\ 
   \hline
b. & year & Year corresponding to the estimate of annual abundance. If abundance has not been calculated for a given series of years, but rather as a general average, then NA has been assigned. & years \\ 
   \hline
b. & breeding\_status & Reproductive status of the individuals. & NA \\ 
   \hline
b. & abundance & Estimate of the annual number of individuals found within the 389 km2 study area located on the southern part of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. The estimates only consider adults, with the exception of lemmings, for which no distinction has been made between juveniles and adults. & individuals \\ 
   \hline
b. & method\_description & Brief overview of the method used to estimate the species abundance. & NA \\ 
   \hline
b. & method\_quality &  Qualitative measure of the method quality based on data available, method used for extrapolation (if necessary),
and in some cases, from the fit of statistical models to estimate density. & NA \\ 
   \hline
c. & species\_en & Common names of species in English. & NA \\ 
   \hline
c. & site & Site where individual body mass measurements were taken. & NA \\ 
   \hline
c. & mean\_body\_mass\_g & Mean individual body mass. & grams \\ 
   \hline
c. & sample\_size & Number of individuals measured. & NA \\ 
   \hline
c. & reference & Reference from which estimate of mean body mass were derived. & NA \\ 
   \hline
d. & species\_en & Common names of species in English. & NA \\ 
   \hline
d. & zone & Sampled zone of the study area (see figure 2 and 3). & NA \\ 
   \hline
d. & mean\_nest\_density\_km2 & Estimate of the mean annual nest density measured within the corresponding zone
of the study area. & nests per square kilometer \\ 
   \hline
d. & sd\_nest\_density\_km2 & Standard deviation of the annual nest density measured within the corresponding
zone of the study area. & nests per square kilometer \\ 
   \hline
d. & number\_years & Number of years consider in the calculation of the nest density. & years \\ 
   \hline
\hline
\label{table:table_variables_definition}
\end{longtable}
\endgroup

### 2024-10-01 MetadataS1/tables/table_variables_format.tex
update table variable definition and complete first draft of the entire document

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Oct  1 10:40:35 2024
\begingroup\fontsize{8pt}{10pt}\selectfont
\begin{longtable}{|p{0.04\textwidth}|p{0.21\textwidth}|p{0.07\textwidth}|p{0.40\textwidth}|p{0.05\textwidth}|p{0.07\textwidth}|}
\caption{Summary of variable storage type, code definition, range and number of digit.} \\ 
  \hline
{\textbf{Data file}} & {\textbf{Variable identity}} & {\textbf{Storage type}} & {\textbf{Definition variable codes}} & {\textbf{Range}} & {\textbf{Number digits}} \\ 
  \hline
a. & class & string & NA & NA & NA \\ 
   \hline
a. & order & string & NA & NA & NA \\ 
   \hline
a. & family & string & NA & NA & NA \\ 
   \hline
a. & genus & string & NA & NA & NA \\ 
   \hline
a. & species\_scientific & string & NA & NA & NA \\ 
   \hline
a. & species\_en & string & NA & NA & NA \\ 
   \hline
a. & species\_fr & string & NA & NA & NA \\ 
   \hline
a. & functional\_group & string & NA & NA & NA \\ 
   \hline
a. & migratory\_status & string & resident: Individuals performing movements within the study area throughout the annual cycle.; partial migrant: A combination of resident and migratory and/or individuals performing long-distance foraging trips outside the study area during the non-breeding period.; migrant: Individuals performing seasonal and highly synchronous movements between the study area and a distant non-breeding ground. & NA & NA \\ 
   \hline
b. & species\_en & string & NA & NA & NA \\ 
   \hline
b. & year & integer & If abundance has not been calculated for a given series of years, but rather as a general average, then NA has been assigned. & 1993-2023 & 0 \\ 
   \hline
b. & breeding\_status & string & undetermined: Individuals present on the study area during the breeding period (June to August) that might have breed or not.; breeding: Individuals present in the study area during the breeding period (June to August) and having attempted and/or completed breeding. & NA & NA \\ 
   \hline
b. & abundance & integer & NA & 0-447630 & 0 \\ 
   \hline
b. & method\_description & string & NA & NA & NA \\ 
   \hline
b. & method\_quality & string & very low: Sampling might not encompasses prime nesting habitat, excludes transient migratory individuals or includes potential non-breeding individuals. If abundance is derived from the abundance estimate of another species based relative abundance, detection probabilities may differ.; low: Abundance is derived from the estimate of another species based on indices of relative abundance.; moderate: Small to intermediate scale sampling with spatial extrapolation.; high: Large scale intensive sampling, with some spatial extrapolation in a few cases. & NA & NA \\ 
   \hline
c. & species\_en & string & NA & NA & NA \\ 
   \hline
c. & site & string & bylot: Southern plain of Bylot Island, Nunavut, Canada.; undetermined: Data were not retrieved from original publications. & NA & NA \\ 
   \hline
c. & mean\_body\_mass\_g & integer & NA & 21 - 6378 & 0 \\ 
   \hline
c. & sample\_size & integer & NA & 1 - 6405 & 0 \\ 
   \hline
c. & reference & string & NA & NA & NA \\ 
   \hline
d. & species\_en & string & NA & NA & NA \\ 
   \hline
d. & zone & string & qarlikturvik (2x1 km plot): Intensive search plot (2 km2) for Lapland Longspur and Baird's sandpiper nests on the south side of the glacial river in the Qarlikturvik valley.;
qarlikturvik (4x2 km plot): Intensive search plot (2 km2) for Sandhill crane, Long-tailed duck, King eider and Rock ptarmigan nests on the south side of the glacial river in the Qarlikturvik valley.;
qarlikturvik valley:Intensive search area (33 km2) for long-tailed jaeger nests on the south side of the glacial river in the Qarlikturvik valley.; whole study area: Entire study area (389 km2) located on the southern plain of Bylot Island. & NA & NA \\ 
   \hline
d. & mean\_nest\_density\_km2 & numeric & NA & 0.001-13.559 & 3 \\ 
   \hline
d. & sd\_nest\_density\_km2 & numeric & NA & 0.001-5.849 & 3 \\ 
   \hline
d. & number\_years & integer & NA & 3-17 & 0 \\ 
   \hline
\hline
\label{table:table_variables_format}
\end{longtable}
\endgroup

### 2024-10-01 script/1_data_cleaning/1.1_clean_year_monitoring_metadata.R

@@ -1,0 +1,92 @@
+#Title: Transform the table of year sampled for nest monitoring of each species into a usable data frame in R
+#Author: Louis Moisan
+#Date: 29 september 2023
+#Last update: June 11 2024
+
+
+#------------------#
+#### Librairies ####
+#------------------#
+#general data manipulation
+library(dplyr)
+library(tidyr)
+#spatial data manipulation
+library(sf)
+
+#--- Extract list of the zone of the study area
+zone_list <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  dplyr::filter(zone != "goose_colony") %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::select(zone) %>% 
+  unlist() %>% 
+  as.vector()
+
+#--- Import raw table of species year monitoring
+sp_monitoring <- read.csv("data/metadata/systematic_nest_monitoring_raw.csv")
+
+#extract a list with year of sampling for each species
+year_sampling <- stringr::str_split(sp_monitoring$year, ",")
+#Divide into multiple parts if species have been sampled non-continuously
+year_sampling <- lapply(year_sampling, function(x) stringr::str_split(x, "-"))
+#Assign species name to each list
+names(year_sampling) <- sp_monitoring$species
+#transform the year format from character to numeric
+for (sp in names(year_sampling)){
+  print(sp)
+  year_sampling[[sp]] <-   lapply(year_sampling[[sp]], function(x) x %>% unlist() %>% as.numeric())
+}
+
+#create an empty data frame to store the output
+year_sampling_df <- data.frame()
+#For each list with more than one year extend the interval of years by integer
+for (list in 1:length(year_sampling)){
+  for (sublist in 1:length(year_sampling[[list]])){
+    if(length(year_sampling[[list]][[sublist]]) > 1){
+      year_sampling[[list]][[sublist]] <- year_sampling[[list]][[sublist]][[1]]: year_sampling[[list]][[sublist]][[2]]}
+    #if list contain only a single value just transform to numeric
+    if(length(year_sampling[[list]][[sublist]]) == 1){
+      year_sampling[[list]][[sublist]] <- as.numeric(year_sampling[[list]][[sublist]])}
+    #transform into a data frame
+    years <- unlist(year_sampling[[list]][[sublist]])
+    year_sampling_df <- year_sampling_df %>% 
[…]

### 2024-10-01 script/1_data_cleaning/1.2_extract_species_body_mass.R

@@ -1,0 +1,108 @@
+#Title: Extract mean species body mass measured directly on Bylot Island or from the literature
+#Author: Louis Moisan
+#Date: NA
+#Last update: June 11 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+
+#--- Import the list of Bylot Island species
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
+  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
+
+
+#--------------------------------#
+#### Measures from literature ####
+#--------------------------------#
+# Wilman, H., Belmaker, J., Simpson, J., de la Rosa, C., Rivadeneira, M. M., & Jetz, W. (2014). EltonTraits 1.0: Species‐level foraging attributes of the world's birds and mammals: Ecological Archives E095‐178. Ecology, 95(7), 2027-2027.
+
+#--- BIRDS
+# Set the download link
+birds_bm_link <- "https://www.esapubs.org/archive/ecol/E095/178/BirdFuncDat.txt"
+#Download the raw data
+download.file(birds_bm_link, "data/raw/body_mass/birds_body_mass.txt")
+#Extract data frame
+birds_bm <-  read.table(file = "data/raw/body_mass/birds_body_mass.txt", header = TRUE, dec = ".", sep="\t", fill = TRUE, quote = "") %>% 
+  #select only column species and body mass
+  dplyr::select(Scientific, BodyMass.Value, BodyMass.Source) %>% 
+  dplyr::rename(species_scientific= Scientific, mean_body_mass_g= BodyMass.Value, reference= BodyMass.Source)
+
+#filter to keep species from Bylot
+birds_bylot <- birds_bm %>% 
+  dplyr::filter(species_scientific %in% sp_taxonomy$species_scientific)
+
+
+#--- MAMMALS
+# Set the download link
+mammals_bm_link <- "https://www.esapubs.org/archive/ecol/E095/178/MamFuncDat.txt"
+#Download the raw data
+download.file(mammals_bm_link, "data/raw/body_mass/mammals_body_mass.txt")
+#Extract data frame
+mammals_bm <-  read.table(file = "data/raw/body_mass/mammals_body_mass.txt", header = TRUE, dec = ".", sep="\t", fill = TRUE, quote = "") %>% 
+  #select only column species and body mass
[…]

### 2024-10-01 script/2_extract_species_abundance/2.1_extract_species_abundance.R

@@ -1,0 +1,766 @@
+#Title: Extract the abundance of every vertebrate species from Bylot Island
+#Date: October 11 2023
+#Last update:  June 17 2024
+
+
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE) #prevent geoprocessing errors
+#Personnal functions to estimate species abundance
+source("script/functions/FUNS.R")
+
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp")
+
+#years each species was monitored
+sampling_year <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv")
+
+#nests
+nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
+nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
+
+#transects
+transect_data <- sf::st_read("data/raw/shapefiles/sampling/vertebrate_count_transects_2010_2023.shp") %>% 
+  #remove groups of shorebirds (>4 individuals) on transects
+  dplyr::filter(!c(nb_ind> 4 & species %in% c("white rumped sandpiper", "bairds sandpiper", "pectoral sandpiper", "buff breasted sandpiper", "red knot", "red phalarope", "ruddy turnstone","black bellied plover", "american golden plover") & status == "inconnu"))
+
+#incidental observation
+incidental_obs <- read.csv("data/raw/sampling/incidental_obs_2007_2019.csv") %>%
+  dplyr::select(-X)
+
+#Index of relative abundance in upland zones
+relative_abundance_upland <- read.csv("data/raw/other/incidental_obs_relative_abundance_upland.csv")
+
+#Index of relative abundance from Gauthier et al., 2024
+relative_abundance_gauthier <- read.csv("data/metadata/relative_abundance_Gauthier_et_al_2023.csv")
+
+#define zones to estimate abundance
+zones <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour", "study area")
+
+#Import the list of Bylot Island species
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
[…]

### 2024-10-01 script/2_extract_species_abundance/species/arctic_fox.R

@@ -1,0 +1,131 @@
+#Title: Estimate density of arctic foxes on the southern plain of Bylot Island from foxes home range size presented in Dulude-de-Broin 2023
+#Original author: Frédéric Dulude-de-Broin
+#Original script: https://datadryad.org/stash/dataset/doi:10.5061/dryad.f1vhhmh30
+#Adapted by: Louis Moisan
+
+
+#------------------#
+#### Librairies ####
+#------------------#
+#data manipulation
+library(dplyr)
+#Generalized linear model
+library(lme4)
+library(AICcmodavg)
+#spatial data manipulation
+library(sf)
+
+#---------------------------------------------------------------#
+#### Code from Dulude et al., 2023 to predict fox home range ####
+#---------------------------------------------------------------#
+# import homerange and covariates
+    dataHR_all <- read.csv(file = "data/raw/sampling/arctic_fox_home_range.csv")
+    
+    # scale all homerange based on the relationships between simulated Argos and GPS data
+    dataHR_all$area <-  0.91570 + 0.72824*dataHR_all$area_argos
+  
+    # compute log home range
+    dataHR_all$l_area.gps <- log(dataHR_all$area)
+    
+    # format column
+    dataHR <- dataHR_all %>% 
+                mutate(sex = factor(sex, levels = c("F", "M")),
+                       rep_status = factor(rep_status, levels= c("Non-breeder", "Breeder")),
+                       goose_category = factor(goose_category, levels = c("outside colony", "outside with overlap", "inside colony")),
+                       centro_in_colony = factor(centro_in_colony, levels = c(0, 1)),
+                      ) %>% 
+                na.omit()
+  
+  
+  #.... fit model to test P1: Effect of primary prey availability on HR size ####
+    mod1 <- lmer(l_area.gps ~ goose_category + sex + rep_status + (1|id_merged_pair) + (1|year), weights = 1/VAR.log, data=dataHR)
+
+  #.... predict values ####
+    # dumy dataset
+    new.data <- unique(dataHR[, c("goose_category", "sex", "rep_status")])
+    
+    # predictions
+      pred <- AICcmodavg::predictSE(mod1, newdata=new.data, se.fit=T, re.form=NA) 
+    
+    # on response scale
+    new.data$pred.fit <- exp(pred$fit)
+    new.data$pred.se <- pred$se.fit
+    new.data$upperCI <- exp(pred$fit+2*pred$se.fit)
+    new.data$lowerCI <- exp(pred$fit-2*pred$se.fit)
[…]

### 2024-10-01 script/2_extract_species_abundance/species/lemmings.R

@@ -1,0 +1,104 @@
+#title: Create dataframe with the time-series for collared and brown lemmings (1995-2022)
+#author: Azenor Bideault
+#date: november 8 2023
+
+#Libraries
+library(dplyr)
+library(sf)
+
+
+#-------------------#
+#### IMPORT DATA ####
+#-------------------#
+#-- For densities from 1995 to 2003
+df_lemming = read.csv('data/raw/sampling/lemming_density_1993-2019.csv', header = TRUE)
+#-- For densities from 2004 to 2022
+df_lemming2 = read.csv('data/raw/sampling/lemming_density_2004-2022.csv', header = TRUE)
+
+
+study_area_df <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::filter(zone=="study area")
+
+
+#-----------------------------------------#
+#### DENSITIES FOR THE YEARS 2004-2022 ####
+#-----------------------------------------#
+#--- Time-series is computed for the valley of camp 1 only
+# The period P2 (july) and P3 (august) are used
+# Filter the data
+filtered_data <- df_lemming2 %>% 
+  filter(site == 'Goose camp', period %in% c("P2", "P3"))
+
+#--- Compute the densities according to the good period to consider for each year
+# 2004 to 2007 -> period 3 is used
+# 2008 to 2018 and 2020 to 2022 -> mean between periods P2 and P3 is used
+# 2019 -> period P2 only is used
+# see document Document_interne_trappage_mortel_version (tables 5 and 6 for details)
+
+# Filter the data for the specified periods and years
+dens_lemming <- df_lemming2 %>% 
+  filter(
+    (year >= 2004 & year <= 2007 & period == "P3") |
+      (year >= 2008 & year <= 2018 & period %in% c("P2", "P3")) |
+      (year == 2019 & period == "P2") |
+      (year >= 2020 & year <= 2022 & period %in% c("P2", "P3"))
+  ) %>% 
+  group_by(species, habitat, grid, year) %>% 
+  summarize(mean_density = mean(density, na.rm = TRUE))
+
+#-- Compute densities for the whole study area according to the percentages of wetland and mesic habitats :
+dens_lemming_area <- dens_lemming %>% 
+  filter((grid == 'LG2' | grid == 'LG1'))  %>% 
+  group_by(species, year) %>% 
+  summarize(
[…]

### 2024-10-01 script/2_extract_species_abundance/species/snow_goose.R

@@ -1,0 +1,276 @@
+#Title: Estimate snow goose abundance at the scale of the study area
+#Author: Louis Moisan
+#Date: NA
+#Last update: 31 May 2024
+
+#------------------#
+#### Librairies ####
+#------------------#
+#General data manipulation
+library(dplyr) 
+library(tidyr)
+#Spatial data manipulation
+library(sf)
+sf::sf_use_s2(FALSE)
+#Visualization
+library(ggplot2)
+
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#Study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") 
+
+#Goose colony outline 2010-2023
+goose_colony_2010_2023 <- study_area %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::filter(zone == "goose_colony") %>% 
+  dplyr::select(-prop_up)
+#Goose colony 1999-2009 (2020-2021)
+goose_colony_1999_2009 <- goose_colony_2010_2023 %>% 
+  dplyr::summarise(area_km2= mean(area_km2), prop_wet= mean(prop_wet), prop_mesic= mean(prop_mesic),zone= "goose_colony") 
+# Replicate the row for each year and add a "year" column
+goose_colony_1999_2009 <- goose_colony_1999_2009[rep(1, length(c(1999:2009, 2020, 2021))), ] %>% 
+dplyr::mutate(year = c(1999:2009, 2020, 2021))
+#combine to extract a single data frame of the goose colony
+goose_colony <- goose_colony_2010_2023 %>% 
+  rbind(goose_colony_1999_2009)
+
+#Systematically monitored wetland plot (peksek)
+peksek <- sf::st_read("data/raw/shapefiles/sampling/snow_goose_systematic_wetland_plot.shp") %>%
+  sf::st_transform(26717) %>% 
+  dplyr::mutate(area_km2= round(as.numeric(sf::st_area(geometry)/ 1000000), digits = 4)) %>% 
+  dplyr::select(area_km2, geometry)
+#Goose nests within the peksek systematic plot
+peksek_nest <- sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp") %>%
+  dplyr::filter(species == "snow goose") %>% 
+  sf::st_intersection(peksek)
+
+#Random plots
+#Number of nest found in random plot
[…]

### 2024-10-01 script/3_tables_figures/3.10_table_expert_based_uncertainty.R

@@ -1,0 +1,42 @@
+#Title: Create a table of the expert based uncertainty described by guess estimate
+#Author: Louis Moisan
+#Date: September 30 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+
+#--- Import the list of Bylot Island species
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
+  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
+
+#--- Import estimates made by experts
+table <- read.csv("data/metadata/expert_based_uncertainty.csv") %>% 
+  dplyr::left_join(sp_taxonomy) %>% 
+  dplyr::rename(Species= species_en, Period= period, `Lowest abundance`= lowest_abundance, `Highest abundance`= highest_abundance, `Mean abundance`= mean_abundance) %>% 
+  dplyr::select(Species, Period, `Lowest abundance`,`Highest abundance`,`Mean abundance`) %>% 
+  dplyr::mutate(Period= ifelse(is.na(Period), "Mean abundance",Period))
+
+#Add above header row
+addtorow <- list()
+addtorow$pos <- list(0, 0)
+addtorow$command <- c("& &\\multicolumn{3}{c}{Annual abundance (ind.)} \\\\\n",
+                      "Species & Period & Lowest & Highest & Mean \\\\\n")
+#as tex
+table_latex <- xtable(table, caption = "Considering the absence of confidence intervals in our abundance estimates, we present below uncertainty intervals on estimated abundance values derived from field expert impressions. The intervals presented represent the minimum and maximum values between which the experts believe the actual abundance values should lie. When annual abundance has been estimated for several years (time series) we present intervals over the minimum and maximum abundance values encountered during the given time series. For other species, the uncertainty interval is estimated on the mean abundance.")
+print(table_latex,
+      include.rownames = FALSE,
+      tabular.environment = "tabularx",
+      width="\\textwidth",
+      align="rXXXXXX",
+      size="\\fontsize{10pt}{8pt}\\selectfont",
+      file = "MetadataS1/tables/table_expert_based_uncertainty.tex",
+      caption.placement = "top",
+      add.to.row = addtorow,
+      include.colnames = FALSE)
+

### 2024-10-01 script/3_tables_figures/3.2_table_species_year_monitoring.R

@@ -1,0 +1,60 @@
+#data manipulation
+library(dplyr)
+library(tidyr)
+#spatial data manipulation
+library(sf)
+#export latex table
+library(xtable)
+
+#--- Import study area 
+#list of zones for which we want to obtain a density
+zone_list <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour", "study area")
+#shapefile of the area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  sf::st_drop_geometry() %>% 
+  #keep only non-nested zones
+  dplyr::filter(zone %in% zone_list)
+
+#--- Filter nest data to keep only nest found during years of systematic survey
+sp_monitoring <- read.csv("data/metadata/systematic_nest_monitoring_raw.csv")
+#reorder names based on defined species list
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
+  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
+# Reorder df1 based on the species meta data order
+sp_monitoring <- sp_taxonomy %>% 
+  dplyr::left_join(sp_monitoring) %>% 
+  na.omit()
+
+#--- Extract standardize and extract a tex table
+table_year_monitoring <- sp_monitoring %>% 
+  dplyr::mutate(zone= dplyr::recode(zone, "camp 1"= "qarlikturvik valley","trapping grids"= "qarlikturvik (trapping grids)", "2x1km plot"= "qarlikturvik (2x1 km plot)", "4x2km plot"= "qarlikturvik (4x2 km plot)","black plateau"= "black plateau", "south plateau"= "south plateau", "study area"= "whole study area")) %>% 
+  dplyr::mutate(number_years= paste("(", number_years, ")", sep="")) %>% 
+  dplyr::rename(Species= species_en, Zone= zone, Years= year, `Number of years`= number_years, Monitoring= monitoring) %>% 
+  dplyr::select(Species, Zone, Years, `Number of years`, Monitoring) %>% 
+  dplyr::arrange(Zone) 
+
+# Convert the zone variable to a factor with the custom order
+table_year_monitoring$Zone <- as.character(factor(table_year_monitoring$Zone, levels = c("qarlikturvik (trapping grids)","qarlikturvik (2x1 km plot)","qarlikturvik (4x2 km plot)","qarlikturvik valley","camp 2","whole study area")))
+
[…]

### 2024-10-01 script/3_tables_figures/3.3_table_species_relative_abundance_upland.R

@@ -1,0 +1,78 @@
+#librairies
+library(dplyr)
+#export latex table
+library(xtable)
+
+#import as .csv
+relat_abund_compared <- read.csv("data/raw/other/incidental_obs_relative_abundance_upland.csv")
+
+#----------------------#
+#### Export as .tex ####
+#----------------------#
+table <- relat_abund_compared %>% 
+  dplyr::mutate(nb_ind_hour_uplands= as.character(round(nb_ind_hour_upland, digits = 3)), 
+                nb_ind_hour_lowland= as.character(round(nb_ind_hour_lowland, digits = 3)),
+                ratio_difference= as.character(ratio_difference)) %>% 
+  dplyr::rename(Uplands=nb_ind_hour_upland,
+                Lowlands=nb_ind_hour_lowland,
+                Ratio= ratio_difference
+  )
+
+
+#reorder names based on defined species list
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
+  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
+# Find the indices of species in df2 in df1
+indices <- match(sp_taxonomy$species, table$species)
+# Reorder df1 based on the indices
+table <- table[indices, , drop = FALSE] 
+
+#update species names
+table <- table %>% 
+  dplyr::left_join(sp_taxonomy) %>% 
+  dplyr::select(species_en, Uplands, Lowlands, Ratio)
+
+
+#Filter some species
+table <- table %>% 
+  dplyr::filter(species_en %in% c("American golden-plover",
+                               "American pipit",
+                               "Arctic hare",
+                               "Baird's sandpiper",
+                               "Black-bellied plover",
+                               "Buff-breasted sandpiper",
+                               "Horned lark",
+                               "Lapland longspur",
+                               "Pectoral sandpiper",
+                               "Red knot",
+                               "Red phalarope",
+                               "Rock ptarmigan",
+                               "Ruddy turnstone",
+                               "Sandhill crane",
+                               "Snow bunting",
+                               "White-rumped sandpiper")) %>% 
+  dplyr::rename(Species = species_en)
+
+custom_row <- c("\\hline", "Custom Header", "\\hline")
+
+#as tex
[…]

### 2024-10-01 script/3_tables_figures/3.4_table_interannual_variation_nest_density.R

@@ -1,0 +1,113 @@
+# This script extract mean and standard deviation of annual nest density monitored in systematic plot
+#Date: April 16 2024
+
+#------------------#
+#### Librairies ####
+#------------------#
+#data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE)
+source("script/functions/FUNS.R")
+library(xtable)
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#--- import study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  dplyr::filter(zone != "goose_colony")
+
+#--- import data frame that specify each species that have been monitored in each zone for which years
+sampling_plot <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv") %>% 
+  dplyr::filter(sampled== "yes", species %in% c("lapland longspur", "bairds sandpiper", "sandhill crane", "long tailed duck", "king eider", "long tailed jaeger", "rock ptarmigan")) 
+
+sampling_study_area <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv") %>%
+  dplyr::filter(sampled== "yes", species %in% c("cackling goose", "glaucous gull", "red throated loon", "pacific loon", "snowy owl", "tundra swan", "common raven", "common ringed plover", "parasitic jaeger"), zone=="study area") 
+
+#--- nest data
+nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
+nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
+
+#--- Import the list of Bylot Island species
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
+  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
+
+
+
+#------------------------------#
+#### Calculate nest density ####
+#------------------------------#
+#Extract nest density for each year (0 are assigned for year not sampled -> those years will be exclude below)
+nest_density <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = c(sampling_study_area$species, sampling_plot$species),
+  zone_list= study_area$zone,
+  year_list= c(1990:2023)) %>% 
[…]

### 2024-10-01 script/3_tables_figures/3.5_table_summary_methods_and_abundance.R

@@ -1,0 +1,90 @@
+#Title: Create a summary table of all data and method use for each species
+#Author: Louis Moisan
+#Date: June 3 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#Metadata
+metadata <- read.csv("data/metadata/summary_methods.csv")
+
+#Import the list of Bylot Island species
+sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
+  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
+
+#Abundance data
+abundance <- read.csv("DataS1/BYLOT-species_abundance.csv") %>% 
+  dplyr::mutate(year= ifelse(is.na(year), "mean", year),
+                species= gsub("'","", gsub("-"," ",tolower(species_en)))) %>%
+  dplyr::group_by(species, method_description) %>% 
+  dplyr::summarise(min= min(abundance), max= max(abundance), mean= as.character(round(mean(abundance), digits=0)), sd= as.character(round(sd(abundance), digits= 0)), n= n()) %>% 
+  dplyr::ungroup() %>% 
+  dplyr::mutate(n= ifelse(n==1, NA, n), min= ifelse(min==max, NA, min), max= ifelse(max==min, NA, max))
+
+
+
+#--------------------------#
+#### Generate the table ####
+#--------------------------#
+#Join data frames together
+table <- metadata %>% 
+  dplyr::left_join(abundance, by= c("species", "method_description")) %>% 
+  dplyr::left_join(sp_taxonomy)
+
+#--- reoreder row based on species taxonomy
+#Extract beginning year of each time series
+table <- table %>% 
+  dplyr::mutate(year= as.numeric(substr(table$period, 1, 4)),
+                species= factor(species, levels= sp_taxonomy$species)) %>% 
+  dplyr::arrange(species, year)
+
+#format table
+table <- table %>% 
+  dplyr::select(species_en, period, method_description, method_quality, justification, min, max, mean, sd, n) %>% 
+  #rename columns
+  dplyr::rename(Species= species_en,  Method= method_description,  `Method quality`= method_quality, Justification=justification, `Lowest abundance`= min, `Highest abundance`= max, `Mean abundance`= mean) %>% 
[…]

### 2024-10-01 script/3_tables_figures/3.7_maps.R

@@ -1,0 +1,168 @@
+#title: Create a map of the study area with the different zones
+#author: Louis Moisan
+#date: October 4 2023
+
+#------------------#
+#### librairies ####
+#------------------#
+library(dplyr)
+library(sf)
+library(ggmap)
+library(ggplot2)
+
+# To extract background satellite map from google, one need to register to have a API key
+#!!! Need a personal API key from google
+# https://developers.google.com/maps
+#You have to create a API key and then under "Key and identifiers" you should see your key with the possibility to click on "show key", copy paste the key in the function below
+ggmap::register_google(key = "", write = TRUE)
+
+# Define the bounding box (latitude and longitude) for your area of interest
+bbox <- c(left = -80.19, bottom = 72.75, right = -78.5, top = 73.3)
+
+#Define only zones of interest
+zones <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour")
+  
+#import study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  dplyr::filter(zone %in% zones) %>% 
+# Transform nc to EPSG 3857 (Pseudo-Mercator, what Google uses)
+ st_transform(3857)
+
+
+#layer of the goose colony
+goose_colony <- sf::st_read("data/raw/shapefiles/study_area.shp") %>%
+  dplyr::filter(year== 2017, zone== "goose_colony") %>% 
+  sf::st_transform(3857)
+
+
+#color palette
+colors_palette <- c("#bd959d","#84a98c", "#94214b", "#463a49","#ad382b","#096377","#63141c", "#003049", "#d4af37", "white")
+
+#Get the google map background satellite
+map <- get_map(location = c(lon = -79.7, lat = 73), maptype = "satellite", source = "google", zoom= 8)
+
+# Define a function to fix the bbox to be in EPSG:3857
+ggmap_bbox <- function(map) {
+  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
+  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
[…]

### 2024-10-01 script/3_tables_figures/3.9_table_variables_format.R

@@ -1,0 +1,80 @@
+#Title: Create a table of all variables and their associated type,code definition and range of values
+#Author: Louis Moisan
+#Date: September 25 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+#Import table structure .csv format
+table <- read.csv("data/metadata/variables_format.csv") 
+
+#Import data set
+species_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv")
+species_abundance <- read.csv("DataS1/BYLOT-species_abundance.csv")
+species_body_mass <- read.csv("DataS1/BYLOT-species_body_mass.csv")
+variation_nest_density <- read.csv("DataS1/BYLOT-interannual_variation_nest_density.csv")
+
+#Sum of columns
+sum(species_abundance$abundance, na.rm = T)
+sum(species_body_mass$mean_body_mass_g, na.rm = T)
+sum(species_body_mass$sample_size, na.rm = T)
+sum(variation_nest_density$mean_nest_density_km2)
+sum(variation_nest_density$sd_nest_density_km2)
+sum(variation_nest_density$number_years)
+
+
+
+
+#Assign range values
+table[table$variable.identity == "year",]$range <- paste0(min(as.numeric(species_abundance$year), na.rm = T), "-", max(as.numeric(species_abundance$year), na.rm = T))
+
+table[table$variable.identity == "abundance",]$range <- paste0(min(species_abundance$abundance), "-", max(species_abundance$abundance))
+
+table[table$variable.identity == "mean_body_mass_g",]$range <- paste0(min(species_body_mass$mean_body_mass_g), " - ", max(species_body_mass$mean_body_mass_g))
+
+table[table$variable.identity == "sample_size",]$range <- paste0(min(species_body_mass$sample_size, na.rm = T), " - ", max(species_body_mass$sample_size, na.rm= T))
+
+table[table$variable.identity == "mean_nest_density_km2",]$range <- paste0(min(variation_nest_density$mean_nest_density_km2), "-", max(variation_nest_density$mean_nest_density_km2))
+
+table[table$variable.identity == "sd_nest_density_km2",]$range <- paste0(min(variation_nest_density$sd_nest_density_km2), "-", max(variation_nest_density$sd_nest_density_km2))
+
+table[table$variable.identity == "number_years",]$range <-paste0(min(variation_nest_density$number_years), "-", max(variation_nest_density$number_years))
+
+#number of digits
[…]

### 2024-10-01 script/data_cleaning/02_extract_species_body_mass.R

@@ -1,97 +1,0 @@
-#Title: Extract mean species body mass measured directly on Bylot Island or from the literature
-#Author: Louis Moisan
-#Date: NA
-#Last update: June 11 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-
-
-
-#--- Import the list of Bylot Island species
-bylot_sp <- read.csv("data/metadata/species_metadata.csv")
-
-
-#--------------------------------#
-#### Measures from literature ####
-#--------------------------------#
-# Wilman, H., Belmaker, J., Simpson, J., de la Rosa, C., Rivadeneira, M. M., & Jetz, W. (2014). EltonTraits 1.0: Species‐level foraging attributes of the world's birds and mammals: Ecological Archives E095‐178. Ecology, 95(7), 2027-2027.
-
-#--- BIRDS
-# Set the download link
-birds_bm_link <- "https://www.esapubs.org/archive/ecol/E095/178/BirdFuncDat.txt"
-#Download the raw data
-download.file(birds_bm_link, "data/raw/body_mass/birds_body_mass.txt")
-#Extract data frame
-birds_bm <-  read.table(file = "data/raw/body_mass/birds_body_mass.txt", header = TRUE, dec = ".", sep="\t", fill = TRUE, quote = "") %>% 
-  #select only column species and body mass
-  dplyr::select(Scientific, BodyMass.Value, BodyMass.Source) %>% 
-  dplyr::rename(scientific= Scientific,body_mass_g= BodyMass.Value, reference= BodyMass.Source)
-
-#filter to keep species from Bylot
-birds_bylot <- birds_bm %>% 
-  dplyr::filter(scientific %in% bylot_sp$scientific)
-
-
-#--- MAMMALS
-# Set the download link
-mammals_bm_link <- "https://www.esapubs.org/archive/ecol/E095/178/MamFuncDat.txt"
-#Download the raw data
-download.file(mammals_bm_link, "data/raw/body_mass/mammals_body_mass.txt")
-#Extract data frame
-mammals_bm <-  read.table(file = "data/raw/body_mass/mammals_body_mass.txt", header = TRUE, dec = ".", sep="\t", fill = TRUE, quote = "") %>% 
-  #select only column species and body mass
-  dplyr::select(Scientific, BodyMass.Value, BodyMass.Source) %>% 
[…]

### 2024-10-01 script/data_cleaning/detection function_american_golden_plover.R

@@ -1,68 +1,0 @@
-#Title: Define detection function parameters of american golden plover based on distance sampling
-#Date: November 1 2023
-#Author: Louis Moisan
-
-#------------------#
-#### Librairies ####
-#------------------#
-#data manipulation
-library(dplyr)
-# spatial data manipulation
-library(sf)
-#Distance sampling (calcul detection function and estimate abundance)
-library(Distance)
-#data visualization
-library(ggplot2)
-
-#---import distance between plovers and transect
-distance_table <- read.csv("data/raw/sampling/american_golden_plover_distance.csv") %>% 
-  dplyr::select(-X)
-
-#--- The code below have been run to determine the best detection function model based on AIC
- aic_model <- data.frame()
- model_plot <- list()
-
-#list of key term
- key_list= c("hn","hr", "unif")
-#list of ajdustement term
- adjust_list= c(NULL, "cos", "herm", "poly")
-
- for (k in key_list) {
-  for (adjust in adjust_list) {
-   print(paste(k, adjust))
-    model <- ds(data=distance_table, truncation = 150,key=k, adjustment=adjust) 
-    summary <-   summary(model)
-    aic <- summary$ds$aic
-    if (is.null(adjust)){
-      adjust= NA
-   }
-    aic_model <- rbind(aic_model, data.frame(key= k, adjustment= adjust,aic= aic))
-
-   plot(model, which =2, showpoints= F, xlab= "Distance (m)")
-   model_plot[[paste(k, adjust)]] <- recordPlot()
-  }
-}
-
-#Dectection function based on combination of parameter that lead to the lowest AIC
-aic_model <- aic_model %>% 
- dplyr::mutate(diff_aic= aic- min(aic)) %>% 
-  dplyr::arrange(aic)
-#Look at the match between predicted and observed value that seem to represent well the detection function. Those with the minimum AIC seem to overfit the function.
-model_plot$`hn cos`
-#Retained paramater
-best_key= "hn"
-best_adjustment <- "cos"
-
-#look at the lowest aic based on number of ajdustement term
- ds_function <-  ds(data=distance_table, truncation = 150, key=best_key, adjustment=best_adjustment)
-#best number adjustment term
-best_nadj=1
-
- ds_function <-  ds(data=distance_table, truncation = 150, key=best_key, adjustment=best_adjustment, nadj=best_nadj)
- ds_function$ddf
-#Look at the fit of the model (Good)
-gof_ds(ds_function)
-
[…]

### 2024-10-01 script/extract_dataset_tables_and_figures.R

@@ -1,27 +1,0 @@
-#---------------------#
-#### Data cleaning ####
-#---------------------#
-source("script/data_cleaning/01_clean_year_monitoring_metadata.R")
-source("script/data_cleaning/02_extract_species_body_mass.R")
-#Run if wanted to look at the AIC model selection for the american golden-plover distance sampling method
-#source("script/data_cleaning/extract_detection function_american_golden_plover.R")
-
-
-#------------------------------------------#
-#### Estimate abundance of each species ####
-#------------------------------------------#
-source("script/extract_species_abundance/extract_species_abundance.R")
-
-#---------------------------------#
-#### Export tables and figures ####
-#---------------------------------#
-source("script/tables_figures/table_species_name_strategy.R")
-source("script/tables_figures/table_species_year_monitoring.R")
-source("script/tables_figures/table_interannual_variation_nest_density.R")
-source("script/tables_figures/table_species_relative_abundance_upland.R")
-source("script/tables_figures/table_body_mass.R")
-source("script/tables_figures/table_summary_methods_data.R")
-source("script/tables_figures/table_species_estimate_uncertainty.R")
-source("script/tables_figures/table_mean_abundance.R")
-source("script/tables_figures/figures_species_temporal_series.R")
-source("script/tables_figures/maps.R")

### 2024-10-01 script/extract_species_abundance/extract_species_abundance.R

@@ -1,759 +1,0 @@
-#Title: Extract the abundance of every vertebrate species from Bylot Island
-#Date: October 11 2023
-#Last update:  June 17 2024
-
-
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#data manipulation
-library(dplyr)
-#spatial data manipulation
-library(sf)
-sf_use_s2(FALSE) #prevent geoprocessing errors
-#Personnal functions to estimate species abundance
-source("script/functions/FUNS.R")
-
-
-#-------------------#
-#### Import data ####
-#-------------------#
-#study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp")
-
-#years each species was monitored
-sampling_year <- read.csv("data/metadata/species_systematic_nest_monitoring_clean.csv")
-
-#nests
-nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
-nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
-
-#transects
-transect_data <- sf::st_read("data/raw/shapefiles/sampling/vertebrate_count_transects_2010_2023.shp") %>% 
-  #remove groups of shorebirds (>4 individuals) on transects
-  dplyr::filter(!c(nb_ind> 4 & species %in% c("white rumped sandpiper", "bairds sandpiper", "pectoral sandpiper", "buff breasted sandpiper", "red knot", "red phalarope", "ruddy turnstone","black bellied plover", "american golden plover") & status == "inconnu"))
-
-#incidental observation
-incidental_obs <- read.csv("data/raw/sampling/incidental_obs_2007_2019.csv") %>%
-  dplyr::select(-X)
-
-#Index of relative abundance in upland zones
-relative_abundance_upland <- read.csv("data/raw/other/incidental_obs_relative_abundance_upland.csv")
-
-#Index of relative abundance from Gauthier et al., 2024
-relative_abundance_gauthier <- read.csv("data/metadata/relative_abundance_Gauthier_et_al_2023.csv")
-
-#define zones to estimate abundance
-zones <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour", "study area")
-
-
-
-#--------------------------------------------------#
-#### 3.1 Pacific loon, red-throated loon, cackling goose, tundra swan, glaucous gull ####
-#-------------------------------------------------#
[…]

### 2024-10-01 script/extract_species_abundance/species/arctic_fox.R

@@ -1,131 +1,0 @@
-#Title: Estimate density of arctic foxes on the southern plain of Bylot Island from foxes home range size presented in Dulude-de-Broin 2023
-#Original author: Frédéric Dulude-de-Broin
-#Original script: https://datadryad.org/stash/dataset/doi:10.5061/dryad.f1vhhmh30
-#Adapted by: Louis Moisan
-
-
-#------------------#
-#### Librairies ####
-#------------------#
-#data manipulation
-library(dplyr)
-#Generalized linear model
-library(lme4)
-library(AICcmodavg)
-#spatial data manipulation
-library(sf)
-
-#---------------------------------------------------------------#
-#### Code from Dulude et al., 2023 to predict fox home range ####
-#---------------------------------------------------------------#
-# import homerange and covariates
-    dataHR_all <- read.csv(file = "data/raw/sampling/arctic_fox_home_range.csv")
-    
-    # scale all homerange based on the relationships between simulated Argos and GPS data
-    dataHR_all$area <-  0.91570 + 0.72824*dataHR_all$area_argos
-  
-    # compute log home range
-    dataHR_all$l_area.gps <- log(dataHR_all$area)
-    
-    # format column
-    dataHR <- dataHR_all %>% 
-                mutate(sex = factor(sex, levels = c("F", "M")),
-                       rep_status = factor(rep_status, levels= c("Non-breeder", "Breeder")),
-                       goose_category = factor(goose_category, levels = c("outside colony", "outside with overlap", "inside colony")),
-                       centro_in_colony = factor(centro_in_colony, levels = c(0, 1)),
-                      ) %>% 
-                na.omit()
-  
-  
-  #.... fit model to test P1: Effect of primary prey availability on HR size ####
-    mod1 <- lmer(l_area.gps ~ goose_category + sex + rep_status + (1|id_merged_pair) + (1|year), weights = 1/VAR.log, data=dataHR)
-
-  #.... predict values ####
-    # dumy dataset
-    new.data <- unique(dataHR[, c("goose_category", "sex", "rep_status")])
-    
-    # predictions
-      pred <- AICcmodavg::predictSE(mod1, newdata=new.data, se.fit=T, re.form=NA) 
-    
-    # on response scale
-    new.data$pred.fit <- exp(pred$fit)
-    new.data$pred.se <- pred$se.fit
-    new.data$upperCI <- exp(pred$fit+2*pred$se.fit)
-    new.data$lowerCI <- exp(pred$fit-2*pred$se.fit)
[…]

### 2024-10-01 script/extract_species_abundance/species/lemmings.R

@@ -1,104 +1,0 @@
-#title: Create dataframe with the time-series for collared and brown lemmings (1995-2022)
-#author: Azenor Bideault
-#date: november 8 2023
-
-#Libraries
-library(dplyr)
-library(sf)
-
-
-#-------------------#
-#### IMPORT DATA ####
-#-------------------#
-#-- For densities from 1995 to 2003
-df_lemming = read.csv('data/raw/sampling/lemming_density_1993-2019.csv', header = TRUE)
-#-- For densities from 2004 to 2022
-df_lemming2 = read.csv('data/raw/sampling/lemming_density_2004-2022.csv', header = TRUE)
-
-
-study_area_df <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  sf::st_drop_geometry() %>% 
-  dplyr::filter(zone=="study area")
-
-
-#-----------------------------------------#
-#### DENSITIES FOR THE YEARS 2004-2022 ####
-#-----------------------------------------#
-#--- Time-series is computed for the valley of camp 1 only
-# The period P2 (july) and P3 (august) are used
-# Filter the data
-filtered_data <- df_lemming2 %>% 
-  filter(site == 'Goose camp', period %in% c("P2", "P3"))
-
-#--- Compute the densities according to the good period to consider for each year
-# 2004 to 2007 -> period 3 is used
-# 2008 to 2018 and 2020 to 2022 -> mean between periods P2 and P3 is used
-# 2019 -> period P2 only is used
-# see document Document_interne_trappage_mortel_version (tables 5 and 6 for details)
-
-# Filter the data for the specified periods and years
-dens_lemming <- df_lemming2 %>% 
-  filter(
-    (year >= 2004 & year <= 2007 & period == "P3") |
-      (year >= 2008 & year <= 2018 & period %in% c("P2", "P3")) |
-      (year == 2019 & period == "P2") |
-      (year >= 2020 & year <= 2022 & period %in% c("P2", "P3"))
-  ) %>% 
-  group_by(species, habitat, grid, year) %>% 
-  summarize(mean_density = mean(density, na.rm = TRUE))
-
-#-- Compute densities for the whole study area according to the percentages of wetland and mesic habitats :
-dens_lemming_area <- dens_lemming %>% 
-  filter((grid == 'LG2' | grid == 'LG1'))  %>% 
-  group_by(species, year) %>% 
-  summarize(
[…]

### 2024-10-01 script/extract_species_abundance/species/snow_goose.R

@@ -1,276 +1,0 @@
-#Title: Estimate snow goose abundance at the scale of the study area
-#Author: Louis Moisan
-#Date: NA
-#Last update: 31 May 2024
-
-#------------------#
-#### Librairies ####
-#------------------#
-#General data manipulation
-library(dplyr) 
-library(tidyr)
-#Spatial data manipulation
-library(sf)
-sf::sf_use_s2(FALSE)
-#Visualization
-library(ggplot2)
-
-
-#-------------------#
-#### Import data ####
-#-------------------#
-#Study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") 
-
-#Goose colony outline 2010-2023
-goose_colony_2010_2023 <- study_area %>% 
-  sf::st_drop_geometry() %>% 
-  dplyr::filter(zone == "goose_colony") %>% 
-  dplyr::select(-prop_up)
-#Goose colony 1999-2009 (2020-2021)
-goose_colony_1999_2009 <- goose_colony_2010_2023 %>% 
-  dplyr::summarise(area_km2= mean(area_km2), prop_wet= mean(prop_wet), prop_mesic= mean(prop_mesic),zone= "goose_colony") 
-# Replicate the row for each year and add a "year" column
-goose_colony_1999_2009 <- goose_colony_1999_2009[rep(1, length(c(1999:2009, 2020, 2021))), ] %>% 
-dplyr::mutate(year = c(1999:2009, 2020, 2021))
-#combine to extract a single data frame of the goose colony
-goose_colony <- goose_colony_2010_2023 %>% 
-  rbind(goose_colony_1999_2009)
-
-#Systematically monitored wetland plot (peksek)
-peksek <- sf::st_read("data/raw/shapefiles/sampling/snow_goose_systematic_wetland_plot.shp") %>%
-  sf::st_transform(26717) %>% 
-  dplyr::mutate(area_km2= round(as.numeric(sf::st_area(geometry)/ 1000000), digits = 4)) %>% 
-  dplyr::select(area_km2, geometry)
-#Goose nests within the peksek systematic plot
-peksek_nest <- sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp") %>%
-  dplyr::filter(species == "snow goose") %>% 
-  sf::st_intersection(peksek)
-
-#Random plots
-#Number of nest found in random plot
[…]

### 2024-10-01 script/functions/FUNS.R

@@ -23,7 +23,7 @@ source("script/functions/nested_functions/spatial_extrapolation_species_density.
 # Extract species density by doing a rule of three with the relative proportion of observation among the selected species in each zone and the known species density of one of the species
 source("script/functions/nested_functions/estimate_density_from_reference_species.R")
 
-#Extract species density in selected zones and years based on the selected parameter of the detection function (see script/data_cleaning/09_extract_american_golden_plover_distance & 10_extract_detection_function_american_golden_plover)
+#Extract species density in selected zones and years based on the selected parameter of the detection function (see script/1_data_cleaning/09_extract_american_golden_plover_distance & 10_extract_detection_function_american_golden_plover)
 source("script/functions/standalone_functions/extract_density_distance_sampling.R")
 
 #---------------------------------------------------------#

### 2024-10-01 script/functions/nested_functions/extend_temporal_series_from_single_zone_missing_coordinates.R

@@ -21,7 +21,7 @@ extend_temporal_series_from_single_zone_nest_missing_coords <- function(
   
 #--- Extract density from nest found in each zone
 sp_density <- nest_data %>% 
-  dplyr::filter(species ==sp) %>% 
+  dplyr::filter(species ==sp, year%in% year_list) %>% 
   group_by(species, zone, year) %>% 
   dplyr::summarise(nb_nest= n()) %>% 
   dplyr::left_join(study_area %>% dplyr::select(zone, area_km2) %>% sf::st_drop_geometry()) %>% 

### 2024-10-01 script/tables_figures/table_body_mass.R

@@ -1,70 +1,0 @@
-#general data manipulation
-library(dplyr)
-library(tidyr)
-#export latex table
-library(xtable)
-
-
-#--- Import as .csv
-df_body_mass <- read.csv("data/clean/species_body_mass.csv") %>% 
-  dplyr::select(-X, -Scientific.name) %>% 
-  dplyr::rename(Species= Species.name, `Body mass g (study area)`= Body.mass.g..study.area., `Body mass g (literature)`= Body.mass.g..literature.)
-
-
-#reorder names based on defined species list
-sp_metadata <- read.csv("data/metadata/species_metadata.csv")
-# Find the indices of species in df2 in df1
-indices <- match(sp_metadata$species, df_body_mass$Species)
-# Reorder df1 based on the indices
-df_body_mass <- df_body_mass[indices, , drop = FALSE] 
-
-#update species names
-df_body_mass <- df_body_mass %>% 
-  dplyr::mutate(Species= dplyr::recode(Species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper", "lapland longspur"= "Lapland longspur"))
-
-
-#Remove decimals
-df_body_mass <- df_body_mass %>% 
-  dplyr::mutate(`Body mass g (study area)`= round(`Body mass g (study area)`, digits = 0),
-                `Body mass g (literature)`= round(`Body mass g (literature)`, digits=0),
-                n= ifelse(is.na(n), NA, paste0("(",n,")")))
-# set NA for ermine and Arctic fox
-df_body_mass[df_body_mass$Species %in% c("ermine", "Arctic fox"),]$n <- "(NA)"
-
-
-
[…]

### 2024-10-01 script/tables_figures/table_interannual_variation_nest_density.R

@@ -1,128 +1,0 @@
-# This script extract mean and standard deviation of annual nest density monitored in systematic plot
-#Date: April 16 2024
-
-#------------------#
-#### Librairies ####
-#------------------#
-#data manipulation
-library(dplyr)
-#spatial data manipulation
-library(sf)
-sf_use_s2(FALSE)
-source("script/functions/FUNS.R")
-library(xtable)
-
-#-------------------#
-#### Import data ####
-#-------------------#
-#--- import study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  dplyr::filter(zone != "goose_colony")
-
-#--- import data frame that specify each species that have been monitored in each zone for which years
-sampling_plot <- read.csv("data/metadata/species_systematic_nest_monitoring_clean.csv") %>% 
-  dplyr::select(-X)%>%
-  dplyr::filter(sampled== "yes", species %in% c("lapland longspur", "bairds sandpiper", "sandhill crane", "long tailed duck", "king eider", "long tailed jaeger", "rock ptarmigan")) 
-
-sampling_study_area <- read.csv("data/metadata/species_systematic_nest_monitoring_clean.csv") %>%
-  dplyr::select(-X) %>% 
-  dplyr::filter(sampled== "yes", species %in% c("cackling goose", "glaucous gull", "red throated loon", "pacific loon", "snowy owl", "tundra swan", "common raven", "common ringed plover", "parasitic jaeger"), zone=="study area") 
-
-#--- nest data
-nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
-nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
-
-#------------------------------#
-#### Calculate nest density ####
-#------------------------------#
-#Extract nest density for each year (0 are assigned for year not sampled -> those years will be exclude below)
-nest_density <- get_nest_density(
-  nest_data= nest_data,
-  zone_sf= study_area,
-  species_list = c(sampling_study_area$species, sampling_plot$species),
-  zone_list= study_area$zone,
-  year_list= c(1990:2023)) %>% 
-  unique() %>% 
-  dplyr::arrange(species, zone, year)
-
-#peregrine falcon and rough legged hawk
-pefa_rlha <- get_nest_density_missing_coordinates(
[…]

### 2024-10-01 script/tables_figures/table_species_year_monitoring.R

@@ -1,64 +1,0 @@
-#data manipulation
-library(dplyr)
-library(tidyr)
-#spatial data manipulation
-library(sf)
-#export latex table
-library(xtable)
-
-#--- Import study area 
-#list of zones for which we want to obtain a density
-zone_list <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour", "study area")
-#shapefile of the area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  sf::st_drop_geometry() %>% 
-  #keep only non-nested zones
-  dplyr::filter(zone %in% zone_list)
-
-#--- Filter nest data to keep only nest found during years of systematic survey
-sp_monitoring <- read.csv("data/metadata/species_systematic_nest_monitoring_raw.csv")
-#reorder names based on defined species list
-sp_metadata <- read.csv("data/metadata/species_metadata.csv") %>% 
-  dplyr::select(species) 
-# Reorder df1 based on the species meta data order
-sp_monitoring <- sp_metadata %>% 
-  dplyr::left_join(sp_monitoring) %>% 
-  na.omit()
-
-#update species names
-sp_monitoring <- sp_monitoring %>% 
-  dplyr::mutate(species= dplyr::recode(species,"bairds sandpiper"="Baird's sandpiper", "long tailed duck"= "long-tailed duck", "long tailed jaeger"= "long-tailed jaeger", "red throated loon"= "red-throated loon", "pacific loon"= "Pacific loon","rough legged hawk"= "rough-legged hawk", "arctic hare"= "Arctic hare", "black bellied plover"= "black-bellied plover", "buff breasted sandpiper"= "buff-breasted sandpiper", "american pipit"= "American pipit", "arctic fox"="Arctic fox", "common ringed plover"= "common-ringed plover", "american golden plover"= "American golden-plover", "white rumped sandpiper"= "white-rumped sandpiper","lapland longspur"= "Lapland longspur"))
-
-#--- Extract standardize and extract a tex table
-table_year_monitoring <- sp_monitoring %>% 
[…]

### 2024-10-01 script/tables_figures/table_summary.R

@@ -1,76 +1,0 @@
-#Title: Create a summary table of all data and method use for each species
-#Author: Louis Moisan
-#Date: June 3 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-#export .tex
-library(xtable)
-
-
-#-------------------#
-#### Import data ####
-#-------------------#
-#Metadata
-metadata <- read.csv("data/metadata/summary_table.csv", sep = "\t")
-
-#Abundance data
-abundance <- read.csv("dataset/BYLOT-species_abundance.csv") %>% 
-  dplyr::mutate(year= ifelse(is.na(year), "mean", year)) %>% 
-  dplyr::group_by(species, method) %>% 
-  dplyr::summarise(min= min(abundance), max= max(abundance), mean= as.character(round(mean(abundance), digits=0)), sd= as.character(round(sd(abundance), digits= 0)), n= n()) %>% 
-  dplyr::ungroup() %>% 
-  dplyr::mutate(n= ifelse(n==1, NA, n), min= ifelse(min==max, NA, min), max= ifelse(max==min, NA, max))
-
-
-#--------------------------#
-#### Generate the table ####
-#--------------------------#
-#Join data frames together
-table <- metadata %>% 
-  dplyr::left_join(abundance, by= c("species", "method"))
-
-#format table
-table <- table %>% 
-  dplyr::select(species, Estimate, Method.description, Method.quality, Justification, min, max, mean, sd, n) %>% 
-  #rename columns
-  dplyr::rename(Species= species,  Method= Method.description,  `Method quality`= Method.quality, `Lowest abundance`= min, `Highest abundance`= max, `Mean abundance`= mean) %>% 
-  #Set mean to NA in column estimate
-  dplyr::mutate(Estimate= ifelse(Estimate == "mean", NA,Estimate)) %>% 
-  #Add years to column n (sample size)
-  dplyr::mutate(n= paste0(n, " (",Estimate,")")) %>% 
-  dplyr::mutate(n= ifelse(n== "NA (NA)", NA, n)) %>% 
-  #Remove column Estimate
-  dplyr::select(-Estimate)
-
-
-#as tex
[…]

### 2024-10-01 table_variables_format.R

@@ -1,39 +1,0 @@
-#Title: Create a table of all variables and their associated type,code definition and range of values
-#Author: Louis Moisan
-#Date: September 25 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-#export .tex
-library(xtable)
-
-#Import table structure .csv format
-table <- read.csv("MetadataS1/tables/table_data_format.csv") 
-
-#Import data set
-species_taxonomy <- read.csv("dataset/BYLOT-species_taxonomy.csv")
-species_abundance <- read.csv("dataset/BYLOT-species_abundance.csv")
-species_body_mass <- read.csv("dataset/BYLOT-species_body_mass.csv")
-variation_nest_density <- read.csv("dataset/BYLOT-interannual_variation_nest_density.csv")
-
-#Assign range values
-table[table$Variable.identity == "year",]$Range <- paste0(min(as.numeric(species_abundance$year), na.rm = T), "-", max(as.numeric(species_abundance$year), na.rm = T))
-table[table$Variable.identity == "abundance",]$Range <- paste0(min(species_abundance$abundance), "-", max(species_abundance$abundance))
-table[table$Variable.identity == "body_mass_g",]$Range <- paste0(min(species_body_mass$body_mass_g), " - ", max(species_body_mass$body_mass_g))
-table[table$Variable.identity == "mean_nest_density_km2",]$Range <- paste0(min(variation_nest_density$mean_nest_density_km2), "-", max(variation_nest_density$mean_nest_density_km2))
-table[table$Variable.identity == "sd_nest_density_km2",]$Range <- paste0(min(variation_nest_density$sd_nest_density_km2), "-", max(variation_nest_density$sd_nest_density_km2))
-table[table$Variable.identity == "sample_size_nest_density_km2",]$Range <-paste0(min(variation_nest_density$sample_size_nest_density_km2), "-", max(variation_nest_density$sample_size_nest_density_km2))
-
-#number of digits
-table[table$Storage.type=="numeric",]$Number.digits <- 3
-table[table$Storage.type=="integer",]$Number.digits <- 0
-
-
-unique(species_taxonomy$annual_cycle_strategy)
-unique(species_abundance$status)
-unique(species_abundance$method)
-unique(species_body_mass$source)
-unique(variation_nest_density$zone)

### 2024-10-03 MetadataS1/MetadataS1.tex
Revision of the entire document and minor changes

@@
 \usepackage{enumitem}
+\usepackage{parskip}
+\setlength{\parskip}{12pt}
+
 %% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Dynamic community modelling can also be applied to address practical challenges, including fishery management \citep{plaganyi2007} and the planning of protected areas \citep{okey2004, dahood2020}.
+
 Modeling food webs requires adjusting trophic flows based on the functional responses of species, which necessitates time series data on the abundance of all species within a community.
-However, accurately determining the abundance of all species is rarely achievable.
+However, determining the abundance of all species within a community is rarely achievable.
 Consequently, empirical community models often reduce taxonomic resolution by grouping species into large functional or taxonomic categories.
@@
 Due to these constraints, empirical datasets with high taxonomic resolution that cover entire communities at broad spatial and temporal scales are rare and often include incomplete or rough estimates.
+
 The composition of ecological communities is influenced by various factors acting at different temporal and spatial scales, leading to noisy data and emphasizing the need for long-term data sets \citep{magurran2010, lindenmayer2012}.
@@
 Therefore, long-term data series are required to untangle the relative effects of diverse abiotic and biotic factors on community composition \citep{magurran2010, lindenmayer2012}.
-Arctic environments are highly valuable systems for studying community patterns and processes due to their relatively low species richness \citep{payer2013, legagneux2014}.
+
+Arctic environments are highly valuable systems for studying community structure and dynamics due to their relatively low species richness \citep{payer2013, legagneux2014}.
 However, logistical challenges in the Arctic limit the number of long-term biodiversity monitoring programs.
@@
 Datasets on terrestrial communities are notably scarce, and this scarcity extends to Arctic communities as well \citep{ims2013}.
+
 Within terrestrial Arctic sites, the south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive biodiversity monitoring programs \citep{gauthier2024a}.
@@
 This approach allows the scaling of local density measurements to the landscape level when required and facilitates the estimation of abundance for less common and rare species.
+
 Previous work based on the tundra community of Bylot Island  has already produced several influential papers \citep{gauthier2011, legagneux2012, legagneux2014,hutchison2020, duchesne2021, gauthier2024b}.
@@
 This synthesis aims to provide readily accessible annual time series (or mean values in some cases) of abundance and biomass for all vertebrate species in a tundra landscape, covering approximately 400 km\textsuperscript{2}.
+
 \begin{figure}[H]
@@
 \end{figure}
+
 \subsection*{Objective}
@@
 This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration.
+We focus on adults, except for lemmings for which we have not distinguished between juveniles and adults.
 Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments.
@@
      \texttt{BYLOT-species\_abundance.csv}\\
-     \texttt{BYLOT-mean\_species\_abundance.csv}\\
-     \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
+     \texttt{BYLOT-species\_body\_mass.csv}\\
      \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\
@@
     \section*{A. Overall project description}
-    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial vertebrate communities
-        \subsection*{2. Originators:}   
+    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial communities
+        \subsection*{2. Originators}   
         \textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
@@
         \subsection*{4. Objectives}
-       i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial vertebrate communities.\\
-       ii) Predict the effects of current global environmental changes on the structure and dynamics of Arctic terrestrial vertebrate communities.
+       i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial communities.\\
+       ii) Predict the effects of current global environmental changes on the structure and dynamics of Arctic terrestrial communities.
 
      \subsection*{5. Abstract}
-        Arctic terrestrial vertebrate communities present low species richness, making those relatively simple communities ideal for studying ecological patterns and dynamics in terrestrial environments.
-Although, they present complex networks of interacting species, extreme seasonal changes of environmental conditions and a large portion of migratory species, which make challenging the identification of the key factors that shape their structure and dynamics.
-In the face of rapid global environmental changes, it is crucial to have a comprehensive understanding of the key processes shaping Arctic terrestrial communities structure and dynamics in order to predict how global changes will impact them in the future.
-Our research emphasizes long-term biodiversity monitoring, a community-wide perspective and food web modeling to achieve this understanding.
+      Arctic terrestrial communities, characterized by relatively low species richness, offer unique opportunities for studying ecological patterns and dynamics in simplified systems.
+Despite their relative simplicity, these ecosystems feature complex species interactions, extreme seasonal environmental changes, and a significant proportion of migratory species, making it difficult to identify the key factors shaping their structure and dynamics.
+As global environmental changes accelerate, it is essential to understand the processes driving these communities to eventually predict future impacts on Arctic ecosystems.
+Our research combines long-term biodiversity monitoring, a community-wide approach, and food web modeling to address these challenges.
 
@@
                 \subsubsection*{b.
-Geography} Our 463 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
+Geography} Our 389 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
                 \subsubsection*{c.
 Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
-                \subsubsection*{d. Geology} See \citet{klassen1993} for a complete and detailed description of the geology of the study area.
+                \subsubsection*{d. Geology} See \citet{klassen1993} for a detailed description of the geology of the study area.
                 \subsubsection*{e.
-Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
+Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Louis-Pierre Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
                 \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                 \subsubsection*{g.
-Climate} The mean temperature in July is 6°C, and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
-The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds.
-\citep{gauthier2024a}.
+Climate} The mean annual air temperature since 1995 is -14.4°C, with mean seasonal temperature of 4.7°C in summer (June to August), -11°C in fall (September to November), -32.4°C in winter (December to February) and -19.4°C in spring (March to May; \citet{bylotclimate}).
+The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds \citep{gauthier2024a}.
+In summer, the study area received on average 77.5 mm of precipitation \citep{bylotclimate}.
+Additionnally, 102 days are frost-free annually on average  \citep{bylotclimate} and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
 \newpage
@@
   \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
-   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada.
-The outline of the snow goose colony area is presented with white dash; we used the outline of the colony in 2017 since it represents an average colony area (74 km\textsuperscript{2}).}
+   \caption{Map of the different zones (colored polygons) of the 389 km\textsuperscript{2} study area located on the south plain of Bylot Island, Nunavut Canada.
+The perimeter of the snow goose colony is delineated by white dashes; we highlighted the perimeter in 2017 since it represents the average colony area (74 km\textsuperscript{2}).}
   \label{figure:zones}
@@
   \vspace{-70pt} % Reduce space between figure and caption
-  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}
+  \caption{Intensive nests searching plots within the Qarlikturvik valley.}
  \label{figure:qarlikturvik_valley}
@@
         \item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}\\
-        Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area.
+        Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik valley, and since 2017, in other zones of the study area.
 This sampling aimed to find all nests of the cackling goose and the glaucous gull.
@@
 Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays.
-We are confident that nest detection probability was high for these species given the open landscape.\\
+We are confident that nest detection probability was high for these species given the open landscape.
 
@@
 The total area covered by the randomly distributed plots averaged 0.79 \textpm{0.37} km\textsuperscript{2} per year.
-From 2010 onwards, except in 2020 and 2021, we traced the approximate boundary of the goose colony using a GPS receiver aboard a helicopter flying along the colony border \citep{duchesne2021}.
+From 2010 onwards, except in 2020 and 2021, we opportunistically traced sections of the approximate boundary of the goose colony using a GPS receiver aboard a helicopter, taking advantage of regular flights across the study area whenever the flight path passed over the colony border \citep{duchesne2021}.
 
@@
 Additionally, since 2012, nests have been recorded across the entire study area by scanning the landscape from hills and ridges during the nesting period \citep{duchesne2021}.
-Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.\\
+Given that snowy owls nest on elevated mounds, exhibit contrasting colors with the landscape, emit alarm calls, and display defensive behaviors, active nesting sites have a high probability of detection.
 
@@
 We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability.
-We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).\\
+We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).
 
@@
 The sampling effort was particularly intensive between 2015 and 2017.
-Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.\\
+Small areas along the coast or on the banks of smaller rivers that could potentially serve as nesting sites may have been overlooked.
 
@@
 Assuming the observer can detect all nests within a 5 or 10 meter radius, analysis of daily GPS tracks shows that the observer covered a minimum area of 0.72 \textpm{0.12} (5 m) or 1.09 \textpm{0.17} km\textsuperscript{2} (10 m) of the core area annually (n= 3 years).
-Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.\\
+Additionally, several other observers conducting related field work in the same zone reported all passerine and shorebird nests found opportunistically.
 
@@
         Since 2005, we also noted the nest location of any other bird species encountered opportunistically during travel or while carrying out the protocols for the previously described species.
-The sampling was particularly intensive in the defined 8 km\textsuperscript{2} (2x4 km plot) area in the Qarlikturvik valley.
+The sampling was particularly intensive in the defined 8 km\textsuperscript{2} area in the Qarlikturvik valley.
 The accuracy of nest monitoring in this plot thus depends on the species detection probability.
@@
         \end{enumerate}
-\input{tables/table_interannual_nest_density_variation.tex}
-%label-> table:table_interannual_nest_density_variation
+
 \newpage
@@
 We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species.
-However, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+Moreover, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands.
@@
 Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}.
-Mean body mass of vertebrate species was determined in priority with measurements from Bylot Island.
 When not available, we extracted mean body mass from the literature \citep{wilman2014}.
@@
 The methods used for each species are summarized in (\textbf{Table \ref{table:table_summary_methods_and_abundance}}).
-\newpage
+
         \begin{enumerate}[label=\alph*.]       
-
      \item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}   \\
 Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022.
-We transformed the number of nests to abundance of individuals by multiplying it by a factor of two.
 We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4).
-Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12406x + 0.13775$).
+Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12409x + 0.13774$).
 However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series.
@@
 We used the fitted model to estimate abundance between 1996 and 2016 when monitoring was less systematic, which could potentially underestimate observed abundance as seen on \textbf{Figure \ref{figure:cackling}}.
-We multiplied nest density by two to obtain the abundance.
+We multiplied nest density by two to obtain the abundance (assuming two individuals per nest).
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=0.5\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
+\includegraphics[width=0.6\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
 \caption{Number of cackling goose nests found across the study area over time.
 The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016.
-The model presents a strong fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area (R\textsuperscript{2}= 0.74, p= 0.15, n= 5).}
+The model presents a strong visual fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area.}
 \label{figure:cackling}
 \end{figure}
+
       \item[] \textit{\textbf{Snow goose}}\\
@@
 We calculated the mean annual density of snow goose nests separately in the mesic and wetland habitats of the area occupied by the goose colony annually.
-We made slight adjustments to the goose colony perimeter defined from helicopter surveys to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}).
+We made slight adjustments to the goose colony perimeter defined from helicopter flights to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}).
 To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled.
@@
 Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).
-                                \newpage
+
         \begin{figure}[ht]
@@
         \caption{Map showing the region occupied by the snow goose colony in 2017 (green polygon) as an example.
-The perimeter was first defined from a helicopter surveys and then slightly adjusted to include all snow goose point counts where at least one breeding pair had been observed in that year (green dots).
+The perimeter was first defined opportunistically using a GPS receiver aboard a helicopter, taking advantage of regular flights across the study area whenever the flight path passed over the colony border.
+The perimeter was then slightly adjusted to include all snow goose point counts where at least one breeding pair had been observed in that year (green dots).
 Snow goose point counts where no breeding geese were observed in that year are presented as red dots.}
@@
         \end{figure}
+
         \begin{figure}[ht]
@@
         \includegraphics[width=\linewidth]{figures/snow_goose_mesic_density.pdf}
-        \caption{Estimates of breeding goose density in mesic habitat of the Bylot Island study area using three independent methods: Nest sampling plot, Snow goose point count, Vertebrate count transect.}
+        \caption{Estimates of breeding goose density in mesic habitat within the Bylot Island snow goose colony using three independent methods: Nest sampling plot, Snow goose point count, Vertebrate count transect.}
         \label{figure:sngo_interannual_variation}
@@
 We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}).
-We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two.
+We transformed nest density to abundance of breeding individuals by multiplying it by a factor of two (assuming two individuals per nest).
 We acknowledge that the opportunistic monitoring of these species likely underestimated their true nest density.
@@
                 We estimated the abundance of breeding rough-legged hawks, peregrine falcons and snowy owls based on systematic nest monitoring conducted throughout the study area for these species.
-To convert the number of nests into breeding abundance, we multiplied it by two.
+To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest).
 For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.99; p < 0.0001, n= 10).
@@
 Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat.
-To convert the number of nests into breeding abundance, we multiplied it by two.
+To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest).
 
@@
 As an alternative approach to determine black-bellied plover abundance, we used the same approach as previously described, but with the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations.
+
                 \item[] \textit{\textbf{Common-ringed plover}}\\
                 To estimate the abundance of common-ringed plovers in the study area, we relied on the total number of nests recorded annually from 2015 to 2017, during which the primary nesting sites underwent intensive sampling.
-We multiplied the total nest count by two to represent the abundance of breeding individuals.
+We multiplied the total nest count by two to represent the abundance of breeding individuals (assuming two individuals per nest).
 
@@
                 \end{figure}
+
                 \item[] \textit{\textbf{Other passerines and sandpipers}}\\
-                We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section \textit{\textbf{Lapland longspur and Baird’s sandpiper}}).
+                We estimated the abundance of other passerines (horned lark, American pipit, and snow bunting) in the lowland zones of the study area with the regression equation between number of individuals per transect and nest density of the Lapland longspur (see section \textit{Lapland longspur and Baird’s sandpiper}).
 We assumed here a similar detection probability for all species.
-We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section \textit{\textbf{Lapland longspur and Baird’s sandpiper}}).
+We used the same approach for other sandpiper species (white-rumped sandpiper, pectoral sandpiper, buff-breasted sandpiper, red knot, ruddy turnstone and red phalarope) based on the regression equation for the Baird's sandpiper (see section \textit{Lapland longspur and Baird’s sandpiper}).
 For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones.
@@
 We therefore derived the absolute abundance of other passerine and sandpiper species from their relative abundances using respectively, the ratios between relative and absolute abundances of Lapland longspur (passerines) and Baird’s sandpiper (sandpipers) as references.
+
                 \item[] \textit{\textbf{Long-tailed jaeger}}\\
@@
 We determined nest density by dividing the annual number of nests recorded by the area of the surveyed zone (33 km\textsuperscript{2}).
-As long-tailed jaegers typically nest in mesic habitat \citep{andersson1971}, we multiplied the area occupied by mesic habitat across the study area by the nest density obtained in the surveyed zone and by two to obtain the total abundance of breeding individuals.
+As long-tailed jaegers typically nest in mesic habitat \citep{andersson1971}, we multiplied the area occupied by mesic habitat across the study area by the nest density obtained in the surveyed zone and by two to obtain the total abundance of breeding individuals (assuming two individuals per nest).
 
@@
 Independently, we estimated the abundance of common ravens with the same approach but using the indices of relative abundance presented by \citet{gauthier2024a}, which was derived from incidental daily observations, rather than observations from the transects.
+
         \item[] \textit{\textbf{Brown and collared lemming}}\\
@@
 \begin{figure}[h]
-\caption{Time series of the estimated annual adult abundance of vertebrate species on the southern plain of Bylot Island (389 km\textsuperscript{2}).
+\caption{Time series of the estimated annual abundance of vertebrate species on the southern plain of Bylot Island (389 km\textsuperscript{2}).
 Estimated abundance represents adult individuals, with the exception of lemmings, for which juveniles were also included in the estimate.
@@
                         \subsubsection*{b. Taxonomy and systematics}
-       Bird scientific names were obtained from the IOC World Bird List 14.2 \citep{gill2024}.
-Mammal scientific names were obtained from the Mammal species of the world: a taxonomic and geographic reference \citep{wilson2005}.
-However, we considered here \textit{Mustela erminea richardsonii} as \textit{Mustela richardsonii} due to recent genetic analysis \citep{colella2021}.
+Birds taxonomy was obtained from the IOC World Bird List 14.2 \citep{gill2024} and mammals taxonomy from the Mammal species of the world: a taxonomic and geographic reference \citep{wilson2005}.
+We considered here \textit{Mustela erminea richardsonii} as \textit{Mustela richardsonii} due to recent genetic analysis \citep{colella2021}.
 
@@
                         \item[] \textit{\textbf{Principal and associated investigators}}\\
-                        Gilles Gauthier, Eric Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux \citep{gauthier2024b}
+                        Gilles Gauthier, Eric Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux (see \citet{gauthier2024b})
                         \item[] \textit{\textbf{Students}}\\
-                        By combining animal and plant ecology, 24 doctoral theses and 56 master theses have been completed in relation to the Bylot Island South Plain study area \citep{gauthier2024b}.
+                        By combining animal and plant ecology, 24 doctoral theses and 56 master theses have been completed in relation to the study area located on the south plain of Bylot Island (see \citet{gauthier2024b}).
                 \end{enumerate}
@@
         \subsection*{3. Metadata status} XXXXXX October 2024
-        \subsection*{4. Data verification} The methods used to extract final species abundance estimates were subject to several rounds of revision by the principal investigators.
+        \subsection*{4. Data verification} The methods employed to estimate species abundance were subject to several rounds of revision by the authors.
    \section*{B. Accessibility}
@@
                         a. 35 records, not including header row (4.3 kB)\\
-                        b. 271 records, not including header row  (34.5 kB)\\
-                        c. 53 records, not including header row  (3.5 kB)\\
+                        b. 271 records, not including header row  (33.0 kB)\\
+                        c. 53 records, not including header row  (3.7 kB)\\
                         d. 18 records, not including header row  (961 B)
@@
          \subsection*{1. Program} R version 4.3.2 (2023-10-31)
-         \subsection*{2. Operating system} Data preparation was performed on x86\_64-pc-linux-gnu (64-bit) with Ubuntu 22.04.3 LTS
+         \subsection*{2. Operating system} Data preparation was performed on x86\_64-pc-linux-gnu (64-bit) with Ubuntu 22.04.3 LTS.
          \subsection*{3.
-Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{Distance}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
+Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{miller2019}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
          \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.

### 2024-10-03 MetadataS1/tables/table_species_name_strategy.tex
Revision of the entire document and minor changes

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Oct  1 09:17:32 2024
+% Thu Oct  3 11:06:53 2024
 \begin{table}[ht]
 \centering
-\caption{Species composition of the vertebrate community of Bylot Island with the corresponding migratory status (i.e., resident, partial migrant or migrant).} 
+\caption{Species of the vertebrate community of Bylot Island and their corresponding migratory status (i.e., resident, partial migrant or migrant).} 
 \label{table:table_species_name_strategy}
 \begingroup\fontsize{10pt}{10pt}\selectfont
-\begin{tabularx}{0.95\textwidth}{llll}
+\begin{tabularx}{0.97\textwidth}{llll}
   \hline
-Functional group & Scientific name & English name & Migratory status \\ 
+{\textbf{Functional group}} & {\textbf{Scientific name}} & {\textbf{English name}} & {\textbf{Migratory status}} \\ 
   \hline
-Ducks and loons  & Gavia pacifica & Pacific loon & migrant \\ 
-  Ducks and loons  & Gavia stellata & Red-throated loon & migrant \\ 
-  Ducks and loons  & Somateria spectabilis & King eider & migrant \\ 
-  Ducks and loons  & Clangula hyemalis & Long-tailed duck & migrant \\ 
-  Geese and swans  & Branta hutchinsii & Cackling goose & migrant \\ 
-  Geese and swans  & Anser caerulescens & Snow goose & migrant \\ 
-  Geese and swans  & Cygnus columbianus & Tundra swan & migrant \\ 
-  Raptors  & Buteo lagopus & Rough-legged hawk & migrant \\ 
-  Raptors  & Falco peregrinus & Peregrine falcon & migrant \\ 
-  Raptors  & Bubo scandiacus & Snowy owl & migrant \\ 
-  Ptarmigans  & Lagopus muta & Rock ptarmigan & resident \\ 
-  Cranes  & Antigone canadensis & Sandhill crane & migrant \\ 
-  Shorebirds  & Pluvialis dominica & American golden-plover & migrant \\ 
-  Shorebirds  & Pluvialis squatarola & Black-bellied plover & migrant \\ 
-  Shorebirds  & Charadrius hiaticula & Common Ringed plover & migrant \\ 
-  Shorebirds  & Arenaria interpres & Ruddy turnstone & migrant \\ 
-  Shorebirds  & Calidris canutus & Red knot & migrant \\ 
-  Shorebirds  & Calidris melanotos & Pectoral sandpiper & migrant \\ 
-  Shorebirds  & Calidris bairdii & Baird's sandpiper & migrant \\ 
-  Shorebirds  & Calidris fuscicollis & White-rumped sandpiper & migrant \\ 
-  Shorebirds  & Calidris subruficollis & Buff-breasted sandpiper & migrant \\ 
-  Shorebirds  & Phalaropus fulicarius & Red phalarope & migrant \\ 
-  Gulls  & Larus hyperboreus & Glaucous gull & migrant \\ 
-  Jaegers  & Stercorarius longicaudus & Long-tailed jaeger & migrant \\ 
-  Jaegers  & Stercorarius parasiticus & Parasitic jaeger & migrant \\ 
-  Ravens  & Corvus corax & Common raven & partial migrant \\ 
-  Passerines & Eremophila alpestris & Horned lark & migrant \\ 
-  Passerines & Anthus rubescens & American pipit & migrant \\ 
-  Passerines & Calcarius lapponicus & Lapland longspur & migrant \\ 
-  Passerines & Plectrophenax nivalis & Snow bunting & migrant \\ 
-  Lemmings  & Lemmus trimucronatus & Nearctic brown lemming & resident \\ 
-  Lemmings  & Dicrostonyx groenlandicus & Nearctic collared lemming & resident \\ 
-  Hares  & Lepus arcticus & Arctic hare & resident \\ 
-  Ermines  & Mustela richardsonii & Ermine & resident \\ 
-  Foxes  & Vulpes lagopus & Arctic fox & partial migrant \\ 
+Ducks and loons  & {\emph{ Gavia pacifica}} & Pacific loon & migrant \\ 
+  Ducks and loons  & {\emph{ Gavia stellata}} & Red-throated loon & migrant \\ 
+  Ducks and loons  & {\emph{ Somateria spectabilis}} & King eider & migrant \\ 
+  Ducks and loons  & {\emph{ Clangula hyemalis}} & Long-tailed duck & migrant \\ 
+  Geese and swans  & {\emph{ Branta hutchinsii}} & Cackling goose & migrant \\ 
+  Geese and swans  & {\emph{ Anser caerulescens}} & Snow goose & migrant \\ 
+  Geese and swans  & {\emph{ Cygnus columbianus}} & Tundra swan & migrant \\ 
+  Raptors  & {\emph{ Buteo lagopus}} & Rough-legged hawk & migrant \\ 
+  Raptors  & {\emph{ Falco peregrinus}} & Peregrine falcon & migrant \\ 
+  Raptors  & {\emph{ Bubo scandiacus}} & Snowy owl & migrant \\ 
+  Ptarmigans  & {\emph{ Lagopus muta}} & Rock ptarmigan & resident \\ 
+  Cranes  & {\emph{ Antigone canadensis}} & Sandhill crane & migrant \\ 
+  Shorebirds  & {\emph{ Pluvialis dominica}} & American golden-plover & migrant \\ 
+  Shorebirds  & {\emph{ Pluvialis squatarola}} & Black-bellied plover & migrant \\ 
+  Shorebirds  & {\emph{ Charadrius hiaticula}} & Common-ringed plover & migrant \\ 
+  Shorebirds  & {\emph{ Arenaria interpres}} & Ruddy turnstone & migrant \\ 
+  Shorebirds  & {\emph{ Calidris canutus}} & Red knot & migrant \\ 
+  Shorebirds  & {\emph{ Calidris melanotos}} & Pectoral sandpiper & migrant \\ 
+  Shorebirds  & {\emph{ Calidris bairdii}} & Baird's sandpiper & migrant \\ 
+  Shorebirds  & {\emph{ Calidris fuscicollis}} & White-rumped sandpiper & migrant \\ 
+  Shorebirds  & {\emph{ Calidris subruficollis}} & Buff-breasted sandpiper & migrant \\ 
+  Shorebirds  & {\emph{ Phalaropus fulicarius}} & Red phalarope & migrant \\ 
+  Gulls  & {\emph{ Larus hyperboreus}} & Glaucous gull & migrant \\ 
+  Jaegers  & {\emph{ Stercorarius longicaudus}} & Long-tailed jaeger & migrant \\ 
+  Jaegers  & {\emph{ Stercorarius parasiticus}} & Parasitic jaeger & migrant \\ 
+  Ravens  & {\emph{ Corvus corax}} & Common raven & partial migrant \\ 
+  Passerines & {\emph{ Eremophila alpestris}} & Horned lark & migrant \\ 
+  Passerines & {\emph{ Anthus rubescens}} & American pipit & migrant \\ 
+  Passerines & {\emph{ Calcarius lapponicus}} & Lapland longspur & migrant \\ 
+  Passerines & {\emph{ Plectrophenax nivalis}} & Snow bunting & migrant \\ 
+  Lemmings  & {\emph{ Lemmus trimucronatus}} & Nearctic brown lemming & resident \\ 
+  Lemmings  & {\emph{ Dicrostonyx groenlandicus}} & Nearctic collared lemming & resident \\ 
+  Hares  & {\emph{ Lepus arcticus}} & Arctic hare & resident \\ 
+  Ermines  & {\emph{ Mustela richardsonii}} & Ermine & resident \\ 
+  Foxes  & {\emph{ Vulpes lagopus}} & Arctic fox & partial migrant \\ 
    \hline

### 2024-10-03 MetadataS1/tables/table_species_year_monitoring.tex
Revision of the entire document and minor changes

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Oct  1 09:17:34 2024
+% Thu Oct  3 13:54:34 2024
 \begin{table}[ht]
@@
   \hline
-Species & Zone & Years & Number of years & Monitoring \\ 
+{\textbf{Species}} & {\textbf{Zone}} & {\textbf{Years}} & {\textbf{Number of years}} & {\textbf{Monitoring}} \\ 
   \hline
-Snow goose & camp 2 & 1999-2019, 2023 & (22) & systematic \\ 
-  Rough-legged hawk & qarlikt., black \& south plat. & 2007-2019, 2022 & (15) & systematic \\ 
-  Peregrine falcon & qarlikt., black \& south plat. & 2007-2019, 2022 & (15) & systematic \\ 
-  Snowy owl & qarlikt., black \& south plat. & 1996-2019, 2023 & (25) & systematic \\ 
-  Baird's sandpiper & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
-  Lapland longspur & qarlikturvik (2x1 km plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
-  King eider & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
-  Long-tailed duck & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
-  Rock ptarmigan & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
-  Sandhill crane & qarlikturvik (4x2 km plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
-  Nearctic brown lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
-  Nearctic collared lemming & qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
-  Pacific loon & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
-  Red-throated loon & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
-  Cackling goose & qarlikturvik valley & 2004-2019, 2022-2023 & (18) & systematic \\ 
-  Tundra swan & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
-  Glaucous gull & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
-  Long-tailed jaeger & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
-  Parasitic jaeger & qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
-  Pacific loon & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
-  Red-throated loon & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
-  Cackling goose & whole study area & 2017-2019, 2022-2023 & (5) & systematic \\ 
-  Tundra swan & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
-  Rough-legged hawk & whole study area & 2013-2019, 2022 & (8) & systematic \\ 
-  Peregrine falcon & whole study area & 2013-2019, 2022  & (8) & systematic \\ 
-  Snowy owl & whole study area & 2012-2019, 2022-2023 & (10) & systematic \\ 
-  Common Ringed plover & whole study area & 2015-2017 & (3) & systematic \\ 
-  Glaucous gull & whole study area & 2017-2019, 2022 & (4) & systematic \\ 
-  Parasitic jaeger & whole study area & 2009-2019, 2022 & (12) & opportunistic \\ 
-  Common raven & whole study area & 2013-2019, 2022 & (8) & systematic \\ 
-  Ermine & whole study area & 1993-2019 & (27) & opportunistic \\ 
-  Arctic fox & whole study area & 2008-2016 & (9) & systematic \\ 
+Pacific loon & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
+  Pacific loon & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
+  Red-throated loon & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
+  Red-throated loon & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
+  King eider & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
+  Long-tailed duck & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
+  Cackling goose & Qarlikturvik valley & 2004-2019, 2022-2023 & (18) & systematic \\ 
+  Cackling goose & Whole study area & 2017-2019, 2022-2023 & (5) & systematic \\ 
+  Snow goose & Camp 2 & 1999-2019, 2023 & (22) & systematic \\ 
+  Tundra swan & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
+  Tundra swan & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
+  Rough-legged hawk & Qarlikt., Black \& South plat. & 2007-2019, 2022 & (15) & systematic \\ 
+  Rough-legged hawk & Whole study area & 2013-2019, 2022 & (8) & systematic \\ 
+  Peregrine falcon & Qarlikt., Black \& South plat. & 2007-2019, 2022 & (15) & systematic \\ 
+  Peregrine falcon & Whole study area & 2013-2019, 2022  & (8) & systematic \\ 
+  Snowy owl & Qarlikt., Black \& South plat. & 1996-2019, 2023 & (25) & systematic \\ 
+  Snowy owl & Whole study area & 2012-2019, 2022-2023 & (10) & systematic \\ 
+  Rock ptarmigan & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
+  Sandhill crane & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
+  Common-ringed plover & Whole study area & 2015-2017 & (3) & systematic \\ 
+  Baird's sandpiper & Qarlikturvik (2 km2 plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
+  Glaucous gull & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
+  Glaucous gull & Whole study area & 2017-2019, 2022 & (4) & systematic \\ 
+  Long-tailed jaeger & South Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
+  Parasitic jaeger & South Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
+  Parasitic jaeger & Whole study area & 2009-2019, 2022 & (12) & opportunistic \\ 
+  Common raven & Whole study area & 2013-2019, 2022 & (8) & systematic \\ 
+  Lapland longspur & Qarlikturvik (2 km2 plot) & 2005-2019, 2022-2023 & (17) & systematic \\ 
+  Nearctic brown lemming & Qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
+  Nearctic collared lemming & Qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
+  Ermine & Whole study area & 1993-2019 & (27) & opportunistic \\ 
+  Arctic fox & Whole study area & 2008-2016 & (9) & systematic \\ 
    \hline

### 2024-10-03 script/2_extract_species_abundance/2.1_extract_species_abundance.R

@@ -108,7 +108,7 @@ pdf(file= "MetadataS1/figures/cackling_goose_nest_exponential.pdf")
 plot(all_cack_nests$year, all_cack_nests$nb_nest, 
      xlab= "Years",
      ylab= "Number of nests",
-     main= "cackling goose",
+     main= "Cackling goose",
      cex= 1.5, pch= 16, col= "steelblue4", cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)+
   lines(c(1996: 2023), exp(predict(cack_model, list(year=c(1996:2023)))), col = "red", lwd= 3)
 dev.off()
@@ -157,7 +157,7 @@ kiei_ltdu_nest <- get_nest_density(
   zone_list= "4x2km plot",
   year_list= sampling_year[sampling_year$species == "long tailed duck" & sampling_year$zone == "4x2km plot" & sampling_year$sampled=="yes",]$year)
 
-#Calculate mean density in the 4x2 plot (8km2 plot)
+#Calculate mean density in the 8km2 plot
 kiei_ltdu_density <- kiei_ltdu_nest %>% 
   dplyr::group_by(species, zone) %>% 
   dplyr::summarise(ind_density_km2= mean(ind_density_km2),
@@ -173,6 +173,7 @@ kiei_ltdu <- kiei_ltdu_density %>%
     breeding_status= "breeding",
     method= "nest sampling (extrapolation habitat)") %>% 
   dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
 
 #---Alternative approach based on Gauhtier et al., 2023 indexes
 kiei_ltdu_alt <-relative_abundance_gauthier %>% 

### 2024-10-03 script/3_tables_figures/3.10_table_expert_based_uncertainty.R

@@ -1,42 +1,0 @@
-#Title: Create a table of the expert based uncertainty described by guess estimate
-#Author: Louis Moisan
-#Date: September 30 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-#export .tex
-library(xtable)
-
-
-#--- Import the list of Bylot Island species
-sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
-  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
-
-#--- Import estimates made by experts
-table <- read.csv("data/metadata/expert_based_uncertainty.csv") %>% 
-  dplyr::left_join(sp_taxonomy) %>% 
-  dplyr::rename(Species= species_en, Period= period, `Lowest abundance`= lowest_abundance, `Highest abundance`= highest_abundance, `Mean abundance`= mean_abundance) %>% 
-  dplyr::select(Species, Period, `Lowest abundance`,`Highest abundance`,`Mean abundance`) %>% 
-  dplyr::mutate(Period= ifelse(is.na(Period), "Mean abundance",Period))
-
-#Add above header row
-addtorow <- list()
-addtorow$pos <- list(0, 0)
-addtorow$command <- c("& &\\multicolumn{3}{c}{Annual abundance (ind.)} \\\\\n",
-                      "Species & Period & Lowest & Highest & Mean \\\\\n")
-#as tex
-table_latex <- xtable(table, caption = "Considering the absence of confidence intervals in our abundance estimates, we present below uncertainty intervals on estimated abundance values derived from field expert impressions. The intervals presented represent the minimum and maximum values between which the experts believe the actual abundance values should lie. When annual abundance has been estimated for several years (time series) we present intervals over the minimum and maximum abundance values encountered during the given time series. For other species, the uncertainty interval is estimated on the mean abundance.")
-print(table_latex,
-      include.rownames = FALSE,
-      tabular.environment = "tabularx",
-      width="\\textwidth",
-      align="rXXXXXX",
-      size="\\fontsize{10pt}{8pt}\\selectfont",
-      file = "MetadataS1/tables/table_expert_based_uncertainty.tex",
-      caption.placement = "top",
-      add.to.row = addtorow,
-      include.colnames = FALSE)
-

### 2024-10-03 script/3_tables_figures/3.3_table_species_relative_abundance_upland.R

@@ -64,6 +64,10 @@ addtorow$pos <- list(0, 0)
 addtorow$command <- c("& \\multicolumn{2}{c}{Individuals/hour} \\\\\n",
                       "Species & Upland & Lowland & Ratio \\\\\n")
 
+
+#function to bold column header
+bold <-  function(x) {paste('{\\textbf{',x,'}}', sep ='')}
+
 print(table_latex,
       hline.after = c(-1, 0, nrow(table)),
       include.rownames = FALSE,
@@ -74,5 +78,6 @@ print(table_latex,
       align="rXXXXXX",
       size="\\fontsize{10pt}{10pt}\\selectfont",
       file = "MetadataS1/tables/table_species_relative_abundance_upland.tex",
+      sanitize.colnames.function=bold,
       caption.placement = "top")
 

### 2024-10-03 script/3_tables_figures/3.4_table_interannual_variation_nest_density.R

@@ -1,113 +1,0 @@
-# This script extract mean and standard deviation of annual nest density monitored in systematic plot
-#Date: April 16 2024
-
-#------------------#
-#### Librairies ####
-#------------------#
-#data manipulation
-library(dplyr)
-#spatial data manipulation
-library(sf)
-sf_use_s2(FALSE)
-source("script/functions/FUNS.R")
-library(xtable)
-
-#-------------------#
-#### Import data ####
-#-------------------#
-#--- import study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  dplyr::filter(zone != "goose_colony")
-
-#--- import data frame that specify each species that have been monitored in each zone for which years
-sampling_plot <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv") %>% 
-  dplyr::filter(sampled== "yes", species %in% c("lapland longspur", "bairds sandpiper", "sandhill crane", "long tailed duck", "king eider", "long tailed jaeger", "rock ptarmigan")) 
-
-sampling_study_area <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv") %>%
-  dplyr::filter(sampled== "yes", species %in% c("cackling goose", "glaucous gull", "red throated loon", "pacific loon", "snowy owl", "tundra swan", "common raven", "common ringed plover", "parasitic jaeger"), zone=="study area") 
-
-#--- nest data
-nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
-nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
-
-#--- Import the list of Bylot Island species
-sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>% 
-  dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
-
-
-
-#------------------------------#
-#### Calculate nest density ####
-#------------------------------#
-#Extract nest density for each year (0 are assigned for year not sampled -> those years will be exclude below)
-nest_density <- get_nest_density(
-  nest_data= nest_data,
-  zone_sf= study_area,
-  species_list = c(sampling_study_area$species, sampling_plot$species),
-  zone_list= study_area$zone,
-  year_list= c(1990:2023)) %>% 
[…]

### 2024-10-03 script/3_tables_figures/3.6_figures_species_temporal_series.R

@@ -1,65 +1,0 @@
-# Extract temporal series of abundance for each species
-#Date: January 16 2024
-
-#------------------#
-#### Librairies ####
-#------------------#
-library(dplyr)
-library(scales)
-library(ggplot2)
-
-#-------------------------#
-#### Data manipulation ####
-#-------------------------#
-#Read species final densities
-sp_abundance <- read.csv("DataS1/BYLOT-species_abundance.csv") %>% 
-  dplyr::filter(is.na(year)== FALSE)
-
-#defined order of graphs
-sp_order <- c("Cackling goose","Snow goose","Rough-legged hawk","Peregrine falcon", "Snowy owl","American golden-plover",
-  "Glaucous gull","Long-tailed jaeger","Nearctic brown lemming", "Nearctic collared lemming", "Ermine")
-
-letters <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K")
-
-#reorder species in data frame
-sp_abundance$species_en <- factor(sp_abundance$species_en, levels = sp_order)
-
-# Order the data frame based on the 'Name' column
-sp_abundance <- sp_abundance[order(sp_abundance$species_en), ]
-
-#Read study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  sf::st_drop_geometry() %>% 
-  dplyr::filter(zone != "goose_colony")
-
-
-#-------------------------------------#
-#### Extract plot for each species ####
-#-------------------------------------#
-for (i in 1:length(unique(sp_abundance$species_en))){
-sp=unique(sp_abundance$species_en)[i]
-
-#Filter data frame for a single species
-df <- sp_abundance %>% 
-    dplyr::filter(species_en ==sp)%>% 
-  mutate(year_diff = c(0, diff(year))) %>% 
-  mutate(group = cumsum(year_diff != 1),
-         year= as.integer(year))
-#Plot
-plot <-
-  ggplot(df, aes(x= year, y= abundance, group=group))+
-  ggtitle(paste0(letters[i], ". ", sp))+
-  geom_point(col= "black", size= 2)+
-  geom_line(col= "black")+
-  labs(x= "Years", y= "Number of individuals")+
-  scale_x_continuous(n.breaks = 6)+
-  scale_y_continuous(n.breaks = 6, labels = scales::number_format())+
-  theme_classic()+
-  guides(color= "none", shape= "none")+
[…]

### 2024-10-03 script/3_tables_figures/3.7_maps.R

@@ -1,168 +1,0 @@
-#title: Create a map of the study area with the different zones
-#author: Louis Moisan
-#date: October 4 2023
-
-#------------------#
-#### librairies ####
-#------------------#
-library(dplyr)
-library(sf)
-library(ggmap)
-library(ggplot2)
-
-# To extract background satellite map from google, one need to register to have a API key
-#!!! Need a personal API key from google
-# https://developers.google.com/maps
-#You have to create a API key and then under "Key and identifiers" you should see your key with the possibility to click on "show key", copy paste the key in the function below
-ggmap::register_google(key = "", write = TRUE)
-
-# Define the bounding box (latitude and longitude) for your area of interest
-bbox <- c(left = -80.19, bottom = 72.75, right = -78.5, top = 73.3)
-
-#Define only zones of interest
-zones <- c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "black plateau", "south plateau", "camp 3 plateau", "dufour")
-  
-#import study area
-study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
-  dplyr::filter(zone %in% zones) %>% 
-# Transform nc to EPSG 3857 (Pseudo-Mercator, what Google uses)
- st_transform(3857)
-
-
-#layer of the goose colony
-goose_colony <- sf::st_read("data/raw/shapefiles/study_area.shp") %>%
-  dplyr::filter(year== 2017, zone== "goose_colony") %>% 
-  sf::st_transform(3857)
-
-
-#color palette
-colors_palette <- c("#bd959d","#84a98c", "#94214b", "#463a49","#ad382b","#096377","#63141c", "#003049", "#d4af37", "white")
-
-#Get the google map background satellite
-map <- get_map(location = c(lon = -79.7, lat = 73), maptype = "satellite", source = "google", zoom= 8)
-
-# Define a function to fix the bbox to be in EPSG:3857
-ggmap_bbox <- function(map) {
-  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
-  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
[…]

### 2024-10-03 script/3_tables_figures/3.7_table_variables_definition.R

@@ -1,0 +1,40 @@
+#Title: Create a summary table of all variables in the data set
+#Author: Louis Moisan
+#Date: September 23 2024
+
+#-----------------#
+#### Libraries ####
+#-----------------#
+#general data manipulation
+library(dplyr)
+#export .tex
+library(xtable)
+
+#Import .csv format
+table <- read.csv("data/metadata/variables_definition.csv") %>% 
+  dplyr::rename(`Data file`= data.file, `Variable identity`= variable.identity, `Variable definition`= variable.definition, Units= units)
+  
+table[is.na(table)] <- "NA"
+
+#as tex
+table_latex <- xtable(table, caption = "Summary of variable definition and unit of measurement.",
+                      label= "table:table_variables_definition",
+                      align= c("p{0.00\\textwidth}", 
+                               "|p{0.08\\textwidth}|", 
+                               "p{0.21\\textwidth}|", 
+                               "p{0.50\\textwidth}|", 
+                               "p{0.08\\textwidth}|" 
+                               ) 
+) 
+#function to bold column header
+bold <-  function(x) {paste('{\\textbf{',x,'}}', sep ='')}
+
+
+print(table_latex,
+      include.rownames = FALSE,
+      tabular.environment = "longtable",
+      hline.after= c(-1,0:nrow(table)),
+      size="\\fontsize{8pt}{10pt}\\selectfont",
+      file = "MetadataS1/tables/table_variables_definition.tex",
+      sanitize.colnames.function=bold,
+      caption.placement = "top")

### 2024-10-03 script/3_tables_figures/3.8_table_variables_definition.R

@@ -1,40 +1,0 @@
-#Title: Create a summary table of all variables in the data set
-#Author: Louis Moisan
-#Date: September 23 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-#export .tex
-library(xtable)
-
-#Import .csv format
-table <- read.csv("data/metadata/variables_definition.csv") %>% 
-  dplyr::rename(`Data file`= data.file, `Variable identity`= variable.identity, `Variable definition`= variable.definition, Units= units)
-  
-table[is.na(table)] <- "NA"
-
-#as tex
-table_latex <- xtable(table, caption = "Summary of variable definition and unit of measurement.",
-                      label= "table:table_variables_definition",
-                      align= c("p{0.00\\textwidth}", 
-                               "|p{0.08\\textwidth}|", 
-                               "p{0.21\\textwidth}|", 
-                               "p{0.50\\textwidth}|", 
-                               "p{0.08\\textwidth}|" 
-                               ) 
-) 
-#function to bold column header
-bold <-  function(x) {paste('{\\textbf{',x,'}}', sep ='')}
-
-
-print(table_latex,
-      include.rownames = FALSE,
-      tabular.environment = "longtable",
-      hline.after= c(-1,0:nrow(table)),
-      size="\\fontsize{8pt}{10pt}\\selectfont",
-      file = "MetadataS1/tables/table_variables_definition.tex",
-      sanitize.colnames.function=bold,
-      caption.placement = "top")

### 2024-10-03 script/3_tables_figures/3.9_table_variables_format.R

@@ -1,80 +1,0 @@
-#Title: Create a table of all variables and their associated type,code definition and range of values
-#Author: Louis Moisan
-#Date: September 25 2024
-
-#-----------------#
-#### Libraries ####
-#-----------------#
-#general data manipulation
-library(dplyr)
-#export .tex
-library(xtable)
-
-#Import table structure .csv format
-table <- read.csv("data/metadata/variables_format.csv") 
-
-#Import data set
-species_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv")
-species_abundance <- read.csv("DataS1/BYLOT-species_abundance.csv")
-species_body_mass <- read.csv("DataS1/BYLOT-species_body_mass.csv")
-variation_nest_density <- read.csv("DataS1/BYLOT-interannual_variation_nest_density.csv")
-
-#Sum of columns
-sum(species_abundance$abundance, na.rm = T)
-sum(species_body_mass$mean_body_mass_g, na.rm = T)
-sum(species_body_mass$sample_size, na.rm = T)
-sum(variation_nest_density$mean_nest_density_km2)
-sum(variation_nest_density$sd_nest_density_km2)
-sum(variation_nest_density$number_years)
-
-
-
-
-#Assign range values
-table[table$variable.identity == "year",]$range <- paste0(min(as.numeric(species_abundance$year), na.rm = T), "-", max(as.numeric(species_abundance$year), na.rm = T))
-
-table[table$variable.identity == "abundance",]$range <- paste0(min(species_abundance$abundance), "-", max(species_abundance$abundance))
-
-table[table$variable.identity == "mean_body_mass_g",]$range <- paste0(min(species_body_mass$mean_body_mass_g), " - ", max(species_body_mass$mean_body_mass_g))
-
-table[table$variable.identity == "sample_size",]$range <- paste0(min(species_body_mass$sample_size, na.rm = T), " - ", max(species_body_mass$sample_size, na.rm= T))
-
-table[table$variable.identity == "mean_nest_density_km2",]$range <- paste0(min(variation_nest_density$mean_nest_density_km2), "-", max(variation_nest_density$mean_nest_density_km2))
-
-table[table$variable.identity == "sd_nest_density_km2",]$range <- paste0(min(variation_nest_density$sd_nest_density_km2), "-", max(variation_nest_density$sd_nest_density_km2))
-
-table[table$variable.identity == "number_years",]$range <-paste0(min(variation_nest_density$number_years), "-", max(variation_nest_density$number_years))
-
-#number of digits
[…]

### 2024-10-07 MetadataS1/MetadataS1.tex
Add star abundance of breeding table 4 and minor change title table 5

@@
 \textbf{Open Research statement}:\\
-The data set is publicly available at \url{https://datadryad.org/} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/}.\\
+The data set is publicly available at \url{https://datadryad.org/XXXXXXX} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/XXXXXXXXXX}.\\
 \endgroup
@@
 \item[] \textit{\textbf{Lemming trapping}}\\
-Since 2004, brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids.
+Since 2004, Nearctic brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids.
 Each grid is made of 144 traps separated by 30 m according to a cartesian plane, one in mesic habitat and the other in wet habitat, located in the Qarlikturvik valley \citep{fauteux2015, gauthier2020lemmings}.
@@
 Bêty and G.
-Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}.
+Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, American ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}.
 When not available, we extracted mean body mass from the literature \citep{wilman2014}.
@@
 Therefore, the estimates we provided for these species include both breeding and potentially non-breeding individuals.
-Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, ermine, and Arctic hare.
+Additionally, we did not distinguish between breeding and non-breeding individuals for mammals such as brown and collared lemmings, Arctic fox, American ermine, and Arctic hare.
 The methods used for each species are summarized in (\textbf{Table \ref{table:table_summary_methods_and_abundance}}).
@@
                 \item[] \textit{\textbf{King eider and long-tailed duck}}\\
-                We estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley.
+                We first estimated the abundance of both king eiders and long-tailed ducks based on the annual nest density of each species found in the 8 km\textsuperscript{2} extensive nest search area located in the Qarlikturvik valley.
 We extrapolated the mean nest density in the wetlands of the Qarlikturvik valley to the wetlands of the study area (38.35 km\textsuperscript{2}).
@@
 However, considering the extensive sampling effort deployed annually within this area, we are confident to obtain a realistic order of magnitude for the number of nests present.
-Because duck sightings are frequent, yet only a few nests are found, we believe there may be non-breeding individuals.
+Because duck sightings are frequent throughout the breeding period, yet only a few nests are found, we believe there may be a significant portion of non-breeding individuals.
 Therefore, we employed an additional method to estimate the overall duck populations without differentiating between breeding and non-breeding individuals.
@@
 We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 
@@
                 \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
-                We applied a distance sampling approach to the observations of American golden-plovers made along vertebrate count transects to estimate the abundance of individuals in the mesic habitat of the study area between 2014 and 2023.
-We used perpendicular distance between detected individuals and transect (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}.
-We applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered).
-We selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term.
-We excluded observations of more than four individuals, as these likely indicated groups of non-breeders passing through the area.
+                We used a distance sampling approach to estimate the abundance of American golden-plovers in the lowland zones of the study area between 2014 and 2023.
+Observations of plovers were made along vertebrate count transects mainly in mesic habitat.
+Perpendicular distance between detected individuals and the transect path were used (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}.
+To determine the detection function, we applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered) and selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term.
+We excluded observations of groups with more than four individuals, as these likely indicated groups of non-breeders.
 We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}.
 We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 
 To determine the abundance of black-bellied plovers, we used the mean number of black-bellied plovers and American golden-plovers observed per transect as an index of relative abundance.
-We assumed that the ratios of between relative and actual abundance are the same (i.e., similar detection probability) among those species.
+We assumed that the ratios between relative and actual abundance are the same (i.e., similar detection probability) among those species.
 This assumption is realistic as those species present similarities in size, color, and reproductive behavior.
@@
 We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.
@@
 For all these species, we estimated the density in the upland zones by applying a correction factor to the mean density in lowland zones.
-This correction factor was determined based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
 Nest density was then converted in number of individuals by multiplying by the area and a factor 2.
@@
                 \item[] \textit{\textbf{Parasitic jaeger}}\\
-                Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 3.75 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}.
+                Based on the opportunistic nest monitoring of parasitic jaegers across the study area, an average of 4 nests is found annually, a small number considering that parasitic jaegers were frequently observed at the study site \citep{gauthier2024a}.
 This suggests that there may be non-breeding individuals present at the study site, or alternatively, individuals may regularly travel long distances, potentially from outside the study area, to forage during the breeding season.
@@
                 Despite the intensive nest searches for raptors in upland zones, we never found more than one common raven nest each year, a small number considering the frequent raven observations at the study site \citep{gauthier2024a}.
-This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging.
+This indicates the potential presence of non-breeding individuals or individuals that breed outside the study area but use it for foraging throughout the breeding period.
 Therefore, we opted for alternative approaches based on individual counts to estimate the abundance of both breeding and non-breeding ravens.
@@
 
-        \item[] \textit{\textbf{Brown and collared lemming}}\\
+        \item[] \textit{\textbf{Nearctic brown and collared lemming}}\\
         Between 1995 and 2003, we used the density estimates derived from the snap-trapping indices obtained in late July in each habitat.
@@
 
-                \item[] \textit{\textbf{Ermine}}\\
+                \item[] \textit{\textbf{American ermine}}\\
    				We estimated the annual abundance of ermines by transforming the annual index of relative abundance provided in \citet{bolduc2023} into individual density.
@@
   \begin{subfigure}{0.45\textwidth}
-    \includegraphics[width=\linewidth]{figures/species_temporal_series/Ermine.pdf}
+    \includegraphics[width=\linewidth]{figures/species_temporal_series/American_ermine.pdf}
   \end{subfigure}
@@
                         \subsubsection*{b. Taxonomy and systematics}
-Birds taxonomy was obtained from the IOC World Bird List 14.2 \citep{gill2024} and mammals taxonomy from the Mammal species of the world: a taxonomic and geographic reference \citep{wilson2005}.
-We considered here \textit{Mustela erminea richardsonii} as \textit{Mustela richardsonii} due to recent genetic analysis \citep{colella2021}.
+Birds taxonomy was obtained from the IOC World Bird List 14.2 \citep{gill2024} and mammals taxonomy from the Mammal species of the world: a taxonomic and geographic reference \citep{MammalDiversityDatabase}.
 
@@
                 \subsection*{1. Storage location and medium}
-The data set is publicly available at \url{https://datadryad.org/}.\\[1.5em]
-Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
+The data set is publicly available at \url{https://datadryad.org/XXXXXXXXXXX}.\\[1.5em]
+Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/XXXXXXXX}.
 
@@
    		\subsection*{2. Location of completed data forms}
-        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://zenodo.org/}.
+        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://zenodo.org/XXXXXXXX}.
         \subsection*{3. Data entry verification procedures}
@@
 Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{miller2019}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
-         \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
+         \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/XXXXXXXX}.
 
     \section*{D. Archiving}
-         \subsection*{1. Archival procedures} Data are publicly available at \url{https://datadryad.org/}.
+         \subsection*{1. Archival procedures} Data are publicly available at \url{https://datadryad.org/XXXXXXXXXX}.
          \subsection*{2. Redundant archival sites} None

### 2024-10-07 MetadataS1/tables/table_species_name_strategy.tex
Add star abundance of breeding table 4 and minor change title table 5

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Oct  3 11:06:53 2024
+% Mon Oct  7 10:12:12 2024
 \begin{table}[ht]
@@
   Hares  & {\emph{ Lepus arcticus}} & Arctic hare & resident \\ 
-  Ermines  & {\emph{ Mustela richardsonii}} & Ermine & resident \\ 
+  Ermines  & {\emph{ Mustela richardsonii}} & American ermine & resident \\ 
   Foxes  & {\emph{ Vulpes lagopus}} & Arctic fox & partial migrant \\ 

### 2024-10-07 MetadataS1/tables/table_species_year_monitoring.tex
Add star abundance of breeding table 4 and minor change title table 5

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Oct  3 13:54:34 2024
+% Mon Oct  7 10:12:16 2024
 \begin{table}[ht]
@@
   Nearctic collared lemming & Qarlikturvik (trapping grids) & 1995-2019, 2021-2022 & (27) & systematic \\ 
-  Ermine & Whole study area & 1993-2019 & (27) & opportunistic \\ 
+  American ermine & Whole study area & 1993-2019 & (27) & opportunistic \\ 
   Arctic fox & Whole study area & 2008-2016 & (9) & systematic \\ 

### 2024-10-07 MetadataS1/tables/table_summary_methods_and_abundance.tex
Add star abundance of breeding table 4 and minor change title table 5

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Oct  1 09:17:43 2024
+% Mon Oct  7 10:13:44 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
 In some cases, two independent approaches have been used to estimate the abundance of the same species as a proxy for uncertainty.
-We provide a qualitative measure of the method quality based on data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.} \\
+We provide a qualitative measure of the method quality based on data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.
+The star (*) refers to the estimate of breeding individuals only.} \\
   \hline
@@
   \hline
-Pacific loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   6 & 4 & 3 & 4 (2017-2019, 2022) \\ 
+Pacific loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 0* & 6* & 4* & 3 & 4 (2017-2019, 2022) \\ 
    \hline
-Red-throated loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  42 &  76 & 64 & 15 & 4 (2017-2019, 2022) \\ 
+Red-throated loon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 42* & 76* & 64* & 15 & 4 (2017-2019, 2022) \\ 
    \hline
-King eider & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & very low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 25 &  &  \\
+King eider & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & very low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 25* &  &  \\
    \hline
@@
    \hline
-Long-tailed duck & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & very low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 20 &  &  \\
+Long-tailed duck & Intensive, but opportunistic nest monitoring (8 km2) extrapolated by habitat & very low & Intensive, but opportunistic monitoring at relatively small spatial scale, but does not include potential non-breeding individuals &  &  & 20* &  &  \\
    \hline
@@
    \hline
-Cackling goose & Extrapolation from exponential model of growth (R2=0.74, p=0.15, n=5) & moderate & Strong correlation with opportunistic nest monitoring &   2 & 158 & 31 & 41 & 23 (1996-2016, 2020-2021) \\
+Cackling goose & Extrapolation from exponential model of growth (strong visual fit with empirical data) & moderate & Strong correlation with opportunistic nest monitoring & 2* & 158* & 31* & 41 & 23 (1996-2016, 2020-2021) \\
    \hline
-Cackling goose & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  80 & 214 & 138 & 50 & 5 (2017-2019, 2022-2023) \\ 
+Cackling goose & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 80* & 214* & 138* & 50 & 5 (2017-2019, 2022-2023) \\ 
    \hline
-Snow goose & Nest monitoring plots extrapolated to mean goose colony area & moderate & Relatively small sample size and uncertainty on goose colony area  & 2505 & 35404 & 18129 & 11037 & 11 (1999-2009) \\
+Snow goose & Nest monitoring plots extrapolated to mean goose colony area & moderate & Relatively small sample size and uncertainty on goose colony area  & 2505* & 35404* & 18129* & 11037 & 11 (1999-2009) \\
    \hline
-Snow goose & Intensive study area-wide monitoring based on a combination of methods (transects, point counts and nest monitoring plots) and annual colony outline & high & Multiple independent methods and annual colony outline & 8687 & 49076 & 31852 & 12092 & 12 (2010-2019, 2022-2023) \\
+Snow goose & Intensive study area-wide monitoring based on a combination of methods (transects, point counts and nest monitoring plots) and annual colony outline & high & Multiple independent methods and annual colony outline & 8687* & 49076* & 31852* & 12092 & 12 (2010-2019, 2022-2023) \\
    \hline
-Tundra swan & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &   2 & 1 & 1 & 4 (2017-2019, 2022) \\ 
+Tundra swan & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 0* & 2* & 1* & 1 & 4 (2017-2019, 2022) \\ 
    \hline
-Rough-legged hawk & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=8) & high & Strong correlation with study area-wide nest density &   3 &  59 & 22 & 21 & 6 (2007-2012) \\
+Rough-legged hawk & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=8) & high & Strong correlation with study area-wide nest density & 3* & 59* & 22* & 21 & 6 (2007-2012) \\
    \hline
-Rough-legged hawk & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 &  66 & 30 & 29 & 8 (2013-2019, 2022) \\ 
+Rough-legged hawk & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 0* & 66* & 30* & 29 & 8 (2013-2019, 2022) \\ 
    \hline
-Peregrine falcon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   8 &  12 & 10 & 1 & 8 (2013-2019, 2022) \\ 
+Peregrine falcon & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 8* & 12* & 10* & 1 & 8 (2013-2019, 2022) \\ 
    \hline
-Snowy owl & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &   0 &  67 & 15 & 24 & 16 (1996-2011) \\ 
+Snowy owl & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &  &  &  &  & NA (1996-2011) \\ 
    \hline
-Snowy owl & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &   0 & 144 & 17 & 45 & 10 (2012-2019, 2022-2023) \\ 
+Snowy owl & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 0* & 144* & 17* & 45 & 10 (2012-2019, 2022-2023) \\ 
    \hline
-Rock ptarmigan & Intensive, but opportunistic nest monitoring (8 km2) extrapolated to study area & very low & Intensive, but opportunistic monitoring at relatively small spatial scale and prime nesting habitat not well sampled &  &  & 24 &  &  \\
+Rock ptarmigan & Intensive, but opportunistic nest monitoring (8 km2) extrapolated to study area & very low & Intensive, but opportunistic monitoring at relatively small spatial scale and prime nesting habitat not well sampled &  &  & 24* &  &  \\
    \hline
-Sandhill crane & Extrapolation from intensive nest monitoring (8 km2) and transect observations (R2= 0.44 p= 0.32, n=8) & moderate & Uncertain relation with large scale indices &  &  & 34 &  &  \\ 
+Sandhill crane & Extrapolation from intensive nest monitoring (8 km2) and transect observations (R2= 0.44 p= 0.32, n=8) & moderate & Uncertain relation with large scale indices &  &  & 34* &  &  \\ 
    \hline
-American golden-plover & Distance sampling throughout lowland (313 km2) & high & Large sample size & 397 & 1725 & 1102 & 432 & 8 (2014-2019, 2022-2023) \\ 
+American golden-plover & Distance sampling throughout lowland (313 km2) & high & Large sample size & 397* & 1725* & 1102* & 432 & 8 (2014-2019, 2022-2023) \\ 
    \hline
-Black-bellied plover & Derived from the abundance estimate of American golden-plover using transects observations & low & Derived from high quality estimate of another species &  &  & 29 &  &  \\ 
+Black-bellied plover & Derived from the abundance estimate of American golden-plover using transects observations & low & Derived from high quality estimate of another species &  &  & 29* &  &  \\ 
    \hline
-Black-bellied plover & Derived from the abundance estimate of American golden-plover using incidental observations & very low & Derived from high quality estimate of another species, but potentially includes transient migratory individuals &  &  & 87 &  &  \\
+Black-bellied plover & Derived from the abundance estimate of American golden-plover using incidental observations & very low & Derived from high quality estimate of another species, but potentially includes transient migratory individuals &  &  & 87* &  &  \\
    \hline
-Common Ringed plover & Nest monitoring on the main breeding sites & moderate & Intensive monitoring, but not exhaustive to study area &  44 &  62 & 55 & 9 & 3 (2015-2017) \\ 
+Common-ringed plover & Nest monitoring on the main breeding sites & moderate & Intensive monitoring, but not exhaustive to study area & 44* & 62* & 55* & 9 & 3 (2015-2017) \\ 
    \hline
-Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40 &  &  \\ 
+Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40* &  &  \\ 
    \hline
-Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 53 &  &  \\
+Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 53* &  &  \\
    \hline
-Red knot & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 66 &  &  \\ 
+Red knot & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 66* &  &  \\ 
    \hline
-Red knot & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 233 &  &  \\
+Red knot & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 233* &  &  \\
    \hline
-Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 80 &  &  \\ 
+Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 80* &  &  \\ 
    \hline
-Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 255 &  &  \\
+Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 255* &  &  \\
    \hline
-Baird's sandpiper & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2448 &  &  \\
+Baird's sandpiper & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2448* &  &  \\
    \hline
-White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 991 &  &  \\ 
+White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 991* &  &  \\ 
    \hline
-White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 1134 &  &  \\
+White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 1134* &  &  \\
    \hline
-Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 6 &  &  \\ 
+Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 6* &  &  \\ 
    \hline
-Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 8 &  &  \\
+Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 8* &  &  \\
    \hline
-Red phalarope & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 140 &  &  \\ 
+Red phalarope & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 140* &  &  \\ 
    \hline
-Red phalarope & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 270 &  &  \\
+Red phalarope & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 270* &  &  \\
    \hline
-Glaucous gull & Extrapolation from intensive nest monitoring (111 km2, R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density &  59 &  80 & 73 & 6 & 13 (2004-2016) \\ 
+Glaucous gull & Extrapolation from intensive nest monitoring (111 km2, R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density & 59* & 80* & 73* & 6 & 13 (2004-2016) \\ 
    \hline
-Glaucous gull & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation &  60 &  80 & 71 & 9 & 4 (2017-2019, 2022) \\ 
+Glaucous gull & Intensive study area-wide nest monitoring (389 km2) & high & No extrapolation & 60* & 80* & 71* & 9 & 4 (2017-2019, 2022) \\ 
    \hline
-Long-tailed jaeger & Intensive nest monitoring (33 km2) extrapolated by habitat & high & Relatively large spatial coverage of sampling &   0 & 900 & 272 & 285 & 17 (2004-2019, 2022) \\ 
+Long-tailed jaeger & Intensive nest monitoring (33 km2) extrapolated by habitat & high & Relatively large spatial coverage of sampling & 0* & 900* & 272* & 285 & 17 (2004-2019, 2022) \\ 
    \hline
@@
    \hline
-Parasitic jaeger & Maximum number of nest found annually during study area-wide opportunistic nest monitoring & very low & Monitoring does not include potential non-breeding individuals &  &  & 8 &  &  \\
+Parasitic jaeger & Maximum number of nest found annually during study area-wide opportunistic nest monitoring & very low & Monitoring does not include potential non-breeding individuals &  &  & 8* &  &  \\
    \hline
@@
    \hline
-Horned lark & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362 &  &  \\ 
+Horned lark & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362* &  &  \\ 
    \hline
-Horned lark & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411 &  &  \\ 
+Horned lark & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411* &  &  \\ 
    \hline
-American pipit & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 53 &  &  \\
+American pipit & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 53* &  &  \\
    \hline
-American pipit & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87 &  &  \\
+American pipit & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87* &  &  \\
    \hline
-Lapland longspur & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110 &  &  \\ 
+Lapland longspur & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110* &  &  \\
    \hline
-Snow bunting & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 18 &  &  \\
+Snow bunting & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 18* &  &  \\
    \hline
-Snow bunting & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276 &  &  \\
+Snow bunting & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276* &  &  \\
    \hline
@@
    \hline
-Ermine & Indices of relative abundance derived from testimonials converted to abundance using home range size & moderate & Indirect indices and uncertainty on ermine home range size estimates &   8 & 156 & 40 & 37 & 27 (1993-2019) \\
+American ermine & Indices of relative abundance derived from testimonials converted to abundance using home range size & moderate & Indirect indices and uncertainty on ermine home range size estimates & 8 & 156 & 40 & 37 & 27 (1993-2019) \\
    \hline

### 2024-10-07 MetadataS1/tables/table_variables_definition.tex
Add star abundance of breeding table 4 and minor change title table 5

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Tue Oct  1 10:40:29 2024
+% Mon Oct  7 10:12:24 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
   \hline
-a. & class & Taxonomic class  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
+a. & class & Taxonomic class  for birds (Gill et al., 2024) and mammals species (Upham et al., 2024). & NA \\ 
    \hline
-a. & order & Taxonomic order  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
+a. & order & Taxonomic order  for birds (Gill et al., 2024) and mammals species (Upham et al., 2024). & NA \\ 
    \hline
-a. & family & Taxonomic family  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
+a. & family & Taxonomic family  for birds (Gill et al., 2024) and mammals species (Upham et al., 2024). & NA \\ 
    \hline
-a. & genus & Taxonomic genus  for birds (Gill et al., 2024) and mammals species (Wilson, 2005). & NA \\ 
+a. & genus & Taxonomic genus  for birds (Gill et al., 2024) and mammals species (Upham et al., 2024). & NA \\ 
    \hline
-a.
-& species\_scientific & Taxonomic species  for birds (Gill et al., 2024) and mammals species (Wilson, 2005).
-We used Mustela richardsonii instead of Mustela erminea richardsonii due to recent genetic analysis (Colella et al., 2021).
-& NA \\
+a. & species\_scientific & Taxonomic species  for birds (Gill et al., 2024) and mammals species (Upham et al., 2024). & NA \\ 
    \hline
@@
    \hline
-a.
-& migratory\_status & Migratory status of each species (i.e., resident, partially migratory or migratory).
-The classification of species migratory status is based on Gauthier et al., (2011)  and Moisan et al.
-(2023).
-& NA \\
+a. & migratory\_status & Migratory status of each species. The classification of species migratory status is based on Gauthier et al., (2011)  and Moisan et al. (2023). & NA \\ 
    \hline
@@
    \hline
-c. & sample\_size & Number of individuals measured. & NA \\ 
+c. & sample\_size & Number of individuals measured. & individuals \\ 
    \hline

### 2024-10-07 script/2_extract_species_abundance/2.1_extract_species_abundance.R

@@ -664,7 +664,7 @@ n = length(df_ermine$Year)
 
 # create df to store the abundances
 df_data_final <- data.frame(
-  species = rep('ermine', n),
+  species = rep('american ermine', n),
   zone = rep('qarlikturvik valley', n),
   year = df_ermine$Year,
   ind_density_km2 = df_ermine$Ermine_estimated_abundance,

### 2024-10-07 script/3_tables_figures/3.4_table_summary_methods_and_abundance.R

@@ -25,7 +25,7 @@ sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>%
 abundance <- read.csv("DataS1/BYLOT-species_abundance.csv") %>% 
   dplyr::mutate(year= ifelse(is.na(year), "mean", year),
                 species= gsub("'","", gsub("-"," ",tolower(species_en)))) %>%
-  dplyr::group_by(species, method_description) %>% 
+  dplyr::group_by(species, method_description, breeding_status) %>% 
   dplyr::summarise(min= min(abundance), max= max(abundance), mean= as.character(round(mean(abundance), digits=0)), sd= as.character(round(sd(abundance), digits= 0)), n= n()) %>% 
   dplyr::ungroup() %>% 
   dplyr::mutate(n= ifelse(n==1, NA, n), min= ifelse(min==max, NA, min), max= ifelse(max==min, NA, max))
@@ -47,6 +47,11 @@ table <- table %>%
                 species= factor(species, levels= sp_taxonomy$species)) %>% 
   dplyr::arrange(species, year)
 
+#Add * for abundance estimate that represent breeding individuals
+table <- table %>% 
+  dplyr::mutate(across(c(min, max, mean), ~ifelse(breeding_status== "breeding", paste0(., "*"), as.character(.)))) %>% 
+  dplyr::mutate(across(c(min, max, mean), ~gsub("NA*",NA,.)))
+
 #format table
 table <- table %>% 
   dplyr::select(species_en, period, method_description, method_quality, justification, min, max, mean, sd, n) %>% 
@@ -62,7 +67,7 @@ table <- table %>%
 
 
 #as tex
-table_latex <- xtable(table, caption = "Summary of the lowest, highest, mean and standard deviation of the estimated abundance of each vertebrate species in the vertebrate community of the southern plain of Bylot Island (389 km2). In some cases, two independent approaches have been used to estimate the abundance of the same species as a proxy for uncertainty. We provide a qualitative measure of the method quality based on data available, method used for extrapolation (if necessary), and in some cases, from the fit of statistical models to estimate density.",
[…]

### 2024-10-07 script/3_tables_figures/3.5_figures_species_temporal_series.R

@@ -17,7 +17,7 @@ sp_abundance <- read.csv("DataS1/BYLOT-species_abundance.csv") %>%
 
 #defined order of graphs
 sp_order <- c("Cackling goose","Snow goose","Rough-legged hawk","Peregrine falcon", "Snowy owl","American golden-plover",
-  "Glaucous gull","Long-tailed jaeger","Nearctic brown lemming", "Nearctic collared lemming", "Ermine")
+  "Glaucous gull","Long-tailed jaeger","Nearctic brown lemming", "Nearctic collared lemming", "American ermine")
 
 letters <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K")
 

### 2024-10-07 script/3_tables_figures/3.9_table_expert_based_uncertainty.R

@@ -25,10 +25,10 @@ table <- read.csv("data/metadata/expert_based_uncertainty.csv") %>%
 #Add above header row
 addtorow <- list()
 addtorow$pos <- list(0, 0)
-addtorow$command <- c("& &\\multicolumn{3}{c}{Annual abundance (ind.)} \\\\\n",
+addtorow$command <- c("& &\\multicolumn{3}{c}{Annual abundance (individuals)} \\\\\n",
                       "Species & Period & Lowest & Highest & Mean \\\\\n")
 #as tex
-table_latex <- xtable(table, caption = "Considering the absence of confidence intervals in our abundance estimates, we present below uncertainty intervals on estimated abundance values derived from field expert impressions. The intervals presented represent the minimum and maximum values between which the experts believe the actual abundance values should lie. When annual abundance has been estimated for several years (time series) we present intervals over the minimum and maximum abundance values encountered during the given time series. For other species, the uncertainty interval is estimated on the mean abundance.")
+table_latex <- xtable(table, caption = "Due to the absence of confidence intervals in our abundance estimates, we present uncertainty intervals based on field expert impressions. These intervals reflect the interval within which experts believe the actual abundance lies. Experts derived these intervals by considering the given abundance estimate, estimates for other species for comparaison, and their field expertise. For species with time series data (several years of estimates), the intervals are presented for the lowest and highest abundance reached within the given time series. For species without time series, the intervals are based on the mean abundance.")
 print(table_latex,
       include.rownames = FALSE,
       tabular.environment = "tabularx",

### 2024-10-10 MetadataS1/tables/table_summary_methods_and_abundance.tex
Reorganize scripts for extract species abundance

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon Oct  7 10:13:44 2024
+% Thu Oct 10 16:07:59 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
    \hline
-Common raven & Derived from the abundance estimate of glaucous gull using incidental observations & very low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 31 &  &  \\
+Common raven & Derived from the abundance estimate of glaucous gull using incidental observations & very low & Derived from moderate quality estimate of another species, but potential difference in detectability between species &  &  & 18 &  &  \\
    \hline

### 2024-10-10 script/1_data_cleaning/1.2_extract_species_body_mass.R

@@ -1,7 +1,7 @@
 #Title: Extract mean species body mass measured directly on Bylot Island or from the literature
 #Author: Louis Moisan
-#Date: NA
-#Last update: June 11 2024
+#Date: summer 2024
+#Last update: October 10 2024
 
 #-----------------#
 #### Libraries ####
@@ -30,7 +30,7 @@ birds_bm <-  read.table(file = "data/raw/body_mass/birds_body_mass.txt", header 
   dplyr::select(Scientific, BodyMass.Value, BodyMass.Source) %>% 
   dplyr::rename(species_scientific= Scientific, mean_body_mass_g= BodyMass.Value, reference= BodyMass.Source)
 
-#filter to keep species from Bylot
+#filter to keep only species from Bylot
 birds_bylot <- birds_bm %>% 
   dplyr::filter(species_scientific %in% sp_taxonomy$species_scientific)
 
@@ -46,16 +46,17 @@ mammals_bm <-  read.table(file = "data/raw/body_mass/mammals_body_mass.txt", hea
   dplyr::select(Scientific, BodyMass.Value, BodyMass.Source) %>% 
   dplyr::rename(species_scientific= Scientific, mean_body_mass_g= BodyMass.Value, reference= BodyMass.Source)
 
-#filter species of bylot
+#filter to keep only species from Bylot
 mammals_bylot <- mammals_bm %>% 
   dplyr::filter(species_scientific %in% sp_taxonomy$species_scientific)
 
 #--- Combined birds and mammals
 wilman_2014 <- rbind(birds_bylot, mammals_bylot)
 
-#Which species are not included
+#Identify which species are not included
 sp_taxonomy$species_scientific[which(sp_taxonomy$species_scientific %in%wilman_2014$species_scientific == FALSE)]
-#create a vector with the different species names
+
+#Create a vector with the species names that present a different taxonomy
 sp_add_scientific <- c("Chen caerulescens","Bubo scandiaca", "Grus canadensis","Tryngites subruficollis", "Mustela erminea")
 
 #Add species to the data frame

### 2024-10-10 script/2_extract_species_abundance/2.1_extract_species_abundance.R

@@ -1,8 +1,6 @@
 #Title: Extract the abundance of every vertebrate species from Bylot Island
 #Date: October 11 2023
-#Last update:  June 17 2024
-
-
+#Last update:  October 10 2024
 
 #-----------------#
 #### Libraries ####
@@ -16,9 +14,9 @@ sf_use_s2(FALSE) #prevent geoprocessing errors
 source("script/functions/FUNS.R")
 
 
-#-------------------#
-#### Import data ####
-#-------------------#
+#-----------------#
+#### Read data ####
+#-----------------#
 #study area
 study_area <- sf::st_read("data/raw/shapefiles/study_area.shp")
 
@@ -52,705 +50,174 @@ sp_taxonomy <- read.csv("DataS1/BYLOT-species_taxonomy.csv") %>%
   dplyr::mutate(species= gsub("'","", gsub("-"," ",tolower(species_en))))
 
 
-#--------------------------------------------------#
-#### 3.1 Pacific loon, red-throated loon, cackling goose, tundra swan, glaucous gull ####
-#-------------------------------------------------#
-#--- Extract nest density for each year and each zones
-wetlands_sp <- get_nest_density(
-  nest_data= nest_data,
-  zone_sf= study_area,
-  species_list = c("cackling goose", "tundra swan","red throated loon", "pacific loon", "glaucous gull"),
-  zone_list= "study area",
-  year_list= c(2017:2023))
+#-------------------------------------#
+#### Pacific and red-throated loon ####
+#-------------------------------------#
+source("script/2_extract_species_abundance/species/loons.R")
 
-#--- Filter to keep only year systematically sampled at the study area
-wetlands_sp <- wetlands_sp %>% 
-  dplyr::semi_join(
-    #Extract combination of year and zone systematically sampled for each species
-    sampling_year %>%
-    dplyr::filter(monitoring == "systematic" & sampled == "yes" & zone == "study area") %>%
-    dplyr::select(species, year)
-    )
-
-#--- Estimate cackling goose abundance between 1996 and 2016 based on a model of exponential increase of the abundance, which presents a better fit than the linear model.
-
-#Extract cackling goose nest 1996, 2017-2019, 2022-2023
-cack_nests <- wetlands_sp %>% 
-  dplyr::filter(species== "cackling goose", zone== "study area") %>% 
-  dplyr::select(year, nb_nest) %>% 
-  #Add first nest found in 1996
-  rbind(data.frame(year=1996, nb_nest=1))
-
[…]

### 2024-10-10 script/2_extract_species_abundance/species/american_ermine.R

@@ -1,0 +1,38 @@
+#Title: Extract the abundance of American ermine on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#--- Import dataframe with indices of ermines' abundances from 
+# https://www.sciencedirect.com/science/article/pii/S1439179122000895
+df_ermine = read.csv('data/raw/sampling/ermine_index.csv', header= TRUE)
+
+# Minimum and maximum ermine abundances
+#https://esajournals.onlinelibrary.wiley.com/doi/abs/10.1890/11-1973.1?casa_token=5nWeE-CMA9gAAAAA%3AcCaOHXXyyZHXxmtxgyzr01ieSct6FZwrzkwQ4SDjkFD-mo1U6uj0oa7_pcyJ-EHBYXy1fPyTXdxOo8NEPg
+# in ind.·km-2
+ermin = 0.02
+ermax = 0.4
+
+#Associate the minimum and maximum scores of relative abundance with the minimum and maximum density of individuals, respectively
+df_ermine$Ermine_estimated_abundance = ermin + (ermax-ermin)*df_ermine$Ermine / max(df_ermine$Ermine)
+
+# number of lines
+n = length(df_ermine$Year)
+
+# create df to store the abundances
+df_data_final <- data.frame(
+  species = rep('american ermine', n),
+  zone = rep('qarlikturvik valley', n),
+  year = df_ermine$Year,
+  ind_density_km2 = df_ermine$Ermine_estimated_abundance,
+  sd_ind_density_km2 = rep(NA, n)
+)
+
+#Assume the same density at the scale of the study area
+ermine <- df_data_final %>% 
+  rbind(df_data_final %>%  dplyr::mutate(zone= "study area"))
+
+#return "ermine" object with estimated density at the scale of the study area
+american_ermine <- ermine %>% 
+  dplyr::filter(zone == "study area") %>% 
+  dplyr::mutate(breeding_status= "undetermined", method= "testimonials") %>% 
+  dplyr::select(-sd_ind_density_km2)

### 2024-10-10 script/2_extract_species_abundance/species/american_golden_plover.R

@@ -1,0 +1,23 @@
+#Title: Extract the abundance of American golden-plover on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#-------------------------------------------------------------------------------#
+#### Extract density of breeding individual abundance with distance sampling ####
+#-------------------------------------------------------------------------------#
+american_golden_plover <- extract_density_distance_sampling(
+  transect_data= transect_data , 
+  distance_data= read.csv("data/raw/sampling/american_golden_plover_distance.csv"), 
+  sp= "american golden plover", 
+  zone_list= zones, 
+  zone_sf= study_area, 
+  year_list= c(2014:2019, 2022:2023), 
+  status_list= c("repro", "non repro", "inconnu"), 
+  key= "hn", 
+  adjustment= "cos",
+  nb_adjust_term=1,
+  transect_width=150,
+  null_density_wetlands=TRUE,
+  relative_abundance_upland= relative_abundance_upland) %>% 
+  dplyr::mutate(breeding_status= "breeding", method= "distance sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status,method)

### 2024-10-10 script/2_extract_species_abundance/species/arctic_fox.R

@@ -128,4 +128,9 @@ fox_density[fox_density$goose_colony == "inside colony", ]$area_km2 <- in_colony
       #add empty column  for standard deviation since it is based on predicted model values
       dplyr::mutate(sd_ind_density_km2= NA)
     
-    
+
+# Format as other data frame
+arctic_fox <- arfo %>% 
+      dplyr::filter(zone== "study area") %>% 
+      dplyr::mutate(breeding_status= "undetermined", method= "individual home range") %>% 
+      dplyr::select(-sd_ind_density_km2)

### 2024-10-10 script/2_extract_species_abundance/species/arctic_hare.R

@@ -1,0 +1,36 @@
+#Title: Extract the abundance of Arctic hares on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#!!! The density of Arctic fox need to extract before running the following code !!!
+# source("script/2_extract_species_abundance/species/arctic_fox.R")
+
+
+#Comparaison of relative abundance with arctic fox based on incidental obs
+source("script/2_extract_species_abundance/species/arctic_fox.R") 
+fox_density <- arctic_fox %>% 
+  dplyr::filter(zone== "study area")
+# Gauthier et al., 2023 -> 0.9 hare/100h observation
+# Gauthier et al., 2023 -> 9.0 fox/100h observation
+
+#Estimate general density in lowland (where most incidental obs are done)
+arha_lowland <- data.frame(
+  zone= c("camp 2", "goose point","malaview","qarlikturvik valley",  "camp 3", "dufour"),
+  ind_density_km2= mean(fox_density$ind_density_km2)*(0.9/9.0))
+
+#Adjust density in plateau based index of relative abundance in incidental obs
+arha <- arha_lowland %>% 
+  rbind(data.frame(zone= c("black plateau", "south plateau", "camp 3 plateau"),
+                   ind_density_km2= unique(arha_lowland$ind_density_km2) *  relative_abundance_upland[relative_abundance_upland$species== "arctic hare",]$ratio_difference) %>% 
+          dplyr::select(zone, ind_density_km2))
+
+#Recalculate density at the study area by combining density in low and upland
+arctic_hare<- arha %>% 
+  dplyr::left_join(study_area %>%
+                     dplyr::select(zone, area_km2) %>%
+                     sf::st_drop_geometry()) %>%
+  dplyr::mutate(species= "arctic hare") %>% 
+  dplyr::group_by(species) %>% 
+  dplyr::summarise(ind_density_km2= sum(ind_density_km2* area_km2) /study_area[study_area$zone=="study area",]$area_km2) %>% 
+  dplyr::mutate(zone= "study area", year= NA, breeding_status= "undetermined", method= "incidental observations (relative abundance)") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)

### 2024-10-10 script/2_extract_species_abundance/species/bairds_sandpiper.R

@@ -1,0 +1,30 @@
+#Title: Extract the abundance of Baird's sandpiper on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#---------------------------------------------------------------------#
+#### Extract regression nest density and individuals per transects ####
+#---------------------------------------------------------------------#
+#Define location to save figure regression nest density and transect
+pdf("MetadataS1/figures/bairds_sandpiper_nest_transect.pdf")
+
+#Correlation of Baird's sandpiper local nest density and proportion of transects with at least one individual observed
+bairds_sandpiper<- spatial_extrapolation_species_density(
+  nest_data= nest_data,
+  transect_data= transect_data,
+  transect_metric= "proportion of transect with individual",
+  zone_sf= study_area,
+  zone_reference_nest= "2x1km plot",
+  zone_reference_transect= "4x2km plot",
+  #list of years to use to extract mean values
+  years_reference= sampling_year[sampling_year$species == "bairds sandpiper" & sampling_year$zone == "2x1km plot" & sampling_year$sampled=="yes",]$year,
+  zone_list= zones,
+  sp="bairds sandpiper",
+  status_list= c("repro", "inconnu", "non repro"),
+  relative_abundance_upland= relative_abundance_upland,
+  null_density_wetlands= FALSE)%>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling (extrapolation transects)", year= NA) %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#Save the figure
+dev.off()

### 2024-10-10 script/2_extract_species_abundance/species/black_bellied_plover.R

@@ -1,0 +1,48 @@
+#Title: Extract the abundance of black-bellied plover on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#!!! The density of American golden plover need to extract before running the following code !!!
+# source("script/2_extract_species_abundance/species/american_golden_plover.R")
+
+#-------------------------------------------------------------------------------#
+#### Extract density based on the relative proportion of observations of each species on transects ####
+#-------------------------------------------------------------------------------#
+black_bellied_plover <- estimate_density_from_reference_species(
+  reference_species_density= american_golden_plover,
+  transect_data= transect_data%>% 
+    #Remove transects with more than 4 individuals pointing towards group of non-breeding individuals
+    dplyr::filter(nb_ind<=4),
+  status_list= c("repro", "non repro", "inconnu"),
+  status_list_ref_sp= c("repro", "non repro", "inconnu"),
+  species_list= c("american golden plover", "black bellied plover"),
+  zone_list= zones,
+  zone_sf= study_area,
+  relative_abundance_upland= relative_abundance_upland,
+  sp_null_density_wetlands=NULL)%>% 
+  dplyr::filter(species== "black bellied plover") %>% 
+  dplyr::mutate(breeding_status= "breeding", method= "transects (relative abundance)", year= NA) %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#-------------------------------------------------------------------#
+#### Alternative approach based on Gauhtier et al., 2023 indexes ####
+#-------------------------------------------------------------------#
+black_bellied_plover_alternative <-relative_abundance_gauthier %>% 
+  dplyr::filter(species == "black bellied plover") %>% 
+  #Add mean density of american golden plover as reference
+  dplyr::left_join(
+    american_golden_plover %>% 
+      dplyr::filter(zone=="study area") %>% 
+      dplyr::group_by(species, zone) %>% 
+      dplyr::summarise(ref_ind_density_km2= mean(ind_density_km2))
+    , by= c("ref_species"="species")) %>% 
+  #Use cross-multiplication to estimate black bellied plover absolute abundance from relative abundance
+  dplyr::mutate(ind_density_km2= ref_ind_density_km2*(index_sp/index_ref_sp), breeding_status= "breeding", method= "incidental observations (relative abundance)", year= NA) %>% 
[…]

### 2024-10-10 script/2_extract_species_abundance/species/common_raven.R

@@ -1,0 +1,52 @@
+#Title: Extract the abundance of Common ravens on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#--------------------------------------------------------------------------#
+#### Estimate density of ravens based on the abundance of glaucous gull ####
+#--------------------------------------------------------------------------#
+#glaucous gull nest density as reference
+glgu <-get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list =  "glaucous gull",
+  zone_list= zones,
+  year_list= c(2017:2023))
+
+#--- Estimate common raven from glaucous gull abundance using mean number of individuals observed per transect as index of relative abundance
+common_raven <- estimate_density_from_reference_species(
+  reference_species_density= glgu,
+  transect_data= transect_data,
+  status_list= c("repro", "inconnu", "non repro"),
+  status_list_ref_sp= c("repro", "inconnu", "non repro"),
+  species_list= "common raven",
+  zone_list= zones,
+  zone_sf= study_area,
+  relative_abundance_upland= relative_abundance_upland,
+  sp_null_density_wetlands=NULL) %>% 
+  dplyr::filter(species== "common raven", zone== "study area") %>% 
+  dplyr::mutate(breeding_status= "undetermined", method= "transects (relative abundance)", year= NA) %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+
+#---------------------------------------------
+#### Alternative approach based on Gauhtier et al., 2023 indexes of relative abundance #### 
+#---------------------------------------------
+common_raven_alternative <-relative_abundance_gauthier %>% 
+  dplyr::filter(species== "common raven") %>% 
+  #Add abundance of glaucous gull as reference
+  dplyr::left_join(
+    glgu %>%
+      dplyr::filter(species== "glaucous gull") %>% 
+      dplyr::group_by(species, zone) %>% 
+      dplyr::summarise(ref_ind_density_km2= mean(ind_density_km2))
+    , by= c("ref_species"="species")) %>% 
+  dplyr::mutate(ind_density_km2= ref_ind_density_km2*(index_sp/index_ref_sp)) %>% 
+  dplyr::mutate(breeding_status= "undetermined", method= "incidental observations (relative abundance)", year= NA) %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#----------------------------#
[…]

### 2024-10-10 script/2_extract_species_abundance/species/glaucous_gull.R

@@ -1,0 +1,39 @@
+#Title: Extract the abundance of Glaucous gulls on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#---------------------------------------------------------------------------------#
+#### Extract nest density between 2017 and 2023 when monitoring was systematic ####
+#---------------------------------------------------------------------------------#
+glaucous_gull <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = c("glaucous gull"),
+  zone_list= "study area",
+  year_list= unique(sampling_year[sampling_year$monitoring =="systematic" & sampling_year$species == "glaucous gull" & sampling_year$sampled == "yes" & sampling_year$zone == "study area",]$year))%>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#------------------------------------------------#
+#### Extend series glaucous gull  (2004-2016) ####
+#------------------------------------------------#
+#--- Extrapolate density of glaucous gull based on density between 2004 - 2016 measured in the qarlikturvik valley
+glaucous_gull_extend <- extend_temporal_series_from_single_zone_nest(
+  nest_data= nest_data,
+  zone_sf=study_area,
+  sp= "glaucous gull",
+  #Extract complete list of year for which monitoring was done in qarlikturvik valley
+  year_list= sampling_year[sampling_year$species == "glaucous gull" & sampling_year$zone == "qarlikturvik valley" & sampling_year$sampled=="yes",]$year,
+  #Extract list of year for which sampling was done at the scale of study area and in the qarlikturvik valley
+  year_compare= sampling_year[sampling_year$species == "glaucous gull" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year,
+  zone_ref= "qarlikturvik valley") %>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling (extrapolation nests)") %>% 
+  #Remove undesired columns
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+
+#----------------------------#
+#### Combined data frames ####
+#----------------------------#
+glaucous_gull <- glaucous_gull %>% 
+  rbind(glaucous_gull_extend)

### 2024-10-10 script/2_extract_species_abundance/species/lemmings.R

@@ -102,3 +102,10 @@ lemming <- final_dens %>%
 
 #Remove row with NA as estimated density
 lemming <- lemming[complete.cases(lemming[, "ind_density_km2"]), ]
+
+#Format as other data frames
+lemming <- lemming %>% 
+  dplyr::filter(zone =="study area") %>% 
+  dplyr::mutate(breeding_status= "undetermined", method= "trapping") %>% 
+  dplyr::select(-sd_ind_density_km2) 
+

### 2024-10-10 script/2_extract_species_abundance/species/loons.R

@@ -1,0 +1,15 @@
+#Title: Extract the abundance of Pacific and red-throated loons on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#---------------------------------------------------------------------------------#
+#### Extract nest density between 2017 and 2023 when monitoring was systematic ####
+#---------------------------------------------------------------------------------#
+loons <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = c("red throated loon", "pacific loon"),
+  zone_list= "study area",
+  year_list= unique(sampling_year[sampling_year$monitoring =="systematic" & sampling_year$species == "red throated loon" & sampling_year$sampled == "yes" & sampling_year$zone == "study area",]$year))%>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)

### 2024-10-10 script/2_extract_species_abundance/species/peregrine_falcon.R

@@ -1,0 +1,19 @@
+#Title: Extract the abundance of peregrine falcons on the Bylot Island study area 
+#Date: October 11 2023
+#Last update: October 10 2024
+
+#---------------------------------------------------------------------------------#
+#### Extract nest density between 2013-2019, 2022 when monitoring was systematic ####
+#---------------------------------------------------------------------------------#
+#--- Extract nest density of rough legged hawk for each year and each zones
+#We could not used the same function as other species because the coordinates are not provided since they are sensitive.
+peregrine_falcon <- get_nest_density_missing_coordinates(
+  nest_data_csv= nest_data_csv, 
+  zone_sf= study_area, 
+  species_list= c("peregrine falcon"), 
+  zone= "study area", 
+  #list of year for which species were sampled systematically at the scale of the study area
+  year_list= sampling_year[sampling_year$species == "peregrine falcon" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year) %>% 
+  dplyr::filter(year %in% sampling_year[sampling_year$species == "peregrine falcon" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year) %>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)

### 2024-10-10 script/2_extract_species_abundance/species/rock_ptarmigan.R

@@ -1,0 +1,20 @@
+#Title: Extract the abundance of Rock ptarmigans on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#--------------------------------------------------#
+#### Extract mean nest density in the 8km2 plot ####
+#--------------------------------------------------#
+rock_ptarmigan_nests <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = "rock ptarmigan",
+  zone_list= "4x2km plot",
+  year_list= sampling_year[sampling_year$species == "rock ptarmigan" & sampling_year$zone == "4x2km plot" & sampling_year$sampled=="yes",]$year)
+
+#Calculate mean density in the qarlikturvik valley
+rock_ptarmigan <- rock_ptarmigan_nests %>% 
+  dplyr::group_by(zone) %>% 
+  dplyr::summarise(species="rock ptarmigan", year= NA, ind_density_km2= mean(ind_density_km2)) %>% 
+  dplyr::mutate(zone="study area", breeding_status= "breeding", method= "nest sampling (extrapolation habitat)") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)

### 2024-10-10 script/2_extract_species_abundance/species/rough_legged_hawk.R

@@ -1,0 +1,42 @@
+#Title: Extract the abundance of Rough-legged hawks on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#---------------------------------------------------------------------------------#
+#### Extract nest density between 2013-2019, 2022 when monitoring was systematic ####
+#---------------------------------------------------------------------------------#
+#--- Extract nest density of rough legged hawk for each year and each zones
+#We could not used the same function as other species because the coordinates are not provided since they are sensitive.
+rough_legged_hawk <- get_nest_density_missing_coordinates(
+  nest_data_csv= nest_data_csv, 
+  zone_sf= study_area, 
+  species_list= c("rough legged hawk"), 
+  zone= "study area", 
+  #list of year for which species were sampled systematically at the scale of the study area
+  year_list= sampling_year[sampling_year$species == "rough legged hawk" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year) %>% 
+  dplyr::filter(year %in% sampling_year[sampling_year$species == "rough legged hawk" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year) %>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#--------------------------------------------#
+#### Extend series snowy owl  (2007-2012) ####
+#--------------------------------------------# 
+#Estimate the density of rough-legged hawks between 2007 and 2012 based on correlation with nest density in the Qarlikturvik valley and nearby plateaus and density at the scale of the study area
+rough_legged_hawk_extend <- extend_temporal_series_from_single_zone_nest_missing_coords(
+  nest_data=read.csv("data/raw/sampling/hawk_falcon_nest_zone.csv") %>%
+    dplyr::select(-X),
+  zone_sf= study_area,
+  sp= "rough legged hawk",
+  #Extract complete list of year for which monitoring was done in qarlikturvik valley
+  year_list= sampling_year[sampling_year$species == "rough legged hawk" & sampling_year$zone == "qarlikturvik valley" & sampling_year$sampled=="yes",]$year,
+  #Extract list of year for which sampling was done at the scale of study area and in the qarlikturvik valley
[…]

### 2024-10-10 script/2_extract_species_abundance/species/sandhill_crane.R

@@ -1,0 +1,31 @@
+#Title: Extract the abundance of Sandhill cranes on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#---------------------------------------------------------------------#
+#### Extract regression nest density and individuals per transects ####
+#---------------------------------------------------------------------#
+#Define location to save figure regression nest density and transect
+pdf("MetadataS1/figures/sandhill_crane_nest_transects.pdf")
+
+#Estimate a mean annual abundance of sandhill cranes based on the mean number of individuals observed per transect in each zone.
+sandhill_crane<- spatial_extrapolation_species_density(
+  nest_data= nest_data,
+  transect_data= transect_data,
+  transect_metric= "mean number of individual per transect",
+  zone_sf= study_area,
+  zone_reference_nest= "4x2km plot",
+  zone_reference_transect= "4x2km plot",
+  #list of years that nest density was measured in the 4x2 km plot
+  years_reference= sampling_year[sampling_year$species == "sandhill crane" & sampling_year$zone == "4x2km plot" & sampling_year$sampled=="yes",]$year,
+  zone_list= zones,
+  sp="sandhill crane",
+  status_list= c("repro", "inconnu", "non repro"),
+  relative_abundance_upland= relative_abundance_upland,
+  null_density_wetlands= FALSE) %>% 
+  dplyr::filter(zone =="study area") %>% 
+  dplyr::mutate(year= NA, breeding_status= "breeding", method= "nest sampling (extrapolation transects)") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#Save figure
+dev.off()  

### 2024-10-10 script/2_extract_species_abundance/species/snowy_owl.R

@@ -1,0 +1,39 @@
+#Title: Extract the abundance of Snowy owls on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#--------------------------------------------------------#
+#### Extract nest density between  2012-2019, 2022-2023 when monitoring was systematic ####
+#-------------------------------------------------------#
+#--- Extract nest density of snowy owls for each year and each zones
+snowy_owl <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = "snowy owl",
+  zone_list= "study area",
+  #keep only year systematically sampled at the study area
+  year_list= sampling_year[sampling_year$species== "snowy owl" & sampling_year$zone== "study area" & sampling_year$sampled=="yes",]$year) %>%
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#--------------------------------------------#
+#### Extend series snowy owl  (1993-2011) ####
+#--------------------------------------------# 
+# From 1993 and 2011, monitoring was only performed in the valley of Qarlikturvik valley and nearby plateaus
+snowy_owl_extend <- extend_temporal_series_from_single_zone_nest(
+  nest_data= nest_data,
+  zone_sf=study_area,
+  sp= "snowy owl",
+  #Extract complete list of year for which monitoring was done in qarlikturvik valley
+  year_list= sampling_year[sampling_year$species == "snowy owl" & sampling_year$zone == "camp 1" & sampling_year$sampled=="yes",]$year,
+  #Extract list of year for which sampling was done at the scale of study area and in the qarlikturvik valley
+  year_compare= sampling_year[sampling_year$species == "snowy owl" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year,
+  zone_ref= c("camp 1", "black plateau", "south plateau")) %>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling (extrapolation nests)") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
+
+#----------------------------#
+#### Combined data frames ####
+#----------------------------#
+snowy_owl <- snowy_owl %>% 
+  rbind(snowy_owl_extend)

### 2024-10-10 script/2_extract_species_abundance/species/tundra_swan.R

@@ -1,0 +1,15 @@
+#Title: Extract the abundance of tundra swan on the Bylot Island study area 
+#Date: October 11 2023
+#Last update:  October 10 2024
+
+#---------------------------------------------------------------------------------#
+#### Extract nest density between 2017 and 2023 when monitoring was systematic ####
+#---------------------------------------------------------------------------------#
+tundra_swan <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = c("tundra swan"),
+  zone_list= "study area",
+  year_list=unique(sampling_year[sampling_year$monitoring =="systematic" & sampling_year$species == "tundra swan" & sampling_year$sampled == "yes" & sampling_year$zone == "study area",]$year))%>% 
+  dplyr::mutate(breeding_status= "breeding", method= "nest sampling") %>% 
+  dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)

### 2024-10-10 script/3_tables_figures/3.8_table_variables_format.R

@@ -28,8 +28,6 @@ sum(variation_nest_density$sd_nest_density_km2)
 sum(variation_nest_density$number_years)
 
 
-
-
 #Assign range values
 table[table$variable.identity == "year",]$range <- paste0(min(as.numeric(species_abundance$year), na.rm = T), "-", max(as.numeric(species_abundance$year), na.rm = T))
 

### 2024-10-16 MetadataS1/MetadataS1.tex
Reorganize scripts for extract species abundance
minor change table uncertainty intervals

@@
                 \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+                \textbf{Dominique Berteaux},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
 
@@
                 \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+                \textbf{Dominique Berteaux},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
 
@@
        i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial communities.\\
-       ii) Predict the effects of current global environmental changes on the structure and dynamics of Arctic terrestrial communities.
+       ii) Predict the effects of environmental changes on the structure and dynamics of Arctic terrestrial communities.
 
@@
                         a. 35 records, not including header row (4.3 kB)\\
-                        b. 271 records, not including header row  (33.0 kB)\\
+                        b. 271 records, not including header row  (33.3 kB)\\
                         c. 53 records, not including header row  (3.7 kB)\\
@@
                         Sums of the numeric columns: \\
-                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2293389\\
+                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2293136\\
                         c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 49617; sample\_size= 13902\\

### 2024-10-16 script/3_tables_figures/3.9_table_expert_based_uncertainty.R

@@ -22,13 +22,17 @@ table <- read.csv("data/metadata/expert_based_uncertainty.csv") %>%
   dplyr::select(Species, Period, `Lowest abundance`,`Highest abundance`,`Mean abundance`) %>% 
   dplyr::mutate(Period= ifelse(is.na(Period), "Mean abundance",Period))
 
+#Reorder rows
+table <- table[match(sp_taxonomy$species_en, table$Species), ]
+
+
 #Add above header row
 addtorow <- list()
 addtorow$pos <- list(0, 0)
 addtorow$command <- c("& &\\multicolumn{3}{c}{Annual abundance (individuals)} \\\\\n",
                       "Species & Period & Lowest & Highest & Mean \\\\\n")
 #as tex
-table_latex <- xtable(table, caption = "Due to the absence of confidence intervals in our abundance estimates, we present uncertainty intervals based on field expert impressions. These intervals reflect the interval within which experts believe the actual abundance lies. Experts derived these intervals by considering the given abundance estimate, estimates for other species for comparaison, and their field expertise. For species with time series data (several years of estimates), the intervals are presented for the lowest and highest abundance reached within the given time series. For species without time series, the intervals are based on the mean abundance.")
+table_latex <- xtable(table, caption = "Due to the absence of confidence intervals in our abundance estimates, we present uncertainty intervals based on field expert judgment. Experts derived these intervals by considering the given abundance estimate, estimates for other species for comparaison, and their field expertise. For species with time series data (several years of estimates), the intervals are presented for the lowest and highest abundance reached within the given time series. These intervals reflect the interval within which the actual abundance lies according to field experts.  For species without time series, the intervals are presented for the mean only.")
 print(table_latex,
       include.rownames = FALSE,
       tabular.environment = "tabularx",

### 2024-10-25 MetadataS1/MetadataS1.tex
version submitted to ecology - Data paper

@@
 \textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
-\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
-\\[1.5em]
+[1.5em]
 \textbf{Open Research statement}:\\
-The data set is publicly available at \url{https://datadryad.org/XXXXXXX} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/XXXXXXXXXX}.\\
+The complete data set is available as Supporting Information at: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
+Metadata document, raw data and codes are also available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
 \endgroup
@@
                 \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
-                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada\\
                 \textbf{Dominique Berteaux},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
@@
    \section*{D. Key words/phrases}
- Bylot Island, Canadian Arctic, Arctic tundra, 1993-2023, long-term monitoring, biodiversity monitoring, community composition, species abundance, species density, species biomass, species body mass, food web
+Arctic tundra, Biodiversity monitoring,  Bylot Island, Canadian Arctic, Community composition, Community structure, Food web, Long-term monitoring, Species abundance, Species biomass, Species body mass
+
   \newpage
@@
     \section*{A. Overall project description}
-    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial communities
+    \subsection*{1. Identity} Structure and functioning of Arctic terrestrial ecosystems
         \subsection*{2. Originators}   
@@
                 \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
-                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada
+                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada\\
                 \textbf{Dominique Berteaux},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
@@
         \subsection*{4. Objectives}
-       i) Understand the factors that shape the structure and drive the dynamics of Arctic terrestrial communities.\\
-       ii) Predict the effects of environmental changes on the structure and dynamics of Arctic terrestrial communities.
+       i) Develop models for the Arctic tundra food web that incorporate seasonality, migratory connectivity and interactions with marine and southern ecosystems.\\
+       ii) Predict the effects of global changes on Arctic terrestrial food webs.
 
@@
 Arctic terrestrial communities, characterized by relatively low species richness, offer unique opportunities for studying ecological patterns and dynamics in simplified systems.
-Despite their relative simplicity, these ecosystems feature complex species interactions, extreme seasonal environmental changes, and a significant proportion of migratory species, making it difficult to identify the key factors shaping their structure and dynamics.
-As global environmental changes accelerate, it is essential to understand the processes driving these communities to eventually predict future impacts on Arctic ecosystems.
+Despite their relative simplicity, these ecosystems feature complex species interactions, extreme seasonal environmental changes, and a significant proportion of migratory species, making it difficult to identify the key factors shaping their structure and functioning.
+As global environmental changes accelerate, it is essential to understand the interacting processes driving these communities to eventually predict future impacts on Arctic ecosystems.
 Our research combines long-term biodiversity monitoring, a community-wide approach, and food web modeling to address these challenges.
@@
       \subsection*{6. Sources of funding}
-Natural Sciences and Engineering Research Council of Canada, Fonds Québécois de Recherche Nature et Technologies, Centre d'études nordiques, Natural Resources Canada (Polar Continental Shelf Program), Network of Centers of Excellence Canada (ArcticNet), Canada First Research Excellence Fund (Sentinel North Program), Polar Knowledge Canada, Environment and Climate Change Canada, Canada Foundation for Innovation, Parks Canada Agency, International Polar Year program of the Government of Canada, Crown–Indigenous Relations and Northern Affairs Canada (Northern Contaminant Program), Duck Unlimited Canada, Kenneth M.
+Natural Sciences and Engineering Research Council of Canada, Fonds Québécois de Recherche Nature et Technologies, Centre d'études nordiques, Natural Resources Canada (Polar Continental Shelf Program), Network of Centers of Excellence Canada (ArcticNet), Canada First Research Excellence Fund (Sentinel North Program), Polar Knowledge Canada, Environment and Climate Change Canada, Canada Foundation for Innovation, Parks Canada Agency, International Polar Year program of the Government of Canada, Crown–Indigenous Relations and Northern Affairs Canada (Northern Contaminant Program), Ducks Unlimited Canada, Kenneth M.
 Molson Foundation (Kenneth M.
@@
 All research involving animals on Bylot Island has been approved by an institutional Animal Care Committee.
-In 1999, the study area became part of Sirmiliik National Park, managed by Parks Canada.
+In 1999, the study area became part of Sirmilik National Park, managed by Parks Canada.
 Since then, all research activities in the park have been approved by a Joint Park Management Committee.
@@
                         \item[] \textit{\textbf{Principal and associated investigators}}\\
-                        Gilles Gauthier, Eric Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux (see \citet{gauthier2024b})
+                        Gilles Gauthier, Austin Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux (see \citet{gauthier2024b})
                         \item[] \textit{\textbf{Students}}\\
@@
     \section*{A. Status}
-                \subsection*{1. Latest update} 30th September 2024
-        \subsection*{2. Latest archive date} XXXXXX October 2024
-        \subsection*{3. Metadata status} XXXXXX October 2024
+                \subsection*{1. Latest update} 25 October 2024
+        \subsection*{2. Latest archive date} 25 October 2024
+        \subsection*{3. Metadata status} 25 October 2024
         \subsection*{4. Data verification} The methods employed to estimate species abundance were subject to several rounds of revision by the authors.
@@
                 \subsection*{1. Storage location and medium}
-The data set is publicly available at \url{https://datadryad.org/XXXXXXXXXXX}.\\[1.5em]
-Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/XXXXXXXX}.
+The complete data set is available as Supporting Information at: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
+Metadata document, raw data and codes are also available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
 
@@
                         Joël Bêty; \textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}; 418 723-1986 \#1701; 300 allée des Ursulines, Rimouski, Québec, Canada, G5L 3A1, Office B-002
-                        \item[] \textit{\textbf{Data and codes}}\\
+                        \item[] \textit{\textbf{Specific subproject description}}\\
                          Louis Moisan, \textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}
@@
                         Sums of the numeric columns: \\
-                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2293136\\
+                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2280165\\
                         c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 49617; sample\_size= 13902\\
@@
    		\subsection*{2. Location of completed data forms}
-        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://zenodo.org/XXXXXXXX}.
+        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
         \subsection*{3. Data entry verification procedures}
@@
 Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{miller2019}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
-         \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/XXXXXXXX}.
+         \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
 
     \section*{D. Archiving}
-         \subsection*{1. Archival procedures} Data are publicly available at \url{https://datadryad.org/XXXXXXXXXX}.
+         \subsection*{1. Archival procedures} Data are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
          \subsection*{2. Redundant archival sites} None
@@
 The presented estimates of species abundance have not been integrated in publications to date.
-Previous estimates of species abundance on the souther plain of Bylot Island were presented by \citet{legagneux2012}, however, the temporal series presented here is longer, the methods are more refined and the taxonomic resolution is higher.
+Previous estimates of species abundance on the southern plain of Bylot Island were presented by \citet{legagneux2012}, however, the temporal series presented here is longer, the methods are refined and the taxonomic resolution is higher.
 

### 2024-10-25 MetadataS1/tables/table_summary_methods_and_abundance.tex
minor change table uncertainty intervals
version submitted to ecology - Data paper

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Oct 10 16:07:59 2024
+% Fri Oct 25 14:16:25 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
    \hline
-Snow goose & Intensive study area-wide monitoring based on a combination of methods (transects, point counts and nest monitoring plots) and annual colony outline & high & Multiple independent methods and annual colony outline & 8687* & 49076* & 31852* & 12092 & 12 (2010-2019, 2022-2023) \\
+Snow goose & Intensive study area-wide monitoring based on a combination of methods (transects, point counts and nest monitoring plots) and annual colony outline & high & Multiple independent methods and annual colony outline & 7982* & 47859* & 30771* & 11962 & 12 (2010-2019, 2022-2023) \\
    \hline

### 2024-10-25 script/2_extract_species_abundance/species/snow_goose.R

@@ -119,7 +119,7 @@ mesic_density_transects <-
   transects_sngo[which((transects_sngo$transect %in% transects_sngo_wet$transect)==FALSE),] %>% 
   #Assign within the goose range
   sf::st_intersection(study_area %>% dplyr::filter(zone=="goose_colony")) %>%
-  dplyr::filter(year ==year.1, nb_ind>0) %>% 
+  dplyr::filter(year ==year.1) %>% 
   sf::st_drop_geometry() %>% 
   dplyr::select(year, transect, nb_ind, area_km2) %>% 
   dplyr::group_by(year) %>% 
@@ -131,7 +131,7 @@ mesic_density_points<-
   sngo_point[which((sngo_point$point %in% points_sngo_wet$point)==FALSE),] %>% 
   #Assign within the goose range
   sf::st_intersection(study_area %>% dplyr::filter(zone=="goose_colony")) %>%
-  dplyr::filter(year ==year.1, n_oie_125>0) %>% 
+  dplyr::filter(year ==year.1) %>% 
   sf::st_drop_geometry() %>% 
   dplyr::select(year, point, n_oie_125, area_km2) %>% 
   dplyr::group_by(year) %>% 

### 2024-10-25 script/3_tables_figures/3.9_table_expert_based_uncertainty.R

@@ -22,10 +22,6 @@ table <- read.csv("data/metadata/expert_based_uncertainty.csv") %>%
   dplyr::select(Species, Period, `Lowest abundance`,`Highest abundance`,`Mean abundance`) %>% 
   dplyr::mutate(Period= ifelse(is.na(Period), "Mean abundance",Period))
 
-#Reorder rows
-table <- table[match(sp_taxonomy$species_en, table$Species), ]
-
-
 #Add above header row
 addtorow <- list()
 addtorow$pos <- list(0, 0)

### 2024-11-08 MetadataS1/MetadataS1.tex
update manuscrit data paper open research statement and minor changes lapland longspur and Baird sandpiper

@@
 \textbf{Open Research statement}:\\
-The complete data set is available as Supporting Information at: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
-Metadata document, raw data and codes are also available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
+The complete data set will be published as supporting information should the manuscript be accepted for publication. Metadata document, raw data and codes will also be archived in Dryad.\\
 \endgroup
@@
 
-        \item[] \textit{\textbf{Long-tailed jaeger and parasitic jaeger}} \\
+        \item[] \textit{\textbf{Long-tailed jaeger, parasitic jaeger and sandhill cranes}} \\
         Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period.
@@
 We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability.
-We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).
+We consider the sampling to be opportunistic for sandhill cranes, as individuals display defensive behaviors only at relatively close distances from their nests (see opportunistic nest monitoring below.
 
@@
 Scores of individual participants were averaged annually as detailed in \citet{bolduc2023}.
+
 \end{enumerate}
@@
 Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022.
-We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4).
+We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.7, p = 0.16, n= 4).
 Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12409x + 0.13774$).
@@
 To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest).
-For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.99; p < 0.0001, n= 10).
+For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.98; p < 0.0001, n= 10).
 We used the same approach for rough-legged hawks ($y=0.49851x$, R\textsuperscript{2}= 0.99, p < 0.0001, n= 8) to extend the time series from 2007 to 2012.
-We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.44, p= 0.27, n=8).
+We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.20, p= 0.27, n=8).
 
@@
                 \item[] \textit{\textbf{Sandhill crane}}\\
-                We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}).
-In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed.
-We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density.
-However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present.
-Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2.
-We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones.
-This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
+               We determined the mean nest density of sandhill cranes from the nest sampling between 2004 and 2023 on the southern side of the glacial river in the Qarlikturvik valley.
+We determined mean nest density by dividing the mean annual number of nests recorded by the area of the surveyed zone (33 km\textsuperscript{2}).
+We estimated total abundance by multiplying mean nest density by the total area of the study area.
+We assumed a uniform density across the study area, as sandhill crane nests have been observed in wetland, mesic, and upland habitats.
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
-\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin).
-The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
-Data points represent annual values.}
-\label{figure:sacr}
-\end{figure}
                 \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
@@
                 \item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}\\
-                We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}).
+                We estimated the mean abundance of Lapland longspurs in the different lowland zones of the study area based on a relation between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}).
 For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected.
 We made this adjustment because this species was less frequently observed.
-In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located.
+In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located (\textbf{Figure \ref{figure:qarlikturvik_valley}}).
 This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically.
-Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2.
+Transects observations in lowland were then converted into nest density using the regressions equation, and then in total number of individuals by multiplying by the area and a factor 2.
 We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones.
 This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).
-We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.
+We acknowledge that the relation for Baird’s sandpiper is weaker; however, it offers some refinement compared to assuming a uniform density throughout the study area.
 
@@
                 \end{subfigure}
-                \caption{a) Linear regression between the nest density of Lapland longspurs and the number of individuals observed per transect (nest density= 2.3422 x number of individuals per transect; regression was forced through the origin).
-The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
-Data points represent annual values.
-b) Linear regression between the nest density of Baird’s sandpiper and the proportion of transect with at least one individual observed (nest density= 34.9248 x proportion of transects with at least one individual; regression was forced through the origin).
-The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots).
-Data points represent annual values.}
+                \caption{a) The red line shows the linear relation between nest density and the number of individuals observed per transect, described by the equation nest density = 2.0031 × number of individuals per transect.
+The relation was forced through the origin.
+Blue dots represent annual values (2014–2019 and 2022–2023) of nest density and mean number of individuals per transect from the core sampled area of the Qarlikturvik Valley.
+We assumed that the relation between nest density and observations on transects is linear.
+b) The red line shows the linear relation between nest density and the proportion of transects with at least an individual, described by the equation nest density= 30.9519 x proportion of transects with at least one individual.
+The relation was forced through the origin.
+Blue dots represent annual values (2014–2019 and 2022–2023) of nest density and mean proportion of transects with at least an individual from the core sampled area of the Qarlikturvik Valley.
+We assumed that the relation between nest density and observations on transects is linear.}
                 \label{figure:lalo_basa}
@@
     \section*{A. Status}
-                \subsection*{1. Latest update} 25 October 2024
-        \subsection*{2. Latest archive date} 25 October 2024
-        \subsection*{3. Metadata status} 25 October 2024
+                \subsection*{1. Latest update} November 8, 2024
+        \subsection*{2. Latest archive date} November 8, 2024
+        \subsection*{3. Metadata status} November 8, 2024
         \subsection*{4. Data verification} The methods employed to estimate species abundance were subject to several rounds of revision by the authors.
@@
                 \subsection*{1. Storage location and medium}
-The complete data set is available as Supporting Information at: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
-Metadata document, raw data and codes are also available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
+The complete data set will be published as supporting information should the manuscript be accepted for publication. Metadata document, raw data and codes will also be archived in Dryad.
 
@@
                         a. 35 records, not including header row (4.3 kB)\\
-                        b. 271 records, not including header row  (33.3 kB)\\
+                        b. 271 records, not including header row  (35.1 kB)\\
                         c. 53 records, not including header row  (3.7 kB)\\
-                        d. 18 records, not including header row  (961 B)
+                        d. 18 records, not including header row  (1.1 kB)
 
@@
                         Sums of the numeric columns: \\
-                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2280165\\
+                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2278553\\
                         c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 49617; sample\_size= 13902\\
-                        d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185
+                        d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 20.447; sd\_nest\_density\_km2= 9.529; sample\_size\_nest\_density\_km2= 186
 
@@
    		\subsection*{2. Location of completed data forms}
-        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
+        Raw data and codes will be archived in Dryad.
         \subsection*{3. Data entry verification procedures}
@@
 Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{miller2019}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
-         \subsection*{4. Codes} Raw data and codes used to extract the presented data set are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
+         \subsection*{4. Codes} Raw data and codes will be archived in Dryad.
 
     \section*{D. Archiving}
-         \subsection*{1. Archival procedures} Data are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
+         \subsection*{1.
+Archival procedures} The complete data set will be published as supporting information should the manuscript be accepted for publication.
+Metadata document, raw data and codes will also be archived in Dryad.
          \subsection*{2. Redundant archival sites} None

### 2024-11-08 MetadataS1/tables/interannual_nest_density_variation.tex
update manuscrit data paper open research statement and minor changes lapland longspur and Baird sandpiper

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Wed Nov  6 17:11:23 2024
\begin{table}[ht]
\centering
\caption{Annual nest density (nests/km2) of selected avian species estimated on different zones of Bylot Island.} 
\label{table:interannual_nest_density_variation}
\begingroup\fontsize{10pt}{10pt}\selectfont
\begin{tabularx}{0.9\textwidth}{lllr}
  \hline
Species & Zone & Mean ± SD & Number of years \\ 
  \hline
Pacific loon & Whole study area & 0.005 ± 0.004 &   4 \\ 
  Red-throated loon & Whole study area & 0.082 ± 0.019 &   4 \\ 
  King eider & Qarlikturvik (4x2 km plot) & 0.115 ± 0.138 &  16 \\ 
  Long-tailed duck & Qarlikturvik (4x2 km plot) & 0.092 ± 0.138 &  16 \\ 
  Cackling goose & Whole study area & 0.177 ± 0.064 &   5 \\ 
  Tundra swan & Whole study area & 0.001 ± 0.001 &   4 \\ 
  Rough-legged hawk & Whole study area & 0.157 ± 0.151 &   8 \\ 
  Peregrine falcon & Whole study area & 0.053 ± 0.007 &   8 \\ 
  Snowy owl & Whole study area & 0.022 ± 0.058 &  10 \\ 
  Rock ptarmigan & Qarlikturvik (4x2 km plot) & 0.031 ± 0.055 &  16 \\ 
  Sandhill crane & Qarlikturvik valley & 0.088 ± 0.055 &  17 \\ 
  Common-ringed plover & Whole study area & 0.070 ± 0.012 &   3 \\ 
  Baird's sandpiper & Qarlikturvik (2x1 km plot) & 5.000 ± 3.558 &  17 \\ 
  Glaucous gull & Whole study area & 0.091 ± 0.011 &   4 \\ 
  Long-tailed jaeger & Qarlikturvik valley & 0.362 ± 0.380 &  17 \\ 
  Parasitic jaeger & Whole study area & 0.010 ± 0.004 &  12 \\ 
  Common raven & Whole study area & 0.003 ± 0.003 &   8 \\ 
  Lapland longspur & Qarlikturvik (2x1 km plot) & 13.559 ± 5.849 &  17 \\ 
   \hline
\end{tabularx}
\endgroup
\end{table}

### 2024-11-08 MetadataS1/tables/table_species_year_monitoring.tex
Reorganize scripts for extract species abundance
minor change table uncertainty intervals
version submitted to ecology - Data paper
update manuscrit data paper open research statement and minor changes lapland longspur and Baird sandpiper

--- 2024-10-03 MetadataS1/tables/table_species_year_monitoring.tex
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu Oct  3 13:54:34 2024
+% Fri Nov  8 08:10:03 2024
 \begin{table}[ht]
@@
   Cackling goose & Whole study area & 2017-2019, 2022-2023 & (5) & systematic \\ 
-  Snow goose & Camp 2 & 1999-2019, 2023 & (22) & systematic \\ 
+  Snow goose & Camp 2 & 1999-2019, 2022-2023 & (23) & systematic \\ 
   Tundra swan & Qarlikturvik valley & 2004-2019, 2022 & (17) & systematic \\ 
@@
   Peregrine falcon & Whole study area & 2013-2019, 2022  & (8) & systematic \\ 
-  Snowy owl & Qarlikt., Black \& South plat. & 1996-2019, 2023 & (25) & systematic \\ 
+  Snowy owl & Qarlikt., Black \& South plat. & 1996-2019, 2022-2023 & (26) & systematic \\ 
   Snowy owl & Whole study area & 2012-2019, 2022-2023 & (10) & systematic \\ 
   Rock ptarmigan & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
-  Sandhill crane & Qarlikturvik (8 km2 plot) & 2005-2019, 2022 & (16) & opportunistic \\ 
+  Sandhill crane & South Qarlikturvik valley & 2004-2019, 2022 & (17) & opportunistic \\ 
   Common-ringed plover & Whole study area & 2015-2017 & (3) & systematic \\ 

### 2024-11-08 MetadataS1/tables/table_summary_methods_and_abundance.tex
update manuscrit data paper open research statement and minor changes lapland longspur and Baird sandpiper

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Fri Oct 25 14:16:25 2024
+% Fri Nov  8 08:10:10 2024
 \begingroup\fontsize{8pt}{10pt}\selectfont
@@
    \hline
-Snowy owl & Extrapolation from intensive nest monitoring (111 km2, R2=0.99, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density &  &  &  &  & NA (1996-2011) \\ 
+Snowy owl & Extrapolation from intensive nest monitoring (111 km2, R2=0.98, p$<$0.0001, n=10) & high & Strong correlation with study area-wide nest density & 0* & 67* & 15* & 24 & 16 (1996-2011) \\ 
    \hline
@@
    \hline
-Sandhill crane & Extrapolation from intensive nest monitoring (8 km2) and transect observations (R2= 0.44 p= 0.32, n=8) & moderate & Uncertain relation with large scale indices &  &  & 34* &  &  \\ 
+Sandhill crane & Intensive nest monitoring (33 km2) extrapolated to study area & moderate & Nest density is extrapolated uniformly across the study area &  &  & 36* &  &  \\ 
    \hline
@@
    \hline
-Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 40* &  &  \\ 
+Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 36* &  &  \\ 
    \hline
-Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 53* &  &  \\
+Ruddy turnstone & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 47* &  &  \\
    \hline
-Red knot & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 66* &  &  \\ 
+Red knot & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 59* &  &  \\ 
    \hline
-Red knot & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 233* &  &  \\
+Red knot & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 206* &  &  \\
    \hline
-Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 80* &  &  \\ 
+Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 71* &  &  \\ 
    \hline
-Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 255* &  &  \\
+Pectoral sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 226* &  &  \\
    \hline
-Baird's sandpiper & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.38, p=0.35, n=8) & moderate & Uncertain relation with large scale indices &  &  & 2448* &  &  \\
+Baird's sandpiper & Extrapolation from intensive nest monitoring (2 km2) and transects observations & moderate & Uncertain relation between nest density and transects observations &  &  & 2170* &  &  \\
    \hline
-White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 991* &  &  \\ 
+White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 878* &  &  \\ 
    \hline
-White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 1134* &  &  \\
+White-rumped sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 1005* &  &  \\
    \hline
-Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 6* &  &  \\ 
+Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 5* &  &  \\ 
    \hline
-Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 8* &  &  \\
+Buff-breasted sandpiper & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 7* &  &  \\
    \hline
-Red phalarope & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 140* &  &  \\ 
+Red phalarope & Derived from the abundance estimate of Bairds sandpiper using transects observations & low & Derived from moderate quality estimate of another species &  &  & 124* &  &  \\ 
    \hline
-Red phalarope & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 270* &  &  \\
+Red phalarope & Derived from the abundance estimate of Bairds sandpiper using incidental observations & very low & Derived from moderate quality estimate of another species, but potentially includes transient migratory individuals &  &  & 240* &  &  \\
    \hline
-Glaucous gull & Extrapolation from intensive nest monitoring (111 km2, R2=0.84, p=0.16, n=4) & high & Strong correlation with study area-wide nest density & 59* & 80* & 73* & 6 & 13 (2004-2016) \\ 
+Glaucous gull & Extrapolation from intensive nest monitoring (111 km2, R2=0.7, p=0.16, n=4) & high & Strong correlation with study area-wide nest density & 59* & 80* & 73* & 6 & 13 (2004-2016) \\ 
    \hline
@@
    \hline
-Horned lark & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 362* &  &  \\ 
+Horned lark & Derived from the abundance estimate of Lapland longspur using transects observations & low & Derived from moderate quality estimate of another species &  &  & 310* &  &  \\ 
    \hline
-Horned lark & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 411* &  &  \\ 
+Horned lark & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species &  &  & 352* &  &  \\ 
    \hline
-American pipit & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 53* &  &  \\
+American pipit & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 46* &  &  \\
    \hline
-American pipit & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 87* &  &  \\
+American pipit & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 74* &  &  \\
    \hline
-Lapland longspur & Extrapolation from intensive nest monitoring (2 km2) and transects observations (R2=0.62, p=0.1, n=8) & moderate & Uncertain relation with large scale indices &  &  & 7110* &  &  \\
+Lapland longspur & Extrapolation from intensive nest monitoring (2 km2) and transects observations & moderate & Uncertain relation between nest density and transects observations &  &  & 6080* &  &  \\
    \hline
-Snow bunting & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 18* &  &  \\
+Snow bunting & Derived from the abundance estimate of Lapland longspur using transects observations & very low & Derived from moderate quality estimate of another species and prime nesting habitat not sampled &  &  & 15* &  &  \\
    \hline
-Snow bunting & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 276* &  &  \\
+Snow bunting & Derived from the abundance estimate of Lapland longspur using incidental observations & low & Derived from moderate quality estimate of another species and prime nesting habitat not well sampled &  &  & 236* &  &  \\
    \hline

### 2024-11-08 script/2_extract_species_abundance/species/sandhill_crane.R

@@ -5,27 +5,16 @@
 #---------------------------------------------------------------------#
 #### Extract regression nest density and individuals per transects ####
 #---------------------------------------------------------------------#
-#Define location to save figure regression nest density and transect
-pdf("MetadataS1/figures/sandhill_crane_nest_transects.pdf")
-
-#Estimate a mean annual abundance of sandhill cranes based on the mean number of individuals observed per transect in each zone.
-sandhill_crane<- spatial_extrapolation_species_density(
-  nest_data= nest_data,
-  transect_data= transect_data,
-  transect_metric= "mean number of individual per transect",
-  zone_sf= study_area,
-  zone_reference_nest= "4x2km plot",
-  zone_reference_transect= "4x2km plot",
-  #list of years that nest density was measured in the 4x2 km plot
-  years_reference= sampling_year[sampling_year$species == "sandhill crane" & sampling_year$zone == "4x2km plot" & sampling_year$sampled=="yes",]$year,
-  zone_list= zones,
-  sp="sandhill crane",
-  status_list= c("repro", "inconnu", "non repro"),
-  relative_abundance_upland= relative_abundance_upland,
-  null_density_wetlands= FALSE) %>% 
-  dplyr::filter(zone =="study area") %>% 
-  dplyr::mutate(year= NA, breeding_status= "breeding", method= "nest sampling (extrapolation transects)") %>% 
+sandhill_crane <- get_nest_density(nest_data= nest_data,
+                 zone_sf= study_area,
+                 species_list= "sandhill crane",
+                 zone_list= "qarlikturvik valley",
+                 year_list= sampling_year[sampling_year$species == "sandhill crane" & sampling_year$zone == "camp 1" & sampling_year$sampled=="yes",]$year) %>% 
+  dplyr::group_by(species) %>% 
+  dplyr::summarise(mean_ind_density_km2= mean(ind_density_km2)) %>% 
+  #rename columns mean to simply ind_density_km2
+  dplyr::rename(ind_density_km2= mean_ind_density_km2) %>% 
+  #Add other column
+  dplyr::mutate(zone= "study area", year= NA, breeding_status= "breeding", method= "nest sampling (extrapolation uniform)") %>% 
+  #reorder columns
   dplyr::select(species, zone, year, ind_density_km2, breeding_status, method)
-
-#Save figure
-dev.off()  

### 2024-11-08 script/2_extract_species_abundance/species/snowy_owl.R

@@ -25,7 +25,7 @@ snowy_owl_extend <- extend_temporal_series_from_single_zone_nest(
   zone_sf=study_area,
   sp= "snowy owl",
   #Extract complete list of year for which monitoring was done in qarlikturvik valley
-  year_list= sampling_year[sampling_year$species == "snowy owl" & sampling_year$zone == "camp 1" & sampling_year$sampled=="yes",]$year,
+  year_list= sampling_year[sampling_year$species == "snowy owl" & sampling_year$zone == "qarlikturvik valley" & sampling_year$sampled=="yes",]$year,
   #Extract list of year for which sampling was done at the scale of study area and in the qarlikturvik valley
   year_compare= sampling_year[sampling_year$species == "snowy owl" & sampling_year$zone == "study area" & sampling_year$sampled=="yes",]$year,
   zone_ref= c("camp 1", "black plateau", "south plateau")) %>% 

### 2024-11-08 script/3_tables_figures/3.10_table_interannual_variation_nest_density.R

@@ -1,0 +1,91 @@
+# This script extract mean and standard deviation of annual nest density monitored in systematic plot
+#Date: November 6 2024
+
+#------------------#
+#### Librairies ####
+#------------------#
+#data manipulation
+library(dplyr)
+#spatial data manipulation
+library(sf)
+sf_use_s2(FALSE)
+source("script/functions/FUNS.R")
+library(xtable)
+
+#-------------------#
+#### Import data ####
+#-------------------#
+#--- import study area
+study_area <- sf::st_read("data/raw/shapefiles/study_area.shp") %>% 
+  dplyr::filter(zone != "goose_colony")
+
+#--- import data frame that specify each species that have been monitored in each zone for which years
+sampling_plot <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv") %>% 
+  dplyr::filter(sampled== "yes", species %in% c("lapland longspur", "bairds sandpiper", "sandhill crane", "long tailed duck", "king eider", "long tailed jaeger", "rock ptarmigan")) 
+
+sampling_study_area <- read.csv("data/metadata/systematic_nest_monitoring_clean.csv") %>%
+  dplyr::filter(sampled== "yes", species %in% c("cackling goose", "glaucous gull", "red throated loon", "pacific loon", "snowy owl", "tundra swan", "common raven", "common ringed plover", "parasitic jaeger"), zone=="study area") 
+
+#--- nest data
+  nest_data <-  sf::st_read("data/raw/shapefiles/sampling/nests_2004-2023.shp")
+nest_data_csv <- read.csv("data/raw/sampling/nests_2004-2023.csv")
+
+#------------------------------#
+#### Calculate nest density ####
+#------------------------------#
+#Extract nest density for each year (0 are assigned for year not sampled -> those years will be exclude below)
+nest_density <- get_nest_density(
+  nest_data= nest_data,
+  zone_sf= study_area,
+  species_list = c(sampling_study_area$species, sampling_plot$species),
+  zone_list= study_area$zone,
+  year_list= c(1990:2023)) %>% 
+  unique() %>% 
+  dplyr::arrange(species, zone, year)
+
+#peregrine falcon and rough legged hawk
+pefa_rlha <- get_nest_density_missing_coordinates(
+  nest_data_csv= nest_data_csv, 
[…]

### 2024-11-08 script/functions/nested_functions/extend_temporal_series_from_single_zone_missing_coordinates.R

@@ -65,7 +65,7 @@ print(plot(sp_density_study_area$ind_density_km2~ sp_density_ref_filter$ind_dens
      xlab= "Density reference zone",
      ylab= "Density study area"))
 
-print(paste("Correlation for", sp, "between", unique(sp_density_ref_filter$zone), "and study area =", round(as.numeric(cor$estimate), digits = 2),",","p=", round(as.numeric(cor$p.value), digits=3),",","n=", nrow(sp_density_ref_filter),sep =" "),  nrow(sp_density_ref_filter))
+print(paste("Correlation for", sp, "between", unique(sp_density_ref_filter$zone), "and study area =", round(as.numeric(cor$estimate^2), digits = 2),",","p=", round(as.numeric(cor$p.value), digits=3),",","n=", nrow(sp_density_ref_filter),sep =" "),  nrow(sp_density_ref_filter))
 
 #Build linear model regression model
 model <- lm(sp_density_study_area$ind_density_km2~sp_density_ref_filter$ind_density_km2)

### 2024-11-08 script/functions/nested_functions/extend_temporal_series_from_single_zone_nest.R

@@ -61,7 +61,7 @@ print(plot(sp_density_study_area$ind_density_km2~ sp_density_ref_filter$ind_dens
      xlab= "Density reference zone",
      ylab= "Density study area"))
 
-print(paste("Correlation for", sp, "between", unique(sp_density_ref_filter$zone), "and study area =", round(as.numeric(cor$estimate), digits = 2),",","p=", round(as.numeric(cor$p.value), digits=3),",","n=", nrow(sp_density_ref_filter),sep =" "),  nrow(sp_density_ref_filter))
+print(paste("Correlation for", sp, "between", unique(sp_density_ref_filter$zone), "and study area =", round(as.numeric(cor$estimate^2), digits = 2),",","p=", round(as.numeric(cor$p.value), digits=3),",","n=", nrow(sp_density_ref_filter),sep =" "),  nrow(sp_density_ref_filter))
 
 #Build linear model regression model
 model <- lm(sp_density_study_area$ind_density_km2~sp_density_ref_filter$ind_density_km2)

### 2024-11-08 script/functions/nested_functions/spatial_extrapolation_species_density.R

@@ -9,6 +9,8 @@ library(dplyr)
 #spatial data manipulation
 library(sf)
 sf_use_s2(FALSE)
+#adjust colors
+library(scales)
 
 #Dependencies on function created in other scripts
 source("script/functions/standalone_functions/number_ind_transects.R")
@@ -114,42 +116,34 @@ transect_obs <- transect_obs %>%
   
 #---------------------------------------------------------------#
 #### Regression nest density - number individual on transect #### 
-#---------------------------------------------------------------#
-# Use point (0,0) and mean nest density and mean number of indivual or mean proportion of transect with at least an individual to fit a regression
-reg_df <- data.frame(
-  nest_density_km2= c(0,mean(sp_density_local$nest_density_km2)),
-  value= c(0, transect_obs[transect_obs$zone == zone_reference_transect,]$value))
-  
+#-------------------------------------------------------------
+#Extract points with nest density and transects
+reg_df <- transect_obs_ref %>% 
+  dplyr::mutate(year= as.integer(year)) %>% 
+  dplyr::left_join(sp_density_local, by= "year") %>% 
+  dplyr::select(year, value, nest_density_km2)
+
 #Build the linear model
-model <- lm(nest_density_km2~value, data= reg_df)
-print(paste("nest density (nest/km2)", "=", as.character(round(model$coefficients[2], digits=4)),"x", transect_metric,"+",as.character(round(model$coefficients[1]), digits= 4),sep= " "))
+model <- lm(nest_density_km2~value -1 , data= reg_df)
+summary(model)
+print(paste("nest density (nest/km2)", "=", as.character(round(model$coefficients, digits=4)),"x", transect_metric))
+
 
 #----------------------------#
 #### Visualize regression ####
 #----------------------------#
-#Keep year with transect and nest data
-trans_obs_filter <- transect_obs_ref[transect_obs_ref$year %in% sp_density_local$year,]$value
-sp_density_local_filter <- sp_density_local[sp_density_local$year %in% transect_obs_ref$year,]$nest_density_km2
-
 #Predict values based on the regression model
-predictions <- predict(model, newdata =data.frame(value= trans_obs_filter))
-
-# Calculate R squared as index of fit
[…]

### 2024-11-19 MetadataS1/MetadataS1.tex
update open research statement with dryad link

@@
 \textbf{Open Research statement}:\\
-The complete data set will be published as supporting information should the manuscript be accepted for publication. Metadata document, raw data and codes will also be archived in Dryad.\\
+The complete data set will be published as supporting information should the manuscript be accepted for publication.
+The data set, raw data, codes and metadata document are also archived in Dryad: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
 \endgroup
@@
                 \subsection*{1. Storage location and medium}
-The complete data set will be published as supporting information should the manuscript be accepted for publication. Metadata document, raw data and codes will also be archived in Dryad.
+The complete data set will be published as supporting information should the manuscript be accepted for publication.
+The data set, raw data, codes and metadata document are also archived in Dryad: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
 
@@
    		\subsection*{2. Location of completed data forms}
-        Raw data and codes will be archived in Dryad.
+        The raw data and codes are archived in Dryad: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
         \subsection*{3. Data entry verification procedures}
@@
 Packages} dplyr \citep{dplyr}, tidyr \citep{tidyr}, sf \citep{sf}, stringr \citep{stringr}, xtable \citep{xtable}, Distance \citep{miller2019}, ggplot2 \citep{ggplot2}, lme4 \citep{lme4}, AICcmodavg \citep{AICcmodavg}, scales\citep{scales}, ggmap \citep{ggmap}
-         \subsection*{4. Codes} Raw data and codes will be archived in Dryad.
+         \subsection*{4. Codes} The raw data and codes are archived in Dryad: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
 
@@
 Archival procedures} The complete data set will be published as supporting information should the manuscript be accepted for publication.
-Metadata document, raw data and codes will also be archived in Dryad.
+The data set, raw data, codes and metadata document are also archived in Dryad: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
          \subsection*{2. Redundant archival sites} None

### 2025-04-08 MetadataS1/MetadataS1.tex
add some comments of reviewer

@@
 
-Within terrestrial Arctic sites, the south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive biodiversity monitoring programs \citep{gauthier2024a}.
+The south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive vertebrate monitoring programs in the High-Arctic \citep{gauthier2024a}.
 Monitoring on Bylot Island began in 1989 with a focus on the snow goose and it gradually expanded to other species over time.
-Currently, the program encompasses all significant vertebrate species in the community with continuous monitoring spanning more than a decade \citep{gauthier2024a}.
+The program currently encompasses all vertebrate species in the community that have a significant influence on the food web dynamics \citep{gauthier2011, legagneux2012}, with continuous monitoring spanning more than a decade \citep{gauthier2024a}.
 Monitoring is also conducted at multiple spatial scales, including intensive and systematic observations conducted across a landscape spanning approximately 400 km\textsuperscript{2}.
@@
 Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August).
-This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration.
+This includes both breeding and non-breeding individuals that remain in the study area for a significant period of time (e.g., territorial breeding or non-breeding foxes), and excludes non-breeding individuals that stop only briefly during migration (e.g., shorebirds using Bylot Island for a short stopover).
+We focus on the breeding season here due to the absence of monitoring during the non-breeding season; however, we note that ecological dynamics during the non-breeding period, for both resident \citep{hutchison2020} and migratory species \citep{moisan2023}, can influence the food web during the breeding season.
 We focus on adults, except for lemmings for which we have not distinguished between juveniles and adults.
@@
                         \subsubsection*{a. Site type}
-  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.
+  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem compared to other sites at similar latitudes in the eastern Canadian High Arctic \citep{gauthier2024a}.
 An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}.
@@
 The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds \citep{gauthier2024a}.
-In summer, the study area received on average 77.5 mm of precipitation \citep{bylotclimate}.
-Additionnally, 102 days are frost-free annually on average  \citep{bylotclimate} and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
+In summer, the study area received on average 77.5 mm of precipitation.
+On average, snow depth is relatively thin in December, January, and February ($\sim$15 cm) and increases from March, reaching its peak in late May ($\sim$27 cm) before decreasing until mid-June ($\sim$3 cm) \citep{bylotclimate}.
+The study area typically remains free of snow from mid-June to late September \citep{gauthier2013} with an average of 102 frost-free days annually \citep{bylotclimate}.
+From September to December, snow depth typically increases continuously \citep{bylotclimate}.
+The summary statistics presented here on the climate of the South Plain of Bylot Island are derived from automatic weather stations located in the Qarlikturvik Valley, which have been recording data on air temperature, thawing degree days, frost-free days, summer precipitation, snow thickness, and wind speed since 1994 \citep{bylotclimate}.
+Raw data from these weather stations are publicly accessible at \url{https://nordicana.cen.ulaval.ca/fr/publication.php?doi=45039SL-EE76C1BDAADC4890}.
 \newpage
@@
 To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled.
-The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts.
+The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts \textbf{Figure \ref{figure:sngo_interannual_variation}}).
 Systematic nest searches were highly precise, but covered a relatively small area, whereas transects and snow goose point counts were less precise but covered larger areas.
 For each method, we calculated the mean density of breeding individuals in mesic habitat by dividing the number of birds (or nests) recorded by the area sampled.
-Despite methodological differences, the three approaches showed similar inter-annual variations, supporting the use of a mean values to estimate nest density in mesic habitat \textbf{Figure \ref{figure:sngo_interannual_variation}}).
+We found a strong and statistically significant correlation between density estimates derived from vertebrate count transects and snow goose point counts (Sperman's rank correlation r\textsubscript{s}= 0.8; p=0.003).
+However, there was no significant correlation between density estimates from nest sampling plots and either the transect data (r\textsubscript{s}=0.37; p=0.24) or the point counts (r\textsubscript{s}=0.22; p= 0.5).
+Despite the lack of correlation between observations (transects and point counts) and nest sampling, we chose to average the estimates from all three methods since observations and nest monitoring present complementary strengths and limitations.
+Transects and point counts offer broader spatial coverage but lower precision, while nest sampling provides higher precision but limited spatial coverage.
+Averaging the three methods allows us to balance these trade-offs and produce estimates that are less precise individually, but likely more accurate overall.
 Lastly, to transform the densities in total abundance, we determined the annual proportion of wetland and mesic habitats within the goose colony and multiplied the area of each habitat by the density of breeding individuals.
@@
 Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat.
+Nest site selection by rock ptarmigan has not been well documented in the Canadian Arctic; however, a study conducted at Sarcpa Lake (n= 7 nests) also highlight that rock ptarmigan can nests in a variety of habitats \citep{montgomerie1983}.
 To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest).
@@
                 We estimated the abundance of Arctic foxes in the study area based on their estimated home range size inside and outside the goose colony.
-We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size.
+We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size of territorial foxes.
 However, here, we did not account for annual variations in lemming density as presented in \citet{dulude2023} in order to obtain mean fox home range.
@@
    \section*{Acknowledgements}
-We are deeply appreciative of the extensive data gathered over the decades by generations students and researchers during their fieldwork at the Bylot Island research station, which made this project possible.
+We are deeply appreciative of the extensive data gathered over the decades by generations of students and researchers during their fieldwork at the Bylot Island research station, which made this project possible.
 Our gratitude also extends to the Center of Northern Studies for providing research facilities, as well as to the Polar Continental Shelf Program and Sirmilik National Park (Parks Canada) for their logistical support throughout the years.

### 2025-04-14 MetadataS1/MetadataS1.tex
minor changes from reviewers comments

@@
         \subsection*{2. Experimental or sampling design}
+Research installations on Bylot Island are typically active from mid-May to the 20th of August due to logistical constraints.
+The period therefore covers the arrival of migratory species and the entire breeding cycle for most species.
+However, the sampling period does not encompasses the departure of some species (e.g., xxxxx).
+
                         \subsubsection*{a. Permanent plots}
@@
 From 2010 to 2023, observers walked 500-meter linear transects where all vertebrate individuals observed within 150 meters on either side were counted (146 to 320 transects per year).
-Transects were distributed across all lowland zones of the study area, typically in mesic habitat, and were carried out during the nesting period (between June 21 and July 14; \citet{duchesne2021}; \textbf{Figure \ref{figure:transects}}).
+Transects were distributed across all lowland zones of the study area, typically in mesic habitat, and were carried out during the nesting period (between June 21 and July 14; \citet{lamarre2017, duchesne2021}; \textbf{Figure \ref{figure:transects}}).
+As a calibration, all observers were trained with a rangefinder to estimate a distance of 150 m before starting data collection.
 Furthermore, specifically for American golden-plovers, we measured the distance of each observed individual to the transect path.
@@
 On average, 613 \textpm{142} point counts were sampled each year, covering an area of 30 \textpm{7} km\textsuperscript{2}.
+Observers were trained with a rangefinder to estimate a distance of 125 m before starting data collection.
 \\
@@
 Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}.
+Observations are made from mid-May to the late August, but the effort is highest during the late June and early July.
 The number of hours spent in the field served as a proxy for the sampling effort.
